### 2019-12-12 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Initial commit with draft of paper (introduction)

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Some Journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Modelling participant interactions in crash severity analysis},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)
% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Modelling participant interactions in crash severity analysis}
    \author[Some University]{Author 1\corref{c1}}
   \ead{a1@example.com} 
   \cortext[c1]{Corresponding Author}
    \author[Some Institute of Technology]{Author 2}
   \ead{a2@example.com}

      \address[Some University]{Department, Street, City, State, Zip}
    \address[Some Institute of Technology]{Department, Street, City, State, Zip}

  \begin{abstract}
  Road safety continues to impose an important burden on health and the
  economy. Numerous efforts to understand the factors that affect road
  safety have been undertaken. One stream of research focus on modelling
  the severity of crashes. Crash severity research is useful to clarify
  the way different factors can influence an outcome where there are no
  injuries, or injuries, or fatalities. The objective of this paper is to
  describe an approach to model the interactions between participants in a
  crash, in the context of crashes involving two parties. Using data from
  Canada's National Collision Database we demonstrate the approach. Three
  levels of crash severity (no injury/injury/fatality) are analyzed using
  ordered logit models and covariates for the participants in the crash
  and the conditions of the crash. The results suggest that there are
  important interactions between the various descriptors of the
  participants in the crash. More generally, the example shows how
  modelling the interactions between participants explicitly leads to
  richer insights into the covariates of crash severity, in addition to
  better model fits. The approach described in this paper is intuitive,
  simple to implement, and can be adapted to advanced modelling methods
  with ease.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Road safety continues to be a concern world-wide. According to a recent
report from the World Health Organization (2019), road accidents are the
8th leading cause of death for all ages, and the number one cause of
death for children and young people between the ages of 5 to 29. Of all
leading causes of death, road accidents are the only non-external cause,
unrelated to disease, disorder, or infection. Road accidents impose a
heavy burden on individuals and society as a whole. Gobally, there has
been an increasing trend in number of vehicles and road accident-related
deaths, even if the rate of death per 100,000 population and 100,000
vehicles have both fallen (World Health Organization, 2019, Figs. 1 and
2). These gains, although they are to be celebrated, cannot distract
from the crippling economic cost of premature death (e.g., Symons et
al., 2019; Wijnen et al., 2019), not to mention the long-term
consequences for survivors, measured in sometimes crushing emotional and
physical pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier
et al., n.d.).

As the statistics suggest, the burden of road accidents is also not
borne evenly. There are important disparities at the international
level, where the odds of death due to a road crash are three times
higher in low-income countries compared to high-income countries; in
fact, no reductions in road accident-related fatalities were appreciated
in low-income countries between 2013 and 2016 (World Health
Organization, 2019). In the case of high-income countries, where
substantial gains in road safety have been observed for years, said
gains have been unevenly distributed; thus, while fatal crashes
involving older adults in the United States and Great Britain declined
between 1997 and 2010 (despite an increase in the number of older
people), the trend remained stable or increased slightly in Australia in
roughly the same period (Thompson et al., 2018). There are also
systematic differences in the impact of road accidents. For example, in
a study in the United States, Obeng (2011) reported that the impact of
covariates of crash severity varied in magnitude by gender, and in some
cases were not even the same. More recently, Regev et al.~(2018) used
adjusted crash risk to find that the risk of crashes in Great Britain
peaked for people 21 to 29 years of age; on the other hand, the risk of
fatal injuries for older drivers was constant, irrespective of the
seriousness of the crash - which highlights the perils of accidents at
older ages. Other studies have concentrated on the consequences of road
accidents on the young (e.g., Peek-Asa et al., 2010), the old (e.g.,
Rakotonirainy et al., 2012), as well as pedestrians and cyclists (e.g.,
Hanson et al., 2013; McArthur et al., 2014).

Given the relevance and cost of the matter, as well as the important
variations of the impacts among different population segments, numerous
efforts efforts exist to better understand the factors that affect road
safety - including the probable consequences of the crash. Consequently,
a stream of research in the analysis of road accidents is concerned with
the severity of crashes. In particular, multivariate analysis of crash
severity is a useful way to clarify the way various factors can affect
the outcome of an incident, to discriminate between various levels of
injury, from no injury (i.e., property damage only), to different
degrees of injury up to and including fatality. This is an active area
of research (Savolainen et al., 2011), and one where methodological
developments have aimed at improving the reliability, accuracy, and
precision of models.

This paper aspires to contribute to the literature on crash severity by
describing an approach to model the interactions between participants in
a crash, in the context of incidents involving two parties. The
importance of these interactions has been recognized in the existing
literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al.,
2017), and in this paper we present a systematic approach to model the
way participants in a crash interact to influence the severity of the
accident for each party. By way of application, we demonstrate the
approach through an empirical analysis of data drawn from Canada's
National Collision Database. Three levels of crash severity (no
injury/injury/fatality) are analyzed using ordered logit models and
covariates for the participants in the crash and the conditions of the
crash. The results suggest that there are important interactions between
the various descriptors of the participants in the crash. More
generally, the example shows how modelling the interactions between
participants explicitly leads to richer insights into the covariates of
crash severity, in addition to better model fits. The approach described
in this paper is intuitive, simple to implement, and can be adapted to
advanced modelling methods with ease.

The rest of this paper is structure as follows. In Section
\ref{sec:review-of-methods} we present a concise review of the methods
used to analyze crash severity, with a particular focus on techniques
that consider the interactions between participants in a crash.

\hypertarget{sec:review-of-methods}{%
\section{Methodological approaches in crash severity
analysis}\label{sec:review-of-methods}}

Methods:

Iranitalab and Khattak (2017); Khan et al. (2015); Chang and Wang
(2006): machine learning Quddus et al. (2010): ordinal models Kadilar
(2016): conditional logistic regression Mooradian et al. (2013): partial
proportional odds models Ma et al. (2008): counts

Interplay between participants:

Chiou et al. (2013)

Lee and Li (2014) consider effect on crash severity of interactions
between different types of vehicles. This they do by subsetting the
dataset and estimating independent models for each subset of data. Since
they consider three types of vehicles, namely cars (C), light trucks
(L), and heavy trucks (H), they work with nine datasets, for each type
of interactions (i.e., C-C, C-L, C-H, and so on).

Li et al. (2017)

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Words go here.

\hypertarget{illustrative-application}{%
\section{Illustrative application}\label{illustrative-application}}

Words go here.

\hypertarget{further-considerations}{%
\section{Further considerations}\label{further-considerations}}

Words go here.

\hypertarget{concluding-remarks}{%
\section{Concluding remarks}\label{concluding-remarks}}

Words go here.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Chang2006analysis}{}%
Chang, L.Y., Wang, H.W., 2006. Analysis of traffic injury severity: An
application of non-parametric classification tree techniques. Accident
Analysis and Prevention 38, 1019--1027.
doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}

\leavevmode\hypertarget{ref-Chiou2013modeling}{}%
Chiou, Y.C., Hwang, C.C., Chang, C.C., Fu, C., 2013. Modeling
two-vehicle crash severity by a bivariate generalized ordered probit
approach. Accident Analysis and Prevention 51, 175--184.
doi:\href{https://doi.org/10.1016/j.aap.2012.11.008}{10.1016/j.aap.2012.11.008}

\leavevmode\hypertarget{ref-Devlin2019road}{}%
Devlin, A., Beck, B., Simpson, P.M., Ekegren, C.L., Giummarra, M.J.,
Edwards, E.R., Cameron, P.A., Liew, S., Oppy, A., Richardson, M., Page,
R., Gabbe, B.J., 2019. The road to recovery for vulnerable road users
hospitalised for orthopaedic injury following an on-road crash. Accident
Analysis and Prevention 132, 10.
doi:\href{https://doi.org/10.1016/j.aap.2019.105279}{10.1016/j.aap.2019.105279}

\leavevmode\hypertarget{ref-Hanson2013severity}{}%
Hanson, C.S., Noland, R.B., Brown, C., 2013. The severity of pedestrian
crashes: An analysis using google street view imagery. Journal of
Transport Geography 33, 42--53.
doi:\href{https://doi.org/10.1016/j.jtrangeo.2013.09.002}{10.1016/j.jtrangeo.2013.09.002}

\leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
Iranitalab, A., Khattak, A., 2017. Comparison of four statistical and
machine learning methods for crash severity prediction. Accident
Analysis and Prevention 108, 27--36.
doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}

\leavevmode\hypertarget{ref-Kadilar2016effect}{}%
Kadilar, G.O., 2016. Effect of driver, roadway, collision, and vehicle
characteristics on crash severity: A conditional logistic regression
approach. International Journal of Injury Control and Safety Promotion
23, 135--144.
doi:\href{https://doi.org/10.1080/17457300.2014.942323}{10.1080/17457300.2014.942323}

\leavevmode\hypertarget{ref-Khan2015exploring}{}%
Khan, G., Bill, A.R., Noyce, D.A., 2015. Exploring the feasibility of
classification trees versus ordinal discrete choice models for analyzing
crash severity. Transportation Research Part C-Emerging Technologies 50,
86--96.
doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}

\leavevmode\hypertarget{ref-Lee2014analysis}{}%
Lee, C., Li, X.C., 2014. Analysis of injury severity of drivers involved
in single- and two-vehicle crashes on highways in ontario. Accident
Analysis and Prevention 71, 286--295.
doi:\href{https://doi.org/10.1016/j.aap.2014.06.008}{10.1016/j.aap.2014.06.008}

\leavevmode\hypertarget{ref-Li2017interplay}{}%
Li, L., Hasnine, M.S., Habib, K.M.N., Persaud, B., Shalaby, A., 2017.
Investigating the interplay between the attributes of at-fault and
not-at-fault drivers and the associated impacts on crash injury
occurrence and severity level. Journal of Transportation Safety \&
Security 9, 439--456.
doi:\href{https://doi.org/10.1080/19439962.2016.1237602}{10.1080/19439962.2016.1237602}

\leavevmode\hypertarget{ref-Ma2008multivariate}{}%
Ma, J.M., Kockelman, K.M., Damien, P., 2008. A multivariate
poisson-lognormal regression model for prediction of crash counts by
severity, using bayesian methods. Accident Analysis and Prevention 40,
964--975.
doi:\href{https://doi.org/10.1016/j.aap.2007.11.002}{10.1016/j.aap.2007.11.002}

\leavevmode\hypertarget{ref-McArthur2014spatial}{}%
McArthur, A., Savolainen, P.T., Gates, T.J., 2014. Spatial analysis of
child pedestrian and bicycle crashes development of safety performance
function for areas adjacent to schools. Transportation Research Record
57--63. doi:\href{https://doi.org/10.3141/2465-08}{10.3141/2465-08}

\leavevmode\hypertarget{ref-Merlin2007stress}{}%
Merlin, E.P.R., Gonzalez-Forteza, C., Lira, L.R., Tapia, J.A.J., 2007.
Post-traumatic stress disorder in patients with non intentional injuries
caused by road traffic accidents. Salud Mental 30, 43--48.

\leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
Mooradian, J., Ivan, J.N., Ravishanker, N., Hu, S., 2013. Analysis of
driver and passenger crash injury severity using partial proportional
odds models. Accident Analysis and Prevention 58, 53--58.
doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}

\leavevmode\hypertarget{ref-Obeng2011gender}{}%
Obeng, K., 2011. Gender differences in injury severity risks in crashes
at signalized intersections. Accident Analysis and Prevention 43,
1521--1531.
doi:\href{https://doi.org/10.1016/j.aap.2011.03.004}{10.1016/j.aap.2011.03.004}

\leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
Peek-Asa, C., Britton, C., Young, T., Pawlovich, M., Falb, S., 2010.
Teenage driver crash incidence and factors influencing crash injury by
rurality. Journal of Safety Research 41, 487--492.
doi:\href{https://doi.org/10.1016/j.jsr.2010.10.002}{10.1016/j.jsr.2010.10.002}

\leavevmode\hypertarget{ref-Pelissier2019medical}{}%
Pelissier, C., Fort, E., Fontana, L., Hours, M., n.d. Medical and
socio-occupational predictive factors of psychological distress 5 years
after a road accident: A prospective study. Social Psychiatry and
Psychiatric Epidemiology 13.
doi:\href{https://doi.org/10.1007/s00127-019-01780-0}{10.1007/s00127-019-01780-0}

\leavevmode\hypertarget{ref-Quddus2010road}{}%
Quddus, M.A., Wang, C., Ison, S.G., 2010. Road traffic congestion and
crash severity: Econometric analysis using ordered response models.
Journal of Transportation Engineering-Asce 136, 424--435.
doi:\href{https://doi.org/10.1061/(asce)te.1943-5436.0000044}{10.1061/(asce)te.1943-5436.0000044}

\leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
Rakotonirainy, A., Steinhardt, D., Delhomme, P., Darvell, M., Schramm,
A., 2012. Older drivers' crashes in queensland, australia. Accident
Analysis and Prevention 48, 423--429.
doi:\href{https://doi.org/10.1016/j.aap.2012.02.016}{10.1016/j.aap.2012.02.016}

\leavevmode\hypertarget{ref-Regev2018crash}{}%
Regev, S., Rolison, J.J., Moutari, S., 2018. Crash risk by driver age,
gender, and time of day using a new exposure methodology. Journal of
Safety Research 66, 131--140.
doi:\href{https://doi.org/10.1016/j.jsr.2018.07.002}{10.1016/j.jsr.2018.07.002}

\leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
Savolainen, P.T., Mannering, F., Lord, D., Quddus, M.A., 2011. The
statistical analysis of highway crash-injury severities: A review and
assessment of methodological alternatives. Accident Analysis and
Prevention 43, 1666--1676.
doi:\href{https://doi.org/10.1016/j.aap.2011.03.025}{10.1016/j.aap.2011.03.025}

\leavevmode\hypertarget{ref-Symons2019reduced}{}%
Symons, J., Howard, E., Sweeny, K., Kumnick, M., Sheehan, P., 2019.
Reduced road traffic injuries for young people: A preliminary investment
analysis. Journal of Adolescent Health 65, S34--S43.
doi:\href{https://doi.org/10.1016/j.jadohealth.2019.01.009}{10.1016/j.jadohealth.2019.01.009}

\leavevmode\hypertarget{ref-Thompson2018trends}{}%
Thompson, J.P., Baldock, M.R.J., Dutschke, J.K., 2018. Trends in the
crash involvement of older drivers in australia. Accident Analysis and
Prevention 117, 262--269.
doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}

\leavevmode\hypertarget{ref-Wijnen2019analysis}{}%
Wijnen, W., Weijermars, W., Schoeters, A., Berghe, W. van den, Bauer,
R., Carnis, L., Elvik, R., Martensen, H., 2019. An analysis of official
road crash cost estimates in european countries. Safety Science 113,
318--327.
doi:\href{https://doi.org/10.1016/j.ssci.2018.12.004}{10.1016/j.ssci.2018.12.004}

\leavevmode\hypertarget{ref-WHO2019global}{}%
World Health Organization, 2019. Global status report on road safety
2018 (2018). Geneva.

\end{document}

### 2019-12-12 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Introduction of manuscript done, now working on background review

@@
 used to analyze crash severity, with a particular focus on techniques
-that consider the interactions between participants in a crash.
+that consider the interactions between participants in a crash. Section
+\ref{sec:methods} describes the data conditions, preprocessing, and
+approach for incorporating participant interactions in models of crash
+severity. An illustrative application of the modelling strategy
+described in this paper appears in \ref{sec:application}. We then
+present some additional thoughts about the applicability of this
+approach in Section \ref{sec:further-considerations} before offering
+some concluding remarks in Section \ref{sec:concluding-remarks}.
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\hypertarget{sec:methods}{%
+\section{Methods}\label{sec:methods}}
+
+\hypertarget{data-considerations}{%
+\subsection{Data considerations}\label{data-considerations}}
+
+Words go here
+
+\hypertarget{model-specification}{%
+\subsection{Model specification}\label{model-specification}}
+
+More words go here
+
+\hypertarget{sec:application}{%
+\section{Illustrative application}\label{sec:application}}
+
+\hypertarget{data-for-the-application}{%
+\subsection{Data for the application}\label{data-for-the-application}}
 
@@
 
-\hypertarget{illustrative-application}{%
-\section{Illustrative application}\label{illustrative-application}}
+\hypertarget{analysis-and-results}{%
+\subsection{Analysis and results}\label{analysis-and-results}}
 
-Words go here.
+Words go here
 
-\hypertarget{further-considerations}{%
-\section{Further considerations}\label{further-considerations}}
+\hypertarget{sec:further-considerations}{%
+\section{Further considerations}\label{sec:further-considerations}}
 
-Words go here.
+Here I plan to discuss the applicability of the modelling strategy to
+advanced modelling techniques (partial proportional odds, heterogeneity,
+hierarchical models, etc.)
 
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
+\hypertarget{sec:concluding-remarks}{%
+\section{Concluding remarks}\label{sec:concluding-remarks}}
 

### 2020-01-02 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Started putting the analysis in the R Markdown document

@@
             pdfauthor={},
-            pdftitle={Modelling participant interactions in crash severity analysis},
+            pdftitle={An assessment of strategies to model opponent effects in crash severity analysis},
             colorlinks=false,
@@
 % Pandoc header
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\usepackage[margin=1in]{geometry}
 
@@
 
-  \title{Modelling participant interactions in crash severity analysis}
+  \title{An assessment of strategies to model opponent effects in crash severity
+analysis}
     \author[Some University]{Author 1\corref{c1}}
@@
   \begin{abstract}
-  Road safety continues to impose an important burden on health and the
-  economy. Numerous efforts to understand the factors that affect road
-  safety have been undertaken. One stream of research focus on modelling
-  the severity of crashes. Crash severity research is useful to clarify
-  the way different factors can influence an outcome where there are no
-  injuries, or injuries, or fatalities. The objective of this paper is to
-  describe an approach to model the interactions between participants in a
-  crash, in the context of crashes involving two parties. Using data from
-  Canada's National Collision Database we demonstrate the approach. Three
+  Road accidents impose an important burden on health and the economy.
+  Numerous efforts to understand the factors that affect road collisions
+  have been undertaken. One stream of research focus on modelling the
+  severity of crashes. Crash severity research is useful to clarify the
+  way different factors can influence the outcome of the event. The
+  objective of this paper is to assess different strategies to model the
+  interactions between participants in a crash, in the context of crashes
+  involving two parties. Towards this objective, a series of models are
+  estimated using data from Canada's National Collision Database. Three
   levels of crash severity (no injury/injury/fatality) are analyzed using
   ordered logit models and covariates for the participants in the crash
-  and the conditions of the crash. The results suggest that there are
-  important interactions between the various descriptors of the
-  participants in the crash. More generally, the example shows how
-  modelling the interactions between participants explicitly leads to
-  richer insights into the covariates of crash severity, in addition to
-  better model fits. The approach described in this paper is intuitive,
-  simple to implement, and can be adapted to advanced modelling methods
-  with ease.
+  and the conditions of the crash. Modelling strategies include different
+  ways of introducing the covariates (e.g., in a single-level or
+  multi-level form), as well as by subsetting the dataset. The models are
+  assessed using predicted shares and outcomes, and the results highlight
+  the importance of considering opponent effects in crash severity
+  analysis. On the other hand, the study suggests that hierarchical (i.e.,
+  multi-level) specifications and subsetting do not perform necessarily
+  better than a relatively simple single-level model with opponent
+  effects.
   \end{abstract}
@@
 death for children and young people between the ages of 5 to 29. Of all
-leading causes of death, road accidents are the only non-external cause,
+leading causes of death, road accidents are the only cause of death
 unrelated to disease, disorder, or infection. Road accidents impose a
-heavy burden on individuals and society as a whole. Gobally, there has
-been an increasing trend in number of vehicles and road accident-related
-deaths, even if the rate of death per 100,000 population and 100,000
-vehicles have both fallen (World Health Organization, 2019, Figs. 1 and
-2). These gains, although they are to be celebrated, cannot distract
-from the crippling economic cost of premature death (e.g., Symons et
-al., 2019; Wijnen et al., 2019), not to mention the long-term
-consequences for survivors, measured in sometimes crushing emotional and
-physical pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier
-et al., n.d.).
+heavy burden on individuals and society as a whole. Gobally, the rate of
+road collision-related deaths per 100,000 population and 100,000
+vehicles have both fallen, even as the number of vehicles has grown
+(World Health Organization, 2019, Figs. 1 and 2). These gains, although
+they are to be celebrated, cannot distract from the crushing economic
+cost of premature death (e.g., Symons et al., 2019; Wijnen et al.,
+2019), not to mention the long-term consequences for survivors, measured
+in sometimes crippling emotional and physical pain (e.g., Merlin et al.,
+2007; Devlin et al., 2019; Pelissier et al., n.d.).
 
-As the statistics suggest, the burden of road accidents is also not
-borne evenly. There are important disparities at the international
-level, where the odds of death due to a road crash are three times
-higher in low-income countries compared to high-income countries; in
-fact, no reductions in road accident-related fatalities were appreciated
-in low-income countries between 2013 and 2016 (World Health
-Organization, 2019). In the case of high-income countries, where
+Evidence from across the world suggests that the burden of road
+accidents is not borne evenly. There are important disparities at the
+international level, where the odds of death due to road crashes are
+three times higher in low-income countries compared to high-income
+countries; in fact, no reductions in road accident-related fatalities
+were appreciated in low-income countries between 2013 and 2016 (World
+Health Organization, 2019). In the case of high-income countries, where
 substantial gains in road safety have been observed for years, said
-gains have been unevenly distributed; thus, while fatal crashes
+gains have also been unevenly distributed; thus, while fatal crashes
 involving older adults in the United States and Great Britain declined
-between 1997 and 2010 (despite an increase in the number of older
-people), the trend remained stable or increased slightly in Australia in
-roughly the same period (Thompson et al., 2018). There are also
-systematic differences in the impact of road accidents. For example, in
-a study in the United States, Obeng (2011) reported that the impact of
-covariates of crash severity varied in magnitude by gender, and in some
-cases were not even the same. More recently, Regev et al.~(2018) used
-adjusted crash risk to find that the risk of crashes in Great Britain
-peaked for people 21 to 29 years of age; on the other hand, the risk of
-fatal injuries for older drivers was constant, irrespective of the
-seriousness of the crash - which highlights the perils of accidents at
-older ages. Other studies have concentrated on the consequences of road
-accidents on the young (e.g., Peek-Asa et al., 2010), the old (e.g.,
-Rakotonirainy et al., 2012), as well as pedestrians and cyclists (e.g.,
-Hanson et al., 2013; McArthur et al., 2014).
+between 1997 and 2010 (despite the graying of the population), the trend
+remained stable or increased slightly in Australia in roughly the same
+period (Thompson et al., 2018). There are also systematic differences in
+the impact of road accidents. For example, in a study in the United
+States, Obeng (2011) reported that the impact of covariates of crash
+severity varied substantially in magnitude by gender. More recently,
+Regev et al.~(2018) used adjusted crash risk to find that the risk of
+crashes in Great Britain peaked for people 21 to 29 years of age; on the
+other hand, the risk of fatal injuries for older drivers was constant,
+irrespective of the seriousness of the crash - which highlights the
+perils of accidents at older ages. Other studies have concentrated on
+the consequences of road accidents for the young (e.g., Peek-Asa et al.,
+2010), the old (e.g., Rakotonirainy et al., 2012), as well as
+pedestrians and cyclists (e.g., Hanson et al., 2013; McArthur et al.,
+2014).
 
-Given the relevance and cost of the matter, as well as the important
+Given the relevance and cost of this matter, as well as the important
 variations of the impacts among different population segments, numerous
-efforts efforts exist to better understand the factors that affect road
-safety - including the probable consequences of the crash. Consequently,
-a stream of research in the analysis of road accidents is concerned with
-the severity of crashes. In particular, multivariate analysis of crash
-severity is a useful way to clarify the way various factors can affect
-the outcome of an incident, to discriminate between various levels of
-injury, from no injury (i.e., property damage only), to different
-degrees of injury up to and including fatality. This is an active area
-of research (Savolainen et al., 2011), and one where methodological
-developments have aimed at improving the reliability, accuracy, and
-precision of models.
+efforts efforts have been conducted to better understand the factors
+that affect road safety - including the probable consequences of
+crashes. Consequently, a stream of research in the analysis of road
+accidents is concerned with the severity of crashes. In particular,
+multivariate analysis of crash severity is a useful way to clarify the
+way various factors can affect the outcome of an incident, to
+discriminate between various levels of injury, from no injury (i.e.,
+property damage only), to different degrees of injury up to and
+including fatality. This is an active area of research (e.g., Savolainen
+et al., 2011), and one where methodological developments have aimed at
+improving the reliability, accuracy, and precision of models.
 
 This paper aspires to contribute to the literature on crash severity by
-describing an approach to model the interactions between participants in
-a crash, in the context of incidents involving two parties. The
-importance of these interactions has been recognized in the existing
-literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al.,
-2017), and in this paper we present a systematic approach to model the
-way participants in a crash interact to influence the severity of the
-accident for each party. By way of application, we demonstrate the
-approach through an empirical analysis of data drawn from Canada's
-National Collision Database. Three levels of crash severity (no
-injury/injury/fatality) are analyzed using ordered logit models and
-covariates for the participants in the crash and the conditions of the
-crash. The results suggest that there are important interactions between
-the various descriptors of the participants in the crash. More
-generally, the example shows how modelling the interactions between
-participants explicitly leads to richer insights into the covariates of
-crash severity, in addition to better model fits. The approach described
-in this paper is intuitive, simple to implement, and can be adapted to
-advanced modelling methods with ease.
+assessing different modelling strategies useful to incorporate opponent
+effects in crash analysis, in the context of incidents involving two
+parties. The importance of these interactions has been recognized in the
+existing literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et
+al., 2017; Tarrao et al., 2014), and a number of different modelling
+strategies have been proposed. In this paper we present a systematic
+assessment of several relevant modelling strategies, ranging from the
+way variables are defined in single-level models, in multi-level models
+(i.e., hierarchical models), as well as using data subsetting
+approaches. For the assessment we use data from Canada's National
+Collision Database, a database that collects all police-reported
+collisions in the country. Using the most recent version of the dataset
+(2017), three levels of crash severity (no injury/injury/fatality) are
+analyzed using ordered logit models and covariates for the participants
+in the crash and the conditions of the crash. For model assessment, we
+conduct an in-sample prediction exercise using the estimation sample
+(i.e., \emph{nowcasting}), and also an out-of-sample prediction exercise
+using the dataset corresponding to 2016 (i.e., \emph{backcasting}). The
+models are assessed using predicted shares and predicted individual
+outcomes using an extensive array of verification statistics. The
+results highlight the importance of considering opponent effects in
+crash severity analysis to improve the goodness-of-fit and predictive
+performance o. On the other hand, the study suggests that hierarchical
+variable specifications and subsetting do not perform necessarily better
+than a relatively simple single-level model with opponent effects.
 
-The rest of this paper is structure as follows. In Section
+The rest of this paper is structured as follows. In Section
 \ref{sec:review-of-methods} we present a concise review of the methods
@@
 that consider the interactions between participants in a crash. Section
-\ref{sec:methods} describes the data conditions, preprocessing, and
-approach for incorporating participant interactions in models of crash
-severity. An illustrative application of the modelling strategy
-described in this paper appears in \ref{sec:application}. We then
-present some additional thoughts about the applicability of this
-approach in Section \ref{sec:further-considerations} before offering
-some concluding remarks in Section \ref{sec:concluding-remarks}.
+\ref{sec:methods} describes the data requirements, data preprocessing,
+and the modelling strategy. Model estimation is presented in
+\ref{sec:application} and the assessment of models is in Section
+\ref{sec:assessment}. We then present some additional thoughts about the
+applicability of this approach in Section
+\ref{sec:further-considerations} before offering some concluding remarks
+in Section \ref{sec:concluding-remarks}.
 
@@
 
-Methods:
+Modelling the outcomes of crashes in terms of the severity of injuries
+to participants has been a preoccupation of transportation researchers,
+planners, auto insurance companies, governments and the public for
+decades. One of the earliests studies to investigate the severity of
+injuries conditional on an accident having occurred was by White and
+Clayton (1972). Kim et al.~(1995) later stated that the ``linkages
+between severity of injury and driver characteristics and behaviors have
+not been thoroughly investigated'' (p.~470). Nowadays, there is a
+burgeoning literature on this subject, including methodological
+developments, case studies, and more niche research with a focus on
+particular situations (e.g., crashes at intersections, Mussone et al.,
+2017; crashes in rural roads, Gong and Fan, 2017) or special populations
+(e.g., crashes involving motorcyclists or active travelers; see Shaheed
+et al., 2013; Salon and McIntyre, 2018).
 
-Iranitalab and Khattak (2017); Khan et al. (2015); Chang and Wang
-(2006): machine learning Quddus et al. (2010): ordinal models Kadilar
-(2016): conditional logistic regression Mooradian et al. (2013): partial
-proportional odds models Ma et al. (2008): counts
+Crash severity is often modelled using models for discrete outcomes,
+including classification techniques from machine learning (e.g.,
+Iranitalab and Khattak, 2017; Chang and Wang, 2006; Effati et al., 2015;
+Khan et al., 2015), Poisson models for counts (e.g., Ma et al., 2008),
+unordered logit/probit models {[}{]}, as well as ordered logit/probit
+models (e.g., Rifaat and Chin, 2007), with numerous variants, such as
+random parameters/mixed logit (e.g., Aziz et al., 2013; Haleem and Gan,
+2013), partial proportional odds models (e.g., Mooradian et al., 2013;
+Sasidharan and Menendez, 2014), and the use of copulas (e.g., Wang et
+al., 2015). Recent reviews of methods include Savolainen et al.~(2011)
+and Shamsunnahar and Eluru (2013). Shamsunnahar and Eluru (2013) in
+particular conducted an extensive comparison of models for discrete
+outcomes and found that while the difference between the performance of
+unordered models and ordered models was so small as to make not
+difference, ordered models are usually more parsimonious since only one
+latent functions needs to be estimated for all outcomes, as opposed to
+one for each outcome in unordered modelling mechanisms.
+
+Irrespective of the modelling framework employed, models of crash
+severity often include variables in several categories, as shown with
+examples in Table \ref{tab:table-variable-categories} (also see Montella
+et al., 2013). Many crash databases (but not by any means all) also
+account for the multievent nature of many crashes. In this way, there
+are crashes that involve a single traffic unit (e.g., Kim et al., 2013;
+Gong and Fan, 2017), others that involve two traffic units (e.g., Tarrao
+et al., 2014; Wang et al., 2015), and more rarely there are
+multi-traffic unit crashes (e.g., Wu et al., 2014; Bogue et al., 2017).
+Likewise, each traffic unit can possibly involve more than one person
+(e.g., driver and passengers). Thus, depending on the incident, each
+record in a crash database may include unique identifier for the crash,
+as well as identifiers for the traffic units (or identifiers for dummy
+objects such as a light pole that was hit by a vehicle), and the people
+involved in the crash.
+
+\begin{table}
+
+\caption{\label{tab:table-variable-categories}\label{tab:table-variable-categories}Categories of variables used in the analysis of crash severity with examples}
+\centering
+\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{22em}}
+\toprule
+Category & Examples\\
+\midrule
+Person-related & Attributes of participants in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
+Traffic unit-related & Attributes of the traffic unit, e.g., type of traffic unit (car, motorcycle, etc.), maneouver, etc.\\
+Crash-related & Attributes of the crash, e.g., location, weather conditions, light conditions, number of parties, etc.\\
+Road-related & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 (L), and heavy trucks (H), they work with nine datasets, for each type
-of interactions (i.e., C-C, C-L, C-H, and so on).
+of interactions (i.e., C-C, C-L, C-H, and so on). Tarrao et al. (2014)
+
+Mannering et al. (2016)
 
 Li et al. (2017)
+
+Salon and McIntyre (2018)
+
+Chen et al. (2019)
+
+Wu et al. (2014) two-vehicle crashes do not really consider the
+interactions.
+
+Many models use a latent-variable approach, whereby the severity of the
+crash (observed) is linked to an underlying latent variable that is a
+function of the variables: \[
+y_{ik}^*=\sum_{l=1}\beta_lp_{ikl} + \sum_{m=1}\gamma_mc_{km} + \epsilon_{ik}
+\]
+
+The left-hand side of the expression above (\(y_{ik}^*\)) is a latent
+(unobservable) variable that is associated with the severity of crash
+\(k\) (\(k=1,\cdots,K\)) for participant \(i\) (\(i = 1,\cdots,n\)). The
+right-hand side of the expression is split in three parts. The first
+part collects \(l=1,\cdots,L\) individual attributes \(p\) for
+participant \(i\) in crash \(k\); these could relate to the person
+(e.g., age and gender) or be individual attributes of the traffic unit
+(e.g., maneuver or vehicle type). The second part collects
+\(m=1,\cdots,M\) attributes \(c\) related to the crash \(k\), including
+crash-related and road-related data. The last part is a random term
+specific to participant \(i\) in crash \(k\).
+
+The latent variable is not observed directly, but it is possible to
+posit a probabilistic relationship with the outcome \(y_{ik}\) (the
+severity of crash \(k\) for participant \(i\)). Depending on the
+characteristics of the data and the assumptions made about the random
+component of the latent function different models can be obtained. For
+example, if crash severity is coded as a binary variable (e.g.,
+non-fatal/fatal), we can relate the latent variable to the outcome as
+follows: \[
+y_{ik} = 
+\begin{cases}
+\text{fatal} & \text{if } y_{ik}^*>0\\
+\text{non-fatal} & \text{if } y_{ik}^*\leq0
+\end{cases}
+\]
+
+Due to the stochastic nature of the latent function, the outcome of the
+crash is not fully determined. However, we can make the following
+probability statement: \[
+P(y_i = \text{fatal}) = P(y_i^* > 0)
+\] In other words, the probability of a fatal accident equals the
+probability that the latent variable is greater than zero. This implies
+(see Maddala, 1986, p. 22): \[
+\begin{array}{rl}\
+P(y_i = \text{fatal}) &= P(\sum_{l=1}\beta_lp_{ikl} + \sum_{m=1}\gamma_mc_{km} + \epsilon_{ik} > 0)\\ 
+&=P(\epsilon_{ik} > -\sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km})
+\end{array}
+\]
+
+It the random terms \(\epsilon_i\) are assumed to follow the logistic
+distribution, the the binary logit model is obtained; if they are
+assumed to follow the normal distribution, the binary probit model is
+obtained.
+
+More often, though, the outcome is recorded using more categories, for
+example property damage only (PDO)/injury/fatality. A similar approach
+can be adopted, with a latent variable that relates to the outcome as
+follows: \[
+y_i = 
+\begin{cases}
+\text{fatality} & \text{if } y_i^*> k_2\\
+\text{injury} & \text{if } k_1< y_i^*<k_2\\
+\text{PDO} & \text{if } y_i^*<k_1
+\end{cases}
+\] where \(k_1\) and \(k_2\) are estimable thresholds. In this case, the
+associated probability statements are as follows: \[
+\begin{array}{rcl}\
+P(y_{ik} = \text{PDO}) &=& 1 - P(y_{ik} = \text{injury}) - P(y_{ik} = \text{fatality})\\ 
+P(y_{ik} = \text{injury}) &=& P(k_1 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} < \epsilon_{ik} < k_2 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} )\\
+P(y_{ik} = \text{fatality}) &=& P(\epsilon_{ik} < k_1 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} )
+\end{array}
+\]
+
+If the random terms are assumed to follow the logistic distribution, the
+ordered logit model is obtained; if the normal distribution, then the
+ordered probit model. Estimation methods for these models are very
+well-established (e.g., Maddala, 1986; Train, 2009)
+
+There are numerous variations of the basic modelling framework above,
+including hierarchical models, bivariate models, multinomial models, and
+Bayesian models, among others (see Savolainen et al., 2011 for a review
+of methods).
 
@@
 
+Montella et al. (2013)
+
 \hypertarget{model-specification}{%
@@
 \hypertarget{sec:application}{%
-\section{Illustrative application}\label{sec:application}}
+\section{Application}\label{sec:application}}
+
+\begin{quote}
+\textbf{Note:} that this paper presents reproducible research. The
+source file is an R Markdown document. All code and data necessary to
+reproduce the analysis are available from the following anonymous Drive
+folder:
+\end{quote}
+
+\begin{quote}
+https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV
+\end{quote}
+
+\begin{quote}
+The source files, code, and data will be publicly available in a GitHub
+repository upon acceptance of the paper for publication
+\end{quote}
 
@@
 
+To assess the performance of the modelling strategies discussed in
+Section \ref{sec:methods}, we use data from Canada's National Collision
+Database (NCDB). This is database contains all motor vehicle collisions
+on public roads in Canada as reported by a police service. Data are
+collected by provinces and territories, and shared with the federal
+government, where data are combined, tracked, and analyzed for reporting
+of deaths, injuries, and collisions in Canada at the national level. The
+NCDB is provided by Transport Canada, the agency of the federal
+government of Canada in charge of transportation policies and programs,
+under the Open Government License - Canada version 2.0
+{[}https://open.canada.ca/en/open-government-licence-canada{]}.
+
+The NCDB is available from 1999. For the purpose of this paper, we use
+the most recent year available as of this writing (2017). Furthermore,
+for assessment we also use the data corresponding to 2016. Similar to
+databases in other jurisdictions {[}see @{]}, the NCDB contains
+information pertaining to the collision, the traffic unit, and the
+person(s) involved.
+
+\begin{tabular}{ll}
+\toprule
+Variable & Details\\
+\midrule
+Age & Details\\
+\bottomrule
+\multicolumn{2}{l}{\textit{Note: }}\\
+\multicolumn{2}{l}{NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\end{tabular}
+
+\hypertarget{model-estimation}{%
+\subsection{Model estimation}\label{model-estimation}}
+
+Words go here
+
+Model specification. See Table:
+
+\begin{table}[H]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{llllll}
+\toprule
+Variable & Notes & Model 1 & Model 2 & Model 3 & Model 4\\
+\midrule
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Individual-level variables}}\\
+\hspace{1em}Age & In decades & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Age Squared &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Sex & Reference: Female & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Use of Safety Devices & 7 levels; Reference: No Safety Device & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Traffic unit-level variables}}\\
+\hspace{1em}Passenger & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\hspace{1em}Pedestrian & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\hspace{1em}Motorcyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Light Truck & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\hspace{1em}Heavy Vehicle & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Opponent variables)}}\\
+\hspace{1em}Age of Opponent & In decades &  & $\checkmark$ & $\checkmark$ & \\
+\hspace{1em}Age of Opponent Squared &  &  & $\checkmark$ & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Sex of Opponent & Reference: Female &  & $\checkmark$ & $\checkmark$ & \\
+\hspace{1em}Opponent: Light Duty Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Opponent: Light Truck & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Opponent: Heavy Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Hierarchical traffic unit variables)}}\\
+\hspace{1em}Age:Light Truck Driver &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Light Truck Driver &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Heavy Vehicle Driver &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Heavy Vehicle Driver &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Light Truck Passenger &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Light Truck Passenger &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Heavy Vehicle Passenger &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Heavy Vehicle Passenger &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Pedestrian &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Pedestrian &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Bicyclist &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Bicyclist &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Age:Motorcyclist &  &  &  & $\checkmark$ & \\
+\hspace{1em}Age Squared:Motorcyclist &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Hierarchical opponent variables)}}\\
+\hspace{1em}Age:Age of Opponent &  &  &  &  & $\checkmark$\\
+\hspace{1em}Age:Age of Female Opponent &  &  &  &  & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Age:Age of Male Opponent Squared &  &  &  &  & $\checkmark$\\
+\hspace{1em}Age:Age of Female Opponent Squared &  &  &  &  & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Age Squared:Age of Male Opponent &  &  &  &  & $\checkmark$\\
+\hspace{1em}Age Squared:Age of Female Opponent &  &  &  &  & $\checkmark$\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Case-level variables)}}\\
+\hspace{1em}Crash Configuration & 19 levels; Reference: Hit a Moving Object & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Road Configuration & 12 levels; Reference: Non-Intersection & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Weather & 9 levels; Reference: Clear and Sunny & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Surface & 11 levels; Reference: Dry & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Road Alignment & 8 levels; Reference: Straight and Level & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Traffic Controls & 19 levels; Reference: Operational Traffic Signals & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Month & 12 levels; Reference: January & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+Notice how there are zero cases of user: BYCICLIST - opponent: Heavy
+Vehicle.
+
+Re-estimate model after subsetting by USER Type of person 1. LDxLD:
+
+Summary of models. See Table \ref{tab:model-summary}.
+
+\begin{table}
+
+\caption{\label{tab:model-summary}\label{tab:model-summary}Summary of model estimation results}
+\centering
+\begin{tabular}[t]{lccc}
+\toprule
+Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC\\
+\midrule
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Full sample models}}\\
+\hspace{1em}Model 1 & 164,511 & 100 & 195,215\\
+\hspace{1em}Model 2 & 164,511 & 106 & 178,943\\
+\rowcolor{gray!6}  \hspace{1em}Model 3 & 164,511 & 116 & 181,333\\
+\hspace{1em}Model 4 & 164,511 & 109 & 179,018\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Model 1 Ensemble (sample subsets by user type vs opponent)}}\\
+\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 95 & 145,396\\
+\hspace{1em}Light duty vehicle vs light truck & 3,237 & 95 & 3,979\\
+\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 95 & 5,913\\
+\hspace{1em}Light truck vs light duty vehicle & 3,121 & 95 & 3,921\\
+\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 95 & 1,230\\
+\hspace{1em}Light truck vs heavy vehicle & 198 & 95 & 354\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,763 & 95 & 4,362\\
+\hspace{1em}Heavy vehicle vs light truck & 180 & 95 & 291\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 95 & 1,193\\
+\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 94 & 2,842\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 94 & 270\\
+\hspace{1em}Pedestrian vs heavy vehicle & 376 & 94 & 473\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 94 & 686\\
+\hspace{1em}Bicyclist vs light truck & 148 & 94 & 192\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
+\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 94 & 1,403\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 94 & 233\\
+\hspace{1em}Motorcyclist vs heavy vehicle & 73 & 94 & 204\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Model 2 Ensemble (sample subsets by user type vs opponent)}}\\
+\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 98 & 143,909\\
+\hspace{1em}Light duty vehicle vs light truck & 3,237 & 98 & 3,963\\
+\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 98 & 5,896\\
+\hspace{1em}Light truck vs light duty vehicle & 3,121 & 98 & 3,913\\
+\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 98 & 1,216\\
+\hspace{1em}Light truck vs heavy vehicle & 198 & 98 & 347\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,763 & 98 & 4,315\\
+\hspace{1em}Heavy vehicle vs light truck & 180 & 98 & 275\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 98 & 1,182\\
+\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 98 & 2,839\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 98 & 270\\
+\hspace{1em}Pedestrian vs heavy vehicle & 376 & 98 & 476\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 98 & 693\\
+\hspace{1em}Bicyclist vs light truck & 148 & 98 & 200\\
+\rowcolor{gray!6}  Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
+\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 98 & 1,411\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 98 & 235\\
+\hspace{1em}Motorcyclist vs heavy vehicle & 73 & 98 & 201\\
+\bottomrule
+\multicolumn{4}{l}{\textit{Note: }}\\
+\multicolumn{4}{l}{There are zero cases of BICYCLIST vs HV in the sample}\\
+\end{tabular}
+\end{table}
+
+\hypertarget{sec:assessment}{%
+\section{Model assessment}\label{sec:assessment}}
+
+\hypertarget{outcome-shares-based-on-probabilities}{%
+\subsection{Outcome shares based on
+probabilities}\label{outcome-shares-based-on-probabilities}}
+
 Words go here.
 
-\hypertarget{analysis-and-results}{%
-\subsection{Analysis and results}\label{analysis-and-results}}
+The datasets are used to predict the probabilities of the outcomes:
 
-Words go here
+The results of calculating the Average Prediction Error appear in Table
+\ref{tab:ape-results}.
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:table-ape-results}\label{tab:ape-results}Predicted shares and average prediction errors (APE) by model (percentages)}
+\centering
+\begin{tabular}[t]{lrrrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{3}{c}{No Injury} & \multicolumn{3}{c}{Injury} & \multicolumn{3}{c}{Fatality} & \multicolumn{1}{c}{} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
+Model & Observed & Predicted & APE & Observed & Predicted & APE & Observed & Predicted & APE & WAPE\\
+\rowcolor{gray!15}
+\midrule
+\addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{In-sample (nowcasting using 2017 dataset, i.e., estimation dataset)}}\\
+\hspace{1em}Model 1 & 78886 & 79029.00 & 0.18 & 84675 & 84533.74 & 0.17 & 950 & 948.26 & 0.18 & 0.17\\
+\rowcolor{gray!15}
+\hspace{1em}Model 1 Ensemble & 62449 & 62439.99 & 0.01 & 83606 & 83614.28 & 0.01 & 933 & 933.73 & 0.08 & 0.01\\
+\hspace{1em}Model 2 & 78886 & 78928.98 & 0.05 & 84675 & 84641.94 & 0.04 & 950 & 940.08 & 1.04 & 0.05\\
+\rowcolor{gray!15}
+\hspace{1em}Model 2 Ensemble & 62449 & 62438.99 & 0.02 & 83606 & 83615.22 & 0.01 & 933 & 933.80 & 0.09 & 0.01\\
+\hspace{1em}Model 3 & 78886 & 79027.29 & 0.18 & 84675 & 84512.50 & 0.19 & 950 & 971.21 & 2.23 & 0.20\\
+\rowcolor{gray!15}
+\hspace{1em}Model 4 & 78886 & 78939.18 & 0.07 & 84675 & 84622.54 & 0.06 & 950 & 949.28 & 0.08 & 0.06\\
+\rowcolor{gray!15}
+\addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 dataset)}}\\
+\hspace{1em}Model 1 & 82812 & 82574.35 & 0.29 & 88586 & 88737.88 & 0.17 & 935 & 1020.77 & 9.17 & 0.28\\
+\rowcolor{gray!15}
+\hspace{1em}Model 1 Ensemble & 64469 & 64525.65 & 0.09 & 87476 & 87291.39 & 0.21 & 909 & 1036.96 & 14.08 & 0.24\\
+\hspace{1em}Model 2 & 82812 & 82900.86 & 0.11 & 88586 & 88432.82 & 0.17 & 935 & 999.32 & 6.88 & 0.18\\
+\rowcolor{gray!15}
+\hspace{1em}Model 2 Ensemble & 64469 & 64541.74 & 0.11 & 87476 & 87261.66 & 0.25 & 909 & 1050.60 & 15.58 & 0.28\\
+\hspace{1em}Model 3 & 82812 & 82948.83 & 0.17 & 88586 & 88340.07 & 0.28 & 935 & 1044.10 & 11.67 & 0.29\\
+\rowcolor{gray!15}
+\hspace{1em}Model 4 & 82812 & 82878.68 & 0.08 & 88586 & 88446.37 & 0.16 & 935 & 1007.94 & 7.80 & 0.16\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
+
+\hypertarget{predicted-outcomes}{%
+\subsection{Predicted outcomes}\label{predicted-outcomes}}
+
+Words go here.
+
+Verification statistics used are summarized in Table
+\ref{tab:verification-statistics}.
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:table-verification-statistics}\label{tab:verification-statistics}Verification statistics}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{30em}>{\raggedright\arraybackslash}p{30em}}
+\toprule
+Statistic & Description & Notes\\
+\midrule
+\rowcolor{gray!6}  Percent Correct ($PC$) & Total hits and correct rejections divided by number of cases & Strongly influenced by most common category\\
+Percent Correct by Class ($PC_c$) & Same as Percent Correct but by category & Strongly influenced by most common category\\
+\rowcolor{gray!6}  Bias ($B$) & Total predicted by category, divided by total observed by category & $B>1$: class is overpredicted; $B<1$: class is underpredicted\\
+Critical Success Index ($CSI$) & Total hits divided by total hits + false alarms + misses & $CSI = 1$: perfect score; $CSI = 0$: no skill\\
+\rowcolor{gray!6}  Probability of False Detection ($F$) & Proportion of no events forecast as yes; sensitive to false alarms but ignores misses & $F = 0$: perfect score\\
+\addlinespace
+Probability of Detection ($POD$) & Total hits divided by total observed by class & $POD = 1$: perfect score\\
+\rowcolor{gray!6}  False Alarm Ratio ($FAR$) & Total false alarms divided by total forecast yes by class; measures fraction of predicted yes that did not occur & $FAR = 0$: perfect score\\
+Heidke Skill Score ($HSS$) & Fraction of correct predictions after removing predictions attributable to chance; measures fractional improvement over random; tends to reward conservative forecasts & $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better\\
+\rowcolor{gray!6}  Peirce Skill Score ($PSS$) & Combines $POD$ and $F$; measures ability to separate yes events from no events; tends to reward conservative forecasts & $PSS = 1$: perfect score; $PSS = 0$: no skill\\
+Gerrity Score ($GS$) & Measures accuracy of predicting the correct category, relative to random; tends to reward correct forecasts of less likely category & $GS = 1$: perfect score; $GS = 0$: no skill\\
+\bottomrule
+\end{tabular}}
+\end{table}
+\end{landscape}
+
+We next evaluate the outcomes of the nowcast using an array of
+verification statistics. See Table
+\ref{tab:nowcast-outcomes-assessment}.
+
+\begin{table}
+
+\caption{\label{tab:nowcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 dataset, i.e., estimation dataset)}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{Observed} & \multicolumn{3}{c}{Predicted Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
+\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
+Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 1}}\\
+\hspace{1em}No Injury & 50652 & 22504 & 150 &  & 69.067 & 0.929 & 0.499 & 0.265 & 0.642 & 0.309 &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62120 & 797 &  & 68.644 & 1.076 & 0.546 & 0.364 & 0.734 & 0.318 &  &  & \\
+
+Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.552} & 99.392 & 0.059 & 0.003 & 0.000 & 0.003 & 0.946 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.372} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.370} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.190}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
+\hspace{1em}No Injury & 34711 & 16429 & 63 &  & 69.909 & 0.820 & 0.440 & 0.195 & 0.556 & 0.322 &  &  & \\
+
+\hspace{1em}Injury & 27738 & 67167 & 830 &  & 69.380 & 1.145 & 0.599 & 0.451 & 0.803 & 0.298 &  &  & \\
+
+Fatality & 0 & 10 & 40 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.338} & 99.386 & 0.054 & 0.042 & 0.000 & 0.043 & 0.200 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.353} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.202}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2}}\\
+\hspace{1em}No Injury & 51530 & 17136 & 85 &  & 72.903 & 0.872 & 0.536 & 0.201 & 0.653 & 0.250 &  &  & \\
+
+\hspace{1em}Injury & 27356 & 67515 & 864 &  & 72.415 & 1.131 & 0.598 & 0.353 & 0.797 & 0.295 &  &  & \\
+
+Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.364} & 99.409 & 0.026 & 0.001 & 0.000 & 0.001 & 0.960 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.447} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.443} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
+\hspace{1em}No Injury & 35473 & 16147 & 60 &  & 70.621 & 0.828 & 0.451 & 0.192 & 0.568 & 0.314 &  &  & \\
+
+\hspace{1em}Injury & 26976 & 67446 & 829 &  & 70.089 & 1.139 & 0.605 & 0.439 & 0.807 & 0.292 &  &  & \\
+
+Fatality & 0 & 13 & 44 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.049} & 99.386 & 0.061 & 0.047 & 0.000 & 0.047 & 0.228 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.379} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.368} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.212}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 3}}\\
+\hspace{1em}No Injury & 51102 & 17297 & 79 &  & 72.549 & 0.868 & 0.531 & 0.203 & 0.648 & 0.254 &  &  & \\
+
+\hspace{1em}Injury & 27784 & 67337 & 868 &  & 72.044 & 1.134 & 0.594 & 0.359 & 0.795 & 0.298 &  &  & \\
+
+Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.996} & 99.399 & 0.046 & 0.003 & 0.000 & 0.003 & 0.932 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.440} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.436} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.224}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 4}}\\
+\hspace{1em}No Injury & 51574 & 17317 & 84 &  & 72.821 & 0.874 & 0.536 & 0.203 & 0.654 & 0.252 &  &  & \\
+
+\hspace{1em}Injury & 27312 & 67335 & 863 &  & 72.333 & 1.128 & 0.597 & 0.353 & 0.795 & 0.295 &  &  & \\
+
+Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.282} & 99.410 & 0.027 & 0.003 & 0.000 & 0.003 & 0.885 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.446} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.441} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\bottomrule
+\multicolumn{14}{l}{\textit{Notes: }}\\
+\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
+\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
+\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
+\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
+\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
+\end{tabular}}
+\end{table}
+
+We next evaluate the outcomes of the nowcast using an array of
+verification statistics. See Table
+\ref{tab:backcast-outcomes-assessment}.
+
+\begin{table}
+
+\caption{\label{tab:backcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 dataset, i.e., estimation dataset)}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{Observed} & \multicolumn{3}{c}{Predicted Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
+\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
+Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 1}}\\
+\hspace{1em}No Injury & 53167 & 23097 & 145 &  & 69.311 & 0.923 & 0.501 & 0.260 & 0.642 & 0.304 &  &  & \\
+
+\hspace{1em}Injury & 29642 & 65455 & 788 &  & 68.920 & 1.082 & 0.550 & 0.363 & 0.739 & 0.317 &  &  & \\
+
+Fatality & 3 & 34 & 2 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.834} & 99.437 & 0.042 & 0.002 & 0.000 & 0.002 & 0.949 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.378} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.375} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.192}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
+\hspace{1em}No Injury & 35564 & 16930 & 71 &  & 69.967 & 0.815 & 0.437 & 0.192 & 0.552 & 0.323 &  &  & \\
+
+\hspace{1em}Injury & 28891 & 70451 & 822 &  & 69.423 & 1.145 & 0.601 & 0.454 & 0.805 & 0.297 &  &  & \\
+
+Fatality & 14 & 95 & 16 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.368} & 99.344 & 0.138 & 0.016 & 0.001 & 0.018 & 0.872 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.362} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.351} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2}}\\
+\hspace{1em}No Injury & 54430 & 17615 & 81 &  & 73.262 & 0.871 & 0.542 & 0.198 & 0.657 & 0.245 &  &  & \\
+
+\hspace{1em}Injury & 28382 & 70954 & 850 &  & 72.806 & 1.131 & 0.602 & 0.349 & 0.801 & 0.292 &  &  & \\
+
+Fatality & 0 & 17 & 4 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.759} & 99.450 & 0.022 & 0.004 & 0.000 & 0.004 & 0.810 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.455} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.451} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.232}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
+\hspace{1em}No Injury & 36234 & 16840 & 73 &  & 70.463 & 0.824 & 0.445 & 0.191 & 0.562 & 0.318 &  &  & \\
+
+\hspace{1em}Injury & 28216 & 70527 & 819 &  & 69.916 & 1.138 & 0.605 & 0.444 & 0.806 & 0.292 &  &  & \\
+
+Fatality & 19 & 109 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.856} & 99.333 & 0.160 & 0.016 & 0.001 & 0.019 & 0.883 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.373} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.362} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.194}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 3}}\\
+\hspace{1em}No Injury & 53985 & 17731 & 83 &  & 72.936 & 0.867 & 0.536 & 0.199 & 0.652 & 0.248 &  &  & \\
+
+\hspace{1em}Injury & 28827 & 70819 & 849 &  & 72.470 & 1.134 & 0.599 & 0.354 & 0.799 & 0.295 &  &  & \\
+
+Fatality & 0 & 36 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.422} & 99.438 & 0.042 & 0.003 & 0.000 & 0.003 & 0.923 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.448} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.444} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.228}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 4}}\\
+\hspace{1em}No Injury & 54493 & 17843 & 79 &  & 73.168 & 0.874 & 0.541 & 0.200 & 0.658 & 0.247 &  &  & \\
+
+\hspace{1em}Injury & 28318 & 70725 & 852 &  & 72.709 & 1.128 & 0.601 & 0.348 & 0.798 & 0.292 &  &  & \\
+
+Fatality & 1 & 18 & 4 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.663} & 99.449 & 0.025 & 0.004 & 0.000 & 0.004 & 0.826 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.453} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.449} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.231}\\
+\bottomrule
+\multicolumn{14}{l}{\textit{Notes: }}\\
+\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
+\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
+\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
+\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
+\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
+\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
+\end{tabular}}
+\end{table}
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Aziz2013exploring}{}%
+Aziz, H.M.A., Ukkusuri, S.V., Hasan, S., 2013. Exploring the
+determinants of pedestrian-vehicle crash severity in new york city.
+Accident Analysis and Prevention 50, 1298--1309.
+doi:\href{https://doi.org/10.1016/j.aap.2012.09.034}{10.1016/j.aap.2012.09.034}
+
+\leavevmode\hypertarget{ref-Bogue2017modified}{}%
+Bogue, S., Paleti, R., Balan, L., 2017. A modified rank ordered logit
+model to analyze injury severity of occupants in multivehicle crashes.
+Analytic Methods in Accident Research 14, 22--40.
+doi:\href{https://doi.org/10.1016/j.amar.2017.03.001}{10.1016/j.amar.2017.03.001}
+
 \leavevmode\hypertarget{ref-Chang2006analysis}{}%
@@
 doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}
+
+\leavevmode\hypertarget{ref-Chen2019investigation}{}%
+Chen, F., Song, M.T., Ma, X.X., 2019. Investigation on the injury
+severity of drivers in rear-end collisions between cars using a random
+parameters bivariate ordered probit model. International Journal of
+Environmental Research and Public Health 16.
+doi:\href{https://doi.org/10.3390/ijerph16142632}{10.3390/ijerph16142632}
 
@@
 
+\leavevmode\hypertarget{ref-Effati2015geospatial}{}%
+Effati, M., Thill, J.C., Shabani, S., 2015. Geospatial and machine
+learning techniques for wicked social science problems: Analysis of
+crash severity on a regional highway corridor. Journal of Geographical
+Systems 17, 107--135.
+doi:\href{https://doi.org/10.1007/s10109-015-0210-x}{10.1007/s10109-015-0210-x}
+
+\leavevmode\hypertarget{ref-Gong2017modeling}{}%
+Gong, L.F., Fan, W.D., 2017. Modeling single-vehicle run-off-road crash
+severity in rural areas: Accounting for unobserved heterogeneity and age
+difference. Accident Analysis and Prevention 101, 124--134.
+doi:\href{https://doi.org/10.1016/j.aap.2017.02.014}{10.1016/j.aap.2017.02.014}
+
+\leavevmode\hypertarget{ref-Haleem2013effect}{}%
+Haleem, K., Gan, A., 2013. Effect of driver's age and side of impact on
+crash severity along urban freeways: A mixed logit approach. Journal of
+Safety Research 46, 67--76.
+doi:\href{https://doi.org/10.1016/j.jsr.2013.04.002}{10.1016/j.jsr.2013.04.002}
+
 \leavevmode\hypertarget{ref-Hanson2013severity}{}%
@@
 
-\leavevmode\hypertarget{ref-Kadilar2016effect}{}%
-Kadilar, G.O., 2016. Effect of driver, roadway, collision, and vehicle
-characteristics on crash severity: A conditional logistic regression
-approach. International Journal of Injury Control and Safety Promotion
-23, 135--144.
-doi:\href{https://doi.org/10.1080/17457300.2014.942323}{10.1080/17457300.2014.942323}
-
 \leavevmode\hypertarget{ref-Khan2015exploring}{}%
@@
 doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}
+
+\leavevmode\hypertarget{ref-Kim2013driver}{}%
+Kim, J.K., Ulfarsson, G.F., Kim, S., Shankar, V.N., 2013. Driver-injury
+severity in single-vehicle crashes in california: A mixed logit analysis
+of heterogeneity due to age and gender. Accident Analysis and Prevention
+50, 1073--1081.
+doi:\href{https://doi.org/10.1016/j.aap.2012.08.011}{10.1016/j.aap.2012.08.011}
+
+\leavevmode\hypertarget{ref-Kim1995personal}{}%
+Kim, K., Nitz, L., Richardson, J., Li, L., 1995. PERSONAL and behavioral
+predictors of automobile crash and injury severity. Accident Analysis
+and Prevention 27, 469--481.
+doi:\href{https://doi.org/10.1016/0001-4575(95)00001-g}{10.1016/0001-4575(95)00001-g}
 
@@
 
+\leavevmode\hypertarget{ref-Maddala1986limited}{}%
+Maddala, G.S., 1986. Limited-dependent and qualitative variables in
+econometrics. Cambridge university press.
+
+\leavevmode\hypertarget{ref-Mannering2016unobserved}{}%
+Mannering, F., Shankar, V., Bhat, C.R., 2016. Unobserved heterogeneity
+and the statistical analysis of highway accident data. Analytic Methods
+in Accident Research 11, 1--16.
+doi:\href{https://doi.org/10.1016/j.amar.2016.04.001}{10.1016/j.amar.2016.04.001}
+
 \leavevmode\hypertarget{ref-McArthur2014spatial}{}%
@@
 
+\leavevmode\hypertarget{ref-Montella2013crash}{}%
+Montella, A., Andreassen, D., Tarko, A.P., Turner, S., Mauriello, F.,
+Imbriani, L.L., Romero, M.A., 2013. Crash databases in australasia, the
+european union, and the united states review and prospects for
+improvement. Transportation Research Record 128--136.
+doi:\href{https://doi.org/10.3141/2386-15}{10.3141/2386-15}
+
 \leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
@@
 doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}
+
+\leavevmode\hypertarget{ref-Mussone2017analysis}{}%
+Mussone, L., Bassani, M., Masci, P., 2017. Analysis of factors affecting
+the severity of crashes in urban road intersections. Accident Analysis
+and Prevention 103, 112--122.
+doi:\href{https://doi.org/10.1016/j.aap.2017.04.007}{10.1016/j.aap.2017.04.007}
 
@@
 
-\leavevmode\hypertarget{ref-Quddus2010road}{}%
-Quddus, M.A., Wang, C., Ison, S.G., 2010. Road traffic congestion and
-crash severity: Econometric analysis using ordered response models.
-Journal of Transportation Engineering-Asce 136, 424--435.
-doi:\href{https://doi.org/10.1061/(asce)te.1943-5436.0000044}{10.1061/(asce)te.1943-5436.0000044}
-
 \leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
@@
 
+\leavevmode\hypertarget{ref-Rifaat2007accident}{}%
+Rifaat, S.M., Chin, H.C., 2007. Accident severity analysis using ordered
+probit model. Journal of Advanced Transportation 41, 91--114.
+doi:\href{https://doi.org/10.1002/atr.5670410107}{10.1002/atr.5670410107}
+
+\leavevmode\hypertarget{ref-Salon2018determinants}{}%
+Salon, D., McIntyre, A., 2018. Determinants of pedestrian and bicyclist
+crash severity by party at fault in san francisco, ca. Accident Analysis
+and Prevention 110, 149--160.
+doi:\href{https://doi.org/10.1016/j.aap.2017.11.007}{10.1016/j.aap.2017.11.007}
+
+\leavevmode\hypertarget{ref-Sasidharan2014partial}{}%
+Sasidharan, L., Menendez, M., 2014. Partial proportional odds model-an
+alternate choice for analyzing pedestrian crash injury severities.
+Accident Analysis and Prevention 72, 330--340.
+doi:\href{https://doi.org/10.1016/j.aap.2014.07.025}{10.1016/j.aap.2014.07.025}
+
 \leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
@@
 
+\leavevmode\hypertarget{ref-Shaheed2013mixed}{}%
+Shaheed, M.S.B., Gkritza, K., Zhang, W., Hans, Z., 2013. A mixed logit
+analysis of two-vehicle crash seventies involving a motorcycle. Accident
+Analysis and Prevention 61, 119--128.
+doi:\href{https://doi.org/10.1016/j.aap.2013.05.028}{10.1016/j.aap.2013.05.028}
+
 \leavevmode\hypertarget{ref-Symons2019reduced}{}%
@@
 
+\leavevmode\hypertarget{ref-Tarrao2014modeling}{}%
+Tarrao, G.A., Coelho, M.C., Rouphail, N.M., 2014. Modeling the impact of
+subject and opponent vehicles on crash severity in two-vehicle
+collisions. Transportation Research Record 53--64.
+doi:\href{https://doi.org/10.3141/2432-07}{10.3141/2432-07}
+
 \leavevmode\hypertarget{ref-Thompson2018trends}{}%
@@
 doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
+
+\leavevmode\hypertarget{ref-Train2009discrete}{}%
+Train, K., 2009. Discrete choice methods with simulation, 2nd Edition.
+ed. Cambridge University Press, Cambridge.
+
+\leavevmode\hypertarget{ref-Wang2015copula}{}%
+Wang, K., Yasmin, S., Konduri, K.C., Eluru, N., Ivan, J.N., 2015.
+Copula-based joint model of injury severity and vehicle damage in
+two-vehicle crashes. Transportation Research Record 158--166.
+doi:\href{https://doi.org/10.3141/2514-17}{10.3141/2514-17}
+
+\leavevmode\hypertarget{ref-White1972effects}{}%
+White, S., Clayton, S., 1972. Some effects of alcohol, age of driver,
+and estimated speed on the likelihood of driver injury. Accident
+Analysis \& Prevention 4.
 
@@
 
+\leavevmode\hypertarget{ref-Wu2014mixed}{}%
+Wu, Q., Chen, F., Zhang, G.H., Liu, X.Y.C., Wang, H., Bogus, S.M., 2014.
+Mixed logit model-based driver injury severity investigations in single-
+and multi-vehicle crashes on rural two-lane highways. Accident Analysis
+and Prevention 72, 105--115.
+doi:\href{https://doi.org/10.1016/j.aap.2014.06.014}{10.1016/j.aap.2014.06.014}
+
+\leavevmode\hypertarget{ref-Shamsunnahar2013evaluating}{}%
+Yasmin, S., Eluru, N., 2013. Evaluating alternate discrete outcome
+frameworks for modeling crash injury severity. Accident Analysis \&
+Prevention 59, 506--521.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
+
 \end{document}

### 2020-01-02 Interactions-Crash-Severity-v1/fixed-polr.R

@@ -1,0 +1,588 @@
+# file MASS/R/polr.R
+# copyright (C) 1994-2008 W. N. Venables and B. D. Ripley
+# Use of transformed intercepts contributed by David Firth
+#
+#  This program is free software; you can redistribute it and/or modify
+#  it under the terms of the GNU General Public License as published by
+#  the Free Software Foundation; either version 2 or 3 of the License
+#  (at your option).
+#
+#  This program is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#
+#  A copy of the GNU General Public License is available at
+#  http://www.r-project.org/Licenses/
+#
+polr <- function(formula, data, weights, start, ..., subset,
+                 na.action, contrasts = NULL, Hess = FALSE,
+                 model = TRUE,
+                 method = c("logistic", "probit", "cloglog", "cauchit"))
+{
+    logit <- function(p) log(p/(1 - p))
+
+    fmin <- function(beta) {
+        theta <- beta[pc + 1L:q]
+        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
+        eta <- offset
+        if (pc > 0)
+            eta <- eta + drop(x %*% beta[1L:pc])
+        pr <- pfun(gamm[y + 1] - eta) - pfun(gamm[y] - eta)
+        if (all(pr > 0))
+            -sum(wt * log(pr))
+        else Inf
+    }
+
+    gmin <- function(beta)
+    {
+        jacobian <- function(theta) { ## dgamma by dtheta matrix
+            k <- length(theta)
+            etheta <- exp(theta)
+            mat <- matrix(0 , k, k)
+            mat[, 1] <- rep(1, k)
+            for (i in 2:k) mat[i:k, i] <- etheta[i]
+            mat
+        }
+        theta <- beta[pc + 1L:q]
+        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
+        eta <- offset
+        if(pc > 0) eta <- eta + drop(x %*% beta[1L:pc])
+        pr <- pfun(gamm[y+1] - eta) - pfun(gamm[y] - eta)
[…]

### 2020-01-15 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.Rmd
Added Model 4 ensemble and verified all models for convergence

--- 2019-12-12 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-\usepackage[hyphens]{url}
+---
+title: An empirical assessment of strategies to model opponent effects in crash severity analysis
+author:
+  - name: Antonio Paez
+    email: paezha@mcmaster.ca
+    affiliation: 
+    footnote: Corresponding Author
+  - name: Hany Hassan
+    email: hassan1@lsu.edu
+    affiliation: Louisiana State University
+  - name: Mark Ferguson
+    email: fergumr@mcmaster.ca
+    affiliation: McMaster University
+  - name: Saiedeh Razavi
+    email: razavi@mcmaster.ca
+    affiliation: McMaster University    
+address:
+  - code: McMaster University
+    address: McMaster Institute for Transportation and Logistics, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1
+  - code: Louisiana State University
+    address: Department of Civil and Environmental Engineering, Baton Rouge, Louisiana, USA 70803
+abstract: |
+  Road accidents impose an important burden on health and the economy.
+Numerous efforts to understand the factors that affect road collisions have been undertaken.
+One stream of research focus on modelling the severity of crashes.
+Crash severity research is useful to clarify the way different factors can influence the outcome of the event.
+The objective of this paper is to assess different strategies to model the interactions between participants in a crash, in the context of crashes involving two parties.
+Towards this objective, a series of models are estimated using data from Canada's National Collision Database.
+Three levels of crash severity (no injury/injury/fatality) are analyzed using ordered logit models and covariates for the participants in the crash and the conditions of the crash.
+Modelling strategies include different ways of introducing the covariates (e.g., in a single-level or multi-level form), as well as by subsetting the dataset.
+The models are assessed using predicted shares and outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis.
+On the other hand, the study suggests that hierarchical (i.e., multi-level) specifications and subsetting do not perform necessarily better than a relatively simple single-level model with opponent effects.
 
-  \journal{Some Journal} % Sets Journal name
+journal: "Some Journal"
+date: "`r Sys.Date()`"
+bibliography: mybibfile.bib
+#linenumbers: true
+numbersections: true
+csl: elsevier-harvard.csl
+output: rticles::elsevier_article
 
-\usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\bibliographystyle{elsarticle-harv}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Modelling participant interactions in crash severity analysis},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-% Pandoc header
-
-\begin{document}
-\begin{frontmatter}
-
-  \title{Modelling participant interactions in crash severity analysis}
-    \author[Some University]{Author 1\corref{c1}}
-   \ead{a1@example.com} 
-   \cortext[c1]{Corresponding Author}
-    \author[Some Institute of Technology]{Author 2}
-   \ead{a2@example.com}
-
-      \address[Some University]{Department, Street, City, State, Zip}
-    \address[Some Institute of Technology]{Department, Street, City, State, Zip}
-
-  \begin{abstract}
-  Road safety continues to impose an important burden on health and the
-  economy. Numerous efforts to understand the factors that affect road
-  safety have been undertaken. One stream of research focus on modelling
-  the severity of crashes. Crash severity research is useful to clarify
-  the way different factors can influence an outcome where there are no
-  injuries, or injuries, or fatalities. The objective of this paper is to
-  describe an approach to model the interactions between participants in a
-  crash, in the context of crashes involving two parties. Using data from
-  Canada's National Collision Database we demonstrate the approach. Three
-  levels of crash severity (no injury/injury/fatality) are analyzed using
-  ordered logit models and covariates for the participants in the crash
-  and the conditions of the crash. The results suggest that there are
-  important interactions between the various descriptors of the
-  participants in the crash. More generally, the example shows how
-  modelling the interactions between participants explicitly leads to
-  richer insights into the covariates of crash severity, in addition to
-  better model fits. The approach described in this paper is intuitive,
-  simple to implement, and can be adapted to advanced modelling methods
-  with ease.
-  \end{abstract}
-
- \end{frontmatter}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Road safety continues to be a concern world-wide. According to a recent
-report from the World Health Organization (2019), road accidents are the
-8th leading cause of death for all ages, and the number one cause of
-death for children and young people between the ages of 5 to 29. Of all
-leading causes of death, road accidents are the only non-external cause,
-unrelated to disease, disorder, or infection. Road accidents impose a
-heavy burden on individuals and society as a whole. Gobally, there has
-been an increasing trend in number of vehicles and road accident-related
-deaths, even if the rate of death per 100,000 population and 100,000
-vehicles have both fallen (World Health Organization, 2019, Figs. 1 and
-2). These gains, although they are to be celebrated, cannot distract
-from the crippling economic cost of premature death (e.g., Symons et
-al., 2019; Wijnen et al., 2019), not to mention the long-term
-consequences for survivors, measured in sometimes crushing emotional and
-physical pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier
-et al., n.d.).
-
-As the statistics suggest, the burden of road accidents is also not
-borne evenly. There are important disparities at the international
-level, where the odds of death due to a road crash are three times
-higher in low-income countries compared to high-income countries; in
-fact, no reductions in road accident-related fatalities were appreciated
-in low-income countries between 2013 and 2016 (World Health
-Organization, 2019). In the case of high-income countries, where
-substantial gains in road safety have been observed for years, said
-gains have been unevenly distributed; thus, while fatal crashes
-involving older adults in the United States and Great Britain declined
-between 1997 and 2010 (despite an increase in the number of older
-people), the trend remained stable or increased slightly in Australia in
-roughly the same period (Thompson et al., 2018). There are also
-systematic differences in the impact of road accidents. For example, in
-a study in the United States, Obeng (2011) reported that the impact of
-covariates of crash severity varied in magnitude by gender, and in some
-cases were not even the same. More recently, Regev et al.~(2018) used
-adjusted crash risk to find that the risk of crashes in Great Britain
-peaked for people 21 to 29 years of age; on the other hand, the risk of
-fatal injuries for older drivers was constant, irrespective of the
-seriousness of the crash - which highlights the perils of accidents at
-older ages. Other studies have concentrated on the consequences of road
-accidents on the young (e.g., Peek-Asa et al., 2010), the old (e.g.,
-Rakotonirainy et al., 2012), as well as pedestrians and cyclists (e.g.,
-Hanson et al., 2013; McArthur et al., 2014).
-
-Given the relevance and cost of the matter, as well as the important
-variations of the impacts among different population segments, numerous
-efforts efforts exist to better understand the factors that affect road
-safety - including the probable consequences of the crash. Consequently,
-a stream of research in the analysis of road accidents is concerned with
-the severity of crashes. In particular, multivariate analysis of crash
-severity is a useful way to clarify the way various factors can affect
-the outcome of an incident, to discriminate between various levels of
-injury, from no injury (i.e., property damage only), to different
-degrees of injury up to and including fatality. This is an active area
-of research (Savolainen et al., 2011), and one where methodological
-developments have aimed at improving the reliability, accuracy, and
-precision of models.
-
-This paper aspires to contribute to the literature on crash severity by
-describing an approach to model the interactions between participants in
-a crash, in the context of incidents involving two parties. The
-importance of these interactions has been recognized in the existing
-literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al.,
-2017), and in this paper we present a systematic approach to model the
-way participants in a crash interact to influence the severity of the
-accident for each party. By way of application, we demonstrate the
-approach through an empirical analysis of data drawn from Canada's
-National Collision Database. Three levels of crash severity (no
-injury/injury/fatality) are analyzed using ordered logit models and
-covariates for the participants in the crash and the conditions of the
-crash. The results suggest that there are important interactions between
-the various descriptors of the participants in the crash. More
-generally, the example shows how modelling the interactions between
-participants explicitly leads to richer insights into the covariates of
-crash severity, in addition to better model fits. The approach described
-in this paper is intuitive, simple to implement, and can be adapted to
-advanced modelling methods with ease.
-
-The rest of this paper is structure as follows. In Section
-\ref{sec:review-of-methods} we present a concise review of the methods
-used to analyze crash severity, with a particular focus on techniques
-that consider the interactions between participants in a crash.
-
-\hypertarget{sec:review-of-methods}{%
-\section{Methodological approaches in crash severity
-analysis}\label{sec:review-of-methods}}
-
-Methods:
-
-Iranitalab and Khattak (2017); Khan et al. (2015); Chang and Wang
-(2006): machine learning Quddus et al. (2010): ordinal models Kadilar
-(2016): conditional logistic regression Mooradian et al. (2013): partial
-proportional odds models Ma et al. (2008): counts
-
-Interplay between participants:
-
-Chiou et al. (2013)
-
-Lee and Li (2014) consider effect on crash severity of interactions
-between different types of vehicles. This they do by subsetting the
-dataset and estimating independent models for each subset of data. Since
-they consider three types of vehicles, namely cars (C), light trucks
-(L), and heavy trucks (H), they work with nine datasets, for each type
-of interactions (i.e., C-C, C-L, C-H, and so on).
-
-Li et al. (2017)
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-Words go here.
-
-\hypertarget{illustrative-application}{%
-\section{Illustrative application}\label{illustrative-application}}
-
-Words go here.
-
-\hypertarget{further-considerations}{%
-\section{Further considerations}\label{further-considerations}}
-
-Words go here.
-
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
-
-Words go here.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Chang2006analysis}{}%
-Chang, L.Y., Wang, H.W., 2006. Analysis of traffic injury severity: An
-application of non-parametric classification tree techniques. Accident
-Analysis and Prevention 38, 1019--1027.
-doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}
-
-\leavevmode\hypertarget{ref-Chiou2013modeling}{}%
-Chiou, Y.C., Hwang, C.C., Chang, C.C., Fu, C., 2013. Modeling
-two-vehicle crash severity by a bivariate generalized ordered probit
-approach. Accident Analysis and Prevention 51, 175--184.
-doi:\href{https://doi.org/10.1016/j.aap.2012.11.008}{10.1016/j.aap.2012.11.008}
-
-\leavevmode\hypertarget{ref-Devlin2019road}{}%
-Devlin, A., Beck, B., Simpson, P.M., Ekegren, C.L., Giummarra, M.J.,
-Edwards, E.R., Cameron, P.A., Liew, S., Oppy, A., Richardson, M., Page,
-R., Gabbe, B.J., 2019. The road to recovery for vulnerable road users
-hospitalised for orthopaedic injury following an on-road crash. Accident
-Analysis and Prevention 132, 10.
-doi:\href{https://doi.org/10.1016/j.aap.2019.105279}{10.1016/j.aap.2019.105279}
-
-\leavevmode\hypertarget{ref-Hanson2013severity}{}%
-Hanson, C.S., Noland, R.B., Brown, C., 2013. The severity of pedestrian
-crashes: An analysis using google street view imagery. Journal of
-Transport Geography 33, 42--53.
-doi:\href{https://doi.org/10.1016/j.jtrangeo.2013.09.002}{10.1016/j.jtrangeo.2013.09.002}
-
-\leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
-Iranitalab, A., Khattak, A., 2017. Comparison of four statistical and
-machine learning methods for crash severity prediction. Accident
-Analysis and Prevention 108, 27--36.
-doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}
-
-\leavevmode\hypertarget{ref-Kadilar2016effect}{}%
-Kadilar, G.O., 2016. Effect of driver, roadway, collision, and vehicle
-characteristics on crash severity: A conditional logistic regression
-approach. International Journal of Injury Control and Safety Promotion
-23, 135--144.
-doi:\href{https://doi.org/10.1080/17457300.2014.942323}{10.1080/17457300.2014.942323}
-
-\leavevmode\hypertarget{ref-Khan2015exploring}{}%
-Khan, G., Bill, A.R., Noyce, D.A., 2015. Exploring the feasibility of
-classification trees versus ordinal discrete choice models for analyzing
-crash severity. Transportation Research Part C-Emerging Technologies 50,
-86--96.
-doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}
-
-\leavevmode\hypertarget{ref-Lee2014analysis}{}%
-Lee, C., Li, X.C., 2014. Analysis of injury severity of drivers involved
-in single- and two-vehicle crashes on highways in ontario. Accident
-Analysis and Prevention 71, 286--295.
-doi:\href{https://doi.org/10.1016/j.aap.2014.06.008}{10.1016/j.aap.2014.06.008}
-
-\leavevmode\hypertarget{ref-Li2017interplay}{}%
-Li, L., Hasnine, M.S., Habib, K.M.N., Persaud, B., Shalaby, A., 2017.
-Investigating the interplay between the attributes of at-fault and
-not-at-fault drivers and the associated impacts on crash injury
-occurrence and severity level. Journal of Transportation Safety \&
-Security 9, 439--456.
-doi:\href{https://doi.org/10.1080/19439962.2016.1237602}{10.1080/19439962.2016.1237602}
-
-\leavevmode\hypertarget{ref-Ma2008multivariate}{}%
-Ma, J.M., Kockelman, K.M., Damien, P., 2008. A multivariate
-poisson-lognormal regression model for prediction of crash counts by
-severity, using bayesian methods. Accident Analysis and Prevention 40,
-964--975.
-doi:\href{https://doi.org/10.1016/j.aap.2007.11.002}{10.1016/j.aap.2007.11.002}
-
-\leavevmode\hypertarget{ref-McArthur2014spatial}{}%
-McArthur, A., Savolainen, P.T., Gates, T.J., 2014. Spatial analysis of
-child pedestrian and bicycle crashes development of safety performance
-function for areas adjacent to schools. Transportation Research Record
-57--63. doi:\href{https://doi.org/10.3141/2465-08}{10.3141/2465-08}
-
-\leavevmode\hypertarget{ref-Merlin2007stress}{}%
-Merlin, E.P.R., Gonzalez-Forteza, C., Lira, L.R., Tapia, J.A.J., 2007.
-Post-traumatic stress disorder in patients with non intentional injuries
-caused by road traffic accidents. Salud Mental 30, 43--48.
-
-\leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
-Mooradian, J., Ivan, J.N., Ravishanker, N., Hu, S., 2013. Analysis of
-driver and passenger crash injury severity using partial proportional
-odds models. Accident Analysis and Prevention 58, 53--58.
-doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}
-
-\leavevmode\hypertarget{ref-Obeng2011gender}{}%
-Obeng, K., 2011. Gender differences in injury severity risks in crashes
-at signalized intersections. Accident Analysis and Prevention 43,
-1521--1531.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.004}{10.1016/j.aap.2011.03.004}
-
-\leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
-Peek-Asa, C., Britton, C., Young, T., Pawlovich, M., Falb, S., 2010.
-Teenage driver crash incidence and factors influencing crash injury by
-rurality. Journal of Safety Research 41, 487--492.
-doi:\href{https://doi.org/10.1016/j.jsr.2010.10.002}{10.1016/j.jsr.2010.10.002}
-
-\leavevmode\hypertarget{ref-Pelissier2019medical}{}%
-Pelissier, C., Fort, E., Fontana, L., Hours, M., n.d. Medical and
-socio-occupational predictive factors of psychological distress 5 years
-after a road accident: A prospective study. Social Psychiatry and
-Psychiatric Epidemiology 13.
-doi:\href{https://doi.org/10.1007/s00127-019-01780-0}{10.1007/s00127-019-01780-0}
-
-\leavevmode\hypertarget{ref-Quddus2010road}{}%
-Quddus, M.A., Wang, C., Ison, S.G., 2010. Road traffic congestion and
-crash severity: Econometric analysis using ordered response models.
-Journal of Transportation Engineering-Asce 136, 424--435.
-doi:\href{https://doi.org/10.1061/(asce)te.1943-5436.0000044}{10.1061/(asce)te.1943-5436.0000044}
-
-\leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
-Rakotonirainy, A., Steinhardt, D., Delhomme, P., Darvell, M., Schramm,
-A., 2012. Older drivers' crashes in queensland, australia. Accident
-Analysis and Prevention 48, 423--429.
-doi:\href{https://doi.org/10.1016/j.aap.2012.02.016}{10.1016/j.aap.2012.02.016}
-
-\leavevmode\hypertarget{ref-Regev2018crash}{}%
-Regev, S., Rolison, J.J., Moutari, S., 2018. Crash risk by driver age,
-gender, and time of day using a new exposure methodology. Journal of
-Safety Research 66, 131--140.
-doi:\href{https://doi.org/10.1016/j.jsr.2018.07.002}{10.1016/j.jsr.2018.07.002}
-
-\leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
-Savolainen, P.T., Mannering, F., Lord, D., Quddus, M.A., 2011. The
-statistical analysis of highway crash-injury severities: A review and
-assessment of methodological alternatives. Accident Analysis and
-Prevention 43, 1666--1676.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.025}{10.1016/j.aap.2011.03.025}
-
-\leavevmode\hypertarget{ref-Symons2019reduced}{}%
-Symons, J., Howard, E., Sweeny, K., Kumnick, M., Sheehan, P., 2019.
-Reduced road traffic injuries for young people: A preliminary investment
-analysis. Journal of Adolescent Health 65, S34--S43.
-doi:\href{https://doi.org/10.1016/j.jadohealth.2019.01.009}{10.1016/j.jadohealth.2019.01.009}
-
-\leavevmode\hypertarget{ref-Thompson2018trends}{}%
-Thompson, J.P., Baldock, M.R.J., Dutschke, J.K., 2018. Trends in the
-crash involvement of older drivers in australia. Accident Analysis and
-Prevention 117, 262--269.
-doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
-
-\leavevmode\hypertarget{ref-Wijnen2019analysis}{}%
-Wijnen, W., Weijermars, W., Schoeters, A., Berghe, W. van den, Bauer,
-R., Carnis, L., Elvik, R., Martensen, H., 2019. An analysis of official
-road crash cost estimates in european countries. Safety Science 113,
-318--327.
-doi:\href{https://doi.org/10.1016/j.ssci.2018.12.004}{10.1016/j.ssci.2018.12.004}
-
-\leavevmode\hypertarget{ref-WHO2019global}{}%
-World Health Organization, 2019. Global status report on road safety
-2018 (2018). Geneva.
-
-\end{document}
+header-includes:
+   - \usepackage[margin=1in]{geometry}
+---

### 2020-01-15 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.tex
Added Model 4 ensemble and verified all models for convergence

--- 2019-12-12 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
@@
 % Pandoc header
+\usepackage[margin=1in]{geometry}
 
@@
 
-  \title{Modelling participant interactions in crash severity analysis}
-    \author[Some University]{Author 1\corref{c1}}
-   \ead{a1@example.com} 
+  \title{An assessment of strategies to model opponent effects in crash severity
+analysis}
+    \author[]{Antonio Paez\corref{c1}}
+   \ead{paezha@mcmaster.ca} 
    \cortext[c1]{Corresponding Author}
-    \author[Some Institute of Technology]{Author 2}
-   \ead{a2@example.com}
+    \author[Louisiana State University]{Hany Hassan}
+   \ead{hassan1@lsu.edu}
 
-      \address[Some University]{Department, Street, City, State, Zip}
-    \address[Some Institute of Technology]{Department, Street, City, State, Zip}
+    \author[McMaster University]{Mark Ferguson}
+   \ead{fergumr@mcmaster.ca}
+
+    \author[McMaster University]{Saiedeh Razavi}
+   \ead{razavi@mcmaster.ca}
+
+      \address[McMaster University]{McMaster Institute for Transportation and Logistics, 1280 Main Street
+West, Hamilton, Ontario, Canada L8S 4K1}
+    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Baton Rouge,
+Louisiana, USA 70803}
 
   \begin{abstract}
-  Road safety continues to impose an important burden on health and the
-  economy. Numerous efforts to understand the factors that affect road
-  safety have been undertaken. One stream of research focus on modelling
-  the severity of crashes. Crash severity research is useful to clarify
-  the way different factors can influence an outcome where there are no
-  injuries, or injuries, or fatalities. The objective of this paper is to
-  describe an approach to model the interactions between participants in a
-  crash, in the context of crashes involving two parties. Using data from
-  Canada's National Collision Database we demonstrate the approach. Three
+  Road accidents impose an important burden on health and the economy.
+  Numerous efforts to understand the factors that affect road collisions
+  have been undertaken. One stream of research focus on modelling the
+  severity of crashes. Crash severity research is useful to clarify the
+  way different factors can influence the outcome of the event. The
+  objective of this paper is to assess different strategies to model the
+  interactions between participants in a crash, in the context of crashes
+  involving two parties. Towards this objective, a series of models are
+  estimated using data from Canada's National Collision Database. Three
   levels of crash severity (no injury/injury/fatality) are analyzed using
   ordered logit models and covariates for the participants in the crash
-  and the conditions of the crash. The results suggest that there are
-  important interactions between the various descriptors of the
-  participants in the crash. More generally, the example shows how
-  modelling the interactions between participants explicitly leads to
-  richer insights into the covariates of crash severity, in addition to
-  better model fits. The approach described in this paper is intuitive,
-  simple to implement, and can be adapted to advanced modelling methods
-  with ease.
+  and the conditions of the crash. Modelling strategies include different
+  ways of introducing the covariates (e.g., in a single-level or
+  multi-level form), as well as by subsetting the dataset. The models are
+  assessed using predicted shares and outcomes, and the results highlight
+  the importance of considering opponent effects in crash severity
+  analysis. On the other hand, the study suggests that hierarchical (i.e.,
+  multi-level) specifications and subsetting do not perform necessarily
+  better than a relatively simple single-level model with opponent
+  effects.
   \end{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Road safety continues to be a concern world-wide. According to a recent
-report from the World Health Organization (2019), road accidents are the
-8th leading cause of death for all ages, and the number one cause of
-death for children and young people between the ages of 5 to 29. Of all
-leading causes of death, road accidents are the only non-external cause,
-unrelated to disease, disorder, or infection. Road accidents impose a
-heavy burden on individuals and society as a whole. Gobally, there has
-been an increasing trend in number of vehicles and road accident-related
-deaths, even if the rate of death per 100,000 population and 100,000
-vehicles have both fallen (World Health Organization, 2019, Figs. 1 and
-2). These gains, although they are to be celebrated, cannot distract
-from the crippling economic cost of premature death (e.g., Symons et
-al., 2019; Wijnen et al., 2019), not to mention the long-term
-consequences for survivors, measured in sometimes crushing emotional and
-physical pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier
-et al., n.d.).
-
-As the statistics suggest, the burden of road accidents is also not
-borne evenly. There are important disparities at the international
-level, where the odds of death due to a road crash are three times
-higher in low-income countries compared to high-income countries; in
-fact, no reductions in road accident-related fatalities were appreciated
-in low-income countries between 2013 and 2016 (World Health
-Organization, 2019). In the case of high-income countries, where
-substantial gains in road safety have been observed for years, said
-gains have been unevenly distributed; thus, while fatal crashes
-involving older adults in the United States and Great Britain declined
-between 1997 and 2010 (despite an increase in the number of older
-people), the trend remained stable or increased slightly in Australia in
-roughly the same period (Thompson et al., 2018). There are also
-systematic differences in the impact of road accidents. For example, in
-a study in the United States, Obeng (2011) reported that the impact of
-covariates of crash severity varied in magnitude by gender, and in some
-cases were not even the same. More recently, Regev et al.~(2018) used
-adjusted crash risk to find that the risk of crashes in Great Britain
-peaked for people 21 to 29 years of age; on the other hand, the risk of
-fatal injuries for older drivers was constant, irrespective of the
-seriousness of the crash - which highlights the perils of accidents at
-older ages. Other studies have concentrated on the consequences of road
-accidents on the young (e.g., Peek-Asa et al., 2010), the old (e.g.,
-Rakotonirainy et al., 2012), as well as pedestrians and cyclists (e.g.,
-Hanson et al., 2013; McArthur et al., 2014).
-
-Given the relevance and cost of the matter, as well as the important
-variations of the impacts among different population segments, numerous
-efforts efforts exist to better understand the factors that affect road
-safety - including the probable consequences of the crash. Consequently,
-a stream of research in the analysis of road accidents is concerned with
-the severity of crashes. In particular, multivariate analysis of crash
-severity is a useful way to clarify the way various factors can affect
-the outcome of an incident, to discriminate between various levels of
-injury, from no injury (i.e., property damage only), to different
-degrees of injury up to and including fatality. This is an active area
-of research (Savolainen et al., 2011), and one where methodological
-developments have aimed at improving the reliability, accuracy, and
-precision of models.
-
-This paper aspires to contribute to the literature on crash severity by
-describing an approach to model the interactions between participants in
-a crash, in the context of incidents involving two parties. The
-importance of these interactions has been recognized in the existing
-literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al.,
-2017), and in this paper we present a systematic approach to model the
-way participants in a crash interact to influence the severity of the
-accident for each party. By way of application, we demonstrate the
-approach through an empirical analysis of data drawn from Canada's
-National Collision Database. Three levels of crash severity (no
-injury/injury/fatality) are analyzed using ordered logit models and
-covariates for the participants in the crash and the conditions of the
-crash. The results suggest that there are important interactions between
-the various descriptors of the participants in the crash. More
-generally, the example shows how modelling the interactions between
-participants explicitly leads to richer insights into the covariates of
-crash severity, in addition to better model fits. The approach described
-in this paper is intuitive, simple to implement, and can be adapted to
-advanced modelling methods with ease.
-
-The rest of this paper is structure as follows. In Section
-\ref{sec:review-of-methods} we present a concise review of the methods
-used to analyze crash severity, with a particular focus on techniques
-that consider the interactions between participants in a crash.
-
-\hypertarget{sec:review-of-methods}{%
-\section{Methodological approaches in crash severity
-analysis}\label{sec:review-of-methods}}
-
-Methods:
-
-Iranitalab and Khattak (2017); Khan et al. (2015); Chang and Wang
-(2006): machine learning Quddus et al. (2010): ordinal models Kadilar
-(2016): conditional logistic regression Mooradian et al. (2013): partial
-proportional odds models Ma et al. (2008): counts
-
-Interplay between participants:
-
-Chiou et al. (2013)
-
-Lee and Li (2014) consider effect on crash severity of interactions
-between different types of vehicles. This they do by subsetting the
-dataset and estimating independent models for each subset of data. Since
-they consider three types of vehicles, namely cars (C), light trucks
-(L), and heavy trucks (H), they work with nine datasets, for each type
-of interactions (i.e., C-C, C-L, C-H, and so on).
-
-Li et al. (2017)
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-Words go here.
-
-\hypertarget{illustrative-application}{%
-\section{Illustrative application}\label{illustrative-application}}
-
-Words go here.
-
-\hypertarget{further-considerations}{%
-\section{Further considerations}\label{further-considerations}}
-
-Words go here.
-
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
-
-Words go here.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Chang2006analysis}{}%
-Chang, L.Y., Wang, H.W., 2006. Analysis of traffic injury severity: An
-application of non-parametric classification tree techniques. Accident
-Analysis and Prevention 38, 1019--1027.
-doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}
-
-\leavevmode\hypertarget{ref-Chiou2013modeling}{}%
-Chiou, Y.C., Hwang, C.C., Chang, C.C., Fu, C., 2013. Modeling
-two-vehicle crash severity by a bivariate generalized ordered probit
-approach. Accident Analysis and Prevention 51, 175--184.
-doi:\href{https://doi.org/10.1016/j.aap.2012.11.008}{10.1016/j.aap.2012.11.008}
-
-\leavevmode\hypertarget{ref-Devlin2019road}{}%
-Devlin, A., Beck, B., Simpson, P.M., Ekegren, C.L., Giummarra, M.J.,
-Edwards, E.R., Cameron, P.A., Liew, S., Oppy, A., Richardson, M., Page,
-R., Gabbe, B.J., 2019. The road to recovery for vulnerable road users
-hospitalised for orthopaedic injury following an on-road crash. Accident
-Analysis and Prevention 132, 10.
-doi:\href{https://doi.org/10.1016/j.aap.2019.105279}{10.1016/j.aap.2019.105279}
-
-\leavevmode\hypertarget{ref-Hanson2013severity}{}%
-Hanson, C.S., Noland, R.B., Brown, C., 2013. The severity of pedestrian
-crashes: An analysis using google street view imagery. Journal of
-Transport Geography 33, 42--53.
-doi:\href{https://doi.org/10.1016/j.jtrangeo.2013.09.002}{10.1016/j.jtrangeo.2013.09.002}
-
-\leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
-Iranitalab, A., Khattak, A., 2017. Comparison of four statistical and
-machine learning methods for crash severity prediction. Accident
-Analysis and Prevention 108, 27--36.
-doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}
-
-\leavevmode\hypertarget{ref-Kadilar2016effect}{}%
-Kadilar, G.O., 2016. Effect of driver, roadway, collision, and vehicle
-characteristics on crash severity: A conditional logistic regression
-approach. International Journal of Injury Control and Safety Promotion
-23, 135--144.
-doi:\href{https://doi.org/10.1080/17457300.2014.942323}{10.1080/17457300.2014.942323}
-
-\leavevmode\hypertarget{ref-Khan2015exploring}{}%
-Khan, G., Bill, A.R., Noyce, D.A., 2015. Exploring the feasibility of
-classification trees versus ordinal discrete choice models for analyzing
-crash severity. Transportation Research Part C-Emerging Technologies 50,
-86--96.
-doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}
-
-\leavevmode\hypertarget{ref-Lee2014analysis}{}%
-Lee, C., Li, X.C., 2014. Analysis of injury severity of drivers involved
-in single- and two-vehicle crashes on highways in ontario. Accident
-Analysis and Prevention 71, 286--295.
-doi:\href{https://doi.org/10.1016/j.aap.2014.06.008}{10.1016/j.aap.2014.06.008}
-
-\leavevmode\hypertarget{ref-Li2017interplay}{}%
-Li, L., Hasnine, M.S., Habib, K.M.N., Persaud, B., Shalaby, A., 2017.
-Investigating the interplay between the attributes of at-fault and
-not-at-fault drivers and the associated impacts on crash injury
-occurrence and severity level. Journal of Transportation Safety \&
-Security 9, 439--456.
-doi:\href{https://doi.org/10.1080/19439962.2016.1237602}{10.1080/19439962.2016.1237602}
-
-\leavevmode\hypertarget{ref-Ma2008multivariate}{}%
-Ma, J.M., Kockelman, K.M., Damien, P., 2008. A multivariate
-poisson-lognormal regression model for prediction of crash counts by
-severity, using bayesian methods. Accident Analysis and Prevention 40,
-964--975.
-doi:\href{https://doi.org/10.1016/j.aap.2007.11.002}{10.1016/j.aap.2007.11.002}
-
-\leavevmode\hypertarget{ref-McArthur2014spatial}{}%
-McArthur, A., Savolainen, P.T., Gates, T.J., 2014. Spatial analysis of
-child pedestrian and bicycle crashes development of safety performance
-function for areas adjacent to schools. Transportation Research Record
-57--63. doi:\href{https://doi.org/10.3141/2465-08}{10.3141/2465-08}
-
-\leavevmode\hypertarget{ref-Merlin2007stress}{}%
-Merlin, E.P.R., Gonzalez-Forteza, C., Lira, L.R., Tapia, J.A.J., 2007.
-Post-traumatic stress disorder in patients with non intentional injuries
-caused by road traffic accidents. Salud Mental 30, 43--48.
-
-\leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
-Mooradian, J., Ivan, J.N., Ravishanker, N., Hu, S., 2013. Analysis of
-driver and passenger crash injury severity using partial proportional
-odds models. Accident Analysis and Prevention 58, 53--58.
-doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}
-
-\leavevmode\hypertarget{ref-Obeng2011gender}{}%
-Obeng, K., 2011. Gender differences in injury severity risks in crashes
-at signalized intersections. Accident Analysis and Prevention 43,
-1521--1531.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.004}{10.1016/j.aap.2011.03.004}
-
-\leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
-Peek-Asa, C., Britton, C., Young, T., Pawlovich, M., Falb, S., 2010.
-Teenage driver crash incidence and factors influencing crash injury by
-rurality. Journal of Safety Research 41, 487--492.
-doi:\href{https://doi.org/10.1016/j.jsr.2010.10.002}{10.1016/j.jsr.2010.10.002}
-
-\leavevmode\hypertarget{ref-Pelissier2019medical}{}%
-Pelissier, C., Fort, E., Fontana, L., Hours, M., n.d. Medical and
-socio-occupational predictive factors of psychological distress 5 years
-after a road accident: A prospective study. Social Psychiatry and
-Psychiatric Epidemiology 13.
-doi:\href{https://doi.org/10.1007/s00127-019-01780-0}{10.1007/s00127-019-01780-0}
-
-\leavevmode\hypertarget{ref-Quddus2010road}{}%
-Quddus, M.A., Wang, C., Ison, S.G., 2010. Road traffic congestion and
-crash severity: Econometric analysis using ordered response models.
-Journal of Transportation Engineering-Asce 136, 424--435.
-doi:\href{https://doi.org/10.1061/(asce)te.1943-5436.0000044}{10.1061/(asce)te.1943-5436.0000044}
-
-\leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
-Rakotonirainy, A., Steinhardt, D., Delhomme, P., Darvell, M., Schramm,
-A., 2012. Older drivers' crashes in queensland, australia. Accident
-Analysis and Prevention 48, 423--429.
-doi:\href{https://doi.org/10.1016/j.aap.2012.02.016}{10.1016/j.aap.2012.02.016}
-
-\leavevmode\hypertarget{ref-Regev2018crash}{}%
-Regev, S., Rolison, J.J., Moutari, S., 2018. Crash risk by driver age,
-gender, and time of day using a new exposure methodology. Journal of
-Safety Research 66, 131--140.
-doi:\href{https://doi.org/10.1016/j.jsr.2018.07.002}{10.1016/j.jsr.2018.07.002}
-
-\leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
-Savolainen, P.T., Mannering, F., Lord, D., Quddus, M.A., 2011. The
-statistical analysis of highway crash-injury severities: A review and
-assessment of methodological alternatives. Accident Analysis and
-Prevention 43, 1666--1676.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.025}{10.1016/j.aap.2011.03.025}
-
-\leavevmode\hypertarget{ref-Symons2019reduced}{}%
-Symons, J., Howard, E., Sweeny, K., Kumnick, M., Sheehan, P., 2019.
-Reduced road traffic injuries for young people: A preliminary investment
-analysis. Journal of Adolescent Health 65, S34--S43.
-doi:\href{https://doi.org/10.1016/j.jadohealth.2019.01.009}{10.1016/j.jadohealth.2019.01.009}
-
-\leavevmode\hypertarget{ref-Thompson2018trends}{}%
-Thompson, J.P., Baldock, M.R.J., Dutschke, J.K., 2018. Trends in the
-crash involvement of older drivers in australia. Accident Analysis and
-Prevention 117, 262--269.
-doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
-
-\leavevmode\hypertarget{ref-Wijnen2019analysis}{}%
-Wijnen, W., Weijermars, W., Schoeters, A., Berghe, W. van den, Bauer,
-R., Carnis, L., Elvik, R., Martensen, H., 2019. An analysis of official
-road crash cost estimates in european countries. Safety Science 113,
-318--327.
-doi:\href{https://doi.org/10.1016/j.ssci.2018.12.004}{10.1016/j.ssci.2018.12.004}
-
-\leavevmode\hypertarget{ref-WHO2019global}{}%
-World Health Organization, 2019. Global status report on road safety
-2018 (2018). Geneva.
-
 \end{document}

### 2020-01-15 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Added Model 4 ensemble and verified all models for convergence

@@
             pdfauthor={},
-            pdftitle={An assessment of strategies to model opponent effects in crash severity analysis},
+            pdftitle={An empirical assessment of strategies to model opponent effects in crash severity analysis},
             colorlinks=false,
@@
 % Pandoc toggle for numbering sections (defaults to be off)
+
 % Pandoc header
+\usepackage[margin=1in]{geometry}
 \usepackage{booktabs}
@@
 
-\usepackage[margin=1in]{geometry}
-
 \begin{document}
@@
 
-  \title{An assessment of strategies to model opponent effects in crash severity
-analysis}
-    \author[Some University]{Author 1\corref{c1}}
+  \title{An empirical assessment of strategies to model opponent effects in crash
+severity analysis}
+    \author[Some University]{Author 1\corref{Corresponding Author}}
    \ead{a1@example.com} 
-   \cortext[c1]{Corresponding Author}
     \author[Some Institute of Technology]{Author 2}
    \ead{a2@example.com} 
-
       \address[Some University]{Department, Street, City, State, Zip}
@@
 that consider the interactions between participants in a crash. Section
-\ref{sec:methods} describes the data requirements, data preprocessing,
-and the modelling strategy. Model estimation is presented in
-\ref{sec:application} and the assessment of models is in Section
-\ref{sec:assessment}. We then present some additional thoughts about the
-applicability of this approach in Section
-\ref{sec:further-considerations} before offering some concluding remarks
-in Section \ref{sec:concluding-remarks}.
+\ref{sec:application} describes the data requirements, data
+preprocessing, and the modelling strategy, along with the results of
+model estimation. The results of assessing the models and the discussion
+of these results is found in Section \ref{sec:assessment}. We then
+present some additional thoughts about the applicability of this
+approach in Section \ref{sec:further-considerations} before offering
+some concluding remarks in Section \ref{sec:concluding-remarks}.
 
@@
 analysis}\label{sec:review-of-methods}}
+
+\hypertarget{general-considerations}{%
+\subsection{General considerations}\label{general-considerations}}
 
@@
 Khan et al., 2015), Poisson models for counts (e.g., Ma et al., 2008),
-unordered logit/probit models {[}{]}, as well as ordered logit/probit
-models (e.g., Rifaat and Chin, 2007), with numerous variants, such as
-random parameters/mixed logit (e.g., Aziz et al., 2013; Haleem and Gan,
-2013), partial proportional odds models (e.g., Mooradian et al., 2013;
-Sasidharan and Menendez, 2014), and the use of copulas (e.g., Wang et
-al., 2015). Recent reviews of methods include Savolainen et al.~(2011)
-and Shamsunnahar and Eluru (2013). Shamsunnahar and Eluru (2013) in
-particular conducted an extensive comparison of models for discrete
-outcomes and found that while the difference between the performance of
-unordered models and ordered models was so small as to make not
-difference, ordered models are usually more parsimonious since only one
-latent functions needs to be estimated for all outcomes, as opposed to
-one for each outcome in unordered modelling mechanisms.
+unordered logit/probit models (e.g., Tay et al., 2011), as well as
+ordered logit/probit models (e.g., Rifaat and Chin, 2007), with numerous
+variants, such as random parameters/mixed logit (e.g., Aziz et al.,
+2013; Haleem and Gan, 2013), partial proportional odds models (e.g.,
+Mooradian et al., 2013; Sasidharan and Menendez, 2014), and the use of
+copulas (e.g., Wang et al., 2015). Recent reviews of methods include
+Savolainen et al.~(2011) and Shamsunnahar and Eluru (2013).
 
@@
 severity often include variables in several categories, as shown with
-examples in Table \ref{tab:table-variable-categories} (also see Montella
-et al., 2013). Many crash databases (but not by any means all) also
-account for the multievent nature of many crashes. In this way, there
-are crashes that involve a single traffic unit (e.g., Kim et al., 2013;
-Gong and Fan, 2017), others that involve two traffic units (e.g., Tarrao
-et al., 2014; Wang et al., 2015), and more rarely there are
+examples in Table \ref{tab:variable-categories} (also see Montella et
+al., 2013). Many crash databases and analyses also account for the
+multievent nature of many crashes. Participants may have had different
+roles in a crash depending on their context, with some acting as
+operators of a vehicle (i.e., drivers, bicyclists), while others were
+passengers. They also may differ depending on what type of traffic unit
+they were, for example occupants of a light duty vehicle or a truck,
+motorcyclists, or pedestrians. The multiplicity of roles makes for
+complicated modelling choices when trying to understand the severity of
+injuries; for example, what is the unit of analysis, the person, the
+traffic unit, or the collision? Not surprisingly, it is possible to find
+examples of studies that adopt different perspectives. A common
+simplifying strategy to specify a model is to consider only
+\emph{drivers} and/or only \emph{single-vehicle} crashes (e.g., Kim et
+al., 2013; Gong and Fan, 2017; Lee and Li, 2014; Osman et al., 2018).
+This strategy reduces the dimensions of the event, and it becomes
+possible, for example, to equate the traffic unit to the person for
+modelling purposes.
+
+The situation becomes more complex when dealing with events that involve
+two traffic units (e.g., Tarrao et al., 2014; Wang et al., 2015) and
 multi-traffic unit crashes (e.g., Wu et al., 2014; Bogue et al., 2017).
-Likewise, each traffic unit can possibly involve more than one person
-(e.g., driver and passengers). Thus, depending on the incident, each
-record in a crash database may include unique identifier for the crash,
-as well as identifiers for the traffic units (or identifiers for dummy
-objects such as a light pole that was hit by a vehicle), and the people
-involved in the crash.
+Different strategies have been developed to study these, more complex
+cases. A number of studies advocate the estimation of separate models
+for different participants and/or situations. In this way, Wang and
+Kockelman (2005) estimated models for single-vehicle and two-vehicle
+crashes, while Savolainen and Mannering (2007) estimated models for
+single-vehicle and multi-vehicle crashes. More recently, Duddu et
+al.~(2018) and Penmetsa et al.~(2017) presented research that estimated
+separate models for at-fault and not-at-fault drivers. The strategy of
+estimating separate models also relies on subsetting the dataset,
+although it is possible to link the relevant models more tightly by
+means of a common covariance structure, as is the case of bivariate
+models (e.g., Chiou et al., 2013; Chen et al., 2019) or models with
+copulas (e.g., Rana et al., 2010; Shamsunnahar et al., 2014; Wang et
+al., 2015).
+
+A related strategy to specify a crash severity model is to organize the
+data in such a way that it is possible to model the influence of the
+attributes of the opponent in a crash. There are numerous examples of
+studies that consider at least some characteristics of the oponents in
+two- or multi-vehicle crashes. For example, Wang and Kockelman (2005)
+considered the type of the opposing vehicle in their model for
+two-vehicle collisions. Similarly, Tarrao et al.~(2014) included in
+their analysis the age, wheelbase, weight, and engine size of the
+opposing vehicle while Bogue et al.~(2017) used the body type of the
+opposing vehicle. Penmetsa et al.~(2017) and Duddu et al.~(2018) are two
+of the most comprehensive examples of using opponent's information, with
+the physical condition, sex, age, and vehicle type of the opponent. The
+twin strategies of subsetting the sample and using the attributes of the
+opponent are not mutually exclusive, but neither are they consistently
+used together, as a scan of the literature reveals.
 
@@
 
-\caption{\label{tab:table-variable-categories}\label{tab:table-variable-categories}Categories of variables used in the analysis of crash severity with examples}
+\caption{\label{tab:table-variable-categories}\label{tab:variable-categories}Categories of variables used in the analysis of crash severity with examples}
 \centering
+\fontsize{7}{9}\selectfont
 \begin{tabular}[t]{l>{\raggedright\arraybackslash}p{22em}}
@@
 Road-related & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
+Opponent-related & Attributes of the opponent, e.g., age of opponent, gender of opponent, opponent vehicle type, etc.\\
 \bottomrule
@@
 
-Interplay between participants:
-
-Chiou et al. (2013)
-
-Lee and Li (2014) consider effect on crash severity of interactions
-between different types of vehicles. This they do by subsetting the
-dataset and estimating independent models for each subset of data. Since
-they consider three types of vehicles, namely cars (C), light trucks
-(L), and heavy trucks (H), they work with nine datasets, for each type
-of interactions (i.e., C-C, C-L, C-H, and so on). Tarrao et al. (2014)
-
-Mannering et al. (2016)
-
-Li et al. (2017)
-
-Salon and McIntyre (2018)
-
-Chen et al. (2019)
-
-Wu et al. (2014) two-vehicle crashes do not really consider the
-interactions.
-
-Many models use a latent-variable approach, whereby the severity of the
-crash (observed) is linked to an underlying latent variable that is a
-function of the variables: \[
-y_{ik}^*=\sum_{l=1}\beta_lp_{ikl} + \sum_{m=1}\gamma_mc_{km} + \epsilon_{ik}
-\]
-
-The left-hand side of the expression above (\(y_{ik}^*\)) is a latent
+\hypertarget{modelling-techniques}{%
+\subsection{Modelling techniques}\label{modelling-techniques}}
+
+With respect to model structures, Shamsunnahar and Eluru (2013)
+conducted an extensive comparison of models for discrete outcomes and
+found only small differences in the performance of unordered models and
+ordered models; however, ordered models are usually more parsimonious
+since only one latent functions needs to be estimated for all outcomes,
+as opposed to one for each outcome in unordered modelling mechanisms.
+Bogue et al.~(2017) also compared unordered and ordered models in the
+form of the mixed multinomial logit and a modified rank ordered logit,
+respectively, and found that the ordered model performed best. To keep
+the empirical assessment managable we will consider only the ordinal
+logit model, and will comment on potential extensions in Section
+\ref{sec:further-considerations}.
+
+The ordinal model is a latent-variable approach, whereby the severity of
+the crash (observed) is linked to an underlying latent variable that is
+a function of the variables of interest, as follows:
+
+\begin{equation}
+\label{eq:latent-function}
+y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
+\end{equation}
+
+The left-hand side of the expression above (\(y_{itk}^*\)) is a latent
 (unobservable) variable that is associated with the severity of crash
-\(k\) (\(k=1,\cdots,K\)) for participant \(i\) (\(i = 1,\cdots,n\)). The
-right-hand side of the expression is split in three parts. The first
+\(k\) (\(k=1,\cdots,K\)) for participant \(i\) in traffic unit \(t\).
+The right-hand side of the expression is split in four parts. The first
 part collects \(l=1,\cdots,L\) individual attributes \(p\) for
-participant \(i\) in crash \(k\); these could relate to the person
-(e.g., age and gender) or be individual attributes of the traffic unit
-(e.g., maneuver or vehicle type). The second part collects
-\(m=1,\cdots,M\) attributes \(c\) related to the crash \(k\), including
-crash-related and road-related data. The last part is a random term
-specific to participant \(i\) in crash \(k\).
+participant \(i\) in traffic unit \(t\) and crash \(k\); these could
+relate to the person (e.g., age, gender, and road user class). The
+second part collects \(m=1,\cdots,M\) attributes \(u\) related to
+traffic unit \(t\) in crash \(k\); these could be items such as maneuver
+or vehicle type. The third part collects \(q=1,\cdots,Q\) attributes
+\(c\) related to the crash \(k\), including crash-related and
+road-related data, such as weather conditions, road alignment, and type
+of surface. Lastly, the fourth element is a random term specific to
+participant \(i\) in traffic unit \(t\) and crash \(k\). The function
+consists of a total of \(Z=L+M+Q\) covariates and associated parameters.
+
+When opponent-related variables are included, the function is augmented
+as follows:
+
+\begin{equation}
+\label{eq:latent-function-with-opponent-variables}
+y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
+\end{equation}
+
+The additional part collects \(r=1,\cdots,R\) attributes \(o\) related
+to individual \(j\) in traffic unit \(v\) and crash \(k\) that opposed
+individual \(i\) in traffic unit \(t\) and crash \(k\). These could be
+individual characteristics of the opponent (such as age and gender)
+and/or characteristics of the opposing vehicle (such as vehicle type or
+weight). To qualify as an opponent, individual \(j\) must have been a
+participant in crash \(k\) but operating traffic unit \(v\ne t\).
+Sometimes the person \emph{is} the traffic unit, as is the case of a
+pedestrian. And we exclude passengers of vehicles as opponents, since
+they do not operate the traffic unit. In case the opponent attributes
+include only characteristics of the traffic unit, the condition for the
+traffic unit to be an opponent is that it participated in crash \(k\)
+and was different from \(t\). After introducing this new set of terms,
+the latent function now consists of a total of \(Z=L+M+Q+R\) covariates
+and associated parameters.
+
+For conciseness, in what follows we will abbreviate the function as
+follows:
+
+\begin{equation}
+\label{eq:latent-function-compact}
+y_{itk}^*=\sum_{z=1}^Z\theta_zx_{itkz} + \epsilon_{itk}
+\end{equation}
 
@@
 non-fatal/fatal), we can relate the latent variable to the outcome as
-follows: \[
-y_{ik} = 
+follows:
+
+\begin{equation}
+\label{eq:latent-function-2-outcomes}
+y_{itk} = 
 \begin{cases}
-\text{fatal} & \text{if } y_{ik}^*>0\\
-\text{non-fatal} & \text{if } y_{ik}^*\leq0
+\text{fatal} & \text{if } y_{itk}^*>0\\
+\text{non-fatal} & \text{if } y_{itk}^*\leq0
 \end{cases}
-\]
+\end{equation}
 
@@
 crash is not fully determined. However, we can make the following
-probability statement: \[
-P(y_i = \text{fatal}) = P(y_i^* > 0)
-\] In other words, the probability of a fatal accident equals the
-probability that the latent variable is greater than zero. This implies
-(see Maddala, 1986, p. 22): \[
+probability statement:
+
+\begin{equation}
+\label{eq:probability-2-outcomes}
+P(y_{itk} = \text{fatal}) = P(y_{itk}^* > 0)
+\end{equation}
+
+In other words, the probability that individual \(i\) in traffic unit
+\(k\) and crash \(k\) was a fatality equals the probability that the
+latent variable is greater than zero. This implies (see Maddala, 1986,
+p. 22):
+
+\begin{equation}
+\label{eq:probability-2-outcomes-2}
 \begin{array}{rl}\
-P(y_i = \text{fatal}) &= P(\sum_{l=1}\beta_lp_{ikl} + \sum_{m=1}\gamma_mc_{km} + \epsilon_{ik} > 0)\\ 
-&=P(\epsilon_{ik} > -\sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km})
+P(y_{itk} = \text{fatal}) &= P(\sum_{z=1}\theta_zp_{itkz} + \epsilon_{itk} > 0)\\ 
+&=P(\epsilon_{itk} > -\sum_{z=1}\theta_zp_{itkz})
 \end{array}
-\]
-
-It the random terms \(\epsilon_i\) are assumed to follow the logistic
-distribution, the the binary logit model is obtained; if they are
-assumed to follow the normal distribution, the binary probit model is
-obtained.
-
-More often, though, the outcome is recorded using more categories, for
-example property damage only (PDO)/injury/fatality. A similar approach
-can be adopted, with a latent variable that relates to the outcome as
-follows: \[
-y_i = 
+\end{equation}
+
+If the random terms \(\epsilon_{itk}\) are assumed to follow the
+logistic distribution, the the binary logit model is obtained; if they
+are assumed to follow the normal distribution, the binary probit model
+is obtained. More often, though, the outcome is recorded using more
+categories, for example property damage only (PDO)/injury/fatality. A
+similar approach can be adopted, with a latent variable that relates to
+the outcome as follows:
+
+\begin{equation}
+\label{eq:latent-function-ordered-outcomes}
+y_{itk} = 
 \begin{cases}
-\text{fatality} & \text{if } y_i^*> k_2\\
-\text{injury} & \text{if } k_1< y_i^*<k_2\\
-\text{PDO} & \text{if } y_i^*<k_1
+\text{fatality} & \text{if } y_{itk}^*> \mu_2\\
+\text{injury} & \text{if } \mu_1< y_{itk}^*< \mu_2\\
+\text{PDO} & \text{if } y_{itk}^*< \mu_1
 \end{cases}
-\] where \(k_1\) and \(k_2\) are estimable thresholds. In this case, the
-associated probability statements are as follows: \[
+\end{equation}
+
+\noindent where \(\mu_1\) and \(\mu_2\) are estimable thresholds. In
+this case, the associated probability statements are as follows:
+
+\begin{equation}
+\label{eq:probability-ordered-outcomes}
 \begin{array}{rcl}\
-P(y_{ik} = \text{PDO}) &=& 1 - P(y_{ik} = \text{injury}) - P(y_{ik} = \text{fatality})\\ 
-P(y_{ik} = \text{injury}) &=& P(k_1 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} < \epsilon_{ik} < k_2 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} )\\
-P(y_{ik} = \text{fatality}) &=& P(\epsilon_{ik} < k_1 - \sum_{l=1}\beta_lp_{ikl} - \sum_{m=1}\gamma_mc_{km} )
+P(y_{itk} = \text{PDO}) &=& 1 - P(y_{itk} = \text{injury}) - P(y_{itk} = \text{fatality})\\ 
+P(y_{itk} = \text{injury}) &=& P(\mu_1 - \sum_{z=1}\theta_zp_{itkz} < \epsilon_{itk} < \mu_2 - \sum_{z=1}\theta_zp_{itkz})\\
+P(y_{itk} = \text{fatality}) &=& P(\epsilon_{itk} < \mu_1 - \sum_{z=1}\theta_zp_{itkz})
 \end{array}
-\]
+\end{equation}
 
@@
 
-\hypertarget{sec:methods}{%
-\section{Methods}\label{sec:methods}}
-
-\hypertarget{data-considerations}{%
-\subsection{Data considerations}\label{data-considerations}}
-
-Words go here
-
-Montella et al. (2013)
-
-\hypertarget{model-specification}{%
-\subsection{Model specification}\label{model-specification}}
-
-More words go here
+\hypertarget{model-specification-strategies}{%
+\subsection{Model specification
+strategies}\label{model-specification-strategies}}
+
+In this paper we consider three general model specification strategies,
+as follows:
+
+\begin{itemize}
+\tightlist
+\item
+  Strategy 1. Introducing opponent-related variables
+\item
+  Strategy 2. Single-level model and multi-level (hierarchical) model
+  specifications
+\item
+  Strategy 3. Full sample and sample subsetting
+\end{itemize}
+
+Introduction of opponent related-variables was explained in the
+preceding subsection. In this way, a base model is given by Equation
+\ref{eq:latent-function}, whereas Strategy 1 (inclusion of
+opponent-related variables) is Equation
+\ref{eq:latent-function-with-opponent-variables}. These two equations
+are also examples of single-level model. Next we describe Strategies 2
+and 3.
+
+\hypertarget{hierarchical-model-specification}{%
+\subsubsection{Hierarchical model
+specification}\label{hierarchical-model-specification}}
+
+We can choose to conceptualize the event leading to the outcome as a
+hierarchical process. There are a few different ways of doing this. For
+example, the hierarchy could be based on individuals in traffic units.
+In this case, we can rewrite the latent function as follows:
+
+\begin{equation}
+\label{eq:latent-function-with-hierarchical-traffic-unit}
+y_{itk}^*=\sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
+\end{equation}
+
+The coefficients of the traffic unit nest the individual attributes as
+follows. For any given coefficient \(q\):
+
+\begin{equation}
+\label{eq:hierarchical-traffic-unit-coefficients}
+\beta_{m}=\sum_{l=1}^L\beta_{ml}p_{itkl} 
+\end{equation}
+
+Therefore, the corresponding term in the latent function becomes
+(assuming that \(p_{itk1} = 1\), i.e., it is a constant term):
+
+\begin{equation}
+\label{eq:hierarchical-traffic-unit-coefficients}
+\begin{array}{rcl}\
+\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL}\big)u_{tkm}\\ 
+ &=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm}
+\end{array}
+\end{equation}
+
+As an alternative, the nesting unit could be the interaction
+person-opponent, in which case the opponent-level attributes are nested
+
+\begin{equation}
+\label{eq:latent-function-with-opponent-variables}
+y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
+\end{equation}
+
+\noindent with any person-level coefficient \(l\) that we wish to expand
+defined as follows:
+
+\begin{equation}
+\label{eq:hierarchical-traffic-unit-coefficients}
+\alpha_{l}=\sum_{r=1}^R\alpha_{lr}o_{jvkr}
+\end{equation}
+
+\noindent with the same conditions as before, that \(j\ne i\) is the
+operator of traffic unit \(v\ne t\). The corresponding term in the
+latent function is now (assuming that \(o_{jvk1}=1\), i.e., it is a
+constant term):
+
+\begin{equation}
+\label{eq:hierarchical-traffic-unit-coefficients}
+\begin{array}{rcl}\
+\alpha_{l}p_{itkl} &=& \big(\alpha_{l1} + \alpha_{l2}o_{jvk2} + \cdots + \alpha_{lR}o_{jvkR} \big)p_{itkl}\\
+ &=& \alpha_{l1}p_{itkl} + \alpha_{l2}o_{jvk2}p_{itkl} + \cdots + \alpha_{lR}o_{jvkR}p_{itkl}
+\end{array} 
+\end{equation}
+
+Alterted readers will identify this model specification strategy as
+Casetti's expansion method (Casetti, 1972). This is a deterministic
+strategy for modelling contextual effects which, when augmented with
+random components, becomes the well-known multi-level modelling method
+(Hedeker and Gibbons, 1994, more on this in Section
+\ref{sec:further-considerations}). It is worthwhile to note that
+higher-order hierarchical effects are possible; for instance, individual
+attributes nested within traffic units, which in turn are nested within
+collisions. We do not explore this further in the current paper.
+
+\hypertarget{sample-subsetting}{%
+\subsubsection{Sample subsetting}\label{sample-subsetting}}
+
+The third model specification strategy that we will consider is
+subsetting the sample. This is applicable in conjunction with any of the
+other strategies discussed above. In essence, we define the latent
+function with restrictions as follows. Consider a continuous variable,
+e.g., age of person, and we wish to concentrate the analysis on older
+adults (e.g., Dissanayake and Lu, 2002). The latent function is defined
+as desired (see above), however, the following restriction might be
+applied to the sample
+
+\begin{equation}
+\label{eq:sampling-age}
+\text{Age of individual } i \text{ in traffic unit } t \text{ in crash } k = 
+\begin{cases}
+\ge 65 & \text{use record } itk\\
+< 65 & \text{do not use record } itk
+\end{cases}
+\end{equation}
+
+Suppose instead that we are interested in crashes by or against a
+specific type of traffic unit (e.g., pedestrians, Amoh-Gyimah et al.,
+2017):
+
+\begin{equation}
+\label{eq:sampling-pedestrian}
+\text{Road user class of individual } i \text{ in traffic unit } t \text{ in crash } k = 
+\begin{cases}
+\text{Pedestrian} & \text{use record } itk\\
+\text{Not pedestrian} & \text{do not use record } itk
+\end{cases}
+\end{equation}
+
+\noindent or:
+
+\begin{equation}
+\label{eq:sampling-pedestrian-opponent}
+\text{Road user class of individual } j \text{ in traffic unit } v \text{ in crash } k = 
+\begin{cases}
+\text{Pedestrian} & \text{use record } jvk\\
+\text{Not pedestrian} & \text{do not use record } jvk
+\end{cases}
+\end{equation}
+
+More generally, for any variable \(x\) of interest:
+
+\begin{equation}
+\label{eq:sampling-general}
+x_{itk} = 
+\begin{cases}
+\text{Condition: TRUE} & \text{use record } itk\\
+\text{Condition: FALSE} & \text{do not use record } itk
+\end{cases}
+\end{equation}
+
+Several conditions can be imposed to subset the sample in any way that
+the analyst deems appropriate or suitable.
 
 \hypertarget{sec:application}{%
-\section{Application}\label{sec:application}}
+\section{Setting for empirical assessment}\label{sec:application}}
+
+In this section we present the setting for the empirical assessment of
+the modelling strategies discussed in Section
+\ref{sec:review-of-methods}, namely matters related to data and model
+estimation.
 
 \begin{quote}
-\textbf{Note:} that this paper presents reproducible research. The
-source file is an R Markdown document. All code and data necessary to
-reproduce the analysis are available from the following anonymous Drive
-folder:
+\textbf{Note:} this paper presents reproducible research. The source
+file is an R Markdown document. All code and data necessary to reproduce
+the analysis are available from the following anonymous Drive folder:
 \end{quote}
@@
 
-\hypertarget{data-for-the-application}{%
-\subsection{Data for the application}\label{data-for-the-application}}
-
-To assess the performance of the modelling strategies discussed in
-Section \ref{sec:methods}, we use data from Canada's National Collision
-Database (NCDB). This is database contains all motor vehicle collisions
-on public roads in Canada as reported by a police service. Data are
-collected by provinces and territories, and shared with the federal
-government, where data are combined, tracked, and analyzed for reporting
-of deaths, injuries, and collisions in Canada at the national level. The
-NCDB is provided by Transport Canada, the agency of the federal
-government of Canada in charge of transportation policies and programs,
-under the Open Government License - Canada version 2.0
+\hypertarget{data-for-empirical-assessment}{%
+\subsection{Data for empirical
+assessment}\label{data-for-empirical-assessment}}
+
+To assess the performance of the various modelling strategies we use
+data from Canada's National Collision Database (NCDB). This is database
+contains all motor vehicle collisions on public roads in Canada as
+reported by a police service. Data are collected by provinces and
+territories, and shared with the federal government, where data are
+combined, tracked, and analyzed for reporting of deaths, injuries, and
+collisions in Canada at the national level. The NCDB is provided by
+Transport Canada, the agency of the federal government of Canada in
+charge of transportation policies and programs, under the Open
+Government License - Canada version 2.0
 {[}https://open.canada.ca/en/open-government-licence-canada{]}.
@@
 The NCDB is available from 1999. For the purpose of this paper, we use
-the most recent year available as of this writing (2017). Furthermore,
-for assessment we also use the data corresponding to 2016. Similar to
-databases in other jurisdictions {[}see @{]}, the NCDB contains
-information pertaining to the collision, the traffic unit, and the
-person(s) involved.
-
-\begin{tabular}{ll}
+the data corresponding to 2017, which is the most recent year available
+as of this writing. Furthermore, for assessment we also use the data
+corresponding to 2016. Similar to databases in other jurisdictions (see
+Montella et al., 2013), the NCDB contains information pertaining to the
+collision, the traffic unit(s), and the person(s) involved in a crash,
+as shown in Tables \ref{tab:ncdb-descriptives-collision},
+\ref{tab:ncdb-descriptives-vehicle}, and
+\ref{tab:ncdb-descriptives-person}. Notice that, compared to Table
+\ref{tab:variable-categories}, crash-related variables and road-related
+variables are collected in a single variable class, namely
+collision-related, since they are unique for each crash.
+
+Data are organized by person; in other words, there is one record per
+participant in a collision, be they drivers, passengers, pedestrians,
+etc. The only variable directly available with respect to opponents in a
+collision is the number of vehicles involved (see models in Bogue et
+al., 2017). Therefore, the data needs to be processed to obtain
+attributes of opponents for each participant in a collision. The
+protocol to do this is described next.
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-collision}\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables}\\
 \toprule
-Variable & Details\\
+Variable & Description & Notes\\
 \midrule
-Age & Details\\
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
 \bottomrule
-\multicolumn{2}{l}{\textit{Note: }}\\
-\multicolumn{2}{l}{NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\end{tabular}
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Preprocessed data for analysis available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  C\_CASE & Unique collision identifier & Unique identifier for collisions\\
+C\_YEAR & Year & Last two digits of year.\\
+\rowcolor{gray!6}  C\_MNTH & Month & 14 levels: January - December; unknown; not reported by jurisdiction.\\
+C\_WDAY & Day of week & 9 levels: Monday - Sunday; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_HOUR & Collision hour & 25 levels: hourly intervals; unknown; not reported by jurisdiction.\\
+\addlinespace
+C\_SEV & Collision severity & 4 levels: collision producing at least one fatality; collision producing non-fatal injury; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_VEHS & Number of vehicles involved in collision & Number of vehicles: 1-98 vehicles involved; 99 or more vehicles involved; unknown; not reported by jurisdiction.\\
+C\_CONF & Collision configuration & 21 levels: SINGLE VEHICLE: Hit a moving object (e.g.
+a person or an animal); Hit a stationary object (e.g.
+a tree); Ran off left shoulder; Ran off right shoulder; Rollover on roadway; Any other single vehicle collision configuration; TWO-VEHICLES SAME DIRECTION OF TRAVEL: Rear-end collision; Side swipe; One vehicle passing to the left of the other, or left turn conflict; One vehicle passing to the right of the other, or right turn conflict; Any other two vehicle - same direction of travel configuration; TWO-VEHICLES DIFFERENT DIRECTION OF TRAVEL: Head-on collision; Approaching side-swipe; Left turn across opposing traffic; Right turn, including turning conflicts; Right angle collision; Any other two-vehicle - different direction of travel configuration; TWO-VEHICLES, HIT A PARKED VEHICLE: Hit a parked motor vehicle; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_RCFG & Roadway configuration & 15 levels: Non-intersection; At an intersection of at least two public roadways; Intersection with parking lot entrance/exit, private driveway or laneway; Railroad level crossing; Bridge, overpass, viaduct; Tunnel or underpass; Passing or climbing lane; Ramp; Traffic circle; Express lane of a freeway system; Collector lane of a freeway system; Transfer lane of a freeway system; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+C\_WTHR & Weather condition & 10 levels: Clear and sunny; Overcast, cloudy but no precipitation; Raining; Snowing, not including drifting snow; Freezing rain, sleet, hail; Visibility limitation; Strong wind; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\addlinespace
+\rowcolor{gray!6}  C\_RSUR & Road surface & 12 levels: Dry, normal; Wet; Snow (fresh, loose snow); Slush, wet snow; Icy, packed snow; Debris on road (e.g., sand/gravel/dirt); Muddy; Oil; Flooded; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+C\_RALN & Road alignment & 9 levels: Straight and level; Straight with gradient; Curved and level; Curved with gradient; Top of hill or gradient; Bottom of hill or gradient; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_TRAF & Traffic control & 21 levels: Traffic signals fully operational; Traffic signals in flashing mode; Stop sign; Yield sign; Warning sign; Pedestrian crosswalk; Police officer; School guard, flagman; School crossing; Reduced speed zone; No passing zone sign; Markings on the road; School bus stopped with school bus signal lights flashing; School bus stopped with school bus signal lights not flashing; Railway crossing with signals, or signals and gates; Railway crossing with signs only; Control device not specified; No control present; Choice is other than the preceding values; unknown; not reported by jurisdiction.\\*
+\end{longtable}
+\endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-vehicle}\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Preprocessed data for analysis available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  V\_ID & Vehicle sequence number & Number of vehicles: 1-98; Pedestrian sequence number: 99; unknown.\\
+V\_TYPE & Vehicle type & 21 levels: Light Duty Vehicle (Passenger car, Passenger van, Light utility vehicles and light duty pick up trucks); Panel/cargo van (<= 4536 KG GVWR   Panel or window type of van designed primarily for carrying goods); Other trucks and vans (<= 4536 KG GVWR); Unit trucks (> 4536 KG GVWR); Road tractor; School bus; Smaller school bus (< 25 passengers); Urban and Intercity Bus; Motorcycle and moped; Off road vehicles; Bicycle; Purpose-built motorhome; Farm equipment; Construction equipment; Fire engine; Snowmobile; Street car; Data element is not applicable  (e.g.
+dummy vehicle record created for pedestrian); Choice is other than the preceding values; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  V\_YEAR & Vehicle model year & Model year; dummy for pedestrians; unknown; not reported by jurisdiction.\\*
+\end{longtable}
+\endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-person}\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Preprocessed data for analysis available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  P\_ID & Person sequence number & Sequence number: 1-99; Not applicable (dummy for parked vehicles); not reported by jurisdiction.\\
+P\_SEX & Person sex & 5 levels: Male; Female; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_AGE & Person age & Age: less than 1 year; 1-98 years old; 99 years or older; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
+P\_PSN & Person position & Person position: Driver; Passenger front row, center; Passenger front row, right outboard (including motorcycle passenger in sidecar); Passenger second row, left outboard, including motorcycle passenger; Passenger second row, center; Passenger second row, right outboard; Passenger third row, left outboard;...; Position unknown, but the person was definitely an occupant; Sitting on someone’s lap; Outside passenger compartment; Pedestrian; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_ISEV & Medical treatment required & 6 levels: No Injury; Injury; Fatality; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\addlinespace
+P\_SAFE & Safety device used & 11 levels: No safety device used; Safety device used; Helmet worn; Reflective clothing worn; Both helmet and reflective clothing used; Other safety device used; No safety device equipped   (e.g.
+buses); Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_USER & Road user class & 6 levels: Motor Vehicle Driver; Motor Vehicle Passenger; Pedestrian; Bicyclist; Motorcyclist; Not stated/Other/Unknown.\\*
+\end{longtable}
+\endgroup{}
+
+\hypertarget{data-preprocessing-initial-filter}{%
+\subsubsection{Data preprocessing: initial
+filter}\label{data-preprocessing-initial-filter}}
+
+The database is scanned to remove all records that are not a person
+(including parked cars and other objects) or that are missing
+information (e.g., runaway vehicles). Next, records missing at least one
+of the next variables are removed: P\_USER (the road user class), P\_SEX
+(sex), P\_AGE (age), and P\_ISEV (individual-level severity of crash).
+This initial filter ensures that all records are complete from the
+perspective of key information.
+
+\hypertarget{data-preprocessing-filter-two-vehicle-collisions}{%
+\subsubsection{Data preprocessing: filter two-vehicle
+collisions}\label{data-preprocessing-filter-two-vehicle-collisions}}
+
+After the initial filter, the database is summarized to find the number
+of individual-level records that correspond to each collision (C\_CASE).
+At this point, there are 32,298 collisions, involving only one (known)
+participant, there are 46,483 collisions involving two participants,
+19,433 involving 3 participants, 8,250 involving four participants,
+3,783 involving 5 participants, 1,789 involving 6 participants, and
+1,491 collisions involving seven or more participants. These
+participants were possibly occupants in different vehicles or were
+otherwise their own traffic units, as follows: the sample includes
+174,741 drivers, 61,403 passengers, 10,798 pedestrians, 5,286
+bicyclists, and 6,564 motorcyclists.
+
+The next step is to remove all collisions involving only one
+participant. This still leaves numerous cases where multiple
+participants could have been in a single vehicle, for instance in a
+collision against a stationary object. Therefore, we proceed to use the
+vehicle sequence number can to find the number of vehicles involved in
+each collision. This reveals that there are 20,732 collisions that
+involve only one vehicle but possibly multiple participants (i.e.,
+driver and one or more passengers). In addition, there are 165,520
+collisions involving two vehicles (and possibly multiple participants).
+Finally, there are 40,242 collisions with three or more vehicles.
+
+Once we have identified the number of vehicles in each collision, we
+proceed to select all cases that involve only two vehicles. The most
+common cases in two-vehicle collisions are those that include drivers
+(40,297 collisions; reflective of the prevalence of single-occupant
+vehicles), followed by cases with passengers (14,120 collisions),
+pedestrians (5,204 collisions), byciclists (2,238 collisions), and
+motorcyclists (1,016 collisions). The distribution of individuals per
+traffic unit is as follows: 80,382 individuals are coded as being in
+V\_ID = 1, 76,523 individuals are coded as being in V\_ID = 2, and 7,932
+individuals are coded as pedestrians. In addition, 683 individuals are
+coded as being in vehicles 3 through 9, despite our earlier filter to
+retain only collisions with two vehicles. We therefore proceed to select
+only individuals assigned to vehicles 1 or 2, as well as pedestrians. As
+a result of this filter a number of cases with only one known
+participant need to be removed.
+
+\hypertarget{data-preprocessing-extract-opponent-information-and-join-to-individual-records}{%
+\subsubsection{Data preprocessing: extract opponent information and join
+to individual
+records}\label{data-preprocessing-extract-opponent-information-and-join-to-individual-records}}
+
+Up to this point, the goal of data preprocessing has been to obtain a
+complete, workable sample of individual records of participants in
+two-vehicle collisions. There are two possible cases for the collisions,
+depending on the traffic units involved: 1) vehicle vs vehicle
+collisions (``vehicle'' includes all motorized vehicles, as well as
+motorcycles/mopeds and bicycles); and 2) vehicle vs pedestrian
+collisions. To identify opponents in a collision, it is convenient to
+classify collisions by pedestrian involvement. Accordingly, we find that
+the database includes 16,636 collisions are vehicle vs pedestrian
+(possibly multiple pedestrians), and 147,594 collisions involving two
+vehicles. After splitting the database in this way, we can now extract
+relevant information about participants in the collision. This involves
+renaming the person-level variables depending so that we can distinguish
+each individual by their role as an individual or an opponent in a given
+collision. Notice that when working with individuals in vehicles, only
+the driver is considered an opponent in a collisions (i.e., passengers
+are never considered opponents).
+
+Once the personal attributes of individuals that can be considered
+opponents in a given collision have ben extracted, their information can
+be joined to the individual records by means of the collision unique
+identifier. As a result of this process, a new set of variables are now
+available for analysis: the age, sex, use of safety device, and road
+user class of the opponent, as well as the type of the opponent vehicle.
+A summary of opponent interactions and outcomes can be found in Table
+\ref{tab:summary-opponent-information}. The information there shows that
+the most commont type of opponent for drivers is other drivers, followed
+by pedestrians. The only opponent of pedestrians, on the other hand, are
+drivers. Bicyclists and motorcyclists, on the other hand, are mostly
+opposed by drivers, but occasionally by other road users as well. In
+terms of outcomes, we observe that virtually all fatalities occur when
+the opponent is a driver, and only very rarely when the opponent is a
+motorcyclist. Injuries are also more common when the opponent is a
+driver, whereas the likelihood of no injuries is higher when the
+opponent is a pedestrian or a bicyclist.
+
+\begin{table}
+
+\caption{\label{tab:summary-opponent-information}\label{tab:summary-opponent-information}Summary of opponent interactions and outcomes by road user class}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{lrrrrrrrlll}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Road User Class of Opponent} & \multicolumn{3}{c}{Outcome} & \multicolumn{3}{c}{Proportion by Road User Class} \\
+\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11}
+Road User Class & Driver & Pedestrian & Bicyclist & Motorcyclist & No Injury & Injury & Fatality & No Injury & Injury & Fatality\\
+\midrule
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{All opponents}}\\
+\hspace{1em}Driver & 97582 & 7880 & 3799 & 2498 & 59180 & 52143 & 436 & 0.52953 & 0.46657 & 0.003901\\
+\hspace{1em}Passenger & 35359 & 1282 & 667 & 818 & 19308 & 18667 & 151 & 0.50643 & 0.48961 & 0.003961\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
+\hspace{1em}Bicyclist & 3799 & 1 & 0 & 40 & 49 & 3760 & 31 & 0.01276 & 0.97917 & 0.008073\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 2498 & 30 & 40 & 338 & 204 & 2598 & 104 & 0.07020 & 0.89401 & 0.035788\\
+\addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{Opponent: Driver}}\\
+\hspace{1em}Driver & 97582 & 0 & 0 & 0 & 45493 & 51657 & 432 & 0.46620 & 0.52937 & 0.004427\\
+\rowcolor{gray!6}  \hspace{1em}Passenger & 35359 & 0 & 0 & 0 & 16672 & 18536 & 151 & 0.47151 & 0.52422 & 0.004270\\
+Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist & 3799 & 0 & 0 & 0 & 43 & 3725 & 31 & 0.01132 & 0.98052 & 0.008160\\
+\hspace{1em}Motorcyclist & 2498 & 0 & 0 & 0 & 98 & 2299 & 101 & 0.03923 & 0.92034 & 0.040432\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{Opponent: Pedestrian}}\\
+\hspace{1em}Driver & 0 & 7880 & 0 & 0 & 7693 & 187 & 0 & 0.97627 & 0.02373 & 0.000000\\
+\hspace{1em}Passenger & 0 & 1282 & 0 & 0 & 1246 & 36 & 0 & 0.97192 & 0.02808 & 0.000000\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
+\hspace{1em}Bicyclist & 0 & 1 & 0 & 0 & 0 & 1 & 0 & 0.00000 & 1.00000 & 0.000000\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 30 & 0 & 0 & 11 & 19 & 0 & 0.36667 & 0.63333 & 0.000000\\
+\addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{Opponent: Bicyclist}}\\
+\hspace{1em}Driver & 0 & 0 & 3799 & 0 & 3706 & 93 & 0 & 0.97552 & 0.02448 & 0.000000\\
+\rowcolor{gray!6}  \hspace{1em}Passenger & 0 & 0 & 667 & 0 & 649 & 18 & 0 & 0.97301 & 0.02699 & 0.000000\\
+\hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{1} -\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & -\\
+\hspace{1em}Motorcyclist & 0 & 0 & 40 & 0 & 16 & 24 & 0 & 0.40000 & 0.60000 & 0.000000\\
+\rowcolor{gray!6}  \addlinespace[0.3em]
+\multicolumn{11}{l}{\textbf{Opponent: Motorcyclist}}\\
+\hspace{1em}Driver & 0 & 0 & 0 & 2498 & 2288 & 206 & 4 & 0.91593 & 0.08247 & 0.001601\\
+\hspace{1em}Passenger & 0 & 0 & 0 & 818 & 741 & 77 & 0 & 0.90587 & 0.09413 & 0.000000\\
+\rowcolor{gray!6}  Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
+\hspace{1em}Bicyclist & 0 & 0 & 0 & 40 & 6 & 34 & 0 & 0.15000 & 0.85000 & 0.000000\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 0 & 0 & 338 & 79 & 256 & 3 & 0.23373 & 0.75740 & 0.008876\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 
-Words go here
-
-Model specification. See Table:
-
-\begin{table}[H]
+Before model estimation, the variables are prepared as follows. First,
+age is scaled from years to decades. Secondly, new variables are defined
+to describe the vehicle type. Three classes of vehicle types are
+considered: 1) light duty vehicles (which in Canada include passenger
+cars, passenger vans, light utility vehicles, and light duty pick up
+trucks); 2) light trucks (all other vehicles \(\le\) 4536 kg in gross
+vehicle weight rating); and heavy vehicles (all other vehicles \(\ge\)
+4536 in gross vehicle weight rating). Furthermore, this classification
+of vehicle types is combined with the road user class to distiguish
+between drivers and passengers of light duty vehicles, light trucks, and
+heavy vehicles, in addition to pedestrians, bicyclists, and
+motorcyclists. This is done for both the individual and the opponent.
+Variable interactions are calculated to produce hierarchical variables.
+For example, for a hierarchical definition of traffic unit-level
+variables, age (and the square of aged) are interacted with gender, road
+user class, and vehicle type. For hierarchical opponent variables, age
+(and the square of age) are interacted with age of opponent (and squared
+aged of opponent). The variables thus obtained are shown in Table
+\ref{tab:model-specification-summary}. As seen in the table, Models 1
+and 2 are single-level models, and the difference between them is that
+Model 2 includes opponent variables. Models 3 and 4, in contrast, are
+hierarchical models. Model 3 considers the hierarchy on the basis of the
+traffic unit, while Model 4 considers the hierarchy on the basis of the
+collision.
+
+Models 1 through 4 are estimated using the full sample. As discussed
+above, a related modelling strategy is to subset the sample (e.g., Islam
+et al., 2014; Lee and Li, 2014; Tarrao et al., 2014; Wu et al., 2014).
+In this case we subset by a combination of traffic unit type of the
+individual (i.e., light duty vehicle, light truck, heavy vehicle,
+pedestrian, bicyclist, and motorcyclist) and vehicle type of the oponent
+(i.e., light duty vehicle, light truck, heavy vehicle). This leads to an
+ensemble of eighteen models to be estimated using subsets of data (see
+Table \ref{tab:model-specification-summary}). By subsetting the sample,
+opponent effects are incorporated implicitly. Models 1 and 2 are
+re-estimated using this strategy, dropping variables as necessary where
+they are irrelevant (for instance, after filtering for pedestrians, no
+other traffic unit types are present in the subset of data).
+
+\begin{table}
+
+\caption{\label{tab:model-specification-summary}\label{tab:model-specification-summary}Summary of variables and model specification}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}{llllll}
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{llllll}
 \toprule
@@
 \rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Opponent variables)}}\\
+\multicolumn{6}{l}{\textbf{Opponent variables}}\\
 \hspace{1em}Age of Opponent & In decades &  & $\checkmark$ & $\checkmark$ & \\
@@
 \rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Hierarchical traffic unit variables)}}\\
-\hspace{1em}Age:Light Truck Driver &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Light Truck Driver &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Heavy Vehicle Driver &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Heavy Vehicle Driver &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Light Truck Passenger &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Light Truck Passenger &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Heavy Vehicle Passenger &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Heavy Vehicle Passenger &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Pedestrian &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Pedestrian &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Bicyclist &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Bicyclist &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Age:Motorcyclist &  &  &  & $\checkmark$ & \\
-\hspace{1em}Age Squared:Motorcyclist &  &  &  & $\checkmark$ & \\
+\multicolumn{6}{l}{\textbf{Hierarchical traffic unit variables}}\\
+\hspace{1em}Light Truck Driver:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Light Truck Driver:Age Squared &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Driver:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Heavy Vehicle Driver:Age Squared &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Light Truck Passenger:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Light Truck Passenger:Age Squared: &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Passenger:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Heavy Vehicle Passenger:Age Squared &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Pedestrian:Age Squared &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Bicyclist:Age Squared &  &  &  & $\checkmark$ & \\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist:Age &  &  &  & $\checkmark$ & \\
+\hspace{1em}Motorcyclist:Age Squared &  &  &  & $\checkmark$ & \\
 \rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Hierarchical opponent variables)}}\\
+\multicolumn{6}{l}{\textbf{Hierarchical opponent variables}}\\
 \hspace{1em}Age:Age of Opponent &  &  &  &  & $\checkmark$\\
@@
 \rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Case-level variables)}}\\
-\hspace{1em}Crash Configuration & 19 levels; Reference: Hit a Moving Object & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Road Configuration & 12 levels; Reference: Non-Intersection & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Weather & 9 levels; Reference: Clear and Sunny & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\multicolumn{6}{l}{\textbf{Collision-level variables}}\\
+\hspace{1em}Crash Configuration & 19 levels; Reference: Hit a moving object & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Road Configuration & 12 levels; Reference: Non-intersection & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Weather & 9 levels; Reference: Clear and sunny & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 \hspace{1em}Surface & 11 levels; Reference: Dry & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Road Alignment & 8 levels; Reference: Straight and Level & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Traffic Controls & 19 levels; Reference: Operational Traffic Signals & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\rowcolor{gray!6}  \hspace{1em}Road Alignment & 8 levels; Reference: Straight and level & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\hspace{1em}Traffic Controls & 19 levels; Reference: Operational traffic signals & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 \rowcolor{gray!6}  \hspace{1em}Month & 12 levels; Reference: January & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
@@
 
-Notice how there are zero cases of user: BYCICLIST - opponent: Heavy
-Vehicle.
-
-Re-estimate model after subsetting by USER Type of person 1. LDxLD:
-
-Summary of models. See Table \ref{tab:model-summary}.
+Table \ref{tab:model-summary} collects some key summary statistics of
+the estimated models. Of interest is the goodness of fit of the models,
+which in the case is measured with Akaike's Information Criterion (AIC).
+This criterion is calculated using the estimated likelihood of the
+model, penalized by the number of coefficients in the model, so that
+other things being equal, it gives preference to more parsimonious
+models. Since smaller values of AIC are better fits, it is possible to
+see that, compared to the base model without opponent variables (Model
+1), there are large and significant improvements in fit to be gained by
+introducing opponent effects. However, the gains are not as large when
+hierarchical specifications are used, even when the number of additional
+coefficients that need to be estimated is not substantially larger
+(recall that the penalty per parameter in AIC is 2).
+
+The likelihood function of the model, and therefore the value of the
+AIC, depend on the size of the sample, which is why AIC is not
+comparable across models estimated with different sample sizes. For this
+reason, the full sample models cannot be compared directly to the models
+estimated with subsets of data. The models in the ensembles, however,
+can be compared, and there we find that introducing opponent variables
+leads to a better fit in most, but not all cases. One model for
+pedestrians (when the opponent is a heavy vehicle), the two models for
+bicyclists, and the three models for motorcyclists are better fits when
+opponent variables are not used.
+
+These results give some preliminary ideas about the relative performance
+of the different modelling strategies. In the next section we delve more
+deeply into this question.
 
@@
 \centering
+\fontsize{7}{9}\selectfont
 \begin{tabular}[t]{lccc}
@@
 \multicolumn{4}{l}{\textbf{Full sample models}}\\
-\hspace{1em}Model 1 & 164,511 & 100 & 195,215\\
-\hspace{1em}Model 2 & 164,511 & 106 & 178,943\\
-\rowcolor{gray!6}  \hspace{1em}Model 3 & 164,511 & 116 & 181,333\\
-\hspace{1em}Model 4 & 164,511 & 109 & 179,018\\
+\hspace{1em}Model 1 & 164,511 & 104 & 195,215\\
+\hspace{1em}Model 2 & 164,511 & 110 & 178,943\\
+\rowcolor{gray!6}  \hspace{1em}Model 3 & 164,511 & 120 & 181,333\\
+\hspace{1em}Model 4 & 164,511 & 113 & 179,018\\
 \rowcolor{gray!6}  \addlinespace[0.3em]
 \multicolumn{4}{l}{\textbf{Model 1 Ensemble (sample subsets by user type vs opponent)}}\\
-\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 95 & 145,396\\
-\hspace{1em}Light duty vehicle vs light truck & 3,237 & 95 & 3,979\\
-\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 95 & 5,913\\
-\hspace{1em}Light truck vs light duty vehicle & 3,121 & 95 & 3,921\\
-\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 95 & 1,230\\
-\hspace{1em}Light truck vs heavy vehicle & 198 & 95 & 354\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,763 & 95 & 4,362\\
-\hspace{1em}Heavy vehicle vs light truck & 180 & 95 & 291\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 95 & 1,193\\
-\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 94 & 2,842\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 94 & 270\\
-\hspace{1em}Pedestrian vs heavy vehicle & 376 & 94 & 473\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 94 & 686\\
-\hspace{1em}Bicyclist vs light truck & 148 & 94 & 192\\
+\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 96 & 145,390\\
+\hspace{1em}Light duty vehicle vs light truck & 3,237 & 81 & 3,943\\
+\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 90 & 5,895\\
+\hspace{1em}Light truck vs light duty vehicle & 3,121 & 81 & 3,885\\
+\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 69 & 1,170\\
+\hspace{1em}Light truck vs heavy vehicle & 198 & 66 & 288\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,726 & 81 & 4,326\\
+\hspace{1em}Heavy vehicle vs light truck & 180 & 66 & 225\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 76 & 1,147\\
+\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 90 & 2,826\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 64 & 202\\
+\hspace{1em}Pedestrian vs heavy vehicle & 376 & 66 & 409\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 82 & 654\\
+\hspace{1em}Bicyclist vs light truck & 116 & 44 & 84\\
 \rowcolor{gray!6}  \hspace{1em}Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
-\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 94 & 1,403\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 94 & 233\\
-\hspace{1em}Motorcyclist vs heavy vehicle & 73 & 94 & 204\\
+\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 80 & 1,367\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 58 & 153\\
+\hspace{1em}Motorcyclist vs heavy vehicle & 62 & 45 & 88\\
 \rowcolor{gray!6}  \addlinespace[0.3em]
 \multicolumn{4}{l}{\textbf{Model 2 Ensemble (sample subsets by user type vs opponent)}}\\
-\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 98 & 143,909\\
-\hspace{1em}Light duty vehicle vs light truck & 3,237 & 98 & 3,963\\
-\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 98 & 5,896\\
-\hspace{1em}Light truck vs light duty vehicle & 3,121 & 98 & 3,913\\
-\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 98 & 1,216\\
-\hspace{1em}Light truck vs heavy vehicle & 198 & 98 & 347\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,763 & 98 & 4,315\\
-\hspace{1em}Heavy vehicle vs light truck & 180 & 98 & 275\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 98 & 1,182\\
-\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 98 & 2,839\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 98 & 270\\
-\hspace{1em}Pedestrian vs heavy vehicle & 376 & 98 & 476\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 98 & 693\\
-\hspace{1em}Bicyclist vs light truck & 148 & 98 & 200\\
+\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 99 & 143,903\\
+\hspace{1em}Light duty vehicle vs light truck & 3,237 & 84 & 3,927\\
+\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 93 & 5,878\\
+\hspace{1em}Light truck vs light duty vehicle & 3,121 & 84 & 3,877\\
+\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 72 & 1,156\\
+\hspace{1em}Light truck vs heavy vehicle & 198 & 69 & 281\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,732 & 86 & 4,283\\
+\hspace{1em}Heavy vehicle vs light truck & 179 & 67 & 205\\
+\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 79 & 1,136\\
+\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 93 & 2,821\\
+\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 67 & 200\\
+\hspace{1em}Pedestrian vs heavy vehicle & 376 & 69 & 410\\
+\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 85 & 659\\
+\hspace{1em}Bicyclist vs light truck & 145 & 59 & 114\\
 \rowcolor{gray!6}  Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
-\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 98 & 1,411\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 98 & 235\\
-\hspace{1em}Motorcyclist vs heavy vehicle & 73 & 98 & 201\\
+\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 83 & 1,373\\
+\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 61 & 153\\
+\hspace{1em}Motorcyclist vs heavy vehicle & 63 & 45 & 90\\
 \bottomrule
 \multicolumn{4}{l}{\textit{Note: }}\\
-\multicolumn{4}{l}{There are zero cases of BICYCLIST vs HV in the sample}\\
+\multicolumn{4}{l}{There are zero cases of Bicyclist vs heavy vehicle in the sample}\\
 \end{tabular}
@@
 
-\hypertarget{outcome-shares-based-on-probabilities}{%
-\subsection{Outcome shares based on
-probabilities}\label{outcome-shares-based-on-probabilities}}
-
-Words go here.
-
-The datasets are used to predict the probabilities of the outcomes:
-
-The results of calculating the Average Prediction Error appear in Table
-\ref{tab:ape-results}.
+In this section we report an in-depth examination of the performance of
+the models. To this end, we use the models to conduct in-sample
+predictions (i.e., nowcasting), using the same sample that was used to
+estimate the models. In addition, we also conduct out-of-sample
+predictions, using the dataset corresponding to the year 2016, processed
+in identical way as the estimation sample (i.e., the 2017 dataset). This
+is an example of backcasting. Broadly, we evaluate the models in two
+ways: first, we compute the estimated shares of each outcome based on
+the predicted probabilities; and secondly, we evaluate the predicted
+outcomes.
+
+\hypertarget{outcome-shares-based-on-estimated-probabilities}{%
+\subsection{Outcome shares based on estimated
+probabilities}\label{outcome-shares-based-on-estimated-probabilities}}
+
+The shares of each outcome are calculated as the sum of the estimated
+probabilities for each observation:
+
+\begin{equation}
+\label{eq:predicted-shares}
+\begin{array}{rcl}
+\hat{S}_{\text{PDO}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{PDO})\\
+\hat{S}_{\text{injury}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{injury})\\
+\hat{S}_{\text{fatality}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{fatality})
+\end{array}
+\end{equation}
+
+\noindent where \(\hat{P}(y_{itk}=h_w)\) is the estimated probability of
+outcome \(h_w\) for individual \(i\) in traffic unit \(t\) and crash
+\(k\). The estimated share of outcome \(h\) is \(\hat{S}_{h_w}\).
+
+The estimated shares can be used to assess the ability of the model to
+forecast for the population the total number of cases of each outcome. A
+summary statistic useful to evaluate the performance is the Average
+Percentage Error, or \(APE\) (see Bogue et al., 2017, p. 31), which is
+calculated for each outcome as follows:
+
+\begin{equation}
+\label{eq:APE}
+APE_{h_w} = \Bigg|\frac{\hat{S}_{h_w} - S_{h_w}}{S_{h_w}}\Bigg|\times 100
+\end{equation}
+
+The Weighted Average Percentage Error (\(WAPE\)) aggregates the \(APE\)
+as follows:
+
+\begin{equation}
+\label{eq:WAPE}
+WAPE = \frac{\sum_w^WAPE_{h_w}\times S_{h_w}}{\sum_w^WS_{h_w}}
+\end{equation}
+
+The results of this exercise are reported in Table
+\ref{tab:ape-results}. Of the four full-sample models (Models 1-4), the
+\(APE\) of Model 2 is lowest in the nowcasting exercise for every
+outcome, with the exception of Fatality, where Model 4 produces a
+considerably lower \(APE\). When the results are aggregated by means of
+the \(WAPE\), Model 2 gives marginally better results than Model 4. It
+is interesting to see that the two Ensemble models have lower \(APE\)
+values across the board in the nowcasting exercise, and much better
+\(WAPE\) than the full sample models. However, when we turn to the
+results of the backcasting (out-of-sample) predictions, these results do
+not hold, as the Average Percentage Errors worsen considerably,
+particularly in the case of Fatality. Excellent in-sample predictions
+but mediocre out-of-sample predictions could constitute evidence of
+overfitting by the Ensemble models. In terms of backcasting, again Model
+4 is marginally better than Model 2, especially due to more accurate
+shares of No Injury.
 
@@
 \centering
+\fontsize{7}{9}\selectfont
 \begin{tabular}[t]{lrrrrrrrrrr}
@@
 \rowcolor{gray!15}
-\hspace{1em}Model 1 Ensemble & 62449 & 62439.99 & 0.01 & 83606 & 83614.28 & 0.01 & 933 & 933.73 & 0.08 & 0.01\\
 \hspace{1em}Model 2 & 78886 & 78928.98 & 0.05 & 84675 & 84641.94 & 0.04 & 950 & 940.08 & 1.04 & 0.05\\
-\rowcolor{gray!15}
-\hspace{1em}Model 2 Ensemble & 62449 & 62438.99 & 0.02 & 83606 & 83615.22 & 0.01 & 933 & 933.80 & 0.09 & 0.01\\
 \hspace{1em}Model 3 & 78886 & 79027.29 & 0.18 & 84675 & 84512.50 & 0.19 & 950 & 971.21 & 2.23 & 0.20\\
@@
 \hspace{1em}Model 4 & 78886 & 78939.18 & 0.07 & 84675 & 84622.54 & 0.06 & 950 & 949.28 & 0.08 & 0.06\\
+\hspace{1em}Model 1 Ensemble & 62413 & 62402.78 & 0.02 & 83564 & 83573.58 & 0.01 & 931 & 931.64 & 0.07 & 0.01\\
+\rowcolor{gray!15}
+\hspace{1em}Model 2 Ensemble & 62417 & 62407.00 & 0.02 & 83595 & 83604.14 & 0.01 & 931 & 931.86 & 0.09 & 0.01\\
+\rowcolor{gray!15}
+\hspace{1em}Model 4 Ensemble & 62405 & 62395.28 & 0.02 & 83578 & 83586.75 & 0.01 & 932 & 932.97 & 0.10 & 0.01\\
 \rowcolor{gray!15}
@@
 \multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 dataset)}}\\
-\hspace{1em}Model 1 & 82812 & 82574.35 & 0.29 & 88586 & 88737.88 & 0.17 & 935 & 1020.77 & 9.17 & 0.28\\
+\hspace{1em}Model 1 & 96860 & 96364.67 & 0.51 & 101605 & 102002.59 & 0.39 & 1109 & 1206.74 & 8.81 & 0.50\\
+\hspace{1em}Model 2 & 96860 & 96361.41 & 0.51 & 101605 & 102112.08 & 0.50 & 1109 & 1100.51 & 0.77 & 0.51\\
 \rowcolor{gray!15}
-\hspace{1em}Model 1 Ensemble & 64469 & 64525.65 & 0.09 & 87476 & 87291.39 & 0.21 & 909 & 1036.96 & 14.08 & 0.24\\
-\hspace{1em}Model 2 & 82812 & 82900.86 & 0.11 & 88586 & 88432.82 & 0.17 & 935 & 999.32 & 6.88 & 0.18\\
+\hspace{1em}Model 3 & 96860 & 96354.01 & 0.52 & 101605 & 102086.18 & 0.47 & 1109 & 1133.82 & 2.24 & 0.51\\
+\hspace{1em}Model 4 & 96860 & 96325.85 & 0.55 & 101605 & 102136.72 & 0.52 & 1109 & 1111.43 & 0.22 & 0.54\\
 \rowcolor{gray!15}
-\hspace{1em}Model 2 Ensemble & 64469 & 64541.74 & 0.11 & 87476 & 87261.66 & 0.25 & 909 & 1050.60 & 15.58 & 0.28\\
-\hspace{1em}Model 3 & 82812 & 82948.83 & 0.17 & 88586 & 88340.07 & 0.28 & 935 & 1044.10 & 11.67 & 0.29\\
-\rowcolor{gray!15}
-\hspace{1em}Model 4 & 82812 & 82878.68 & 0.08 & 88586 & 88446.37 & 0.16 & 935 & 1007.94 & 7.80 & 0.16\\
+\hspace{1em}Model 1 Ensemble & 77457 & 76822.49 & 0.82 & 100013 & 100580.60 & 0.57 & 1072 & 1138.91 & 6.24 & 0.71\\
+\hspace{1em}Model 2 Ensemble & 77459 & 76799.11 & 0.85 & 100049 & 100630.48 & 0.58 & 1071 & 1149.41 & 7.32 & 0.74\\
+\hspace{1em}Model 4 Ensemble & 77461 & 76766.08 & 0.90 & 100029 & 100630.21 & 0.60 & 1070 & 1163.71 & 8.76 & 0.78\\
 \bottomrule
@@
 \resizebox{\linewidth}{!}{
+\fontsize{7}{9}\selectfont
 \begin{tabular}[t]{l>{\raggedright\arraybackslash}p{30em}>{\raggedright\arraybackslash}p{30em}}
@@
 verification statistics. See Table
-\ref{tab:nowcast-outcomes-assessment}.
-
-\begin{table}
+\ref{tab:nowcast-outcomes-assessment}. We next evaluate the outcomes of
+the nowcast using an array of verification statistics. See Table
+\ref{tab:backcast-outcomes-assessment}.
+
+\begin{table}[!h]
 
@@
 \resizebox{\linewidth}{!}{
+\fontsize{7}{9}\selectfont
 \begin{tabular}[t]{lrrrrrrrrrrrrr}
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 50652 & 22504 & 150 &  & 69.067 & 0.929 & 0.499 & 0.265 & 0.642 & 0.309 &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62120 & 797 &  & 68.644 & 1.076 & 0.546 & 0.364 & 0.734 & 0.318 &  &  & \\
+\hspace{1em}No Injury & 50652 & 22503 & 150 &  & 69.068 & 0.929 & 0.499 & 0.265 & 0.642 & 0.309 &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62121 & 797 &  & 68.645 & 1.076 & 0.546 & 0.364 & 0.734 & 0.318 &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 34711 & 16429 & 63 &  & 69.909 & 0.820 & 0.440 & 0.195 & 0.556 & 0.322 &  &  & \\
-
-\hspace{1em}Injury & 27738 & 67167 & 830 &  & 69.380 & 1.145 & 0.599 & 0.451 & 0.803 & 0.298 &  &  & \\
-
-Fatality & 0 & 10 & 40 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.338} & 99.386 & 0.054 & 0.042 & 0.000 & 0.043 & 0.200 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.353} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.202}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51530 & 17136 & 85 &  & 72.903 & 0.872 & 0.536 & 0.201 & 0.653 & 0.250 &  &  & \\
-
-\hspace{1em}Injury & 27356 & 67515 & 864 &  & 72.415 & 1.131 & 0.598 & 0.353 & 0.797 & 0.295 &  &  & \\
+\hspace{1em}No Injury & 51531 & 17137 & 85 &  & 72.903 & 0.872 & 0.536 & 0.201 & 0.653 & 0.250 &  &  & \\
+
+\hspace{1em}Injury & 27355 & 67514 & 864 &  & 72.415 & 1.131 & 0.598 & 0.353 & 0.797 & 0.295 &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35473 & 16147 & 60 &  & 70.621 & 0.828 & 0.451 & 0.192 & 0.568 & 0.314 &  &  & \\
-
-\hspace{1em}Injury & 26976 & 67446 & 829 &  & 70.089 & 1.139 & 0.605 & 0.439 & 0.807 & 0.292 &  &  & \\
-
-Fatality & 0 & 13 & 44 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.049} & 99.386 & 0.061 & 0.047 & 0.000 & 0.047 & 0.228 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.379} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.368} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.212}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51102 & 17297 & 79 &  & 72.549 & 0.868 & 0.531 & 0.203 & 0.648 & 0.254 &  &  & \\
-
-\hspace{1em}Injury & 27784 & 67337 & 868 &  & 72.044 & 1.134 & 0.594 & 0.359 & 0.795 & 0.298 &  &  & \\
+\hspace{1em}No Injury & 51101 & 17296 & 79 &  & 72.549 & 0.868 & 0.531 & 0.203 & 0.648 & 0.254 &  &  & \\
+
+\hspace{1em}Injury & 27785 & 67338 & 868 &  & 72.044 & 1.134 & 0.594 & 0.359 & 0.795 & 0.298 &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 51574 & 17317 & 84 &  & 72.821 & 0.874 & 0.536 & 0.203 & 0.654 & 0.252 &  &  & \\
-
-\hspace{1em}Injury & 27312 & 67335 & 863 &  & 72.333 & 1.128 & 0.597 & 0.353 & 0.795 & 0.295 &  &  & \\
-
-Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.282} & 99.410 & 0.027 & 0.003 & 0.000 & 0.003 & 0.885 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.446} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.441} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\hspace{1em}No Injury & 51575 & 17316 & 84 &  & 72.822 & 0.874 & 0.536 & 0.203 & 0.654 & 0.252 &  &  & \\
+
+\hspace{1em}Injury & 27311 & 67336 & 863 &  & 72.334 & 1.128 & 0.597 & 0.353 & 0.795 & 0.295 &  &  & \\
+
+Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.283} & 99.410 & 0.027 & 0.003 & 0.000 & 0.003 & 0.885 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.446} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.441} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
+\hspace{1em}No Injury & 34664 & 16434 & 63 &  & 69.882 & 0.820 & 0.439 & 0.195 & 0.555 & 0.322 &  &  & \\
+
+\hspace{1em}Injury & 27749 & 67120 & 829 &  & 69.354 & 1.145 & 0.599 & 0.451 & 0.803 & 0.299 &  &  & \\
+
+Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.311} & 99.386 & 0.053 & 0.041 & 0.000 & 0.042 & 0.204 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.352} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.201}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
+\hspace{1em}No Injury & 35443 & 16145 & 60 &  & 70.615 & 0.827 & 0.451 & 0.192 & 0.568 & 0.314 &  &  & \\
+
+\hspace{1em}Injury & 26974 & 67437 & 829 &  & 70.083 & 1.139 & 0.605 & 0.439 & 0.807 & 0.292 &  &  & \\
+
+Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.042} & 99.386 & 0.059 & 0.044 & 0.000 & 0.045 & 0.236 &  &  & \multirow{-3}{*}{\raggedleft\arraybackslash 0.211}\\
+
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
+\hspace{1em}No Injury & 35553 & 16297 & 59 &  & 70.590 & 0.832 & 0.451 & 0.194 & 0.570 & 0.315 &  &  & \\
+
+\hspace{1em}Injury & 26852 & 67271 & 827 &  & 70.060 & 1.136 & 0.605 & 0.437 & 0.805 & 0.292 &  &  & \\
+
+Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.020} & 99.390 & 0.060 & 0.049 & 0.000 & 0.049 & 0.179 & \multirow{-6}{*}{\raggedleft\arraybackslash 0.378} & \multirow{-6}{*}{\raggedleft\arraybackslash 0.368} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.213}\\
 \bottomrule
@@
 
-We next evaluate the outcomes of the nowcast using an array of
-verification statistics. See Table
-\ref{tab:backcast-outcomes-assessment}.
-
-\begin{table}
+\begin{table}[!h]
 
@@
 \centering
+\resizebox{\linewidth}{!}{
+\fontsize{7}{9}\selectfont
 \resizebox{\linewidth}{!}{
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 53167 & 23097 & 145 &  & 69.311 & 0.923 & 0.501 & 0.260 & 0.642 & 0.304 &  &  & \\
-
-\hspace{1em}Injury & 29642 & 65455 & 788 &  & 68.920 & 1.082 & 0.550 & 0.363 & 0.739 & 0.317 &  &  & \\
-
-Fatality & 3 & 34 & 2 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.834} & 99.437 & 0.042 & 0.002 & 0.000 & 0.002 & 0.949 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.378} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.375} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.192}\\
+\hspace{1em}No Injury & 61684 & 27447 & 184 &  & 68.529 & 0.922 & 0.495 & 0.269 & 0.637 & 0.309 &  &  & \\
+
+\hspace{1em}Injury & 35171 & 74073 & 915 &  & 68.123 & 1.084 & 0.538 & 0.368 & 0.729 & 0.328 &  &  & \\
+
+Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.028} & 99.404 & 0.090 & 0.008 & 0.000 & 0.009 & 0.900 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.360} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 2}}\\
+\hspace{1em}No Injury & 62735 & 21013 & 106 &  & 72.319 & 0.866 & 0.532 & 0.206 & 0.648 & 0.252 &  &  & \\
+
+\hspace{1em}Injury & 34125 & 80569 & 996 &  & 71.862 & 1.139 & 0.589 & 0.358 & 0.793 & 0.304 &  &  & \\
+
+Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.808} & 99.436 & 0.027 & 0.006 & 0.000 & 0.006 & 0.767 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.437} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.434} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.224}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 3}}\\
+\hspace{1em}No Injury & 62248 & 21133 & 107 &  & 72.014 & 0.862 & 0.527 & 0.207 & 0.643 & 0.254 &  &  & \\
+
+\hspace{1em}Injury & 34610 & 80433 & 996 &  & 71.550 & 1.142 & 0.586 & 0.363 & 0.792 & 0.307 &  &  & \\
+
+Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.496} & 99.427 & 0.042 & 0.005 & 0.000 & 0.005 & 0.872 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.431} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.427} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.221}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 4}}\\
+\hspace{1em}No Injury & 62788 & 21247 & 102 &  & 72.230 & 0.869 & 0.531 & 0.208 & 0.648 & 0.254 &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80331 & 1000 &  & 71.767 & 1.136 & 0.588 & 0.358 & 0.791 & 0.304 &  &  & \\
+
+Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.716} & 99.434 & 0.032 & 0.006 & 0.000 & 0.006 & 0.800 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.435} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.432} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.223}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 35564 & 16930 & 71 &  & 69.967 & 0.815 & 0.437 & 0.192 & 0.552 & 0.323 &  &  & \\
-
-\hspace{1em}Injury & 28891 & 70451 & 822 &  & 69.423 & 1.145 & 0.601 & 0.454 & 0.805 & 0.297 &  &  & \\
-
-Fatality & 14 & 95 & 16 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.368} & 99.344 & 0.138 & 0.016 & 0.001 & 0.018 & 0.872 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.362} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.351} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 54430 & 17615 & 81 &  & 73.262 & 0.871 & 0.542 & 0.198 & 0.657 & 0.245 &  &  & \\
-
-\hspace{1em}Injury & 28382 & 70954 & 850 &  & 72.806 & 1.131 & 0.602 & 0.349 & 0.801 & 0.292 &  &  & \\
-
-Fatality & 0 & 17 & 4 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.759} & 99.450 & 0.022 & 0.004 & 0.000 & 0.004 & 0.810 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.455} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.451} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.232}\\
+\hspace{1em}No Injury & 42896 & 20230 & 95 &  & 69.259 & 0.816 & 0.439 & 0.201 & 0.554 & 0.321 &  &  & \\
+
+\hspace{1em}Injury & 34546 & 79692 & 962 &  & 68.731 & 1.152 & 0.588 & 0.452 & 0.797 & 0.308 &  &  & \\
+
+Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.669} & 99.349 & 0.113 & 0.013 & 0.001 & 0.014 & 0.876 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.354} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.345} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.183}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 36234 & 16840 & 73 &  & 70.463 & 0.824 & 0.445 & 0.191 & 0.562 & 0.318 &  &  & \\
-
-\hspace{1em}Injury & 28216 & 70527 & 819 &  & 69.916 & 1.138 & 0.605 & 0.444 & 0.806 & 0.292 &  &  & \\
-
-Fatality & 19 & 109 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.856} & 99.333 & 0.160 & 0.016 & 0.001 & 0.019 & 0.883 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.373} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.362} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.194}\\
+\hspace{1em}No Injury & 43486 & 19937 & 95 &  & 69.758 & 0.820 & 0.446 & 0.198 & 0.561 & 0.315 &  &  & \\
+
+\hspace{1em}Injury & 33953 & 80009 & 961 &  & 69.227 & 1.149 & 0.593 & 0.445 & 0.800 & 0.304 &  &  & \\
+
+Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.163} & 99.340 & 0.129 & 0.013 & 0.001 & 0.014 & 0.891 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.364} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.355} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
 \cmidrule{1-14}
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 53985 & 17731 & 83 &  & 72.936 & 0.867 & 0.536 & 0.199 & 0.652 & 0.248 &  &  & \\
-
-\hspace{1em}Injury & 28827 & 70819 & 849 &  & 72.470 & 1.134 & 0.599 & 0.354 & 0.799 & 0.295 &  &  & \\
-
-Fatality & 0 & 36 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.422} & 99.438 & 0.042 & 0.003 & 0.000 & 0.003 & 0.923 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.448} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.444} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.228}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 54493 & 17843 & 79 &  & 73.168 & 0.874 & 0.541 & 0.200 & 0.658 & 0.247 &  &  & \\
-
-\hspace{1em}Injury & 28318 & 70725 & 852 &  & 72.709 & 1.128 & 0.601 & 0.348 & 0.798 & 0.292 &  &  & \\
-
-Fatality & 1 & 18 & 4 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.663} & 99.449 & 0.025 & 0.004 & 0.000 & 0.004 & 0.826 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.453} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.449} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.231}\\
+\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
+\hspace{1em}No Injury & 43560 & 20160 & 94 &  & 69.671 & 0.824 & 0.446 & 0.200 & 0.561 & 0.317 &  &  & \\
+
+\hspace{1em}Injury & 33876 & 79762 & 959 &  & 69.141 & 1.146 & 0.591 & 0.444 & 0.800 & 0.304 &  &  & \\
+
+Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.074} & 99.336 & 0.139 & 0.014 & 0.001 & 0.014 & 0.886 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.354} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.189}\\
 \bottomrule
@@
 \multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
-\end{tabular}}
+\end{tabular}}}
 \end{table}
+
+\hypertarget{sec:insights}{%
+\section{Best model: insights}\label{sec:insights}}
+
+Create a table with the results of Model 2.
 
@@
 
-Words go here.
+Different modelling strategies can be used to model complex
+hierarchical, multievent outcomes such as the severity of injuries
+following a collision. The objective of this paper was to assess the
+performance of different strategies to model opponent effects in
+two-vehicle crashes. In broad terms, three strategies were considered:
+1) incorporating opponent-level variables in the model; 2) single-
+versus multi-level model specifications; and 3) sample subsetting and
+estimation of separate models for different types of individual-opponent
+interactions.
+
+The results of the empirical assessment strongly suggest that
+incorporating opponent effects can greatly improve the fit and
+predictive performance of a model. There is some evidence that
+subsetting the sample can improve the results in some isolated
+situations (e.g., when modelling the severity of crashes involving
+active travellers or motorcyclists), possibly at the risk of
+overfitting. In this paper we did not compare individual models in our
+Ensemble approach, but we suggest that this is an avenue for future
+research.
+
+In terms of the full sample models, the evidence was not conclusive in
+favor of a single-level model with opponent variables (Model 2), or a
+hierarchical model with individual-opponent interactions (Model 4). On
+the one hand, the AIC tended to favor the
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Amoh2017effect}{}%
+Amoh-Gyimah, R., Aidoo, E.N., Akaateba, M.A., Appiah, S.K., 2017. The
+effect of natural and built environmental characteristics on
+pedestrian-vehicle crash severity in ghana. International Journal of
+Injury Control and Safety Promotion 24, 459--468.
+doi:\href{https://doi.org/10.1080/17457300.2016.1232274}{10.1080/17457300.2016.1232274}
+
 \leavevmode\hypertarget{ref-Aziz2013exploring}{}%
@@
 doi:\href{https://doi.org/10.1016/j.amar.2017.03.001}{10.1016/j.amar.2017.03.001}
+
+\leavevmode\hypertarget{ref-Casetti1972generating}{}%
+Casetti, E., 1972. Generating models by the expansion method:
+Applications to geographic research. Geographical Analysis 4, 81--91.
 
@@
 
+\leavevmode\hypertarget{ref-Dissanayake2002factors}{}%
+Dissanayake, S., Lu, J.J., 2002. Factors influential in making an injury
+severity difference to older drivers involved in fixed object-passenger
+car crashes. Accident Analysis and Prevention 34, 609--618.
+doi:\href{https://doi.org/10.1016/s0001-4575(01)00060-4}{10.1016/s0001-4575(01)00060-4}
+
+\leavevmode\hypertarget{ref-Duddu2018modeling}{}%
+Duddu, V.R., Penmetsa, P., Pulugurtha, S.S., 2018. Modeling and
+comparing injury severity of at-fault and not at-fault drivers in
+crashes. Accident Analysis and Prevention 120, 55--63.
+doi:\href{https://doi.org/10.1016/j.aap.2018.07.036}{10.1016/j.aap.2018.07.036}
+
 \leavevmode\hypertarget{ref-Effati2015geospatial}{}%
@@
 
+\leavevmode\hypertarget{ref-Hedeker1994random}{}%
+Hedeker, D., Gibbons, R.D., 1994. A random-effects ordinal
+regression-model for multilevel analysis. Biometrics 50, 933--944.
+
 \leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
@@
 doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}
+
+\leavevmode\hypertarget{ref-Islam2014comprehensive}{}%
+Islam, S., Jones, S.L., Dye, D., 2014. Comprehensive analysis of single-
+and multi-vehicle large truck at-fault crashes on rural and urban
+roadways in alabama. Accident Analysis \& Prevention 67, 148--158.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2014.02.014}{https://doi.org/10.1016/j.aap.2014.02.014}
 
@@
 
-\leavevmode\hypertarget{ref-Mannering2016unobserved}{}%
-Mannering, F., Shankar, V., Bhat, C.R., 2016. Unobserved heterogeneity
-and the statistical analysis of highway accident data. Analytic Methods
-in Accident Research 11, 1--16.
-doi:\href{https://doi.org/10.1016/j.amar.2016.04.001}{10.1016/j.amar.2016.04.001}
-
 \leavevmode\hypertarget{ref-McArthur2014spatial}{}%
@@
 
+\leavevmode\hypertarget{ref-Osman2018injury}{}%
+Osman, M., Mishra, S., Paleti, R., 2018. Injury severity analysis of
+commercially-licensed drivers in single-vehicle crashes: Accounting for
+unobserved heterogeneity and age group differences. Accident Analysis
+and Prevention 118, 289--300.
+doi:\href{https://doi.org/10.1016/j.aap.2018.05.004}{10.1016/j.aap.2018.05.004}
+
 \leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
@@
 
+\leavevmode\hypertarget{ref-Penmetsa2017examining}{}%
+Penmetsa, P., Pulugurtha, S.S., Duddu, V.R., 2017. Examining injury
+severity of not-at-fault drivers in two-vehicle crashes. Transportation
+Research Record 164--173.
+doi:\href{https://doi.org/10.3141/2659-18}{10.3141/2659-18}
+
 \leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
@@
 
+\leavevmode\hypertarget{ref-Rana2010copula}{}%
+Rana, T.A., Sikder, S., Pinjari, A.R., 2010. Copula-based method for
+addressing endogeneity in models of severity of traffic crash injuries
+application to two-vehicle crashes. Transportation Research Record
+75--87. doi:\href{https://doi.org/10.3141/2147-10}{10.3141/2147-10}
+
 \leavevmode\hypertarget{ref-Regev2018crash}{}%
@@
 doi:\href{https://doi.org/10.1016/j.aap.2014.07.025}{10.1016/j.aap.2014.07.025}
+
+\leavevmode\hypertarget{ref-Savolainen2007probabilistic}{}%
+Savolainen, P., Mannering, F., 2007. Probabilistic models of
+motorcyclists' injury severities in single- and multi-vehicle crashes.
+Accident Analysis and Prevention 39, 955--963.
+doi:\href{https://doi.org/10.1016/j.aap.2006.12.016}{10.1016/j.aap.2006.12.016}
 
@@
 
+\leavevmode\hypertarget{ref-Shamsunnahar2014examining}{}%
+Shamsunnahar, Y., Eluru, N., Pinjari, A.R., Tay, R., 2014. Examining
+driver injury severity in two vehicle crashes - a copula based approach.
+Accident Analysis and Prevention 66, 120--135.
+doi:\href{https://doi.org/10.1016/j.aap.2014.01.018}{10.1016/j.aap.2014.01.018}
+
 \leavevmode\hypertarget{ref-Symons2019reduced}{}%
@@
 
+\leavevmode\hypertarget{ref-Tay2011multinomial}{}%
+Tay, R., Choi, J., Kattan, L., Khan, A., 2011. A multinomial logit model
+of pedestrian-vehicle crash severity. International Journal of
+Sustainable Transportation 5, 233--249.
+doi:\href{https://doi.org/10.1080/15568318.2010.497547}{10.1080/15568318.2010.497547}
+
 \leavevmode\hypertarget{ref-Thompson2018trends}{}%
@@
 doi:\href{https://doi.org/10.3141/2514-17}{10.3141/2514-17}
+
+\leavevmode\hypertarget{ref-Wang2005use}{}%
+Wang, X.K., Kockelman, K.M., 2005. Use of heteroscedastic ordered logit
+model to study severity of occupant injury - distinguishing effects of
+vehicle weight and type, in: Statistical Methods; Highway Safety Data,
+Analysis, and Evaluation; Occupant Protection; Systematic Reviews and
+Meta-Analysis, Transportation Research Record. pp. 195--204.
 

### 2020-01-15 Interactions-Crash-Severity-v1/fixed-polr.R

@@ -1,588 +1,0 @@
-# file MASS/R/polr.R
-# copyright (C) 1994-2008 W. N. Venables and B. D. Ripley
-# Use of transformed intercepts contributed by David Firth
-#
-#  This program is free software; you can redistribute it and/or modify
-#  it under the terms of the GNU General Public License as published by
-#  the Free Software Foundation; either version 2 or 3 of the License
-#  (at your option).
-#
-#  This program is distributed in the hope that it will be useful,
-#  but WITHOUT ANY WARRANTY; without even the implied warranty of
-#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-#  GNU General Public License for more details.
-#
-#  A copy of the GNU General Public License is available at
-#  http://www.r-project.org/Licenses/
-#
-polr <- function(formula, data, weights, start, ..., subset,
-                 na.action, contrasts = NULL, Hess = FALSE,
-                 model = TRUE,
-                 method = c("logistic", "probit", "cloglog", "cauchit"))
-{
-    logit <- function(p) log(p/(1 - p))
-
-    fmin <- function(beta) {
-        theta <- beta[pc + 1L:q]
-        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
-        eta <- offset
-        if (pc > 0)
-            eta <- eta + drop(x %*% beta[1L:pc])
-        pr <- pfun(gamm[y + 1] - eta) - pfun(gamm[y] - eta)
-        if (all(pr > 0))
-            -sum(wt * log(pr))
-        else Inf
-    }
-
-    gmin <- function(beta)
-    {
-        jacobian <- function(theta) { ## dgamma by dtheta matrix
-            k <- length(theta)
-            etheta <- exp(theta)
-            mat <- matrix(0 , k, k)
-            mat[, 1] <- rep(1, k)
-            for (i in 2:k) mat[i:k, i] <- etheta[i]
-            mat
-        }
-        theta <- beta[pc + 1L:q]
-        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
-        eta <- offset
-        if(pc > 0) eta <- eta + drop(x %*% beta[1L:pc])
-        pr <- pfun(gamm[y+1] - eta) - pfun(gamm[y] - eta)
[…]

### 2020-01-16 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.tex
Completed work with Model 3 Ensemble

@@
 % Pandoc toggle for numbering sections (defaults to be off)
+
 % Pandoc header
@@
 
-  \title{An assessment of strategies to model opponent effects in crash severity
-analysis}
-    \author[]{Antonio Paez\corref{c1}}
+  \title{An empirical assessment of strategies to model opponent effects in crash
+severity analysis}
+    \author[]{Antonio Paez\corref{Corresponding Author}}
    \ead{paezha@mcmaster.ca} 
-   \cortext[c1]{Corresponding Author}
     \author[Louisiana State University]{Hany Hassan}
    \ead{hassan1@lsu.edu} 
-
     \author[McMaster University]{Mark Ferguson}
    \ead{fergumr@mcmaster.ca} 
-
     \author[McMaster University]{Saiedeh Razavi}
    \ead{razavi@mcmaster.ca} 
-
       \address[McMaster University]{McMaster Institute for Transportation and Logistics, 1280 Main Street

### 2020-01-16 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Completed work with Model 3 Ensemble

@@
 severity analysis}
-    \author[Some University]{Author 1\corref{Corresponding Author}}
-   \ead{a1@example.com} 
-    \author[Some Institute of Technology]{Author 2}
-   \ead{a2@example.com} 
-      \address[Some University]{Department, Street, City, State, Zip}
-    \address[Some Institute of Technology]{Department, Street, City, State, Zip}
+    \author[]{Antonio Paez\corref{Corresponding Author}}
+   \ead{paezha@mcmaster.ca} 
+    \author[Louisiana State University]{Hany Hassan}
+   \ead{hassan1@lsu.edu} 
+    \author[McMaster University]{Mark Ferguson}
+   \ead{fergumr@mcmaster.ca} 
+    \author[McMaster University]{Saiedeh Razavi}
+   \ead{razavi@mcmaster.ca} 
+      \address[McMaster University]{McMaster Institute for Transportation and Logistics, 1280 Main Street
+West, Hamilton, Ontario, Canada L8S 4K1}
+    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Baton Rouge,
+Louisiana, USA 70803}
 
@@
 \rowcolor{gray!15}
+\hspace{1em}Model 3 Ensemble & 62411 & 62401.23 & 0.02 & 83596 & 83604.71 & 0.01 & 933 & 934.06 & 0.11 & 0.01\\
 \hspace{1em}Model 4 Ensemble & 62405 & 62395.28 & 0.02 & 83578 & 83586.75 & 0.01 & 932 & 932.97 & 0.10 & 0.01\\
@@
 \hspace{1em}Model 1 & 96860 & 96364.67 & 0.51 & 101605 & 102002.59 & 0.39 & 1109 & 1206.74 & 8.81 & 0.50\\
+\rowcolor{gray!15}
 \hspace{1em}Model 2 & 96860 & 96361.41 & 0.51 & 101605 & 102112.08 & 0.50 & 1109 & 1100.51 & 0.77 & 0.51\\
+\hspace{1em}Model 3 & 96860 & 96354.01 & 0.52 & 101605 & 102086.18 & 0.47 & 1109 & 1133.82 & 2.24 & 0.51\\
 \rowcolor{gray!15}
-\hspace{1em}Model 3 & 96860 & 96354.01 & 0.52 & 101605 & 102086.18 & 0.47 & 1109 & 1133.82 & 2.24 & 0.51\\
 \hspace{1em}Model 4 & 96860 & 96325.85 & 0.55 & 101605 & 102136.72 & 0.52 & 1109 & 1111.43 & 0.22 & 0.54\\
+\hspace{1em}Model 1 Ensemble & 77457 & 76822.49 & 0.82 & 100013 & 100580.60 & 0.57 & 1072 & 1138.91 & 6.24 & 0.71\\
 \rowcolor{gray!15}
-\hspace{1em}Model 1 Ensemble & 77457 & 76822.49 & 0.82 & 100013 & 100580.60 & 0.57 & 1072 & 1138.91 & 6.24 & 0.71\\
 \hspace{1em}Model 2 Ensemble & 77459 & 76799.11 & 0.85 & 100049 & 100630.48 & 0.58 & 1071 & 1149.41 & 7.32 & 0.74\\
-\hspace{1em}Model 4 Ensemble & 77461 & 76766.08 & 0.90 & 100029 & 100630.21 & 0.60 & 1070 & 1163.71 & 8.76 & 0.78\\
+\hspace{1em}Model 3 Ensemble & 77459 & 76786.76 & 0.87 & 100050 & 100644.29 & 0.59 & 1072 & 1149.95 & 7.27 & 0.75\\
+\rowcolor{gray!15}
+Model 4 Ensemble & 77461 & 76766.08 & 0.90 & 100029 & 100630.21 & 0.60 & 1070 & 1163.71 & 8.76 & 0.78\\
 \bottomrule
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 50652 & 22503 & 150 &  & 69.068 & 0.929 & 0.499 & 0.265 & 0.642 & 0.309 &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62121 & 797 &  & 68.645 & 1.076 & 0.546 & 0.364 & 0.734 & 0.318 &  &  & \\
-
-Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.552} & 99.392 & 0.059 & 0.003 & 0.000 & 0.003 & 0.946 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.372} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.370} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.190}\\
+\hspace{1em}No Injury & 50652 & 22503 & 150 &  & 69.0677 & 0.9293 & 0.4988 & 0.2646 & 0.6421 & 0.3090 &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62121 & 797 &  & 68.6446 & 1.0765 & 0.5463 & 0.3636 & 0.7336 & 0.3185 &  &  & \\
+
+Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.5523} & 99.3921 & 0.0589 & 0.0030 & 0.0003 & 0.0032 & 0.9464 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3725} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3696} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1902}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51531 & 17137 & 85 &  & 72.903 & 0.872 & 0.536 & 0.201 & 0.653 & 0.250 &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67514 & 864 &  & 72.415 & 1.131 & 0.598 & 0.353 & 0.797 & 0.295 &  &  & \\
-
-Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.364} & 99.409 & 0.026 & 0.001 & 0.000 & 0.001 & 0.960 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.447} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.443} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\hspace{1em}No Injury & 51531 & 17137 & 85 &  & 72.9033 & 0.8715 & 0.5362 & 0.2011 & 0.6532 & 0.2505 &  &  & \\
+
+\hspace{1em}Injury & 27355 & 67514 & 864 &  & 72.4152 & 1.1306 & 0.5980 & 0.3535 & 0.7973 & 0.2948 &  &  & \\
+
+Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.3636} & 99.4086 & 0.0263 & 0.0010 & 0.0001 & 0.0011 & 0.9600 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4474} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4429} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2265}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51101 & 17296 & 79 &  & 72.549 & 0.868 & 0.531 & 0.203 & 0.648 & 0.254 &  &  & \\
-
-\hspace{1em}Injury & 27785 & 67338 & 868 &  & 72.044 & 1.134 & 0.594 & 0.359 & 0.795 & 0.298 &  &  & \\
-
-Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.996} & 99.399 & 0.046 & 0.003 & 0.000 & 0.003 & 0.932 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.440} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.436} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.224}\\
+\hspace{1em}No Injury & 51101 & 17296 & 79 &  & 72.5489 & 0.8680 & 0.5309 & 0.2029 & 0.6478 & 0.2537 &  &  & \\
+
+\hspace{1em}Injury & 27785 & 67338 & 868 &  & 72.0444 & 1.1336 & 0.5942 & 0.3589 & 0.7953 & 0.2985 &  &  & \\
+
+Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.9964} & 99.3994 & 0.0463 & 0.0030 & 0.0003 & 0.0032 & 0.9318 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4400} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4356} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2239}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 51575 & 17316 & 84 &  & 72.822 & 0.874 & 0.536 & 0.203 & 0.654 & 0.252 &  &  & \\
-
-\hspace{1em}Injury & 27311 & 67336 & 863 &  & 72.334 & 1.128 & 0.597 & 0.353 & 0.795 & 0.295 &  &  & \\
-
-Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.283} & 99.410 & 0.027 & 0.003 & 0.000 & 0.003 & 0.885 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.446} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.441} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.227}\\
+\hspace{1em}No Injury & 51575 & 17318 & 84 &  & 72.8207 & 0.8744 & 0.5356 & 0.2032 & 0.6538 & 0.2523 &  &  & \\
+
+\hspace{1em}Injury & 27311 & 67334 & 863 &  & 72.3332 & 1.1279 & 0.5967 & 0.3529 & 0.7952 & 0.2950 &  &  & \\
+
+Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.2821} & 99.4104 & 0.0274 & 0.0031 & 0.0001 & 0.0032 & 0.8846 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4458} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4414} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2268}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 34664 & 16434 & 63 &  & 69.882 & 0.820 & 0.439 & 0.195 & 0.555 & 0.322 &  &  & \\
-
-\hspace{1em}Injury & 27749 & 67120 & 829 &  & 69.354 & 1.145 & 0.599 & 0.451 & 0.803 & 0.299 &  &  & \\
-
-Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.311} & 99.386 & 0.053 & 0.041 & 0.000 & 0.042 & 0.204 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.352} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.201}\\
+\hspace{1em}No Injury & 34664 & 16434 & 63 &  & 69.8818 & 0.8197 & 0.4393 & 0.1952 & 0.5554 & 0.3225 &  &  & \\
+
+\hspace{1em}Injury & 27749 & 67120 & 829 &  & 69.3536 & 1.1452 & 0.5985 & 0.4512 & 0.8032 & 0.2986 &  &  & \\
+
+Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.3107} & 99.3860 & 0.0526 & 0.0414 & 0.0001 & 0.0419 & 0.2041 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3626} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3521} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2010}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35443 & 16145 & 60 &  & 70.615 & 0.827 & 0.451 & 0.192 & 0.568 & 0.314 &  &  & \\
-
-\hspace{1em}Injury & 26974 & 67437 & 829 &  & 70.083 & 1.139 & 0.605 & 0.439 & 0.807 & 0.292 &  &  & \\
-
-Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.042} & 99.386 & 0.059 & 0.044 & 0.000 & 0.045 & 0.236 &  &  & \multirow{-3}{*}{\raggedleft\arraybackslash 0.211}\\
-
+\hspace{1em}No Injury & 35443 & 16145 & 60 &  & 70.6151 & 0.8275 & 0.4508 & 0.1917 & 0.5678 & 0.3138 &  &  & \\
+
+\hspace{1em}Injury & 26974 & 67437 & 829 &  & 70.0830 & 1.1393 & 0.6054 & 0.4389 & 0.8067 & 0.2919 &  &  & \\
+
+Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0421} & 99.3862 & 0.0591 & 0.0445 & 0.0001 & 0.0451 & 0.2364 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3784} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3678} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2106}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
+\hspace{1em}No Injury & 35498 & 16204 & 60 &  & 70.6159 & 0.8294 & 0.4512 & 0.1924 & 0.5688 & 0.3142 &  &  & \\
+
+\hspace{1em}Injury & 26913 & 67379 & 828 &  & 70.0844 & 1.1379 & 0.6052 & 0.4379 & 0.8060 & 0.2916 &  &  & \\
+
+Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0436} & 99.3868 & 0.0622 & 0.0476 & 0.0001 & 0.0482 & 0.2241 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3786} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3681} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2123}\\
+\cmidrule{1-14}
 \addlinespace[0.3em]
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 35553 & 16297 & 59 &  & 70.590 & 0.832 & 0.451 & 0.194 & 0.570 & 0.315 &  &  & \\
-
-\hspace{1em}Injury & 26852 & 67271 & 827 &  & 70.060 & 1.136 & 0.605 & 0.437 & 0.805 & 0.292 &  &  & \\
-
-Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.020} & 99.390 & 0.060 & 0.049 & 0.000 & 0.049 & 0.179 & \multirow{-6}{*}{\raggedleft\arraybackslash 0.378} & \multirow{-6}{*}{\raggedleft\arraybackslash 0.368} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.213}\\
+\hspace{1em}No Injury & 35553 & 16297 & 59 &  & 70.5898 & 0.8318 & 0.4514 & 0.1935 & 0.5697 & 0.3151 &  &  & \\
+
+\hspace{1em}Injury & 26852 & 67271 & 827 &  & 70.0602 & 1.1361 & 0.6046 & 0.4370 & 0.8049 & 0.2915 &  &  & \\
+
+Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0201} & 99.3901 & 0.0601 & 0.0488 & 0.0001 & 0.0494 & 0.1786 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3783} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3679} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2127}\\
 \bottomrule
@@
 
-\caption{\label{tab:backcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 dataset, i.e., estimation dataset)}
+\caption{\label{tab:backcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
 \centering
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 61684 & 27447 & 184 &  & 68.529 & 0.922 & 0.495 & 0.269 & 0.637 & 0.309 &  &  & \\
-
-\hspace{1em}Injury & 35171 & 74073 & 915 &  & 68.123 & 1.084 & 0.538 & 0.368 & 0.729 & 0.328 &  &  & \\
-
-Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.028} & 99.404 & 0.090 & 0.008 & 0.000 & 0.009 & 0.900 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.360} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
+\hspace{1em}No Injury & 61684 & 27447 & 184 &  & 68.5295 & 0.9221 & 0.4955 & 0.2690 & 0.6368 & 0.3094 &  &  & \\
+
+\hspace{1em}Injury & 35171 & 74073 & 915 &  & 68.1231 & 1.0842 & 0.5380 & 0.3683 & 0.7290 & 0.3276 &  &  & \\
+
+Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.0284} & 99.4042 & 0.0902 & 0.0083 & 0.0005 & 0.0090 & 0.9000 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3628} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3604} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1882}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 62735 & 21013 & 106 &  & 72.319 & 0.866 & 0.532 & 0.206 & 0.648 & 0.252 &  &  & \\
-
-\hspace{1em}Injury & 34125 & 80569 & 996 &  & 71.862 & 1.139 & 0.589 & 0.358 & 0.793 & 0.304 &  &  & \\
-
-Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.808} & 99.436 & 0.027 & 0.006 & 0.000 & 0.006 & 0.767 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.437} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.434} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.224}\\
+\hspace{1em}No Injury & 62735 & 21013 & 106 &  & 72.3190 & 0.8657 & 0.5317 & 0.2056 & 0.6477 & 0.2519 &  &  & \\
+
+\hspace{1em}Injury & 34125 & 80569 & 996 &  & 71.8616 & 1.1386 & 0.5893 & 0.3585 & 0.7930 & 0.3036 &  &  & \\
+
+Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.8085} & 99.4363 & 0.0271 & 0.0062 & 0.0001 & 0.0063 & 0.7667 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4372} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4335} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2241}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 62248 & 21133 & 107 &  & 72.014 & 0.862 & 0.527 & 0.207 & 0.643 & 0.254 &  &  & \\
-
-\hspace{1em}Injury & 34610 & 80433 & 996 &  & 71.550 & 1.142 & 0.586 & 0.363 & 0.792 & 0.307 &  &  & \\
-
-Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.496} & 99.427 & 0.042 & 0.005 & 0.000 & 0.005 & 0.872 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.431} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.427} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.221}\\
+\hspace{1em}No Injury & 62248 & 21133 & 107 &  & 72.0144 & 0.8619 & 0.5271 & 0.2068 & 0.6427 & 0.2544 &  &  & \\
+
+\hspace{1em}Injury & 34610 & 80433 & 996 &  & 71.5504 & 1.1421 & 0.5862 & 0.3634 & 0.7916 & 0.3068 &  &  & \\
+
+Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.4958} & 99.4268 & 0.0424 & 0.0052 & 0.0002 & 0.0054 & 0.8723 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4310} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4274} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2205}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 62788 & 21247 & 102 &  & 72.230 & 0.869 & 0.531 & 0.208 & 0.648 & 0.254 &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80331 & 1000 &  & 71.767 & 1.136 & 0.588 & 0.358 & 0.791 & 0.304 &  &  & \\
-
-Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.716} & 99.434 & 0.032 & 0.006 & 0.000 & 0.006 & 0.800 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.435} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.432} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.223}\\
+\hspace{1em}No Injury & 62788 & 21246 & 102 &  & 72.2309 & 0.8686 & 0.5312 & 0.2078 & 0.6482 & 0.2537 &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80332 & 1000 &  & 71.7679 & 1.1358 & 0.5878 & 0.3580 & 0.7906 & 0.3039 &  &  & \\
+
+Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.7163} & 99.4338 & 0.0316 & 0.0062 & 0.0001 & 0.0063 & 0.8000 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4355} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4318} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2233}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 42896 & 20230 & 95 &  & 69.259 & 0.816 & 0.439 & 0.201 & 0.554 & 0.321 &  &  & \\
-
-\hspace{1em}Injury & 34546 & 79692 & 962 &  & 68.731 & 1.152 & 0.588 & 0.452 & 0.797 & 0.308 &  &  & \\
-
-Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.669} & 99.349 & 0.113 & 0.013 & 0.001 & 0.014 & 0.876 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.354} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.345} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.183}\\
+\hspace{1em}No Injury & 42896 & 20230 & 95 &  & 69.2588 & 0.8162 & 0.4387 & 0.2011 & 0.5538 & 0.3215 &  &  & \\
+
+\hspace{1em}Injury & 34546 & 79692 & 962 &  & 68.7306 & 1.1519 & 0.5880 & 0.4522 & 0.7968 & 0.3082 &  &  & \\
+
+Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.6690} & 99.3486 & 0.1129 & 0.0127 & 0.0006 & 0.0140 & 0.8760 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3539} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3447} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1831}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 43486 & 19937 & 95 &  & 69.758 & 0.820 & 0.446 & 0.198 & 0.561 & 0.315 &  &  & \\
-
-\hspace{1em}Injury & 33953 & 80009 & 961 &  & 69.227 & 1.149 & 0.593 & 0.445 & 0.800 & 0.304 &  &  & \\
-
-Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.163} & 99.340 & 0.129 & 0.013 & 0.001 & 0.014 & 0.891 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.364} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.355} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.188}\\
+\hspace{1em}No Injury & 43486 & 19937 & 95 &  & 69.7585 & 0.8200 & 0.4461 & 0.1981 & 0.5614 & 0.3154 &  &  & \\
+
+\hspace{1em}Injury & 33953 & 80009 & 961 &  & 69.2271 & 1.1487 & 0.5928 & 0.4446 & 0.7997 & 0.3038 &  &  & \\
+
+Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.1627} & 99.3398 & 0.1289 & 0.0126 & 0.0007 & 0.0140 & 0.8913 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3644} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3551} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1883}\\
 \cmidrule{1-14}
 \addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
+\hspace{1em}No Injury & 43526 & 20033 & 92 &  & 69.7291 & 0.8217 & 0.4460 & 0.1990 & 0.5619 & 0.3162 &  &  & \\
+
+\hspace{1em}Injury & 33915 & 79915 & 964 &  & 69.1938 & 1.1474 & 0.5923 & 0.4441 & 0.7988 & 0.3038 &  &  & \\
+
+Fatality & 18 & 102 & 16 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.1322} & 99.3415 & 0.1269 & 0.0134 & 0.0007 & 0.0149 & 0.8824 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3639} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3546} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1885}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 43560 & 20160 & 94 &  & 69.671 & 0.824 & 0.446 & 0.200 & 0.561 & 0.317 &  &  & \\
-
-\hspace{1em}Injury & 33876 & 79762 & 959 &  & 69.141 & 1.146 & 0.591 & 0.444 & 0.800 & 0.304 &  &  & \\
-
-Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.074} & 99.336 & 0.139 & 0.014 & 0.001 & 0.014 & 0.886 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.363} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.354} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.189}\\
+\hspace{1em}No Injury & 43561 & 20160 & 94 &  & 69.6718 & 0.8238 & 0.4458 & 0.2003 & 0.5624 & 0.3174 &  &  & \\
+
+\hspace{1em}Injury & 33875 & 79762 & 959 &  & 69.1415 & 1.1456 & 0.5914 & 0.4436 & 0.7974 & 0.3040 &  &  & \\
+
+Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.0748} & 99.3364 & 0.1393 & 0.0141 & 0.0007 & 0.0159 & 0.8859 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3629} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3538} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1886}\\
 \bottomrule

### 2020-01-23 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Completed Section: Setting for empirical assessment

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{lineno}
+\linenumbers
 \usepackage{booktabs}
@@
 
-Road safety continues to be a concern world-wide. According to a recent
+Road safety continues to be a world-wide concern. According to a recent
 report from the World Health Organization (2019), road accidents are the
@@
 
-Table \ref{tab:model-summary} collects some key summary statistics of
+Tables \ref{tab:model-full-sample-summary} and
+\ref{tab:model-subsample-summary} collect some key summary statistics of
 the estimated models. Of interest is the goodness of fit of the models,
@@
 This criterion is calculated using the estimated likelihood of the
-model, penalized by the number of coefficients in the model, so that
+model, penalized by the number of coefficients in the model, so that,
 other things being equal, it gives preference to more parsimonious
-models. Since smaller values of AIC are better fits, it is possible to
-see that, compared to the base model without opponent variables (Model
-1), there are large and significant improvements in fit to be gained by
-introducing opponent effects. However, the gains are not as large when
-hierarchical specifications are used, even when the number of additional
-coefficients that need to be estimated is not substantially larger
-(recall that the penalty per parameter in AIC is 2).
-
-The likelihood function of the model, and therefore the value of the
-AIC, depend on the size of the sample, which is why AIC is not
-comparable across models estimated with different sample sizes. For this
-reason, the full sample models cannot be compared directly to the models
-estimated with subsets of data. The models in the ensembles, however,
-can be compared, and there we find that introducing opponent variables
-leads to a better fit in most, but not all cases. One model for
-pedestrians (when the opponent is a heavy vehicle), the two models for
-bicyclists, and the three models for motorcyclists are better fits when
-opponent variables are not used.
+models. The objective is to minimize the AIC, and therefore smaller
+values of this criterion represent better model fits. Turning our
+attention to the models estimated using the full sample (Table
+\ref{tab:model-full-sample-summary}), it is possible to see that,
+compared to the base (single level) model without opponent variables
+(Model 1), there are large and significant improvements in fit to be
+gained by introducing opponent effects. However, the gains are not as
+large when hierarchical specifications are used, even when the number of
+additional coefficients that need to be estimated is not substantially
+larger (recall that the penalty per parameter in AIC is 2). The best
+model according to this measure of goodness of fit is Model 2
+(single-level with opponent effects), followed by Model 4 (hierarchical
+opponent variables), Model 3 (hierarchical traffic unit variables with
+opponent effects), and finally Model 1 (single level without opponent
+effects).
+
+It is important to note that the likelihood function of the model, and
+therefore the value of the AIC, both depend on the size of the sample,
+which is why AIC is not comparable across models estimated with
+different sample sizes. For this reason, the full sample models cannot
+be compared directly to the models estimated with subsets of data. The
+models in the ensembles, however, can be compared to each other (Table
+\ref{tab:model-subsample-summary}). As seen in the table, introducing
+opponent variables leads to a better fit in the case of most, but not
+all models. The simplest model (single level without opponent effects)
+is clearly the best fitting candidate in the case of bicycle vs light
+truck collisions, bicycle vs heavy vehicle collisions, motorcyclist vs
+light duty vehicle collisions, and motorcyclist vs heavy vehicle
+collisions. Model 1 is a statistical toss for best performance with
+other model(s) in the case of pedestrian vs heavy vehicle collisions.
+The relative likelihood of Model 1 compared to Models 2 and 3 in this
+case is 0.56, which means that either of these two competing models are
+0.56 times as probable as Model 1 to minimize the information loss.
+
+Model 2 is the best fit in the case of light truck vs heavy vehicle
+collisions. This model is also tied for best fit with Model 2 in the
+case of pedestrian vs light duty vehicle and pedestrian vs light truck
+collisions, and is a statistical toss with Model 4 in the case of heavy
+vehicle vs heavy vehicle collisions (relative likelihood is 0.592).
+Model 3 is the best fit in the case of light duty vehicle vs light duty
+vehicle collisions and heavy vehicle vs light duty vehicle. Model 4 is
+the best fit in the case of light duty vehicle vs light truck
+collisions, light duty vehicle vs heavy vehicle collisions, light truck
+vs light duty vehicle collisions, heavy vehicle vs light truck
+collisions, and motorcycle vs light truck collisions. This model is a
+statistical toss with Model 2 in the case of light truck vs light truck
+collisions, with a relative likelihood of 0.791.
 
@@
 
-\caption{\label{tab:model-summary}\label{tab:model-summary}Summary of model estimation results}
+\caption{\label{tab:model-full-sample-summary}\label{tab:model-full-sample-summary}Summary of model estimation results: Full sample models}
 \centering
@@
 \midrule
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Full sample models}}\\
-\hspace{1em}Model 1 & 164,511 & 104 & 195,215\\
-\hspace{1em}Model 2 & 164,511 & 110 & 178,943\\
-\rowcolor{gray!6}  \hspace{1em}Model 3 & 164,511 & 120 & 181,333\\
-\hspace{1em}Model 4 & 164,511 & 113 & 179,018\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Model 1 Ensemble (sample subsets by user type vs opponent)}}\\
-\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 96 & 145,390\\
-\hspace{1em}Light duty vehicle vs light truck & 3,237 & 81 & 3,943\\
-\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 90 & 5,895\\
-\hspace{1em}Light truck vs light duty vehicle & 3,121 & 81 & 3,885\\
-\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 69 & 1,170\\
-\hspace{1em}Light truck vs heavy vehicle & 198 & 66 & 288\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,726 & 81 & 4,326\\
-\hspace{1em}Heavy vehicle vs light truck & 180 & 66 & 225\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 76 & 1,147\\
-\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 90 & 2,826\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 64 & 202\\
-\hspace{1em}Pedestrian vs heavy vehicle & 376 & 66 & 409\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 82 & 654\\
-\hspace{1em}Bicyclist vs light truck & 116 & 44 & 84\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
-\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 80 & 1,367\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 58 & 153\\
-\hspace{1em}Motorcyclist vs heavy vehicle & 62 & 45 & 88\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Model 2 Ensemble (sample subsets by user type vs opponent)}}\\
-\hspace{1em}Light duty vehicle vs light duty vehicle & 114,841 & 99 & 143,903\\
-\hspace{1em}Light duty vehicle vs light truck & 3,237 & 84 & 3,927\\
-\rowcolor{gray!6}  \hspace{1em}Light duty vehicle vs heavy vehicle & 5,013 & 93 & 5,878\\
-\hspace{1em}Light truck vs light duty vehicle & 3,121 & 84 & 3,877\\
-\rowcolor{gray!6}  \hspace{1em}Light truck vs light truck & 809 & 72 & 1,156\\
-\hspace{1em}Light truck vs heavy vehicle & 198 & 69 & 281\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs light duty vehicle & 4,732 & 86 & 4,283\\
-\hspace{1em}Heavy vehicle vs light truck & 179 & 67 & 205\\
-\rowcolor{gray!6}  \hspace{1em}Heavy vehicle vs heavy vehicle & 779 & 79 & 1,136\\
-\hspace{1em}Pedestrian vs light duty vehicle & 7,176 & 93 & 2,821\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian vs light truck & 328 & 67 & 200\\
-\hspace{1em}Pedestrian vs heavy vehicle & 376 & 69 & 410\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist vs light duty vehicle & 3,521 & 85 & 659\\
-\hspace{1em}Bicyclist vs light truck & 145 & 59 & 114\\
-\rowcolor{gray!6}  Bicyclist vs heavy vehicle & NA & NA & \vphantom{1} NA\\
-\hspace{1em}Motorcyclist vs light duty vehicle & 2,298 & 83 & 1,373\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist vs light truck & 127 & 61 & 153\\
-\hspace{1em}Motorcyclist vs heavy vehicle & 63 & 45 & 90\\
+\rowcolor{gray!6}  Model 1 & 164,511 & 102 & 195,215\\
+Model 2 & 164,511 & 108 & 178,943\\
+\rowcolor{gray!6}  Model 3 & 164,511 & 118 & 181,333\\
+Model 4 & 164,511 & 111 & 179,018\\
 \bottomrule
-\multicolumn{4}{l}{\textit{Note: }}\\
-\multicolumn{4}{l}{There are zero cases of Bicyclist vs heavy vehicle in the sample}\\
 \end{tabular}
 \end{table}
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:model-subsample-summary}\label{tab:model-summary}Summary of model estimation results: Subsample Models}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{lccccccccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} & \multicolumn{2}{c}{Model 4} \\
+\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
+Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC\\
+\midrule
+\rowcolor{gray!6}  Light duty vehicle vs light duty vehicle & 114,841 & 94 & 145,390 & 97 & 143,903 & 100 & 143,896 & 100 & 144,004\\
+Light duty vehicle vs light truck & 3,237 & 79 & 3,943 & 82 & 3,927 & 85 & 3,937 & 85 & 3,922\\
+\rowcolor{gray!6}  Light duty vehicle vs heavy vehicle & 5,013 & 88 & 5,895 & 91 & 5,878 & 94 & 5,888 & 94 & 5,864\\
+Light truck vs light duty vehicle & 3,121 & 79 & 3,885 & 82 & 3,877 & 85 & 3,881 & 85 & 3,875\\
+\rowcolor{gray!6}  Light truck vs light truck & 809 & 67 & 1,170 & 70 & 1,156 & 73 & 1,162 & 73 & 1,155\\
+\addlinespace
+Light truck vs heavy vehicle & 198 & 64 & 288 & 67 & 281 & 70 & 287 & 70 & 286\\
+\rowcolor{gray!6}  Heavy vehicle vs light duty vehicle & 4,726 & 79 & 4,326 & 84 & 4,283 & 86 & 4,268 & 87 & 4,287\\
+Heavy vehicle vs light truck & 180 & 64 & 225 & 65 & 205 & 67 & 207 & 66 & 187\\
+\rowcolor{gray!6}  Heavy vehicle vs heavy vehicle & 779 & 74 & 1,147 & 77 & 1,136 & 80 & 1,141 & 80 & 1,137\\
+Pedestrian vs light duty vehicle & 7,176 & 88 & 2,826 & 91 & 2,821 & 91 & 2,821 & 93 & 2,827\\
+\addlinespace
+\rowcolor{gray!6}  Pedestrian vs light truck & 328 & 62 & 202 & 65 & 200 & 65 & 200 & 68 & 206\\
+Pedestrian vs heavy vehicle & 376 & 64 & 409 & 67 & 410 & 67 & 410 & 70 & 417\\
+\rowcolor{gray!6}  Bicyclist vs light duty vehicle & 3,521 & 80 & 654 & 83 & 659 & 83 & 659 & 85 & 657\\
+Bicyclist vs light truck & 116 & 42 & 84 & 57 & 114 & 57 & 114 & 54 & 108\\
+\rowcolor{gray!6}  Bicyclist vs heavy vehicle & - & - & - & - & - & - & - & - & -\\
+\addlinespace
+Motorcyclist vs light duty vehicle & 2,298 & 78 & 1,367 & 81 & 1,373 & 81 & 1,373 & 84 & 1,373\\
+\rowcolor{gray!6}  Motorcyclist vs light truck & 127 & 56 & 153 & 59 & 153 & 59 & 153 & 47 & 94\\
+Motorcyclist vs heavy vehicle & 62 & 43 & 88 & 45 & 90 & 46 & 92 & 51 & 102\\
+\bottomrule
+\multicolumn{10}{l}{\textit{Note: }}\\
+\multicolumn{10}{l}{There are zero cases of Bicyclist vs heavy vehicle in the sample}\\
+\end{tabular}}
+\end{table}
+\end{landscape}
 
@@
 
-\hypertarget{outcome-shares-based-on-estimated-probabilities}{%
-\subsection{Outcome shares based on estimated
-probabilities}\label{outcome-shares-based-on-estimated-probabilities}}
+\hypertarget{outcome-shares-based-on-predicted-probabilities}{%
+\subsection{Outcome shares based on predicted
+probabilities}\label{outcome-shares-based-on-predicted-probabilities}}
 
@@
 
-\hypertarget{predicted-outcomes}{%
-\subsection{Predicted outcomes}\label{predicted-outcomes}}
+\hypertarget{outcome-frequency-based-on-predicted-classes}{%
+\subsection{Outcome frequency based on predicted
+classes}\label{outcome-frequency-based-on-predicted-classes}}
 
@@
 
-\hypertarget{sec:insights}{%
-\section{Best model: insights}\label{sec:insights}}
-
-Create a table with the results of Model 2.
-
 \hypertarget{sec:further-considerations}{%

### 2020-01-24 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Started working on the section on model assessment. Reorganized some of the results and created new tables to present more effectively the verification statistics.

@@
 severity analysis}
-    \author[]{Antonio Paez\corref{Corresponding Author}}
+    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
    \ead{paezha@mcmaster.ca} 
@@
   severity of crashes. Crash severity research is useful to clarify the
-  way different factors can influence the outcome of the event. The
+  way different factors can influence the outcome of an event. The
   objective of this paper is to assess different strategies to model the
-  interactions between participants in a crash, in the context of crashes
+  interactions between participants in a crash in the context of crashes
   involving two parties. Towards this objective, a series of models are
@@
   multi-level form), as well as by subsetting the dataset. The models are
-  assessed using predicted shares and outcomes, and the results highlight
-  the importance of considering opponent effects in crash severity
-  analysis. On the other hand, the study suggests that hierarchical (i.e.,
-  multi-level) specifications and subsetting do not perform necessarily
+  assessed using predicted shares and classes of outcomes, and the results
+  highlight the importance of considering opponent effects in crash
+  severity analysis. The study also suggests that hierarchical (i.e.,
+  multi-level) specifications and subsetting do not necessarily perform
   better than a relatively simple single-level model with opponent
-  effects.
+  effects. The results of this study provide insights regarding the
+  performace of different modelling strategies, and should be informative
+  to researchers working with crash severity models.
   \end{abstract}
@@
 leading causes of death, road accidents are the only cause of death
-unrelated to disease, disorder, or infection. Road accidents impose a
-heavy burden on individuals and society as a whole. Gobally, the rate of
-road collision-related deaths per 100,000 population and 100,000
-vehicles have both fallen, even as the number of vehicles has grown
-(World Health Organization, 2019, Figs. 1 and 2). These gains, although
-they are to be celebrated, cannot distract from the crushing economic
-cost of premature death (e.g., Symons et al., 2019; Wijnen et al.,
-2019), not to mention the long-term consequences for survivors, measured
-in sometimes crippling emotional and physical pain (e.g., Merlin et al.,
-2007; Devlin et al., 2019; Pelissier et al., n.d.).
+unrelated to disease, health disorder, or infection. For this reason,
+road accidents impose a heavy burden on individuals and society as a
+whole. Gobally, the rate of road collision-related deaths per 100,000
+population and 100,000 vehicles have both fallen, even as the number of
+vehicles has grown (World Health Organization, 2019, Figs. 1 and 2).
+These gains, although they are to be celebrated, cannot distract from
+the crushing economic cost of premature death (e.g., Symons et al.,
+2019; Wijnen et al., 2019), not to mention the long-term consequences
+for survivors, measured in sometimes crippling emotional and physical
+pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier et al.,
+n.d.).
 
@@
 variations of the impacts among different population segments, numerous
-efforts efforts have been conducted to better understand the factors
-that affect road safety - including the probable consequences of
-crashes. Consequently, a stream of research in the analysis of road
-accidents is concerned with the severity of crashes. In particular,
-multivariate analysis of crash severity is a useful way to clarify the
-way various factors can affect the outcome of an incident, to
-discriminate between various levels of injury, from no injury (i.e.,
-property damage only), to different degrees of injury up to and
-including fatality. This is an active area of research (e.g., Savolainen
-et al., 2011), and one where methodological developments have aimed at
-improving the reliability, accuracy, and precision of models.
+efforts have been conducted to better understand the factors that affect
+road safety - including the probable consequences of crashes. Along
+these lines, a stream of research in the analysis of road accidents is
+concerned with the severity of crashes. In particular, multivariate
+analysis of crash severity is a useful way to clarify the way various
+factors can affect the outcome of an incident, and to discriminate
+between various levels of injury, from no injury (i.e., property damage
+only), to different degrees of injury up to and including death. This is
+an active area of research, and one where methodological developments
+have aimed at improving the reliability, accuracy, and precision of
+models (e.g., Savolainen et al., 2011; Bogue et al., 2017; Shamsunnahar
+and Eluru, 2013).
 
@@
 using the dataset corresponding to 2016 (i.e., \emph{backcasting}). The
-models are assessed using predicted shares and predicted individual
-outcomes using an extensive array of verification statistics. The
-results highlight the importance of considering opponent effects in
-crash severity analysis to improve the goodness-of-fit and predictive
-performance o. On the other hand, the study suggests that hierarchical
-variable specifications and subsetting do not perform necessarily better
-than a relatively simple single-level model with opponent effects.
+models are assessed using predicted shares and predicted classes of
+outcomes at the individual level, using an extensive array of
+verification statistics. The results highlight the importance of
+considering opponent effects in crash severity analysis to improve the
+goodness-of-fit and predictive performance o. On the other hand, the
+study suggests that hierarchical variable specifications and subsetting
+do not perform necessarily better than a relatively simple single-level
+model with opponent effects.
 
@@
 used to analyze crash severity, with a particular focus on techniques
-that consider the interactions between participants in a crash. Section
+that consider the effect of opponents in a crash. Section
 \ref{sec:application} describes the data requirements, data
-preprocessing, and the modelling strategy, along with the results of
+preprocessing, and the modelling strategies, along with the results of
 model estimation. The results of assessing the models and the discussion
@@
 to participants has been a preoccupation of transportation researchers,
-planners, auto insurance companies, governments and the public for
-decades. One of the earliests studies to investigate the severity of
+planners, auto insurance companies, governments, and the general public
+for decades. One of the earliests studies to investigate the severity of
 injuries conditional on an accident having occurred was by White and
@@
 particular situations (e.g., crashes at intersections, Mussone et al.,
-2017; crashes in rural roads, Gong and Fan, 2017) or special populations
-(e.g., crashes involving motorcyclists or active travelers; see Shaheed
-et al., 2013; Salon and McIntyre, 2018).
-
-Crash severity is often modelled using models for discrete outcomes,
-including classification techniques from machine learning (e.g.,
-Iranitalab and Khattak, 2017; Chang and Wang, 2006; Effati et al., 2015;
-Khan et al., 2015), Poisson models for counts (e.g., Ma et al., 2008),
-unordered logit/probit models (e.g., Tay et al., 2011), as well as
-ordered logit/probit models (e.g., Rifaat and Chin, 2007), with numerous
-variants, such as random parameters/mixed logit (e.g., Aziz et al.,
-2013; Haleem and Gan, 2013), partial proportional odds models (e.g.,
-Mooradian et al., 2013; Sasidharan and Menendez, 2014), and the use of
-copulas (e.g., Wang et al., 2015). Recent reviews of methods include
-Savolainen et al.~(2011) and Shamsunnahar and Eluru (2013).
+2017; crashes in rural roads, Gong and Fan, 2017), and crashes involving
+special population groups (e.g., crashes involving motorcyclists or
+active travelers; see Shaheed et al., 2013; Salon and McIntyre, 2018).
+
+Crash severity is often modelled using models for discrete outcomes. An
+analyst interested in crash severity has at their disposal an ample menu
+of models to choose from, including classification techniques from
+machine learning (e.g., Iranitalab and Khattak, 2017; Chang and Wang,
+2006; Effati et al., 2015; Khan et al., 2015), Poisson models for counts
+(e.g., Ma et al., 2008), unordered logit/probit models (e.g., Tay et
+al., 2011), as well as ordered logit/probit models (e.g., Rifaat and
+Chin, 2007), with numerous variants, such as random parameters/mixed
+logit (e.g., Aziz et al., 2013; Haleem and Gan, 2013), partial
+proportional odds models (e.g., Mooradian et al., 2013; Sasidharan and
+Menendez, 2014), and the use of copulas (e.g., Wang et al., 2015).
+Recent reviews of methods include Savolainen et al.~(2011), Shamsunnahar
+and Eluru (2013), and Mannering et al.~(2016).
 
@@
 motorcyclists, or pedestrians. The multiplicity of roles makes for
-complicated modelling choices when trying to understand the severity of
-injuries; for example, what is the unit of analysis, the person, the
+complicated modelling decisions when trying to understand the severity
+of injuries; for example, what is the unit of analysis, the person, the
 traffic unit, or the collision? Not surprisingly, it is possible to find
 examples of studies that adopt different perspectives. A common
-simplifying strategy to specify a model is to consider only
+simplifying strategy in model specification is to consider only
 \emph{drivers} and/or only \emph{single-vehicle} crashes (e.g., Kim et
@@
 their analysis the age, wheelbase, weight, and engine size of the
-opposing vehicle while Bogue et al.~(2017) used the body type of the
+opposing vehicle, while Bogue et al.~(2017) used the body type of the
 opposing vehicle. Penmetsa et al.~(2017) and Duddu et al.~(2018) are two
-of the most comprehensive examples of using opponent's information, with
-the physical condition, sex, age, and vehicle type of the opponent. The
-twin strategies of subsetting the sample and using the attributes of the
-opponent are not mutually exclusive, but neither are they consistently
-used together, as a scan of the literature reveals.
+of the most comprehensive examples of using opponent's information, as
+they used individual attributes of opponents (their physical condition,
+sex, and age), as well as characteristics of the opposing traffic unit
+(the vehicle type of the opponent). The twin strategies of subsetting
+the sample and using the attributes of the opponent are not mutually
+exclusive, but neither are they used consistently together, as a scan of
+the literature reveals.
 
@@
 The latent variable is not observed directly, but it is possible to
-posit a probabilistic relationship with the outcome \(y_{ik}\) (the
-severity of crash \(k\) for participant \(i\)). Depending on the
-characteristics of the data and the assumptions made about the random
-component of the latent function different models can be obtained. For
-example, if crash severity is coded as a binary variable (e.g.,
-non-fatal/fatal), we can relate the latent variable to the outcome as
-follows:
+posit a probabilistic relationship with the outcome \(y_{itk}\) (the
+severity of crash \(k\) for participant \(i\) in traffic unit \(t\)).
+Depending on the characteristics of the data and the assumptions made
+about the random component of the latent function different models can
+be obtained. For example, if crash severity is coded as a binary
+variable (e.g., non-fatal/fatal), we can relate the latent variable to
+the outcome as follows:
 
@@
 In other words, the probability that individual \(i\) in traffic unit
-\(k\) and crash \(k\) was a fatality equals the probability that the
+\(t\) and crash \(k\) was a fatality equals the probability that the
 latent variable is greater than zero. This implies (see Maddala, 1986,
@@
 If the random terms \(\epsilon_{itk}\) are assumed to follow the
-logistic distribution, the the binary logit model is obtained; if they
+logistic distribution, then the binary logit model is obtained; if they
 are assumed to follow the normal distribution, the binary probit model
@@
 ordered probit model. Estimation methods for these models are very
-well-established (e.g., Maddala, 1986; Train, 2009)
-
-There are numerous variations of the basic modelling framework above,
-including hierarchical models, bivariate models, multinomial models, and
-Bayesian models, among others (see Savolainen et al., 2011 for a review
-of methods).
+well-established (e.g., Maddala, 1986; Train, 2009). There are numerous
+variations of the basic modelling framework above, including
+hierarchical models, bivariate models, multinomial models, and Bayesian
+models, among others (see Savolainen et al., 2011 for a review of
+methods).
 
@@
 Introduction of opponent related-variables was explained in the
-preceding subsection. In this way, a base model is given by Equation
-\ref{eq:latent-function}, whereas Strategy 1 (inclusion of
-opponent-related variables) is Equation
-\ref{eq:latent-function-with-opponent-variables}. These two equations
-are also examples of single-level model. Next we describe Strategies 2
-and 3.
-
-\hypertarget{hierarchical-model-specification}{%
-\subsubsection{Hierarchical model
-specification}\label{hierarchical-model-specification}}
+preceding subsection. In this way, a base model that ignores opponent
+effects is given by Equation \ref{eq:latent-function}. Strategy 1, in
+contrast, is Equation \ref{eq:latent-function-with-opponent-variables},
+which includes opponent-related variables. These two equations are also
+examples of single-level models. Next we describe Strategies 2 and 3.
+
+\hypertarget{strategy-2-hierarchical-model-specification}{%
+\subsubsection{Strategy 2: hierarchical model
+specification}\label{strategy-2-hierarchical-model-specification}}
 
@@
 person-opponent, in which case the opponent-level attributes are nested
+in the following fashion:
 
@@
 Alterted readers will identify this model specification strategy as
-Casetti's expansion method (Casetti, 1972). This is a deterministic
-strategy for modelling contextual effects which, when augmented with
-random components, becomes the well-known multi-level modelling method
-(Hedeker and Gibbons, 1994, more on this in Section
+Casetti's expansion method (Casetti, 1972; Roorda et al., 2010). This is
+a deterministic strategy for modelling contextual effects which, when
+augmented with random components, becomes the well-known multi-level
+modelling method (Hedeker and Gibbons, 1994, more on this in Section
 \ref{sec:further-considerations}). It is worthwhile to note that
@@
 attributes nested within traffic units, which in turn are nested within
-collisions. We do not explore this further in the current paper.
-
-\hypertarget{sample-subsetting}{%
-\subsubsection{Sample subsetting}\label{sample-subsetting}}
+collisions. We do not explore higher-level hierarchies further in the
+current paper.
+
+\hypertarget{strategy-3-sample-subsetting}{%
+\subsubsection{Strategy 3: sample
+subsetting}\label{strategy-3-sample-subsetting}}
 
@@
 function with restrictions as follows. Consider a continuous variable,
-e.g., age of person, and we wish to concentrate the analysis on older
-adults (e.g., Dissanayake and Lu, 2002). The latent function is defined
-as desired (see above), however, the following restriction might be
-applied to the sample
+e.g., age of person, and imagine that we wish to concentrate the
+analysis on older adults (e.g., Dissanayake and Lu, 2002). The latent
+function is defined as desired (see above), however, the following
+restriction is applied to the sample:
 
@@
 To assess the performance of the various modelling strategies we use
-data from Canada's National Collision Database (NCDB). This is database
-contains all motor vehicle collisions on public roads in Canada as
-reported by a police service. Data are collected by provinces and
-territories, and shared with the federal government, where data are
-combined, tracked, and analyzed for reporting of deaths, injuries, and
-collisions in Canada at the national level. The NCDB is provided by
-Transport Canada, the agency of the federal government of Canada in
-charge of transportation policies and programs, under the Open
-Government License - Canada version 2.0
+data from Canada's National Collision Database (NCDB). This database
+contains all police-reported motor vehicle collisions on public roads in
+Canada. Data are originally collected by provinces and territories, and
+shared with the federal government, that proceeds to combine, track, and
+analyze them for reporting deaths, injuries, and collisions in Canada at
+the national level. The NCDB is provided by Transport Canada, the agency
+of the federal government of Canada in charge of transportation policies
+and programs, under the Open Government License - Canada version 2.0
 {[}https://open.canada.ca/en/open-government-licence-canada{]}.
@@
 \ref{tab:variable-categories}, crash-related variables and road-related
-variables are collected in a single variable class, namely
+variables are collected under a single variable class, namely
 collision-related, since they are unique for each crash.
@@
 \multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Preprocessed data for analysis available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
 \endlastfoot
@@
 
-The database is scanned to remove all records that are not a person
-(including parked cars and other objects) or that are missing
-information (e.g., runaway vehicles). Next, records missing at least one
-of the next variables are removed: P\_USER (the road user class), P\_SEX
-(sex), P\_AGE (age), and P\_ISEV (individual-level severity of crash).
-This initial filter ensures that all records are complete from the
-perspective of key information.
+We apply an initial filter, whereby we scan the database to remove all
+records that are not a person (including parked cars and other objects)
+or that are missing information (as is the case when a participant in a
+crash is a runaway vehicle). Next, records missing at least one of the
+next variables are removed: P\_USER (the road user class), P\_SEX (sex),
+P\_AGE (age), and P\_ISEV (individual-level severity of crash). This
+initial filter ensures that all records are complete from the
+perspective of key information for analysis.
 
@@
 participant, there are 46,483 collisions involving two participants,
-19,433 involving 3 participants, 8,250 involving four participants,
-3,783 involving 5 participants, 1,789 involving 6 participants, and
-1,491 collisions involving seven or more participants. These
-participants were possibly occupants in different vehicles or were
-otherwise their own traffic units, as follows: the sample includes
-174,741 drivers, 61,403 passengers, 10,798 pedestrians, 5,286
-bicyclists, and 6,564 motorcyclists.
-
-The next step is to remove all collisions involving only one
-participant. This still leaves numerous cases where multiple
-participants could have been in a single vehicle, for instance in a
-collision against a stationary object. Therefore, we proceed to use the
-vehicle sequence number can to find the number of vehicles involved in
-each collision. This reveals that there are 20,732 collisions that
-involve only one vehicle but possibly multiple participants (i.e.,
-driver and one or more passengers). In addition, there are 165,520
-collisions involving two vehicles (and possibly multiple participants).
-Finally, there are 40,242 collisions with three or more vehicles.
+19,433 collisions with three participants, 8,250 collisions involving
+four participants, 3,783 collisions with five participants, 1,789
+collisions with six participants, and 1,491 collisions involving seven
+or more participants. These participants were possibly occupants in
+different vehicles or were otherwise their own traffic units.
+Accordingly, the sample includes 174,741 drivers, 61,403 passengers,
+10,798 pedestrians, 5,286 bicyclists, and 6,564 motorcyclists.
+
+The next step in our data preprocessing protocol is to remove all
+collisions that involve only one participant. This still leaves numerous
+cases where multiple participants could have been in a single vehicle,
+for instance in a collision against a stationary object. Therefore, we
+proceed to use the vehicle sequence number to find the number of
+vehicles involved in each collision. This reveals that there are 20,732
+collisions that involve only one vehicle but possibly multiple
+participants (i.e., driver and one or more passengers). In addition,
+there are 165,520 collisions involving two vehicles (and possibly
+multiple participants). Finally, there are 40,242 collisions with three
+or more vehicles.
 
@@
 common cases in two-vehicle collisions are those that include drivers
-(40,297 collisions; reflective of the prevalence of single-occupant
-vehicles), followed by cases with passengers (14,120 collisions),
-pedestrians (5,204 collisions), byciclists (2,238 collisions), and
-motorcyclists (1,016 collisions). The distribution of individuals per
-traffic unit is as follows: 80,382 individuals are coded as being in
-V\_ID = 1, 76,523 individuals are coded as being in V\_ID = 2, and 7,932
-individuals are coded as pedestrians. In addition, 683 individuals are
-coded as being in vehicles 3 through 9, despite our earlier filter to
-retain only collisions with two vehicles. We therefore proceed to select
-only individuals assigned to vehicles 1 or 2, as well as pedestrians. As
-a result of this filter a number of cases with only one known
-participant need to be removed.
+(40,297 collisions; this is reflective of the prevalence of
+single-occupant vehicles). This is followed by cases with passengers
+(14,120 collisions), pedestrians (5,204 collisions), byciclists (2,238
+collisions), and motorcyclists (1,016 collisions). The distribution of
+individuals per traffic unit is as follows: 80,382 individuals are coded
+as being in V\_ID = 1, 76,523 individuals are coded as being in V\_ID =
+2, and 7,932 individuals are coded as pedestrians. In addition, 683
+individuals are coded as being in vehicles 3 through 9, despite our
+earlier filter to retain only collisions with two vehicles. We therefore
+proceed to select only individuals assigned to vehicles 1 or 2, as well
+as pedestrians. As a result of this filter a number of cases with only
+one known participant need to be removed.
 
@@
 depending on the traffic units involved: 1) vehicle vs vehicle
-collisions (``vehicle'' includes all motorized vehicles, as well as
-motorcycles/mopeds and bicycles); and 2) vehicle vs pedestrian
+collisions (``vehicle'' is all motorized vehicles, including
+motorcycles/mopeds, as well as bicycles); and 2) vehicle vs pedestrian
 collisions. To identify opponents in a collision, it is convenient to
-classify collisions by pedestrian involvement. Accordingly, we find that
-the database includes 16,636 collisions are vehicle vs pedestrian
-(possibly multiple pedestrians), and 147,594 collisions involving two
-vehicles. After splitting the database in this way, we can now extract
-relevant information about participants in the collision. This involves
-renaming the person-level variables depending so that we can distinguish
-each individual by their role as an individual or an opponent in a given
-collision. Notice that when working with individuals in vehicles, only
-the driver is considered an opponent in a collisions (i.e., passengers
-are never considered opponents).
+classify collisions by pedestrian involvement. In this way, we find that
+the database includes 16,636 collisions that are vehicle vs pedestrian
+(possibly multiple pedestrians), and 147,594 collisions that involve two
+vehicles. After splitting the database according to pedestrian
+involvement, we can now extract relevant information about participants
+in the collision. This involves renaming the person-level variables so
+that we can distinguish each individual by their role as an individual
+or an opponent in a given record. Notice that when working with
+individuals in vehicles, only the driver is considered a legitimate
+opponent in a collision (i.e., passengers are never considered
+opponents).
 
 Once the personal attributes of individuals that can be considered
-opponents in a given collision have ben extracted, their information can
-be joined to the individual records by means of the collision unique
+opponents in a given collision have been extracted, their information is
+joined to the individual records by means of the collision unique
 identifier. As a result of this process, a new set of variables are now
-available for analysis: the age, sex, use of safety device, and road
-user class of the opponent, as well as the type of the opponent vehicle.
-A summary of opponent interactions and outcomes can be found in Table
+available for analysis: the age, sex, and road user class of the
+opponent, as well as the type of the opposing vehicle. A summary of
+opponent interactions and outcomes can be found in Table
 \ref{tab:summary-opponent-information}. The information there shows that
 the most commont type of opponent for drivers is other drivers, followed
-by pedestrians. The only opponent of pedestrians, on the other hand, are
-drivers. Bicyclists and motorcyclists, on the other hand, are mostly
+by pedestrians. The only opponents of pedestrians, on the other hand,
+are drivers. Bicyclists and motorcyclists, on the other hand, are mostly
 opposed by drivers, but occasionally by other road users as well. In
@@
 motorcyclist. Injuries are also more common when the opponent is a
-driver, whereas the likelihood of no injuries is higher when the
+driver, whereas no injuries are relatively more frequent when the
 opponent is a pedestrian or a bicyclist.
@@
 vehicle weight rating); and heavy vehicles (all other vehicles \(\ge\)
-4536 in gross vehicle weight rating). Furthermore, this classification
-of vehicle types is combined with the road user class to distiguish
-between drivers and passengers of light duty vehicles, light trucks, and
-heavy vehicles, in addition to pedestrians, bicyclists, and
+4536 in gross vehicle weight rating). Furthermore, this typology of
+vehicle is combined with the road user class of the individual to
+distiguish between drivers and passengers of light duty vehicles, light
+trucks, and heavy vehicles, in addition to pedestrians, bicyclists, and
 motorcyclists. This is done for both the individual and the opponent.
@@
 For example, for a hierarchical definition of traffic unit-level
-variables, age (and the square of aged) are interacted with gender, road
-user class, and vehicle type. For hierarchical opponent variables, age
-(and the square of age) are interacted with age of opponent (and squared
-aged of opponent). The variables thus obtained are shown in Table
+variables, age (and the square of age to account for possible
+non-monotonic effects) are interacted with gender, road user class, and
+vehicle type. For hierarchical opponent variables, age (and the square
+of age) are interacted with the age of opponent (and the corresponding
+square). The variables thus obtained are shown in Table
 \ref{tab:model-specification-summary}. As seen in the table, Models 1
@@
 traffic unit, while Model 4 considers the hierarchy on the basis of the
-collision.
+opponent.
 
@@
 Table \ref{tab:model-specification-summary}). By subsetting the sample,
-opponent effects are incorporated implicitly. Models 1 and 2 are
-re-estimated using this strategy, dropping variables as necessary where
-they are irrelevant (for instance, after filtering for pedestrians, no
-other traffic unit types are present in the subset of data).
+at least \emph{some} opponent effects are incorporated implicitly.
+Models 1 and 2 are re-estimated using this strategy, dropping variables
+as necessary whenever they become irrelevant (for instance, after
+filtering for pedestrians, no other traffic unit types are present in
+the subset of data). In addition to variables that are no longer
+relevant in some subsamples, it is important to note that when using
+some subsamples a few variables had to be occasionally dropped to avoid
+convergence issues. This tended to happen particularly with smaller
+subsamples where some particular combination of attributes was rare as a
+result of subsampling (e.g., in 2017 there were few or no collisions
+that involved a motorcyclist and a heavy vehicle in a bridge, or
+overpass, or viaduct). The process of estimation carefully paid
+attention to convergence issues to ensure the validity of the models
+reported here.
+
+The resuls of model estimation are discussed in the following section.
 
@@
 
-Tables \ref{tab:model-full-sample-summary} and
-\ref{tab:model-subsample-summary} collect some key summary statistics of
-the estimated models. Of interest is the goodness of fit of the models,
-which in the case is measured with Akaike's Information Criterion (AIC).
-This criterion is calculated using the estimated likelihood of the
-model, penalized by the number of coefficients in the model, so that,
-other things being equal, it gives preference to more parsimonious
-models. The objective is to minimize the AIC, and therefore smaller
-values of this criterion represent better model fits. Turning our
-attention to the models estimated using the full sample (Table
-\ref{tab:model-full-sample-summary}), it is possible to see that,
-compared to the base (single level) model without opponent variables
-(Model 1), there are large and significant improvements in fit to be
-gained by introducing opponent effects. However, the gains are not as
-large when hierarchical specifications are used, even when the number of
-additional coefficients that need to be estimated is not substantially
-larger (recall that the penalty per parameter in AIC is 2). The best
-model according to this measure of goodness of fit is Model 2
-(single-level with opponent effects), followed by Model 4 (hierarchical
-opponent variables), Model 3 (hierarchical traffic unit variables with
-opponent effects), and finally Model 1 (single level without opponent
-effects).
-
-It is important to note that the likelihood function of the model, and
-therefore the value of the AIC, both depend on the size of the sample,
-which is why AIC is not comparable across models estimated with
-different sample sizes. For this reason, the full sample models cannot
-be compared directly to the models estimated with subsets of data. The
-models in the ensembles, however, can be compared to each other (Table
-\ref{tab:model-subsample-summary}). As seen in the table, introducing
-opponent variables leads to a better fit in the case of most, but not
-all models. The simplest model (single level without opponent effects)
-is clearly the best fitting candidate in the case of bicycle vs light
-truck collisions, bicycle vs heavy vehicle collisions, motorcyclist vs
-light duty vehicle collisions, and motorcyclist vs heavy vehicle
-collisions. Model 1 is a statistical toss for best performance with
-other model(s) in the case of pedestrian vs heavy vehicle collisions.
-The relative likelihood of Model 1 compared to Models 2 and 3 in this
-case is 0.56, which means that either of these two competing models are
-0.56 times as probable as Model 1 to minimize the information loss.
+\hypertarget{sec:assessment}{%
+\section{Model assessment}\label{sec:assessment}}
+
+In this section we report an in-depth examination of the performance of
+the models. To this end, first inspect the goodness of fit of the
+models. Next, we use the models to conduct in-sample predictions (i.e.,
+\emph{nowcasting}), using the same sample that was used to estimate the
+models. In addition, we also conduct out-of-sample predictions, using
+the dataset corresponding to the year 2016, processed in identical way
+as the estimation sample. This is an example of \emph{backcasting}.
+Using these predictions we evaluate the models in two ways: first, we
+compute the estimated shares of each outcome based on the predicted
+probabilities; and secondly, we evaluate the classes of outcomes of the
+individual-level predictions.
+
+\hypertarget{goodness-of-fit-of-models}{%
+\subsection{Goodness of fit of models}\label{goodness-of-fit-of-models}}
+
+We begin our empirical assessment by examining the results of estimating
+the models described above. Tables \ref{tab:model-full-sample-summary}
+and \ref{tab:model-subsample-summary} collect some key summary
+statistics of the estimated models. Of interest is the goodness of fit
+of the models, which in the case is measured with Akaike's Information
+Criterion (\(AIC\)). This criterion is calculated as follows:
+
+\begin{equation}
+\label{eq:aic}
+AIC = 2Z - 2\ln{\hat{L}}
+\end{equation}
+
+\noindent where \(Z\) is the number of coefficients estimated by the
+model, and \(\hat{L}\) the maximized likelihood of the model. Since
+\(AIC\) penalizes the model fit by means of the number of coefficients,
+this criterion gives preference to more parsimonious models. The
+objective is to minimize the \(AIC\), and therefore smaller values of
+this criterion represent better model fits. Model comparison can be
+conducted using the relative likelihood. Suppose that we have two
+models, say Model \(a\) and Model \(b\), with \(AIC_{a} \le AIC_{b}\).
+The relative likelihood is calculated as:
+
+\begin{equation}
+\label{eq:relative-likelihood}
+e^{(AIC_b - AIC_b)/2}
+\end{equation}
+
+The relative likelihood is interpreted as the probability that Model
+\(b\) minimizes the information loss as well as Model \(a\).
+
+Turning our attention to the models estimated using the full sample
+(Table \ref{tab:model-full-sample-summary}), it is possible to see that,
+compared to the base (single-level) model without opponent variables
+(Model 1), there are large and significant improvements in goodness of
+fit to be gained by introducing opponent effects. However, the gains are
+not as large when hierarchical specifications are used, even when the
+number of additional coefficients that need to be estimated is not
+substantially larger (recall that the penalty per coefficient in \(AIC\)
+is 2). The best model according to this measure of goodness of fit is
+Model 2 (single-level with opponent effects), followed by Model 4
+(hierarchical opponent variables), Model 3 (hierarchical traffic unit
+variables with opponent effects), and finally Model 1 (single-level
+without opponent effects).
+
+It is important to note that the likelihood function of a model, and
+therefore the value of its \(AIC\), both depend on the size of the
+sample, which is why \(AIC\) is not comparable across models estimated
+with different sample sizes. For this reason, the full sample models
+cannot be compared directly to the models estimated with subsets of
+data. The models in the ensembles, however, can be compared to each
+other (Table \ref{tab:model-subsample-summary}). As seen in the table,
+introducing opponent variables leads to a better fit in the case of
+most, but not all models. The simplest model (single-level without
+opponent effects) is clearly the best fitting candidate in the case of
+bicycle vs light truck collisions, bicycle vs heavy vehicle collisions,
+motorcyclist vs light duty vehicle collisions, and motorcyclist vs heavy
+vehicle collisions. Model 1 is a statistical toss for best performance
+with two competing models in the case of pedestrian vs heavy vehicle
+collisions. The relative likelihood of Model 1 compared to Models 2 and
+3 in this case is 0.56, which means that these two models are 0.56 times
+as probable as Model 1 to minimize the information loss.
 
@@
 These results give some preliminary ideas about the relative performance
-of the different modelling strategies. In the next section we delve more
-deeply into this question.
+of the different modelling strategies. In the next subsection we delve
+more deeply into this question by examining the predictive performance
+of the various modelling strategies. The results up to this point
+indicate that different model specification strategies might work best
+when combined with subsampling strategies. For space reasons, from this
+point onwards, we will consider the ensembles of models for predictions
+and will not compare individual models within the ensembles; this we
+suggest is a matter for future research.
 
@@
 
-\hypertarget{sec:assessment}{%
-\section{Model assessment}\label{sec:assessment}}
-
-In this section we report an in-depth examination of the performance of
-the models. To this end, we use the models to conduct in-sample
-predictions (i.e., nowcasting), using the same sample that was used to
-estimate the models. In addition, we also conduct out-of-sample
-predictions, using the dataset corresponding to the year 2016, processed
-in identical way as the estimation sample (i.e., the 2017 dataset). This
-is an example of backcasting. Broadly, we evaluate the models in two
-ways: first, we compute the estimated shares of each outcome based on
-the predicted probabilities; and secondly, we evaluate the predicted
-outcomes.
-
 \hypertarget{outcome-shares-based-on-predicted-probabilities}{%
@@
 probabilities}\label{outcome-shares-based-on-predicted-probabilities}}
+
+In this, and the following section, \emph{backcasting} refers to the
+prediction of probabilities and classes of outcomes using the 2016
+dataset. When conducting backcasting, the dataset is preprocessed in
+identical manner as the 2017 dataset. In addition, the variables used in
+backcasting match exactly those in the models. This means that some
+variables were dropped when they were present in the 2016 dataset but
+not in the models. This tended to happen in the case of relatively rare
+outcomes (e.g., in 2016, there was at least one collision between a
+heavy vehicle and a light duty vehicle in a school crossing zone; no
+such event was observed in 2017).
 
@@
 the \(WAPE\), Model 2 gives marginally better results than Model 4. It
-is interesting to see that the two Ensemble models have lower \(APE\)
+is interesting to see that the four ensemble models have lower \(APE\)
 values across the board in the nowcasting exercise, and much better
-\(WAPE\) than the full sample models. However, when we turn to the
-results of the backcasting (out-of-sample) predictions, these results do
-not hold, as the Average Percentage Errors worsen considerably,
-particularly in the case of Fatality. Excellent in-sample predictions
-but mediocre out-of-sample predictions could constitute evidence of
-overfitting by the Ensemble models. In terms of backcasting, again Model
-4 is marginally better than Model 2, especially due to more accurate
-shares of No Injury.
-
-\begin{landscape}\begin{table}
+\(WAPE\) than the full sample models. However, once we turn to the
+results of the backcasting exercise, these results do not hold, and it
+is possible to see that the Average Percentage Errors of the ensemble
+worsen considerably, particularly in the case of Fatality. The Weighted
+Average Prediction Error of the ensemble models in the backcasting
+exercise is also worse than for any of the full sample models. Excellent
+in-sample predictions but mediocre out-of-sample predictions are often
+evidence of overfitting, as in the case of the ensemble models here.
+
+In terms of backcasting, full sample Model 1 is marginally better than
+full sample Models 2 and 3, and better than full sample Model 4. The
+reason for this is the lower \(APE\) of Model 1 when predicting Injury,
+the most frequent outcome. However, the performance of Model 1 with
+respect to Fatality (the least frequent outcome) is the worst of all
+models. Whereas Model 4 has the best performance predicting Fatality,
+its performance with respect to other classes of outcomes is less
+impressive. Model 3 does better than Model 2 with respect to Injury, but
+performs relatively poorly when backcasting Fatality. Overall, Model 2
+appears to be the most balanced, with good in-sample performance and
+competitive out-of-sample performance that is also balanced with respect
+to the various classes of outcomes.
+
+\begin{table}
 
@@
 \end{table}
-\end{landscape}
 
@@
 
-Words go here.
-
-Verification statistics used are summarized in Table
-\ref{tab:verification-statistics}.
+\(APE\) and \(WAPE\) are summary measures of the performance of models
+at the aggregated level. Aggregate-level predictions (i.e., shares of
+outcomes) are of interest from a population health perspective. In other
+cases, an analyst might be interested in the predicted outcomes at the
+individual level. In this section we examine the frequency of outcomes
+based on predicted classes, using the same two settings as above:
+nowcasting and backcasting.
+
+The individual-level outcomes are examined using a battery of
+verification statistics. Verification statistics are widely used in the
+evaluation of predective approaches were the outcomes are categorical,
+and are often based on the analysis of \emph{confusion matrices} (e.g.,
+Provost and Kohavi, 1998; Beguería, 2006). Confusion matrices are
+cross-tabulations of \emph{observed} and \emph{predicted} classes. In a
+two-by-two confusion matrix there are four possible combinations of
+observed to predicted classes: hits, misses, false alarms, and correct
+non-events, as shown in Table \ref{tab:confusion-matrix}. When the
+outcome has more than two classes, the confusion matrix is converted to
+a two-by-two table to calculate verification statistics.
+
+\begin{table}[!h]
+
+\caption{\label{tab:example-confusion-matrix}\label{tab:confusion-matrix}Example of a two-by-two confusion matrix}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{lccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Observed} & \multicolumn{1}{c}{ } \\
+\cmidrule(l{3pt}r{3pt}){2-3}
+Predicted & Yes & No & Marginal Total\\
+\midrule
+\rowcolor{gray!6}  Yes & Hit & False Alarm & Predicted Yes\\
+No & Miss & Correct Non-event & Predicted No\\
+\rowcolor{gray!6}  Marginal Total & Observed Yes & Observed No & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The statistics used in our assessment are summarized in Table
+\ref{tab:verification-statistics}, including brief descriptions of their
+interpretation. The statistics evaluate the performance of models from
+different perspectives. Take as examples Percent Correct (\(PC\)) is the
+sum of hits and correct rejections divided by the number of cases; Bias
+(\(B\)), which measures whether a class has been over- or
+under-predicted; and the probability of detection (\(POD\)) is the
+fraction of hits relative to the total observed for the corresponding
+class. The results of calculating the battery of verification statistics
+are shown in Table \ref{tab:nowcast-outcomes-assessment} for the
+nowcasting exercise, and Table \ref{tab:backcast-outcomes-assessment}
+for the backcasting exercise.
 
@@
 
-We next evaluate the outcomes of the nowcast using an array of
-verification statistics. See Table
-\ref{tab:nowcast-outcomes-assessment}. We next evaluate the outcomes of
-the nowcast using an array of verification statistics. See Table
-\ref{tab:backcast-outcomes-assessment}.
+\hypertarget{nowcasting}{%
+\subsubsection{Nowcasting}\label{nowcasting}}
+
+First, some general remarks regarding the results are in order. It is
+clear that no model performs consistently well from every perspective.
+Recalling Box's maxim, all models are wrong - in this case it just so
+happens that the models are wrong in subtly different ways.
+
+The model that performs best in terms of Percent Correct and is Model 2,
+followed by Model 4. The worst performer from this perspective is Model
+1, with a \(PC\) score several percentage points below the top models.
+The second score is Percent Correct by Class (\(PC_c\)). This score is
+calculated individually for each outcome class. Model 2, again, has the
+best performance for outcomes No Injury and Injury, and the second best
+score for Fatality. Model 4 has the best score for Fatality, and is
+second best for No Injury and Injury. Model 1 (full sample) has worst
+scores for No Injury and Injury whereas its ensemble version has the
+worst score for Fatality. It is important to note that Percent Correct
+and Percent Correct by Class are heavily influenced by the most common
+category. This can be appreciated in the scores, particularly for
+Fatality, which has generally high values of the score, despite the fact
+that the number of hits in that class are relatively low; the high
+values in this case are due to the high ocurrence of correct rejections
+elsewhere in the table.
+
+Bias, in contrast, measures for each outcome class, the proportion of
+total predictions by category (e.g., hits as well as false alarms)
+relative to the total observed for that class. Overall, predictions can
+have low bias (values closer to 1) but still do poorly in terms of hits.
+The models with the best performance in terms of Bias are Model 1 for No
+Injury and Injury, and Model 3 (ensemble) for Fatality. Model 4 is the
+second best performer for No Injury and Injury, and Model 4 (ensemble)
+is second best performer for Fatality. Model 1 (ensemble) has the worst
+bias for No Injury and Injury, whereas Model 2 has the worst bias for
+Fatality.
+
+Critical Success Index (\(CSI\)) is a verification statistic that
+assumes that the number of correct non-events is inconsequential for
+determining forecasting skill. For this reason, the statistic is
+calculated as the proportion of hits relative to the sum of hits plus
+false alarms plus misses. No model performs uniformly best from this
+perspective. Model 1 has the best \(CSI\) for No Injury, Model 2
+(ensemble) has the best score for Injury, and Model 4 (ensemble) the
+best score for Fatality. Model 2 has the worst score for Fatality - this
+indicates that Model 4 is not particularly skilled at predicting
+Fatalities correctly, given the frequency with which it gives false
+alarms for this class, or misses it.
+
+The next statistic is Probability of False Detection (\(F\)). This is
+the proportion of false alarms relative to the total observed no events.
+This statistic measure the frequency with which the model incorrectly
+predicts an event, but not when it incorrectly misses it. The
 
@@
 \fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lrrrrrrrrrrrrr}
+\begin{tabular}[t]{lrrrllllllllll}
 \toprule
-\multicolumn{1}{c}{Observed} & \multicolumn{3}{c}{Predicted Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
+\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 50652 & 22503 & 150 &  & 69.0677 & 0.9293 & 0.4988 & 0.2646 & 0.6421 & 0.3090 &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62121 & 797 &  & 68.6446 & 1.0765 & 0.5463 & 0.3636 & 0.7336 & 0.3185 &  &  & \\
-
-Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.5523} & 99.3921 & 0.0589 & 0.0030 & 0.0003 & 0.0032 & 0.9464 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3725} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3696} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1902}\\
+\hspace{1em}No Injury & 50652 & 22503 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62121 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
+
+Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.55}} & \textcolor{black}{99.39} & \textcolor{black}{0.0589} & \textcolor{black}{0.003} & \textcolor{red}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9464} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3725}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3696}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1902}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51531 & 17137 & 85 &  & 72.9033 & 0.8715 & 0.5362 & 0.2011 & 0.6532 & 0.2505 &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67514 & 864 &  & 72.4152 & 1.1306 & 0.5980 & 0.3535 & 0.7973 & 0.2948 &  &  & \\
-
-Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.3636} & 99.4086 & 0.0263 & 0.0010 & 0.0001 & 0.0011 & 0.9600 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4474} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4429} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2265}\\
+\hspace{1em}No Injury & 51531 & 17137 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
+
+\hspace{1em}Injury & 27355 & 67514 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
+
+Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{72.36}}} & \textcolor{black}{\underline{99.41}} & \textcolor{red}{0.0263} & \textcolor{red}{0.001} & \textcolor{black}{1e-04} & \textcolor{red}{0.0011} & \textcolor{red}{0.96} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4474}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4429}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2265}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51101 & 17296 & 79 &  & 72.5489 & 0.8680 & 0.5309 & 0.2029 & 0.6478 & 0.2537 &  &  & \\
-
-\hspace{1em}Injury & 27785 & 67338 & 868 &  & 72.0444 & 1.1336 & 0.5942 & 0.3589 & 0.7953 & 0.2985 &  &  & \\
-
-Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.9964} & 99.3994 & 0.0463 & 0.0030 & 0.0003 & 0.0032 & 0.9318 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4400} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4356} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2239}\\
+\hspace{1em}No Injury & 51101 & 17296 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.868} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
+
+\hspace{1em}Injury & 27785 & 67338 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7953} & \textcolor{black}{0.2985} &  &  & \\
+
+Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4356}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 51575 & 17318 & 84 &  & 72.8207 & 0.8744 & 0.5356 & 0.2032 & 0.6538 & 0.2523 &  &  & \\
-
-\hspace{1em}Injury & 27311 & 67334 & 863 &  & 72.3332 & 1.1279 & 0.5967 & 0.3529 & 0.7952 & 0.2950 &  &  & \\
-
-Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedleft\arraybackslash 72.2821} & 99.4104 & 0.0274 & 0.0031 & 0.0001 & 0.0032 & 0.8846 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4458} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4414} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2268}\\
+\hspace{1em}No Injury & 51575 & 17318 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
+
+\hspace{1em}Injury & 27311 & 67334 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.1279}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
+
+Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{72.28}}} & \textcolor{black}{\textbf{99.41}} & \textcolor{black}{0.0274} & \textcolor{black}{0.0031} & \textcolor{black}{1e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.8846} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4458}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4414}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2268}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 34664 & 16434 & 63 &  & 69.8818 & 0.8197 & 0.4393 & 0.1952 & 0.5554 & 0.3225 &  &  & \\
-
-\hspace{1em}Injury & 27749 & 67120 & 829 &  & 69.3536 & 1.1452 & 0.5985 & 0.4512 & 0.8032 & 0.2986 &  &  & \\
-
-Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.3107} & 99.3860 & 0.0526 & 0.0414 & 0.0001 & 0.0419 & 0.2041 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3626} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3521} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2010}\\
+\hspace{1em}No Injury & 34664 & 16434 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1952} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\
+
+\hspace{1em}Injury & 27749 & 67120 & 829 &  & \textcolor{black}{69.35} & \textcolor{red}{1.1452} & \textcolor{black}{0.5985} & \textcolor{red}{0.4512} & \textcolor{black}{0.8032} & \textcolor{black}{0.2986} &  &  & \\
+
+Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.31}} & \textcolor{red}{99.39} & \textcolor{black}{0.0526} & \textcolor{black}{0.0414} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0419} & \textcolor{black}{\underline{0.2041}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3626}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3521}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.201}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35443 & 16145 & 60 &  & 70.6151 & 0.8275 & 0.4508 & 0.1917 & 0.5678 & 0.3138 &  &  & \\
-
-\hspace{1em}Injury & 26974 & 67437 & 829 &  & 70.0830 & 1.1393 & 0.6054 & 0.4389 & 0.8067 & 0.2919 &  &  & \\
-
-Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0421} & 99.3862 & 0.0591 & 0.0445 & 0.0001 & 0.0451 & 0.2364 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3784} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3678} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2106}\\
-\cmidrule{1-14}
+\hspace{1em}No Injury & 35443 & 16145 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
+
+\hspace{1em}Injury & 26974 & 67437 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
+
+Fatality & 0 & 13 & 42 &  & \textcolor{black}{99.39} & \textcolor{black}{0.0591} & \textcolor{black}{0.0445} & \textcolor{black}{1e-04} & \textcolor{black}{0.0451} & \textcolor{black}{0.2364} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3784}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3678}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2106}}\\
+
 \addlinespace[0.3em]
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 35498 & 16204 & 60 &  & 70.6159 & 0.8294 & 0.4512 & 0.1924 & 0.5688 & 0.3142 &  &  & \\
-
-\hspace{1em}Injury & 26913 & 67379 & 828 &  & 70.0844 & 1.1379 & 0.6052 & 0.4379 & 0.8060 & 0.2916 &  &  & \\
-
-Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0436} & 99.3868 & 0.0622 & 0.0476 & 0.0001 & 0.0482 & 0.2241 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3786} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3681} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2123}\\
+\hspace{1em}No Injury & 35498 & 16204 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
+
+\hspace{1em}Injury & 26913 & 67379 & 828 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
+
+Fatality & 0 & 13 & 45 & \multirow{-6}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 35553 & 16297 & 59 &  & 70.5898 & 0.8318 & 0.4514 & 0.1935 & 0.5697 & 0.3151 &  &  & \\
-
-\hspace{1em}Injury & 26852 & 67271 & 827 &  & 70.0602 & 1.1361 & 0.6046 & 0.4370 & 0.8049 & 0.2915 &  &  & \\
-
-Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedleft\arraybackslash 70.0201} & 99.3901 & 0.0601 & 0.0488 & 0.0001 & 0.0494 & 0.1786 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3783} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3679} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2127}\\
+\hspace{1em}No Injury & 35553 & 16297 & 59 &  & \textcolor{black}{70.59} & \textcolor{black}{0.8318} & \textcolor{black}{0.4514} & \textcolor{black}{0.1935} & \textcolor{black}{0.5697} & \textcolor{black}{0.3151} &  &  & \\
+
+\hspace{1em}Injury & 26852 & 67271 & 827 &  & \textcolor{black}{70.06} & \textcolor{black}{1.1361} & \textcolor{black}{0.6046} & \textcolor{black}{0.437} & \textcolor{black}{0.8049} & \textcolor{black}{\textbf{0.2915}} &  &  & \\
+
+Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.02}} & \textcolor{black}{99.39} & \textcolor{black}{\underline{0.0601}} & \textcolor{black}{\textbf{0.0488}} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{\textbf{0.0494}} & \textcolor{black}{\textbf{0.1786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3783}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3679}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2127}}\\
 \bottomrule
-\multicolumn{14}{l}{\textit{Notes: }}\\
+\multicolumn{14}{l}{\textit{Note: }}\\
+\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
 \multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrrrrrrrrrrrrr}
+\begin{tabular}[t]{lrrrllllllllll}
 \toprule
-\multicolumn{1}{c}{Observed} & \multicolumn{3}{c}{Predicted Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
+\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 61684 & 27447 & 184 &  & 68.5295 & 0.9221 & 0.4955 & 0.2690 & 0.6368 & 0.3094 &  &  & \\
-
-\hspace{1em}Injury & 35171 & 74073 & 915 &  & 68.1231 & 1.0842 & 0.5380 & 0.3683 & 0.7290 & 0.3276 &  &  & \\
-
-Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.0284} & 99.4042 & 0.0902 & 0.0083 & 0.0005 & 0.0090 & 0.9000 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3628} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3604} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1882}\\
+\hspace{1em}No Injury & 61684 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
+
+\hspace{1em}Injury & 35171 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3683} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
+
+Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.03}} & \textcolor{black}{99.4} & \textcolor{black}{0.0902} & \textcolor{black}{0.0083} & \textcolor{black}{5e-04} & \textcolor{black}{0.009} & \textcolor{red}{0.9} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3628}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3604}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1882}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 62735 & 21013 & 106 &  & 72.3190 & 0.8657 & 0.5317 & 0.2056 & 0.6477 & 0.2519 &  &  & \\
-
-\hspace{1em}Injury & 34125 & 80569 & 996 &  & 71.8616 & 1.1386 & 0.5893 & 0.3585 & 0.7930 & 0.3036 &  &  & \\
-
-Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.8085} & 99.4363 & 0.0271 & 0.0062 & 0.0001 & 0.0063 & 0.7667 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4372} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4335} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2241}\\
+\hspace{1em}No Injury & 62735 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5317}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
+
+\hspace{1em}Injury & 34125 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
+
+Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4372}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4335}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 62248 & 21133 & 107 &  & 72.0144 & 0.8619 & 0.5271 & 0.2068 & 0.6427 & 0.2544 &  &  & \\
-
-\hspace{1em}Injury & 34610 & 80433 & 996 &  & 71.5504 & 1.1421 & 0.5862 & 0.3634 & 0.7916 & 0.3068 &  &  & \\
-
-Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.4958} & 99.4268 & 0.0424 & 0.0052 & 0.0002 & 0.0054 & 0.8723 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4310} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4274} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2205}\\
+\hspace{1em}No Injury & 62248 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.8619} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
+
+\hspace{1em}Injury & 34610 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
+
+Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{71.5}} & \textcolor{black}{99.43} & \textcolor{black}{0.0424} & \textcolor{red}{0.0052} & \textcolor{black}{2e-04} & \textcolor{red}{0.0054} & \textcolor{black}{0.8723} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.431}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4274}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2205}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 62788 & 21246 & 102 &  & 72.2309 & 0.8686 & 0.5312 & 0.2078 & 0.6482 & 0.2537 &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80332 & 1000 &  & 71.7679 & 1.1358 & 0.5878 & 0.3580 & 0.7906 & 0.3039 &  &  & \\
-
-Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedleft\arraybackslash 71.7163} & 99.4338 & 0.0316 & 0.0062 & 0.0001 & 0.0063 & 0.8000 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4355} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.4318} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.2233}\\
+\hspace{1em}No Injury & 62788 & 21246 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80332 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5878} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
+
+Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4355}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 42896 & 20230 & 95 &  & 69.2588 & 0.8162 & 0.4387 & 0.2011 & 0.5538 & 0.3215 &  &  & \\
-
-\hspace{1em}Injury & 34546 & 79692 & 962 &  & 68.7306 & 1.1519 & 0.5880 & 0.4522 & 0.7968 & 0.3082 &  &  & \\
-
-Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 68.6690} & 99.3486 & 0.1129 & 0.0127 & 0.0006 & 0.0140 & 0.8760 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3539} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3447} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1831}\\
+\hspace{1em}No Injury & 42896 & 20230 & 95 &  & \textcolor{black}{69.26} & \textcolor{red}{0.8162} & \textcolor{red}{0.4387} & \textcolor{black}{0.2011} & \textcolor{red}{0.5538} & \textcolor{red}{0.3215} &  &  & \\
+
+\hspace{1em}Injury & 34546 & 79692 & 962 &  & \textcolor{black}{68.73} & \textcolor{red}{1.1519} & \textcolor{black}{0.588} & \textcolor{red}{0.4522} & \textcolor{black}{0.7968} & \textcolor{black}{0.3082} &  &  & \\
+
+Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{68.67}} & \textcolor{black}{99.35} & \textcolor{black}{0.1129} & \textcolor{black}{0.0127} & \textcolor{black}{6e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.876} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3539}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3447}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1831}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 43486 & 19937 & 95 &  & 69.7585 & 0.8200 & 0.4461 & 0.1981 & 0.5614 & 0.3154 &  &  & \\
-
-\hspace{1em}Injury & 33953 & 80009 & 961 &  & 69.2271 & 1.1487 & 0.5928 & 0.4446 & 0.7997 & 0.3038 &  &  & \\
-
-Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.1627} & 99.3398 & 0.1289 & 0.0126 & 0.0007 & 0.0140 & 0.8913 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3644} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3551} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1883}\\
+\hspace{1em}No Injury & 43486 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.4461} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
+
+\hspace{1em}Injury & 33953 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
+
+Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3551}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 43526 & 20033 & 92 &  & 69.7291 & 0.8217 & 0.4460 & 0.1990 & 0.5619 & 0.3162 &  &  & \\
-
-\hspace{1em}Injury & 33915 & 79915 & 964 &  & 69.1938 & 1.1474 & 0.5923 & 0.4441 & 0.7988 & 0.3038 &  &  & \\
-
-Fatality & 18 & 102 & 16 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.1322} & 99.3415 & 0.1269 & 0.0134 & 0.0007 & 0.0149 & 0.8824 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3639} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3546} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1885}\\
+\hspace{1em}No Injury & 43526 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8217} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
+
+\hspace{1em}Injury & 33915 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\
+
+Fatality & 18 & 102 & 16 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.13}} & \textcolor{black}{99.34} & \textcolor{black}{0.1269} & \textcolor{black}{\underline{0.0134}} & \textcolor{black}{7e-04} & \textcolor{black}{\underline{0.0149}} & \textcolor{black}{0.8824} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3639}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3546}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1885}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 43561 & 20160 & 94 &  & 69.6718 & 0.8238 & 0.4458 & 0.2003 & 0.5624 & 0.3174 &  &  & \\
-
-\hspace{1em}Injury & 33875 & 79762 & 959 &  & 69.1415 & 1.1456 & 0.5914 & 0.4436 & 0.7974 & 0.3040 &  &  & \\
-
-Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedleft\arraybackslash 69.0748} & 99.3364 & 0.1393 & 0.0141 & 0.0007 & 0.0159 & 0.8859 & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3629} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.3538} & \multirow{-3}{*}{\raggedleft\arraybackslash 0.1886}\\
+\hspace{1em}No Injury & 43561 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5624} & \textcolor{black}{0.3174} &  &  & \\
+
+\hspace{1em}Injury & 33875 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
+
+Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.07}} & \textcolor{red}{99.34} & \textcolor{black}{\textbf{0.1393}} & \textcolor{black}{\textbf{0.0141}} & \textcolor{red}{7e-04} & \textcolor{black}{\textbf{0.0159}} & \textcolor{black}{0.8859} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3629}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3538}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1886}}\\
 \bottomrule
-\multicolumn{14}{l}{\textit{Notes: }}\\
+\multicolumn{14}{l}{\textit{Note: }}\\
+\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
 \multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
@@
 hierarchical model with individual-opponent interactions (Model 4). On
-the one hand, the AIC tended to favor the
+the one hand, the \(AIC\) tended to favor the
 
@@
 
+\leavevmode\hypertarget{ref-Begueria2006validation}{}%
+Beguería, S., 2006. Validation and evaluation of predictive models in
+hazard assessment and risk management. Natural Hazards 37, 315--329.
+doi:\href{https://doi.org/10.1007/s11069-005-5182-6}{10.1007/s11069-005-5182-6}
+
 \leavevmode\hypertarget{ref-Bogue2017modified}{}%
@@
 
+\leavevmode\hypertarget{ref-Mannering2016unobserved}{}%
+Mannering, F., Shankar, V., Bhat, C.R., 2016. Unobserved heterogeneity
+and the statistical analysis of highway accident data. Analytic Methods
+in Accident Research 11, 1--16.
+doi:\href{https://doi.org/10.1016/j.amar.2016.04.001}{10.1016/j.amar.2016.04.001}
+
 \leavevmode\hypertarget{ref-McArthur2014spatial}{}%
@@
 
+\leavevmode\hypertarget{ref-Provost1998glossary}{}%
+Provost, F., Kohavi, R., 1998. Glossary of terms. Journal of Machine
+Learning 30, 271--274.
+
 \leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
@@
 doi:\href{https://doi.org/10.1002/atr.5670410107}{10.1002/atr.5670410107}
+
+\leavevmode\hypertarget{ref-Roorda2010trip}{}%
+Roorda, M.J., Paez, A., Morency, C., Mercado, R., Farber, S., 2010. Trip
+generation of vulnerable populations in three canadian cities: A spatial
+ordered probit approach. Transportation 37, 525--548.
+doi:\href{https://doi.org/10.1007/s11116-010-9263-3}{10.1007/s11116-010-9263-3}
 
@@
 
+\leavevmode\hypertarget{ref-Shamsunnahar2013evaluating}{}%
+Shamsunnahar, Y., Eluru, N., 2013. Evaluating alternate discrete outcome
+frameworks for modeling crash injury severity. Accident Analysis \&
+Prevention 59, 506--521.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
+
 \leavevmode\hypertarget{ref-Shamsunnahar2014examining}{}%
@@
 
-\leavevmode\hypertarget{ref-Shamsunnahar2013evaluating}{}%
-Yasmin, S., Eluru, N., 2013. Evaluating alternate discrete outcome
-frameworks for modeling crash injury severity. Accident Analysis \&
-Prevention 59, 506--521.
-doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
-
 \end{document}

### 2020-01-25 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Completed first full draft of paper, to send to co-authors

@@
    \ead{razavi@mcmaster.ca} 
-      \address[McMaster University]{McMaster Institute for Transportation and Logistics, 1280 Main Street
-West, Hamilton, Ontario, Canada L8S 4K1}
-    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Baton Rouge,
-Louisiana, USA 70803}
+      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
+University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
+    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
+University, Baton Rouge, Louisiana, USA 70803}
 
@@
 
-\hypertarget{general-considerations}{%
-\subsection{General considerations}\label{general-considerations}}
+\hypertarget{sec:general-considerations}{%
+\subsection{General considerations}\label{sec:general-considerations}}
 
@@
 
-\hypertarget{modelling-techniques}{%
-\subsection{Modelling techniques}\label{modelling-techniques}}
+\hypertarget{sec:techniques}{%
+\subsection{Modelling techniques}\label{sec:techniques}}
 
@@
 
-\hypertarget{model-specification-strategies}{%
-\subsection{Model specification
-strategies}\label{model-specification-strategies}}
+\hypertarget{sec:strategies}{%
+\subsection{Model specification strategies}\label{sec:strategies}}
 
@@
 
-\hypertarget{data-for-empirical-assessment}{%
-\subsection{Data for empirical
-assessment}\label{data-for-empirical-assessment}}
+\hypertarget{sec:data}{%
+\subsection{Data for empirical assessment}\label{sec:data}}
 
@@
 
-\hypertarget{model-estimation}{%
-\subsection{Model estimation}\label{model-estimation}}
+\hypertarget{sec:model-estimation}{%
+\subsection{Model estimation}\label{sec:model-estimation}}
 
@@
 
-\hypertarget{goodness-of-fit-of-models}{%
-\subsection{Goodness of fit of models}\label{goodness-of-fit-of-models}}
+\hypertarget{sec:goodness-of-fit}{%
+\subsection{Goodness of fit of models}\label{sec:goodness-of-fit}}
 
@@
 \label{eq:relative-likelihood}
-e^{(AIC_b - AIC_b)/2}
+e^{(AIC_a - AIC_b)/2}
 \end{equation}
@@
 
-\hypertarget{outcome-shares-based-on-predicted-probabilities}{%
+\hypertarget{sec:outcomes-shares}{%
 \subsection{Outcome shares based on predicted
-probabilities}\label{outcome-shares-based-on-predicted-probabilities}}
+probabilities}\label{sec:outcomes-shares}}
 
@@
 
-\hypertarget{outcome-frequency-based-on-predicted-classes}{%
-\subsection{Outcome frequency based on predicted
-classes}\label{outcome-frequency-based-on-predicted-classes}}
+\hypertarget{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}{%
+\subsection{Outcome frequency based on predicted classes
+\{sec:outcomes-classes\}}\label{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}}
 
@@
 \ref{tab:verification-statistics}, including brief descriptions of their
-interpretation. The statistics evaluate the performance of models from
-different perspectives. Take as examples Percent Correct (\(PC\)) is the
-sum of hits and correct rejections divided by the number of cases; Bias
-(\(B\)), which measures whether a class has been over- or
-under-predicted; and the probability of detection (\(POD\)) is the
-fraction of hits relative to the total observed for the corresponding
-class. The results of calculating the battery of verification statistics
-are shown in Table \ref{tab:nowcast-outcomes-assessment} for the
-nowcasting exercise, and Table \ref{tab:backcast-outcomes-assessment}
-for the backcasting exercise.
+interpretation (more on this below). The statistics evaluate different
+aspects of the performance of a model. Some are concerned with the
+ability of the model to be right. For example, Percent Correct (\(PC\))
+is the sum of hits and correct rejections divided by the number of
+cases. Others are concerned with the ability of the model to match the
+observed outcomes, including Bias (\(B\)), which measures whether a
+class has been over- or under-predicted. Whereas others measure the
+ability of the model to not be wrong, such as the probability of false
+detection (\(F\)) and the False Alarm Ration (\(FAR\)). We discuss the
+results of calculating the battery of verification statistics, first for
+the nowcasting case (are shown in (Table
+\ref{tab:nowcast-outcomes-assessment}) and subsequently for the
+backcasting case (Table \ref{tab:backcast-outcomes-assessment}).
 
@@
 
-\hypertarget{nowcasting}{%
-\subsubsection{Nowcasting}\label{nowcasting}}
-
-First, some general remarks regarding the results are in order. It is
-clear that no model performs consistently well from every perspective.
-Recalling Box's maxim, all models are wrong - in this case it just so
-happens that the models are wrong in subtly different ways.
-
-The model that performs best in terms of Percent Correct and is Model 2,
-followed by Model 4. The worst performer from this perspective is Model
-1, with a \(PC\) score several percentage points below the top models.
-The second score is Percent Correct by Class (\(PC_c\)). This score is
-calculated individually for each outcome class. Model 2, again, has the
-best performance for outcomes No Injury and Injury, and the second best
-score for Fatality. Model 4 has the best score for Fatality, and is
-second best for No Injury and Injury. Model 1 (full sample) has worst
+\hypertarget{nowcasting-verification-statistics}{%
+\subsubsection{Nowcasting: verification
+statistics}\label{nowcasting-verification-statistics}}
+
+At first glance, the results of the verification statistics (Table
+\ref{tab:nowcast-outcomes-assessment}) make it clear that no model under
+comparison is consistently a top performer from every aspect of
+prediction. Recalling Box's aphorism, all models are wrong but some are
+useful - in this case it just so happens that some models are more wrong
+than others in subtly different ways. That said, it is noticeable that
+the worst scores across the board tend to accrue to Model 1 in its full
+sample and ensemble versions. On the other hand, Model 2 (full sample)
+concentrates most of the best scores and second best scores of all the
+models, but also some of the worst scores for Fatality. Model 4, in
+contrast, has most of the second best scores and a few top scores, but
+not a single worst score.
+
+Of all the models, Model 2 (full sample) performs best in terms of
+Percent Correct, followed by Model 4 (full sample). The worst performer
+from this perspective is Model 1 (full sample), with a \(PC\) score
+several percentage points below the top models. The second score is
+Percent Correct by Class (\(PC_c\)). This score is calculated
+individually for each outcome class. Model 2 (full sample), has the best
+performance for outcomes No Injury and Injury, and the second best score
+for Fatality. Model 4 (full sample) has the best score for Fatality, and
+is second best for No Injury and Injury. Model 1 (full sample) has worst
 scores for No Injury and Injury whereas its ensemble version has the
-worst score for Fatality. It is important to note that Percent Correct
-and Percent Correct by Class are heavily influenced by the most common
-category. This can be appreciated in the scores, particularly for
-Fatality, which has generally high values of the score, despite the fact
-that the number of hits in that class are relatively low; the high
-values in this case are due to the high ocurrence of correct rejections
+worst score for Fatality. It is important to note that \(PC\) and
+\(PC_c\) are heavily influenced by the most common category, something
+that can be particularly appreciated in the scores for Fatality. The
+scores for this class of outcome are generally high, despite the fact
+that the number of hits are relatively low; the high values of \(PC_c\)
+in this case are due to the high ocurrence of correct rejections
 elsewhere in the table.
 
-Bias, in contrast, measures for each outcome class, the proportion of
-total predictions by category (e.g., hits as well as false alarms)
-relative to the total observed for that class. Overall, predictions can
-have low bias (values closer to 1) but still do poorly in terms of hits.
-The models with the best performance in terms of Bias are Model 1 for No
-Injury and Injury, and Model 3 (ensemble) for Fatality. Model 4 is the
-second best performer for No Injury and Injury, and Model 4 (ensemble)
-is second best performer for Fatality. Model 1 (ensemble) has the worst
-bias for No Injury and Injury, whereas Model 2 has the worst bias for
-Fatality.
-
-Critical Success Index (\(CSI\)) is a verification statistic that
-assumes that the number of correct non-events is inconsequential for
-determining forecasting skill. For this reason, the statistic is
-calculated as the proportion of hits relative to the sum of hits plus
-false alarms plus misses. No model performs uniformly best from this
-perspective. Model 1 has the best \(CSI\) for No Injury, Model 2
-(ensemble) has the best score for Injury, and Model 4 (ensemble) the
-best score for Fatality. Model 2 has the worst score for Fatality - this
-indicates that Model 4 is not particularly skilled at predicting
-Fatalities correctly, given the frequency with which it gives false
-alarms for this class, or misses it.
+Bias (\(B\)) measures for each outcome class the proportion of total
+predictions by class (e.g., hits as well as false alarms) relative to
+the total number of cases observed for that class. For this reason, it
+is possible for predictions to have low bias (values closer to 1) but
+still do poorly in terms of hits. The models with the best performance
+in terms of \(B\) are Model 1 (full sample) for No Injury and Injury,
+and Model 3 (ensemble) for Fatality. Model 4 (full sample) is the second
+best performer for No Injury and Injury, and Model 4 (ensemble) is
+second best performer for Fatality. Model 1 (ensemble) has the worst
+bias for No Injury and Injury, whereas Model 2 (full sample) has the
+worst bias for Fatality.
+
+Critical Success Index (\(CSI\)) evaluates forecasting skill while
+assuming that the number of correct non-events is inconsequential skill.
+Accordingly, the statistic is calculated as the proportion of hits
+relative to the sum of hits plus false alarms plus misses. No model
+performs uniformly best from this perspective. Model 2 (full sample) has
+the best \(CSI\) for No Injury, Model 2 (ensemble) has the best score
+for Injury, and Model 4 (ensemble) the best score for Fatality. On the
+other hand, Model 1 (ensemble) has the worst score for No Injury, Model
+1 (full sample) the worst score for Injury, and Model 2 the worst score
+for Fatality. These scores indicates that the models are not
+particularly skilled at predicting the corresponding classes correctly,
+given the frequency with which they give false alarms or miss the class.
 
 The next statistic is Probability of False Detection (\(F\)). This is
-the proportion of false alarms relative to the total observed no events.
-This statistic measure the frequency with which the model incorrectly
-predicts an event, but not when it incorrectly misses it. The
-
-\begin{table}[!h]
+the proportion of false alarms relative to the total number of cases
+that the event is not observed. This statistic measure the frequency
+with which the model incorrectly predicts an event, but not when it
+incorrectly misses it. The lowest probability of false detection in the
+case of No Injury is 19.17\% for Model 2 (ensemble), with every other
+model having values lower than 21\%, with the exception of Model 1 (full
+sample) that has a score of 26.46\%. With respect to Injury, the lowest
+probabilities range 35.8\% and 35.29\% and 35.35\% for Models 4 (full
+sample) and 2 (full sample). In contrast, the highest probability of
+false detection for Injury is 45.12\% for Model 1 (ensemble). The scores
+for \(F\) for Fatality are all extremely low as a consequence of the
+very low frequency of this class of outcome in the sample.
+
+The Probability of Detection (\(POD\)) measures the frequency with which
+the model correctly predicts a class, relative to the number of cases of
+that class. As with some other verification statistics, no model is
+consistently a best performer. Model 4 (full sample) has the highest
+probability of detection for No Injury (65.38\%), followed by Model 2
+(65.32\%), whereas the worst probability of detection is by Model 1
+(ensemble) with a score of 55.54\%. In terms of Injury, all models have
+\(POD\) higher than 79\%, and the highest score is 80.67\% for Model 2
+(ensemble). The exception is Model 1 (full sample), which has a
+considerably lower \(POD\) of Injury with a score of 73.36\%. Lastly, in
+terms of Fatalities, all models have very low probabilities of
+detection, ranging from a high of 4.94\% in the case of Model 4
+(ensemble) to a worst score of 0.11\% in the case of Model 2 (full
+sample).
+
+The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
+that were false alarms, which means that lower scores are better. Model
+2 (full sample) has the best \(FAR\) statistic for No Injury, as only
+25.05\% of predictions for this class are false alarms. The next best
+score is by Model 4 (full sample), with only 25.23\% of No Injury
+predictions being false alarms. The worst performance in this class is
+by Model 1 (ensemble), which produces almost a third of false alarms in
+its predictions of No Injury. In the case of Injury, the False Alarm
+Ratio ranges from a low of 29.15\% by Model (ensemble), with every other
+model having scores lower than 30\% except Model 1 (full sample), that
+gives almost 32\% of false alarms. In terms of Fatality, the lowest
+\(FAR\) is also for Model 4 (ensemble) with only 17.86\% of false
+alarms, whereas the worst performance is by Model 2 (full sample), which
+produces over 95\% of false alarms.
+
+The next three verification statistics are skill scores that
+simultaneously consider different aspects of prediction, and are
+therefore overall indicators of prediction skill. Heidke's Skill Score
+(\(HSS\)) is the fraction of correct predictions above those that could
+be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
+Probability of Detection (\(POD\)) of a model and its Probability of
+False Detection (\(F\)) to measure the skill of a model to discriminate
+the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
+the model's skill predicting the correct classes that tends to reward
+correct forecasts of the least frequent class. These scores help to
+remove some of the ambiguity regarding the overall performance of a
+model. In this way, we know that Model 2 (full sample) does not do
+particularly well with the class Fatality - however, of all models, it
+tends to have the best overall performance. Its \(HSS\), for example,
+suggests that it achieves 44.74\% of correct predictions after removing
+correct predictions attributable to chance. In contrast, the lowest
+score is for Model 1 (ensemble), which only achieves 36.26\% correct
+predictions after removing those attributable to chance. Model 2 (full
+sample) also has the highest \(PSS\) and the second highest \(GS\).
+Model 4 (full sample) has the highest \(GS\) and the second highest
+\(HSS\) and \(PSS\).
+
+\begin{table}
 
@@
 
-\begin{table}[!h]
-
-\caption{\label{tab:backcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
+\hypertarget{backcasting-verification-statistics}{%
+\subsubsection{Backcasting: verification
+statistics}\label{backcasting-verification-statistics}}
+
+Table \ref{tab:backcast-outcomes-assessment} presents the results of the
+verification exercise for the case of our out-of-sample predictions
+(i.e., backasting). Qualitatively, the results are similar to those of
+the nowcasting experiments, but with a somewhat weaker performance of
+the ensemble models. This, again, supports the idea that these models
+might be overfitting the process, as discussed in reference to the
+aggregate forecasts (see Section \ref{sec:outcomes-shares}). Models 2
+(full sample) and 4 (full sample) are again identified as the best
+overall performers, and particularly Model 2 (full sample) performs
+somewhat more adroitly with respect to Fatality in backcasting than it
+did in nowcasting.
+
+\begin{table}
+
+\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
 \centering
@@
 
-Here I plan to discuss the applicability of the modelling strategy to
-advanced modelling techniques (partial proportional odds, heterogeneity,
-hierarchical models, etc.)
+As discussed in Section \ref{sec:techniques}, there is a rich selection
+of modelling approaches that are applicable to crash severity analysis.
+Based on the literature, we limited our empirical assessment of
+modelling strategies to only one model, namely the ordinal probit. On
+the other hand, since the modelling strategies discussed here all relate
+to the specification of the latent function and data subsetting, it is a
+relatively simple matter to extend them to other modelling approaches.
+For example, take Expression
+\ref{eq:latent-function-with-opponent-variables} and add a random
+component \(\mu_{k}\) as follows:
+
+\begin{equation}
+\label{eq:latent-function-with-opponent-variables-and-random-component}
+y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \mu_{k} + \epsilon_{itk}
+\end{equation}
+
+The addition of the random component in this fashion would help to
+capture, when appropriate, unobserved heterogeneity at the level of the
+crash (this is similar to the random intercepts approach in multi-level
+modelling; also see Mannering et al., 2016). As a second example, take
+Expressions \ref{eq:latent-function-with-hierarchical-traffic-unit} to
+\ref{eq:hierarchical-traffic-unit-coefficients} and add a random
+component to a hierarchical coefficient, to obtain:
+
+\begin{equation}
+\label{eq:hierarchical-traffic-unit-coefficients-with-random-component}
+\begin{array}{rcl}\
+\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL} + \mu_{mk}\big)u_{tkm}\\ 
+&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm} + \mu_{mk}u_{tkm}
+\end{array}
+\end{equation}
+
+This is similar to the random slopes strategy in multi-level modelling.
+
+We do not report results regarding other modelling strategies. On the
+one hand, more sophisticated modelling frameworks are generally capable
+of improving the performance of a model. On the other hand, there are
+well-known challenges in the estimation of more sophisticated models
+(see Lenguerrand et al., 2006, p. 47, for a discussion of convergence
+issues in models with mixed effects; Mannering et al., 2016, p. 13, for
+some considerations regarding the complexity and cost of estimating more
+complex models; and Bogue et al., 2017, p. 27, on computational demands
+of models with random components). The additional cost and complexity of
+more sophisticated modelling approaches would, in our view, have greatly
+complicated our empirical assessment, particularly considering the large
+size of the sample involved in this research (a dataset with over
+164,000 records in the case of the full sample models). That said, we
+experimented with a model with random components using monthly subsets
+of data to find that, indeed, estimation takes considerably longer, is
+more demanding in terms of fixing potential estimation quirks, and in
+the end resulted in variance components that could not be reliably
+estimated as different from zero (results can be consulted in the source
+R Notebook). For this reason, we choose to leave the application of more
+sophisticated models as a matter for future research.
 
@@
 
-Different modelling strategies can be used to model complex
-hierarchical, multievent outcomes such as the severity of injuries
-following a collision. The objective of this paper was to assess the
-performance of different strategies to model opponent effects in
-two-vehicle crashes. In broad terms, three strategies were considered:
-1) incorporating opponent-level variables in the model; 2) single-
-versus multi-level model specifications; and 3) sample subsetting and
-estimation of separate models for different types of individual-opponent
-interactions.
+The study of crash severity is an important component of accident
+research, as seen from a large and vibrant literature and numerous
+applications. Part of this literature covers different modelling
+strategies that can be used to model complex hierarchical, multievent
+outcomes such as the severity of injuries following a collision. In this
+paper, our objective has been to assess the performance of different
+strategies to model opponent effects in two-vehicle crashes. In broad
+terms, three strategies were considered: 1) incorporating opponent-level
+variables in the model; 2) single- versus multi-level model
+specifications; and 3) sample subsetting and estimation of separate
+models for different types of individual-opponent interactions. The
+empirical evaluation was based on data from Canada's National Crash
+Database and the application of ordered probit models. A suite of models
+that implemented the various strategies considered was estimated using
+data from 2017. We then assessed the performance of the models using one
+information criterion (AIC). Furthermore, the predictive performance of
+the models was assessed in terms of both nowcasting (in-sample
+predictions) and backcasting (out-of-sample predictions), the latter
+using data from 2016.
 
 The results of the empirical assessment strongly suggest that
-incorporating opponent effects can greatly improve the fit and
-predictive performance of a model. There is some evidence that
-subsetting the sample can improve the results in some isolated
-situations (e.g., when modelling the severity of crashes involving
-active travellers or motorcyclists), possibly at the risk of
-overfitting. In this paper we did not compare individual models in our
-Ensemble approach, but we suggest that this is an avenue for future
-research.
-
-In terms of the full sample models, the evidence was not conclusive in
-favor of a single-level model with opponent variables (Model 2), or a
-hierarchical model with individual-opponent interactions (Model 4). On
-the one hand, the \(AIC\) tended to favor the
+incorporating opponent effects can greatly improve the goodness-of-fit
+and predictive performance of a model. Two modelling strategies appear
+to outperform the rest: a relatively simple single-level modelling
+approach that incorporates opponent effects, and a hierarchical
+modelling approach with nested opponent effects. There was some evidence
+that subsetting the sample can improve the results in some cases s(e.g.,
+when modelling the severity of crashes involving active travellers or
+motorcyclists), but possibly at the risk of overfitting the process and
+leading to worse out-of-sample performance. In this paper we did not
+compare individual models in our ensemble approach, but we suggest that
+this is an avenue for future research.
+
+The results of this research should be informative to analysts
+interested in crashes involving two parties, since it provides some
+useful guidelines regarding the specification of opponent effects.
+
+The analysis also opens up a few avenues for future research. First, for
+reasons discussed in Section \{sec:further-considerations\}, we did not
+consider more sophisticated modelling approaches, such as models with
+random components, partial proportional odds, ranked ordered models, or
+multinomial models, to mention just a few possibilities. Secondly, we
+only considered the performance of the models when making predictions
+for the full sample. That is, the submodels in the ensembles were not
+compared in detail, just their aggregate results when predicting the
+full sample. However, the goodness-of-fit was not uniformly better for
+any one modelling strategy when the data were subset, and it is possible
+that individual models perform better for a certain subset than
+competitors that are part of a better ensemble, overall. For this
+reason, we suggest that additional work with ensemble approaches is
+warranted. Finally, it is clear that the models do not generally do well
+when predicting the least frequent class of outcome, namely Fatality. It
+would be worthwhile to further investigate approaches for so-called
+imbalanced learning, a task where Tarrao et al.~(2014) have already made
+some headway.
+
+Finally, as an aside, this paper is, to the best of our knowledge, the
+first example of reproducible research in crash severity analysis. By
+providing the data \emph{and} code for the analysis, it is our hope that
+this will allow other researchers to easily verify the results, and to
+extend them. A common practice in the machine learning community is to
+use canonical dataset to demonstrate the performance of new techniques.
+Sharing code and data has remained relatively rare in transportation
+research, and we would like to suggest that the datasets used in this
+research could constitute one such canonical dataset for future
+methodological developments.
+
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+To be added in the final version of the paper.
 
@@
 
+\leavevmode\hypertarget{ref-Lenguerrand2006modelling}{}%
+Lenguerrand, E., Martin, J.L., Laumon, B., 2006. Modelling the
+hierarchical structure of road crash data - application to severity
+analysis. Accident Analysis and Prevention 38, 43--53.
+doi:\href{https://doi.org/10.1016/j.aap.2005.06.021}{10.1016/j.aap.2005.06.021}
+
 \leavevmode\hypertarget{ref-Li2017interplay}{}%

### 2020-01-25 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.Rmd
Anonymized the title page of the manuscript

@@
     email: paezha@mcmaster.ca
-    affiliation: 
+    affiliation: McMaster University
     footnote: Corresponding Author
@@
   - code: McMaster University
-    address: McMaster Institute for Transportation and Logistics, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1
+    address: McMaster Institute for Transportation and Logistics, McMaster University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1
   - code: Louisiana State University
-    address: Department of Civil and Environmental Engineering, Baton Rouge, Louisiana, USA 70803
+    address: Department of Civil and Environmental Engineering, Louisiana State University, Baton Rouge, Louisiana, USA 70803
 abstract: |
@@
 One stream of research focus on modelling the severity of crashes.
-Crash severity research is useful to clarify the way different factors can influence the outcome of the event.
-The objective of this paper is to assess different strategies to model the interactions between participants in a crash, in the context of crashes involving two parties.
+Crash severity research is useful to clarify the way different factors can influence the outcome of an event.
+The objective of this paper is to assess different strategies to model the interactions between participants in a crash in the context of crashes involving two parties.
 Towards this objective, a series of models are estimated using data from Canada's National Collision Database.
@@
 Modelling strategies include different ways of introducing the covariates (e.g., in a single-level or multi-level form), as well as by subsetting the dataset.
-The models are assessed using predicted shares and outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis.
-On the other hand, the study suggests that hierarchical (i.e., multi-level) specifications and subsetting do not perform necessarily better than a relatively simple single-level model with opponent effects.
+The models are assessed using predicted shares and classes of outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis.
+The study also suggests that hierarchical (i.e., multi-level) specifications and subsetting do not necessarily perform better than a relatively simple single-level model with opponent effects.
+The results of this study provide insights regarding the performace of different modelling strategies, and should be informative to researchers working with crash severity models.
 
@@
    - \usepackage[margin=1in]{geometry}
+   - \usepackage{lineno}
+   - \linenumbers
 ---

### 2020-01-25 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
Anonymized the title page of the manuscript

@@
 severity analysis}
-    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
-   \ead{paezha@mcmaster.ca} 
-    \author[Louisiana State University]{Hany Hassan}
-   \ead{hassan1@lsu.edu} 
-    \author[McMaster University]{Mark Ferguson}
-   \ead{fergumr@mcmaster.ca} 
-    \author[McMaster University]{Saiedeh Razavi}
-   \ead{razavi@mcmaster.ca} 
-      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
-University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
-    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
-University, Baton Rouge, Louisiana, USA 70803}
+    \author[Some University]{Author 1\corref{Corresponding Author}}
+   \ead{author1@address.edu} 
+    \author[Another University]{Author 2}
+   \ead{author2@address.edu} 
+      \address[Some University]{Address}
+    \address[Another University]{Address}
 

### 2020-02-17 Interactions-Crash-Severity-v1/Cover-Letter-v1.Rmd
First version for submission, after comments from HH and MF

---
title: " "
author: "Antonio Paez"
date: "2/17/2020"
output: pdf_document
---

Dear Editor:

Please find attached our jointly authored submission entitled "An empirical assessment of strategies to model opponent effects in crash severity analysis". We wish to submit this paper for publication consideration in Analytical Methods in Accident Research.

Modelling opponent effects in crash severity analysis is nothing new. However, of the various strategies often employed to this end, there is no systematic evaluation in the literature about their relative performance. Therefore, our objective with this paper is to empirically evaluate the following modelling strategies:

1. Inclusion of opponent-level effects.
2. Data subsetting.
3. Use of hierarchical variable specification.

We conduct an extensive evaluation based on analysis of the latest edition of Canada's National Crash Database (2017), and also use the 2016 edition for validation purposes in extensive _backcasting_ experiments.

The contributions of this research are as follows:

1. We systematize the analysis of opponent effects in crash severity models. Although every strategy discussed has been used in past research, we organize it in a way that clarifies its operation.

2. We present a data management workflow to prepare a dataset to implement analysis of opponent effects.

3. We provide evidence of the performance of different modelling strategies. In particular, the importance of considering opponent-level effects and the suitability of single-level models.

4. Last, but not least, we present our paper as an example of reproducible research: all data and code are publicly available from the beginning of the peer-review process, and will be available upon publication.

We anticipate that the findings reported here will be useful to guide future research on crash severity analysis. And by making it an early example of reproducible research, we expect that the impact of this paper will extend beyond opponent effects, since other researchers will be able, if so they wish, to take the data and our results as a starting point for their own accident research.

It is our hope that you will find the topic of the research to be a good fit to Analytical Methods in Accident Research, and that your expert reviewers will be able to provide valuable feedback to improve our research.

Looking forward to hearing back from you in due course, we remain sincerely yours.

The Authors

### 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.Rmd
First version for submission, after comments from HH and MF

@@
   Road accidents impose an important burden on health and the economy.
-Numerous efforts to understand the factors that affect road collisions have been undertaken.
-One stream of research focus on modelling the severity of crashes.
-Crash severity research is useful to clarify the way different factors can influence the outcome of an event.
-The objective of this paper is to assess different strategies to model the interactions between participants in a crash in the context of crashes involving two parties.
-Towards this objective, a series of models are estimated using data from Canada's National Collision Database.
+Numerous efforts have been undertaken to understand the factors that affect road collisions.
+One stream of research is on modelling the severity of crashes.
+In this literature several strategies have been proposed to deal with the way the different parties in a crash interact to influence the severity of the individual outcomes, including the use of opponent variables, data subsetting, and estimating models with hierarchical components.
+However, no systematic assessment has been conducted of the performance of these strategies.
+The objective of this paper is to empirically evaluate these modelling strategies in the context of crashes involving two parties.
+To this end, a series of models are estimated using data from Canada's National Collision Database.
 Three levels of crash severity (no injury/injury/fatality) are analyzed using ordered logit models and covariates for the participants in the crash and the conditions of the crash.
-Modelling strategies include different ways of introducing the covariates (e.g., in a single-level or multi-level form), as well as by subsetting the dataset.
 The models are assessed using predicted shares and classes of outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis.
 The study also suggests that hierarchical (i.e., multi-level) specifications and subsetting do not necessarily perform better than a relatively simple single-level model with opponent effects.
-The results of this study provide insights regarding the performace of different modelling strategies, and should be informative to researchers working with crash severity models.
+The results of this study provide insights regarding the performance of different modelling strategies, and should be informative to researchers working with crash severity models.
 

### 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
First version for submission, after comments from HH and MF

@@
   Road accidents impose an important burden on health and the economy.
-  Numerous efforts to understand the factors that affect road collisions
-  have been undertaken. One stream of research focus on modelling the
-  severity of crashes. Crash severity research is useful to clarify the
-  way different factors can influence the outcome of an event. The
-  objective of this paper is to assess different strategies to model the
-  interactions between participants in a crash in the context of crashes
-  involving two parties. Towards this objective, a series of models are
-  estimated using data from Canada's National Collision Database. Three
-  levels of crash severity (no injury/injury/fatality) are analyzed using
-  ordered logit models and covariates for the participants in the crash
-  and the conditions of the crash. Modelling strategies include different
-  ways of introducing the covariates (e.g., in a single-level or
-  multi-level form), as well as by subsetting the dataset. The models are
-  assessed using predicted shares and classes of outcomes, and the results
-  highlight the importance of considering opponent effects in crash
-  severity analysis. The study also suggests that hierarchical (i.e.,
-  multi-level) specifications and subsetting do not necessarily perform
-  better than a relatively simple single-level model with opponent
-  effects. The results of this study provide insights regarding the
-  performace of different modelling strategies, and should be informative
-  to researchers working with crash severity models.
+  Numerous efforts have been undertaken to understand the factors that
+  affect road collisions in general, and the severity of crashes in
+  particular. In this literature several strategies have been proposed to
+  model opponent effects, that is, ways to deal with the way the different
+  parties in a crash interact to influence the severity of the individual
+  outcomes. These strategies include the use of opponent variables, data
+  subsetting, and estimating models with hierarchical components. Since,
+  no systematic assessment has been conducted of the performance of these
+  strategies, they appear to be used in an ad-hoc fashion in the
+  literature. The objective of this paper is to empirically evaluate ways
+  to model opponent effects in the context of crashes involving two
+  parties. To this end, a series of models are estimated using data from
+  Canada's National Collision Database. Three levels of crash severity (no
+  injury/injury/fatality) are analyzed using ordered logit models and
+  covariates for the participants in the crash and the conditions of the
+  crash. The models are assessed using predicted shares and classes of
+  outcomes, and the results highlight the importance of considering
+  opponent effects in crash severity analysis. The study also suggests
+  that hierarchical (i.e., multi-level) specifications and subsetting do
+  not necessarily perform better than a relatively simple single-level
+  model with opponent-related factors. The results of this study provide
+  insights regarding the performance of different modelling strategies,
+  and should be informative to researchers in the field of crash severity.
   \end{abstract}
@@
 \section{Introduction}\label{introduction}}
-
-Road safety continues to be a world-wide concern. According to a recent
-report from the World Health Organization (2019), road accidents are the
-8th leading cause of death for all ages, and the number one cause of
-death for children and young people between the ages of 5 to 29. Of all
-leading causes of death, road accidents are the only cause of death
-unrelated to disease, health disorder, or infection. For this reason,
-road accidents impose a heavy burden on individuals and society as a
-whole. Gobally, the rate of road collision-related deaths per 100,000
-population and 100,000 vehicles have both fallen, even as the number of
-vehicles has grown (World Health Organization, 2019, Figs. 1 and 2).
-These gains, although they are to be celebrated, cannot distract from
-the crushing economic cost of premature death (e.g., Symons et al.,
-2019; Wijnen et al., 2019), not to mention the long-term consequences
-for survivors, measured in sometimes crippling emotional and physical
-pain (e.g., Merlin et al., 2007; Devlin et al., 2019; Pelissier et al.,
-n.d.).
 
@@
 three times higher in low-income countries compared to high-income
-countries; in fact, no reductions in road accident-related fatalities
-were appreciated in low-income countries between 2013 and 2016 (World
-Health Organization, 2019). In the case of high-income countries, where
-substantial gains in road safety have been observed for years, said
-gains have also been unevenly distributed; thus, while fatal crashes
-involving older adults in the United States and Great Britain declined
-between 1997 and 2010 (despite the graying of the population), the trend
-remained stable or increased slightly in Australia in roughly the same
-period (Thompson et al., 2018). There are also systematic differences in
-the impact of road accidents. For example, in a study in the United
-States, Obeng (2011) reported that the impact of covariates of crash
-severity varied substantially in magnitude by gender. More recently,
-Regev et al.~(2018) used adjusted crash risk to find that the risk of
-crashes in Great Britain peaked for people 21 to 29 years of age; on the
-other hand, the risk of fatal injuries for older drivers was constant,
-irrespective of the seriousness of the crash - which highlights the
-perils of accidents at older ages. Other studies have concentrated on
-the consequences of road accidents for the young (e.g., Peek-Asa et al.,
-2010), the old (e.g., Rakotonirainy et al., 2012), as well as
+countries (World Health Organization, 2019). In the case of high-income
+countries, where substantial gains in road safety have been observed for
+years, said gains have also been unevenly distributed; thus, while fatal
+crashes involving older adults in the United States and Great Britain
+declined between 1997 and 2010 (despite the graying of the population),
+the trend remained stable or increased slightly in Australia in roughly
+the same period (Thompson et al., 2018). The impact of road accidents
+also varies considerably. For example, in a study in the United States,
+Obeng (2011) reported that the factors affecting crash severity varied
+substantially in magnitude by gender. More recently, Regev et al.~(2018)
+used adjusted crash risk to find that the risk of fatal injuries for
+older drivers remained high even for minor collisions - which highlights
+the perils of accidents at older ages. Other studies have concentrated
+on the consequences of road accidents for the young (e.g., Peek-Asa et
+al., 2010), the old (e.g., Rakotonirainy et al., 2012), as well as
 pedestrians and cyclists (e.g., Hanson et al., 2013; McArthur et al.,
@@
 
-Given the relevance and cost of this matter, as well as the important
-variations of the impacts among different population segments, numerous
-efforts have been conducted to better understand the factors that affect
-road safety - including the probable consequences of crashes. Along
-these lines, a stream of research in the analysis of road accidents is
-concerned with the severity of crashes. In particular, multivariate
-analysis of crash severity is a useful way to clarify the way various
-factors can affect the outcome of an incident, and to discriminate
-between various levels of injury, from no injury (i.e., property damage
-only), to different degrees of injury up to and including death. This is
-an active area of research, and one where methodological developments
-have aimed at improving the reliability, accuracy, and precision of
-models (e.g., Savolainen et al., 2011; Bogue et al., 2017; Shamsunnahar
-and Eluru, 2013).
-
-This paper aspires to contribute to the literature on crash severity by
-assessing different modelling strategies useful to incorporate opponent
-effects in crash analysis, in the context of incidents involving two
-parties. The importance of these interactions has been recognized in the
-existing literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et
-al., 2017; Tarrao et al., 2014), and a number of different modelling
-strategies have been proposed. In this paper we present a systematic
-assessment of several relevant modelling strategies, ranging from the
-way variables are defined in single-level models, in multi-level models
-(i.e., hierarchical models), as well as using data subsetting
-approaches. For the assessment we use data from Canada's National
-Collision Database, a database that collects all police-reported
-collisions in the country. Using the most recent version of the dataset
-(2017), three levels of crash severity (no injury/injury/fatality) are
-analyzed using ordered logit models and covariates for the participants
-in the crash and the conditions of the crash. For model assessment, we
-conduct an in-sample prediction exercise using the estimation sample
-(i.e., \emph{nowcasting}), and also an out-of-sample prediction exercise
-using the dataset corresponding to 2016 (i.e., \emph{backcasting}). The
-models are assessed using predicted shares and predicted classes of
-outcomes at the individual level, using an extensive array of
-verification statistics. The results highlight the importance of
-considering opponent effects in crash severity analysis to improve the
-goodness-of-fit and predictive performance o. On the other hand, the
-study suggests that hierarchical variable specifications and subsetting
-do not perform necessarily better than a relatively simple single-level
-model with opponent effects.
+Given the relevance and cost of this matter, as well as the unevenness
+of impacts among different population segments, numerous efforts have
+been conducted to better understand the factors that affect road safety
+- including the probable consequences of crashes. Along these lines, a
+stream of research in the analysis of road accidents is concerned with
+the severity of crashes. In particular, multivariate analysis of crash
+severity is a useful way to clarify the way various factors can affect
+the outcome of an incident, and to discriminate between various levels
+of injury. This is an active area of research, and one where
+methodological developments have aimed at improving the reliability,
+accuracy, and precision of models (e.g., Savolainen et al., 2011; Bogue
+et al., 2017; Shamsunnahar and Eluru, 2013). Of interest in the
+literature on crash severity is the way different parties in a crash
+interact to influence the severity of individual outcomes. The
+importance of these interactions has been recognized in the existing
+literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al., 2017;
+Torrao et al., 2014), and a number of different modelling strategies
+have been proposed. These include the use of \emph{data subsetting},
+\emph{opponent variables}, and estimating models with \emph{hierarchical
+components}. A gap in the literature, however, is a systematic
+assessment of the performance of these various strategies.
+
+The objective of this paper is to empirically evaluate different
+strategies to model opponent effects in crash severity, in the context
+of incidents involving two parties. Therefore, we present a systematic
+assessment of several relevant modelling strategies. For the assessment
+we use data from Canada's National Collision Database, a database that
+collects all police-reported collisions in the country. Using the most
+recent version of the data set (2017). Three levels of crash severity
+(no injury/injury/fatality) are analyzed using ordered logit models, and
+covariates for the participants in the crash and the conditions of the
+crash. For model assessment, we conduct an in-sample prediction exercise
+using the estimation sample (i.e., \emph{nowcasting}), and also an
+out-of-sample prediction exercise using the data set corresponding to
+2016 (i.e., \emph{backcasting}). The models are assessed using predicted
+shares and predicted classes of outcomes at the individual level, using
+an extensive array of verification statistics.
 
 The rest of this paper is structured as follows. In Section
-\ref{sec:review-of-methods} we present a concise review of the methods
-used to analyze crash severity, with a particular focus on techniques
-that consider the effect of opponents in a crash. Section
-\ref{sec:application} describes the data requirements, data
-preprocessing, and the modelling strategies, along with the results of
-model estimation. The results of assessing the models and the discussion
-of these results is found in Section \ref{sec:assessment}. We then
-present some additional thoughts about the applicability of this
-approach in Section \ref{sec:further-considerations} before offering
-some concluding remarks in Section \ref{sec:concluding-remarks}.
-
-\hypertarget{sec:review-of-methods}{%
-\section{Methodological approaches in crash severity
-analysis}\label{sec:review-of-methods}}
-
-\hypertarget{sec:general-considerations}{%
-\subsection{General considerations}\label{sec:general-considerations}}
-
-Modelling the outcomes of crashes in terms of the severity of injuries
-to participants has been a preoccupation of transportation researchers,
-planners, auto insurance companies, governments, and the general public
-for decades. One of the earliests studies to investigate the severity of
-injuries conditional on an accident having occurred was by White and
-Clayton (1972). Kim et al.~(1995) later stated that the ``linkages
-between severity of injury and driver characteristics and behaviors have
-not been thoroughly investigated'' (p.~470). Nowadays, there is a
-burgeoning literature on this subject, including methodological
-developments, case studies, and more niche research with a focus on
-particular situations (e.g., crashes at intersections, Mussone et al.,
-2017; crashes in rural roads, Gong and Fan, 2017), and crashes involving
-special population groups (e.g., crashes involving motorcyclists or
-active travelers; see Shaheed et al., 2013; Salon and McIntyre, 2018).
-
-Crash severity is often modelled using models for discrete outcomes. An
-analyst interested in crash severity has at their disposal an ample menu
-of models to choose from, including classification techniques from
-machine learning (e.g., Iranitalab and Khattak, 2017; Chang and Wang,
-2006; Effati et al., 2015; Khan et al., 2015), Poisson models for counts
-(e.g., Ma et al., 2008), unordered logit/probit models (e.g., Tay et
-al., 2011), as well as ordered logit/probit models (e.g., Rifaat and
+\ref{sec:background} we discuss some background matters, and follow this
+with a concise review of the modelling strategies used to analyze crash
+severity \ref{sec:methods}. Section \ref{sec:application} describes the
+data requirements, data preprocessing, and the modelling strategies,
+along with the results of model estimation. The results of assessing the
+models and the discussion of these results is found in Section
+\ref{sec:assessment}. We then present some additional thoughts about the
+applicability of this approach in Section
+\ref{sec:further-considerations} before offering some concluding remarks
+in Section \ref{sec:concluding-remarks}.
+
+\hypertarget{sec:background}{%
+\section{Background}\label{sec:background}}
+
+Modelling the severity of injuries to participants in road crashes has
+been a preoccupation of transportation researchers, planners, auto
+insurance companies, governments, and the general public for decades.
+One of the earliest studies to investigate the severity of injuries
+conditional on an accident having occurred was by White and Clayton
+(1972). Kim et al.~(1995) later stated that the ``linkages between
+severity of injury and driver characteristics and behaviors have not
+been thoroughly investigated'' (p.~470). Nowadays, there is a burgeoning
+literature on this subject, including methodological developments, case
+studies, and more niche research with a focus on particular situations
+(e.g., crashes at intersections, Mussone et al., 2017; crashes in rural
+roads, Gong and Fan, 2017), and crashes involving special population
+groups (e.g., crashes involving motorcyclists or active travelers; see
+Shaheed et al., 2013; Salon and McIntyre, 2018).
+
+Crash severity is often modeled using models for discrete outcomes. An
+analyst interested in crash severity has at their disposal an ample
+repertoire of models to choose from, including classification techniques
+from machine learning (e.g., Iranitalab and Khattak, 2017; Chang and
+Wang, 2006; Effati et al., 2015; Khan et al., 2015), Poisson models for
+counts (e.g., Ma et al., 2008), unordered logit/probit models (e.g., Tay
+et al., 2011), as well as ordered logit/probit models (e.g., Rifaat and
 Chin, 2007), with numerous variants, such as random parameters/mixed
@@
 passengers. They also may differ depending on what type of traffic unit
-they were, for example occupants of a light duty vehicle or a truck,
-motorcyclists, or pedestrians. The multiplicity of roles makes for
-complicated modelling decisions when trying to understand the severity
-of injuries; for example, what is the unit of analysis, the person, the
-traffic unit, or the collision? Not surprisingly, it is possible to find
-examples of studies that adopt different perspectives. A common
-simplifying strategy in model specification is to consider only
-\emph{drivers} and/or only \emph{single-vehicle} crashes (e.g., Kim et
-al., 2013; Gong and Fan, 2017; Lee and Li, 2014; Osman et al., 2018).
-This strategy reduces the dimensions of the event, and it becomes
-possible, for example, to equate the traffic unit to the person for
-modelling purposes.
+they were, for example pedestrians, or operators/passengers of a
+vehicle. The multiplicity of roles makes for complicated modelling
+decisions when trying to understand the severity of injuries; for
+example, what is the unit of analysis, the person, the traffic unit, or
+the collision? Not surprisingly, it is possible to find examples of
+studies that adopt different perspectives. A common simplifying strategy
+in model specification is to consider only \emph{drivers} and/or only
+\emph{single-vehicle} crashes (e.g., Kim et al., 2013; Gong and Fan,
+2017; Lee and Li, 2014; Osman et al., 2018). This strategy reduces the
+dimensions of the event, and it becomes possible, for example, to equate
+the traffic unit to the person for modelling purposes.
 
 The situation becomes more complex when dealing with events that involve
-two traffic units (e.g., Tarrao et al., 2014; Wang et al., 2015) and
+two traffic units (e.g., Torrao et al., 2014; Wang et al., 2015) and
 multi-traffic unit crashes (e.g., Wu et al., 2014; Bogue et al., 2017).
 Different strategies have been developed to study these, more complex
-cases. A number of studies advocate the estimation of separate models
+events. A number of studies advocate the estimation of separate models
 for different participants and/or situations. In this way, Wang and
@@
 separate models for at-fault and not-at-fault drivers. The strategy of
-estimating separate models also relies on subsetting the dataset,
+estimating separate models relies on \emph{subsetting} the data set,
 although it is possible to link the relevant models more tightly by
@@
 
-A related strategy to specify a crash severity model is to organize the
-data in such a way that it is possible to model the influence of the
-attributes of the opponent in a crash. There are numerous examples of
-studies that consider at least some characteristics of the oponents in
-two- or multi-vehicle crashes. For example, Wang and Kockelman (2005)
-considered the type of the opposing vehicle in their model for
-two-vehicle collisions. Similarly, Tarrao et al.~(2014) included in
-their analysis the age, wheelbase, weight, and engine size of the
-opposing vehicle, while Bogue et al.~(2017) used the body type of the
-opposing vehicle. Penmetsa et al.~(2017) and Duddu et al.~(2018) are two
-of the most comprehensive examples of using opponent's information, as
-they used individual attributes of opponents (their physical condition,
-sex, and age), as well as characteristics of the opposing traffic unit
-(the vehicle type of the opponent). The twin strategies of subsetting
-the sample and using the attributes of the opponent are not mutually
-exclusive, but neither are they used consistently together, as a scan of
-the literature reveals.
+A related strategy to specify crash severity models is to organize the
+data in such a way that it is possible to introduce \emph{opponent
+effects}. There are numerous examples of studies that consider at least
+some characteristics of the opponents in two- or multi-vehicle crashes.
+For example, Wang and Kockelman (2005) considered the type of the
+opposing vehicle in their model for two-vehicle collisions. Similarly,
+Torrao et al.~(2014) included in their analysis the age, wheelbase,
+weight, and engine size of the opposing vehicle, while Bogue et
+al.~(2017) used the body type of the opposing vehicle. Penmetsa et
+al.~(2017) and Duddu et al.~(2018) are two of the most comprehensive
+examples of using opponent's information, as they used individual
+attributes of opponents (their physical condition, sex, and age), as
+well as characteristics of the opposing traffic unit (the vehicle type
+of the opponent). The twin strategies of subsetting the sample and using
+the attributes of the opponent are not mutually exclusive, but neither
+are they used consistently together, as a scan of the literature
+reveals. Another strategy is to introduce \emph{hierarchical components}
+in the model, a technique widely used in the hierarchical or multi-level
+modelling literature.
+
+In this paper we consider three general modelling strategies, as
+follows:
+
+\begin{itemize}
+\tightlist
+\item
+  Strategy 1. Introducing opponent-related factors
+\item
+  Strategy 2. Single-level model and multi-level (hierarchical) model
+  specifications
+\item
+  Strategy 3. Full sample and sample subsetting
+\end{itemize}
+
+These are discussed in more detail in the following section.
 
@@
 \midrule
-Person-related & Attributes of participants in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
-Traffic unit-related & Attributes of the traffic unit, e.g., type of traffic unit (car, motorcycle, etc.), maneouver, etc.\\
-Crash-related & Attributes of the crash, e.g., location, weather conditions, light conditions, number of parties, etc.\\
-Road-related & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
-Opponent-related & Attributes of the opponent, e.g., age of opponent, gender of opponent, opponent vehicle type, etc.\\
+Human-factors & Attributes of participants in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
+Traffic unit-factors & Attributes of the traffic unit, e.g., type of traffic unit (pedestrian, car, motorcycle, etc.), maneuver, etc.\\
+Environmental-factors & Attributes of the crash, e.g., location, weather conditions, light conditions, number of parties, etc.\\
+Road-factors & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
+Opponent-related factors & Attributes of the opponent, e.g., age of opponent, gender of opponent, opponent vehicle type, etc.\\
 \bottomrule
@@
 
-\hypertarget{sec:techniques}{%
-\subsection{Modelling techniques}\label{sec:techniques}}
-
-With respect to model structures, Shamsunnahar and Eluru (2013)
-conducted an extensive comparison of models for discrete outcomes and
-found only small differences in the performance of unordered models and
-ordered models; however, ordered models are usually more parsimonious
-since only one latent functions needs to be estimated for all outcomes,
-as opposed to one for each outcome in unordered modelling mechanisms.
-Bogue et al.~(2017) also compared unordered and ordered models in the
-form of the mixed multinomial logit and a modified rank ordered logit,
-respectively, and found that the ordered model performed best. To keep
-the empirical assessment managable we will consider only the ordinal
+\hypertarget{sec:methods}{%
+\section{Methods}\label{sec:methods}}
+
+\hypertarget{choice-of-model}{%
+\subsection{Choice of model}\label{choice-of-model}}
+
+Shamsunnahar and Eluru (2013) conducted an extensive comparison of
+models for discrete outcomes in crash severity analysis, and found only
+small differences in the performance of unordered models and ordered
+models; however, ordered models are usually more parsimonious since only
+one latent functions needs to be estimated for all outcomes, as opposed
+to one for each outcome in unordered models. Bogue et al.~(2017) also
+compared unordered and ordered models in the form of the mixed
+multinomial logit and a modified rank ordered logit, respectively, and
+found that the ordered model performed best. To keep the empirical
+assessment manageable, in this paper we will consider only the ordinal
 logit model, and will comment on potential extensions in Section
@@
 The right-hand side of the expression is split in four parts. The first
-part collects \(l=1,\cdots,L\) individual attributes \(p\) for
-participant \(i\) in traffic unit \(t\) and crash \(k\); these could
-relate to the person (e.g., age, gender, and road user class). The
-second part collects \(m=1,\cdots,M\) attributes \(u\) related to
-traffic unit \(t\) in crash \(k\); these could be items such as maneuver
-or vehicle type. The third part collects \(q=1,\cdots,Q\) attributes
-\(c\) related to the crash \(k\), including crash-related and
-road-related data, such as weather conditions, road alignment, and type
-of surface. Lastly, the fourth element is a random term specific to
-participant \(i\) in traffic unit \(t\) and crash \(k\). The function
-consists of a total of \(Z=L+M+Q\) covariates and associated parameters.
+part collects \(l=1,\cdots,L\) human-factors \(p\) for participant \(i\)
+in traffic unit \(t\) and crash \(k\); these could relate to the person
+(e.g., age, gender, and road user class). The second part collects
+\(m=1,\cdots,M\) attributes \(u\) related to traffic unit \(t\) in crash
+\(k\); these could be items such as maneuver or vehicle type. The third
+part collects \(q=1,\cdots,Q\) attributes \(c\) related to the crash
+\(k\), including environmental-factors and road-factors, such as weather
+conditions, road alignment, and type of surface. Lastly, the fourth
+element is a random term specific to participant \(i\) in traffic unit
+\(t\) and crash \(k\). The function consists of a total of \(Z=L+M+Q\)
+covariates and associated parameters.
+
+For conciseness, in what follows we will abbreviate the function as
+follows:
+
+\begin{equation}
+\label{eq:latent-function-compact}
+y_{itk}^*=\sum_{z=1}^Z\theta_zx_{itkz} + \epsilon_{itk}
+\end{equation}
+
+The latent variable is not observed directly, but it is possible to
+posit a probabilistic relationship with the outcome \(y_{itk}\) (the
+severity of crash \(k\) for participant \(i\) in traffic unit \(t\)).
+Depending on the characteristics of the data and the assumptions made
+about the random component of the latent function different models can
+be obtained. For example, if crash severity is coded as a variable with
+three outcomes (e.g., property damage only/injury/fatal), we can relate
+the latent variable to the outcome as follows:
+
+\begin{equation}
+\label{eq:latent-function-ordered-outcomes}
+y_{itk} = 
+\begin{cases}
+\text{fatality} & \text{if } y_{itk}^*> \mu_2\\
+\text{injury} & \text{if } \mu_1< y_{itk}^*< \mu_2\\
+\text{PDO} & \text{if } y_{itk}^*< \mu_1
+\end{cases}
+\end{equation}
+
+\noindent where \(\mu_1\) and \(\mu_2\) are estimable thresholds. Due to
+the stochastic nature of the latent function, the outcome of the crash
+is not fully determined. However, we can make the following probability
+statements:
+
+\begin{equation}
+\label{eq:probability-ordered-outcomes}
+\begin{array}{rcl}\
+P(y_{itk} = \text{PDO}) &=& 1 - P(y_{itk} = \text{injury}) - P(y_{itk} = \text{fatality})\\ 
+P(y_{itk} = \text{injury}) &=& P(\mu_1 - \sum_{z=1}\theta_zp_{itkz} < \epsilon_{itk} < \mu_2 - \sum_{z=1}\theta_zp_{itkz})\\
+P(y_{itk} = \text{fatality}) &=& P(\epsilon_{itk} < \mu_1 - \sum_{z=1}\theta_zp_{itkz})
+\end{array}
+\end{equation}
+
+If the random terms are assumed to follow the logistic distribution, the
+ordered logit model is obtained; if the normal distribution, then the
+ordered probit model. Estimation methods for these models are very
+well-established (e.g., Maddala, 1986; Train, 2009). There are numerous
+variations of the basic modelling framework above, including
+hierarchical models, bivariate models, multinomial models, and Bayesian
+models, among others (see Savolainen et al., 2011 for a review of
+methods).
+
+\hypertarget{strategy-1-opponent-related-factors}{%
+\subsection{Strategy 1: opponent-related
+factors}\label{strategy-1-opponent-related-factors}}
 
@@
 
-For conciseness, in what follows we will abbreviate the function as
-follows:
-
-\begin{equation}
-\label{eq:latent-function-compact}
-y_{itk}^*=\sum_{z=1}^Z\theta_zx_{itkz} + \epsilon_{itk}
-\end{equation}
-
-The latent variable is not observed directly, but it is possible to
-posit a probabilistic relationship with the outcome \(y_{itk}\) (the
-severity of crash \(k\) for participant \(i\) in traffic unit \(t\)).
-Depending on the characteristics of the data and the assumptions made
-about the random component of the latent function different models can
-be obtained. For example, if crash severity is coded as a binary
-variable (e.g., non-fatal/fatal), we can relate the latent variable to
-the outcome as follows:
-
-\begin{equation}
-\label{eq:latent-function-2-outcomes}
-y_{itk} = 
-\begin{cases}
-\text{fatal} & \text{if } y_{itk}^*>0\\
-\text{non-fatal} & \text{if } y_{itk}^*\leq0
-\end{cases}
-\end{equation}
-
-Due to the stochastic nature of the latent function, the outcome of the
-crash is not fully determined. However, we can make the following
-probability statement:
-
-\begin{equation}
-\label{eq:probability-2-outcomes}
-P(y_{itk} = \text{fatal}) = P(y_{itk}^* > 0)
-\end{equation}
-
-In other words, the probability that individual \(i\) in traffic unit
-\(t\) and crash \(k\) was a fatality equals the probability that the
-latent variable is greater than zero. This implies (see Maddala, 1986,
-p. 22):
-
-\begin{equation}
-\label{eq:probability-2-outcomes-2}
-\begin{array}{rl}\
-P(y_{itk} = \text{fatal}) &= P(\sum_{z=1}\theta_zp_{itkz} + \epsilon_{itk} > 0)\\ 
-&=P(\epsilon_{itk} > -\sum_{z=1}\theta_zp_{itkz})
-\end{array}
-\end{equation}
-
-If the random terms \(\epsilon_{itk}\) are assumed to follow the
-logistic distribution, then the binary logit model is obtained; if they
-are assumed to follow the normal distribution, the binary probit model
-is obtained. More often, though, the outcome is recorded using more
-categories, for example property damage only (PDO)/injury/fatality. A
-similar approach can be adopted, with a latent variable that relates to
-the outcome as follows:
-
-\begin{equation}
-\label{eq:latent-function-ordered-outcomes}
-y_{itk} = 
-\begin{cases}
-\text{fatality} & \text{if } y_{itk}^*> \mu_2\\
-\text{injury} & \text{if } \mu_1< y_{itk}^*< \mu_2\\
-\text{PDO} & \text{if } y_{itk}^*< \mu_1
-\end{cases}
-\end{equation}
-
-\noindent where \(\mu_1\) and \(\mu_2\) are estimable thresholds. In
-this case, the associated probability statements are as follows:
-
-\begin{equation}
-\label{eq:probability-ordered-outcomes}
-\begin{array}{rcl}\
-P(y_{itk} = \text{PDO}) &=& 1 - P(y_{itk} = \text{injury}) - P(y_{itk} = \text{fatality})\\ 
-P(y_{itk} = \text{injury}) &=& P(\mu_1 - \sum_{z=1}\theta_zp_{itkz} < \epsilon_{itk} < \mu_2 - \sum_{z=1}\theta_zp_{itkz})\\
-P(y_{itk} = \text{fatality}) &=& P(\epsilon_{itk} < \mu_1 - \sum_{z=1}\theta_zp_{itkz})
-\end{array}
-\end{equation}
-
-If the random terms are assumed to follow the logistic distribution, the
-ordered logit model is obtained; if the normal distribution, then the
-ordered probit model. Estimation methods for these models are very
-well-established (e.g., Maddala, 1986; Train, 2009). There are numerous
-variations of the basic modelling framework above, including
-hierarchical models, bivariate models, multinomial models, and Bayesian
-models, among others (see Savolainen et al., 2011 for a review of
-methods).
-
-\hypertarget{sec:strategies}{%
-\subsection{Model specification strategies}\label{sec:strategies}}
-
-In this paper we consider three general model specification strategies,
-as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  Strategy 1. Introducing opponent-related variables
-\item
-  Strategy 2. Single-level model and multi-level (hierarchical) model
-  specifications
-\item
-  Strategy 3. Full sample and sample subsetting
-\end{itemize}
-
-Introduction of opponent related-variables was explained in the
-preceding subsection. In this way, a base model that ignores opponent
-effects is given by Equation \ref{eq:latent-function}. Strategy 1, in
-contrast, is Equation \ref{eq:latent-function-with-opponent-variables},
-which includes opponent-related variables. These two equations are also
-examples of single-level models. Next we describe Strategies 2 and 3.
-
 \hypertarget{strategy-2-hierarchical-model-specification}{%
-\subsubsection{Strategy 2: hierarchical model
+\subsection{Strategy 2: hierarchical model
 specification}\label{strategy-2-hierarchical-model-specification}}
@@
 
-Alterted readers will identify this model specification strategy as
+Discerning readers will identify this model specification strategy as
 Casetti's expansion method (Casetti, 1972; Roorda et al., 2010). This is
@@
 \hypertarget{strategy-3-sample-subsetting}{%
-\subsubsection{Strategy 3: sample
+\subsection{Strategy 3: sample
 subsetting}\label{strategy-3-sample-subsetting}}
@@
 In this section we present the setting for the empirical assessment of
-the modelling strategies discussed in Section
-\ref{sec:review-of-methods}, namely matters related to data and model
-estimation.
+the modelling strategies discussed in Section \ref{sec:methods}, namely
+matters related to data and model estimation.
 
@@
 Montella et al., 2013), the NCDB contains information pertaining to the
-collision, the traffic unit(s), and the person(s) involved in a crash,
-as shown in Tables \ref{tab:ncdb-descriptives-collision},
+collision, the traffic unit(s), and the person(s) involved in a crash.
+The definitions of variables in this database are shown in the Appendix
+at the end of this document, in Tables
+\ref{tab:ncdb-descriptives-collision},
 \ref{tab:ncdb-descriptives-vehicle}, and
 \ref{tab:ncdb-descriptives-person}. Notice that, compared to Table
-\ref{tab:variable-categories}, crash-related variables and road-related
-variables are collected under a single variable class, namely
-collision-related, since they are unique for each crash.
+\ref{tab:variable-categories}, environmental-factors variables and
+road-factors variables are collected under a single variable class,
+namely collision-related, since they are unique for each crash.
 
@@
 protocol to do this is described next.
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-collision}\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  C\_CASE & Unique collision identifier & Unique identifier for collisions\\
-C\_YEAR & Year & Last two digits of year.\\
-\rowcolor{gray!6}  C\_MNTH & Month & 14 levels: January - December; unknown; not reported by jurisdiction.\\
-C\_WDAY & Day of week & 9 levels: Monday - Sunday; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_HOUR & Collision hour & 25 levels: hourly intervals; unknown; not reported by jurisdiction.\\
-\addlinespace
-C\_SEV & Collision severity & 4 levels: collision producing at least one fatality; collision producing non-fatal injury; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_VEHS & Number of vehicles involved in collision & Number of vehicles: 1-98 vehicles involved; 99 or more vehicles involved; unknown; not reported by jurisdiction.\\
-C\_CONF & Collision configuration & 21 levels: SINGLE VEHICLE: Hit a moving object (e.g.
-a person or an animal); Hit a stationary object (e.g.
-a tree); Ran off left shoulder; Ran off right shoulder; Rollover on roadway; Any other single vehicle collision configuration; TWO-VEHICLES SAME DIRECTION OF TRAVEL: Rear-end collision; Side swipe; One vehicle passing to the left of the other, or left turn conflict; One vehicle passing to the right of the other, or right turn conflict; Any other two vehicle - same direction of travel configuration; TWO-VEHICLES DIFFERENT DIRECTION OF TRAVEL: Head-on collision; Approaching side-swipe; Left turn across opposing traffic; Right turn, including turning conflicts; Right angle collision; Any other two-vehicle - different direction of travel configuration; TWO-VEHICLES, HIT A PARKED VEHICLE: Hit a parked motor vehicle; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_RCFG & Roadway configuration & 15 levels: Non-intersection; At an intersection of at least two public roadways; Intersection with parking lot entrance/exit, private driveway or laneway; Railroad level crossing; Bridge, overpass, viaduct; Tunnel or underpass; Passing or climbing lane; Ramp; Traffic circle; Express lane of a freeway system; Collector lane of a freeway system; Transfer lane of a freeway system; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-C\_WTHR & Weather condition & 10 levels: Clear and sunny; Overcast, cloudy but no precipitation; Raining; Snowing, not including drifting snow; Freezing rain, sleet, hail; Visibility limitation; Strong wind; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\addlinespace
-\rowcolor{gray!6}  C\_RSUR & Road surface & 12 levels: Dry, normal; Wet; Snow (fresh, loose snow); Slush, wet snow; Icy, packed snow; Debris on road (e.g., sand/gravel/dirt); Muddy; Oil; Flooded; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-C\_RALN & Road alignment & 9 levels: Straight and level; Straight with gradient; Curved and level; Curved with gradient; Top of hill or gradient; Bottom of hill or gradient; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_TRAF & Traffic control & 21 levels: Traffic signals fully operational; Traffic signals in flashing mode; Stop sign; Yield sign; Warning sign; Pedestrian crosswalk; Police officer; School guard, flagman; School crossing; Reduced speed zone; No passing zone sign; Markings on the road; School bus stopped with school bus signal lights flashing; School bus stopped with school bus signal lights not flashing; Railway crossing with signals, or signals and gates; Railway crossing with signs only; Control device not specified; No control present; Choice is other than the preceding values; unknown; not reported by jurisdiction.\\*
-\end{longtable}
-\endgroup{}
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-vehicle}\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  V\_ID & Vehicle sequence number & Number of vehicles: 1-98; Pedestrian sequence number: 99; unknown.\\
-V\_TYPE & Vehicle type & 21 levels: Light Duty Vehicle (Passenger car, Passenger van, Light utility vehicles and light duty pick up trucks); Panel/cargo van (<= 4536 KG GVWR   Panel or window type of van designed primarily for carrying goods); Other trucks and vans (<= 4536 KG GVWR); Unit trucks (> 4536 KG GVWR); Road tractor; School bus; Smaller school bus (< 25 passengers); Urban and Intercity Bus; Motorcycle and moped; Off road vehicles; Bicycle; Purpose-built motorhome; Farm equipment; Construction equipment; Fire engine; Snowmobile; Street car; Data element is not applicable  (e.g.
-dummy vehicle record created for pedestrian); Choice is other than the preceding values; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  V\_YEAR & Vehicle model year & Model year; dummy for pedestrians; unknown; not reported by jurisdiction.\\*
-\end{longtable}
-\endgroup{}
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-person}\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/dataset/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Preprocessed data for analysis available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  P\_ID & Person sequence number & Sequence number: 1-99; Not applicable (dummy for parked vehicles); not reported by jurisdiction.\\
-P\_SEX & Person sex & 5 levels: Male; Female; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_AGE & Person age & Age: less than 1 year; 1-98 years old; 99 years or older; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
-P\_PSN & Person position & Person position: Driver; Passenger front row, center; Passenger front row, right outboard (including motorcycle passenger in sidecar); Passenger second row, left outboard, including motorcycle passenger; Passenger second row, center; Passenger second row, right outboard; Passenger third row, left outboard;...; Position unknown, but the person was definitely an occupant; Sitting on someone’s lap; Outside passenger compartment; Pedestrian; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_ISEV & Medical treatment required & 6 levels: No Injury; Injury; Fatality; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\addlinespace
-P\_SAFE & Safety device used & 11 levels: No safety device used; Safety device used; Helmet worn; Reflective clothing worn; Both helmet and reflective clothing used; Other safety device used; No safety device equipped   (e.g.
-buses); Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_USER & Road user class & 6 levels: Motor Vehicle Driver; Motor Vehicle Passenger; Pedestrian; Bicyclist; Motorcyclist; Not stated/Other/Unknown.\\*
-\end{longtable}
-\endgroup{}
 
@@
 single-occupant vehicles). This is followed by cases with passengers
-(14,120 collisions), pedestrians (5,204 collisions), byciclists (2,238
+(14,120 collisions), pedestrians (5,204 collisions), bicyclists (2,238
 collisions), and motorcyclists (1,016 collisions). The distribution of
@@
 \ref{tab:summary-opponent-information}. The information there shows that
-the most commont type of opponent for drivers is other drivers, followed
+the most common type of opponent for drivers is other drivers, followed
 by pedestrians. The only opponents of pedestrians, on the other hand,
@@
 vehicle is combined with the road user class of the individual to
-distiguish between drivers and passengers of light duty vehicles, light
+distinguish between drivers and passengers of light duty vehicles, light
 trucks, and heavy vehicles, in addition to pedestrians, bicyclists, and
@@
 above, a related modelling strategy is to subset the sample (e.g., Islam
-et al., 2014; Lee and Li, 2014; Tarrao et al., 2014; Wu et al., 2014).
+et al., 2014; Lee and Li, 2014; Torrao et al., 2014; Wu et al., 2014).
 In this case we subset by a combination of traffic unit type of the
 individual (i.e., light duty vehicle, light truck, heavy vehicle,
-pedestrian, bicyclist, and motorcyclist) and vehicle type of the oponent
-(i.e., light duty vehicle, light truck, heavy vehicle). This leads to an
-ensemble of eighteen models to be estimated using subsets of data (see
-Table \ref{tab:model-specification-summary}). By subsetting the sample,
-at least \emph{some} opponent effects are incorporated implicitly.
-Models 1 and 2 are re-estimated using this strategy, dropping variables
-as necessary whenever they become irrelevant (for instance, after
-filtering for pedestrians, no other traffic unit types are present in
-the subset of data). In addition to variables that are no longer
-relevant in some subsamples, it is important to note that when using
-some subsamples a few variables had to be occasionally dropped to avoid
-convergence issues. This tended to happen particularly with smaller
-subsamples where some particular combination of attributes was rare as a
-result of subsampling (e.g., in 2017 there were few or no collisions
-that involved a motorcyclist and a heavy vehicle in a bridge, or
-overpass, or viaduct). The process of estimation carefully paid
+pedestrian, bicyclist, and motorcyclist) and vehicle type of the
+opponent (i.e., light duty vehicle, light truck, heavy vehicle). This
+leads to an ensemble of eighteen models to be estimated using subsets of
+data (see Table \ref{tab:model-specification-summary}). By subsetting
+the sample, at least \emph{some} opponent effects are incorporated
+implicitly. Models 1 and 2 are re-estimated using this strategy,
+dropping variables as necessary whenever they become irrelevant (for
+instance, after filtering for pedestrians, no other traffic unit types
+are present in the subset of data). In addition to variables that are no
+longer relevant in some data subsets, it is important to note that when
+using some data subsets a few variables had to be occasionally dropped
+to avoid convergence issues. This tended to happen particularly with
+smaller subsets where some particular combination of attributes was rare
+as a result of subsampling (e.g., in 2017 there were few or no
+collisions that involved a motorcyclist and a heavy vehicle in a bridge,
+or overpass, or viaduct). The process of estimation carefully paid
 attention to convergence issues to ensure the validity of the models
@@
 
-The resuls of model estimation are discussed in the following section.
+The results of model estimation are discussed in the following section.
 
@@
 In this section we report an in-depth examination of the performance of
-the models. To this end, first inspect the goodness of fit of the
-models. Next, we use the models to conduct in-sample predictions (i.e.,
+the models. We begin by inspecting the statistical goodness of fit of
+the models by means of Akaike's Information Criterion (\(AIC\)). Next,
+we use the models to conduct in-sample predictions (i.e.,
 \emph{nowcasting}), using the same sample that was used to estimate the
-models. In addition, we also conduct out-of-sample predictions, using
-the dataset corresponding to the year 2016, processed in identical way
-as the estimation sample. This is an example of \emph{backcasting}.
-Using these predictions we evaluate the models in two ways: first, we
-compute the estimated shares of each outcome based on the predicted
-probabilities; and secondly, we evaluate the classes of outcomes of the
-individual-level predictions.
+models, and out-of-sample predictions, using the data set corresponding
+to the year 2016 (i.e., \emph{backcasting}). Predictions are commonly
+evaluated in two different in the literature. Some researchers analyze
+the outcome shares based on the predicted probabilities (e.g., Bogue et
+al., 2017; Shamsunnahar and Eluru, 2013). This is a form of aggregate
+forecasting. Other researchers, in contrast, evaluate the classes of
+outcomes based on the individual-level predictions (e.g., Tang et al.,
+2019; Torrao et al., 2014). This is a form of disaggregate forecasting.
 
@@
 
-\caption{\label{tab:model-subsample-summary}\label{tab:model-summary}Summary of model estimation results: Subsample Models}
+\caption{\label{tab:model-subsample-summary}\label{tab:model-summary}Summary of model estimation results: Data Subset Models}
 \centering
@@
 In this, and the following section, \emph{backcasting} refers to the
-prediction of probabilities and classes of outcomes using the 2016
-dataset. When conducting backcasting, the dataset is preprocessed in
-identical manner as the 2017 dataset. In addition, the variables used in
-backcasting match exactly those in the models. This means that some
-variables were dropped when they were present in the 2016 dataset but
+prediction of probabilities and classes of outcomes using the 2016 data
+set. When conducting backcasting, the data set is preprocessed in
+identical manner as the 2017 data set. In addition, the variables used
+in backcasting match exactly those in the models. This means that some
+variables were dropped when they were present in the 2016 data set but
 not in the models. This tended to happen in the case of relatively rare
@@
 \addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{In-sample (nowcasting using 2017 dataset, i.e., estimation dataset)}}\\
+\multicolumn{11}{l}{\textbf{In-sample (nowcasting using 2017 data set, i.e., estimation data set)}}\\
 \hspace{1em}Model 1 & 78886 & 79029.00 & 0.18 & 84675 & 84533.74 & 0.17 & 950 & 948.26 & 0.18 & 0.17\\
@@
 \addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 dataset)}}\\
+\multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 data set)}}\\
 \hspace{1em}Model 1 & 96860 & 96364.67 & 0.51 & 101605 & 102002.59 & 0.39 & 1109 & 1206.74 & 8.81 & 0.50\\
@@
 
-\hypertarget{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}{%
-\subsection{Outcome frequency based on predicted classes
-\{sec:outcomes-classes\}}\label{outcome-frequency-based-on-predicted-classes-secoutcomes-classes}}
+\hypertarget{sec:outcomes-classes}{%
+\subsection{Outcome frequency based on predicted
+classes}\label{sec:outcomes-classes}}
 
@@
 
-The individual-level outcomes are examined using a battery of
+The individual-level outcomes are examined using an extensive array of
 verification statistics. Verification statistics are widely used in the
-evaluation of predective approaches were the outcomes are categorical,
+evaluation of predictive approaches were the outcomes are categorical,
 and are often based on the analysis of \emph{confusion matrices} (e.g.,
@@
 that the number of hits are relatively low; the high values of \(PC_c\)
-in this case are due to the high ocurrence of correct rejections
+in this case are due to the high occurrence of correct rejections
 elsewhere in the table.
@@
 
-\caption{\label{tab:nowcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 dataset, i.e., estimation dataset)}
+\caption{\label{tab:nowcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 data set, i.e., estimation data set)}
 \centering
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51531 & 17137 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67514 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
+\hspace{1em}No Injury & 51531 & 17136 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
+
+\hspace{1em}Injury & 27355 & 67515 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51101 & 17296 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.868} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
-
-\hspace{1em}Injury & 27785 & 67338 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7953} & \textcolor{black}{0.2985} &  &  & \\
+\hspace{1em}No Injury & 51102 & 17297 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
+
+\hspace{1em}Injury & 27784 & 67337 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7952} & \textcolor{black}{0.2985} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35443 & 16145 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
-
-\hspace{1em}Injury & 26974 & 67437 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
+\hspace{1em}No Injury & 35443 & 16146 & 60 &  & \textcolor{black}{70.61} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
+
+\hspace{1em}Injury & 26974 & 67436 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 35498 & 16204 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
-
-\hspace{1em}Injury & 26913 & 67379 & 828 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
+\hspace{1em}No Injury & 35498 & 16203 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
+
+\hspace{1em}Injury & 26913 & 67380 & 828 &  & \textcolor{black}{70.09} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
 
@@
 
-\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 dataset)}
+\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 data set)}
 \centering
@@
 \multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 61684 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
-
-\hspace{1em}Injury & 35171 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3683} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
+\hspace{1em}No Injury & 61683 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
+
+\hspace{1em}Injury & 35172 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3684} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 62735 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5317}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
-
-\hspace{1em}Injury & 34125 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
-
-Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4372}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4335}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
+\hspace{1em}No Injury & 62736 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5318}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
+
+\hspace{1em}Injury & 34124 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
+
+Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4373}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4336}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 62248 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.8619} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
-
-\hspace{1em}Injury & 34610 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
+\hspace{1em}No Injury & 62249 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.862} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
+
+\hspace{1em}Injury & 34609 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 62788 & 21246 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80332 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5878} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
-
-Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4355}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
+\hspace{1em}No Injury & 62788 & 21248 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8687}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2079} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80330 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
+
+Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4354}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 43526 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8217} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
-
-\hspace{1em}Injury & 33915 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\
+\hspace{1em}No Injury & 43526 & 20034 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8218} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
+
+\hspace{1em}Injury & 33915 & 79914 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7987}} & \textcolor{black}{0.3038} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 43561 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5624} & \textcolor{black}{0.3174} &  &  & \\
-
-\hspace{1em}Injury & 33875 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
+\hspace{1em}No Injury & 43560 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5623} & \textcolor{black}{0.3174} &  &  & \\
+
+\hspace{1em}Injury & 33876 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
 
@@
 
-As discussed in Section \ref{sec:techniques}, there is a rich selection
-of modelling approaches that are applicable to crash severity analysis.
+As discussed in Section \ref{sec:methods}, there is a rich selection of
+modelling approaches that are applicable to crash severity analysis.
 Based on the literature, we limited our empirical assessment of
@@
 complicated our empirical assessment, particularly considering the large
-size of the sample involved in this research (a dataset with over
+size of the sample involved in this research (a data set with over
 164,000 records in the case of the full sample models). That said, we
@@
 that subsetting the sample can improve the results in some cases s(e.g.,
-when modelling the severity of crashes involving active travellers or
+when modelling the severity of crashes involving active travelers or
 motorcyclists), but possibly at the risk of overfitting the process and
@@
 interested in crashes involving two parties, since it provides some
-useful guidelines regarding the specification of opponent effects.
+useful guidelines regarding the specification of opponent effects. Not
+only do opponent effects improve the goodness of fit and performance of
+models, they also add rich insights into their effects. The focus on
+this paper was on performance, and for space reasons it is not possible
+to include an examination of the best model without failing to do it
+justice. We plan to report the results of the best-fitting model in a
+future paper.
 
 The analysis also opens up a few avenues for future research. First, for
-reasons discussed in Section \{sec:further-considerations\}, we did not
-consider more sophisticated modelling approaches, such as models with
-random components, partial proportional odds, ranked ordered models, or
-multinomial models, to mention just a few possibilities. Secondly, we
-only considered the performance of the models when making predictions
-for the full sample. That is, the submodels in the ensembles were not
-compared in detail, just their aggregate results when predicting the
-full sample. However, the goodness-of-fit was not uniformly better for
-any one modelling strategy when the data were subset, and it is possible
-that individual models perform better for a certain subset than
-competitors that are part of a better ensemble, overall. For this
-reason, we suggest that additional work with ensemble approaches is
-warranted. Finally, it is clear that the models do not generally do well
-when predicting the least frequent class of outcome, namely Fatality. It
-would be worthwhile to further investigate approaches for so-called
-imbalanced learning, a task where Tarrao et al.~(2014) have already made
-some headway.
+reasons discussed in Section \ref{sec:further-considerations}, we did
+not consider more sophisticated modelling approaches, such as models
+with random components, partial proportional odds, ranked ordered
+models, or multinomial models, to mention just a few possibilities.
+Secondly, we only considered the performance of the models when making
+predictions for the full sample. That is, the submodels in the ensembles
+were not compared in detail, just their aggregate results when
+predicting the full sample. However, the goodness-of-fit was not
+uniformly better for any one modelling strategy when the data were
+subset, and it is possible that individual models perform better for a
+certain subset than competitors that are part of a better ensemble,
+overall. For this reason, we suggest that additional work with ensemble
+approaches is warranted. Finally, it is clear that the models do not
+generally do well when predicting the least frequent class of outcome,
+namely Fatality. It would be worthwhile to further investigate
+approaches for so-called imbalanced learning, a task where Torrao et
+al.~(2014) have already made some headway.
 
@@
 extend them. A common practice in the machine learning community is to
-use canonical dataset to demonstrate the performance of new techniques.
+use canonical data set to demonstrate the performance of new techniques.
 Sharing code and data has remained relatively rare in transportation
-research, and we would like to suggest that the datasets used in this
-research could constitute one such canonical dataset for future
+research, and we would like to suggest that the data sets used in this
+research could constitute one such canonical data set for future
 methodological developments.
@@
 To be added in the final version of the paper.
+
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
+
+Variable definitions in Canada's National Collision Database.
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-collision}\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  C\_CASE & Unique collision identifier & Unique identifier for collisions\\
+C\_YEAR & Year & Last two digits of year.\\
+\rowcolor{gray!6}  C\_MNTH & Month & 14 levels: January - December; unknown; not reported by jurisdiction.\\
+C\_WDAY & Day of week & 9 levels: Monday - Sunday; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_HOUR & Collision hour & 25 levels: hourly intervals; unknown; not reported by jurisdiction.\\
+\addlinespace
+C\_SEV & Collision severity & 4 levels: collision producing at least one fatality; collision producing non-fatal injury; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_VEHS & Number of vehicles involved in collision & Number of vehicles: 1-98 vehicles involved; 99 or more vehicles involved; unknown; not reported by jurisdiction.\\
+C\_CONF & Collision configuration & 21 levels: SINGLE VEHICLE: Hit a moving object (e.g.
+a person or an animal); Hit a stationary object (e.g.
+a tree); Ran off left shoulder; Ran off right shoulder; Rollover on roadway; Any other single vehicle collision configuration; TWO-VEHICLES SAME DIRECTION OF TRAVEL: Rear-end collision; Side swipe; One vehicle passing to the left of the other, or left turn conflict; One vehicle passing to the right of the other, or right turn conflict; Any other two vehicle - same direction of travel configuration; TWO-VEHICLES DIFFERENT DIRECTION OF TRAVEL: Head-on collision; Approaching side-swipe; Left turn across opposing traffic; Right turn, including turning conflicts; Right angle collision; Any other two-vehicle - different direction of travel configuration; TWO-VEHICLES, HIT A PARKED VEHICLE: Hit a parked motor vehicle; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_RCFG & Roadway configuration & 15 levels: Non-intersection; At an intersection of at least two public roadways; Intersection with parking lot entrance/exit, private driveway or laneway; Railroad level crossing; Bridge, overpass, viaduct; Tunnel or underpass; Passing or climbing lane; Ramp; Traffic circle; Express lane of a freeway system; Collector lane of a freeway system; Transfer lane of a freeway system; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+C\_WTHR & Weather condition & 10 levels: Clear and sunny; Overcast, cloudy but no precipitation; Raining; Snowing, not including drifting snow; Freezing rain, sleet, hail; Visibility limitation; Strong wind; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\addlinespace
+\rowcolor{gray!6}  C\_RSUR & Road surface & 12 levels: Dry, normal; Wet; Snow (fresh, loose snow); Slush, wet snow; Icy, packed snow; Debris on road (e.g., sand/gravel/dirt); Muddy; Oil; Flooded; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+C\_RALN & Road alignment & 9 levels: Straight and level; Straight with gradient; Curved and level; Curved with gradient; Top of hill or gradient; Bottom of hill or gradient; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
+\rowcolor{gray!6}  C\_TRAF & Traffic control & 21 levels: Traffic signals fully operational; Traffic signals in flashing mode; Stop sign; Yield sign; Warning sign; Pedestrian crosswalk; Police officer; School guard, flagman; School crossing; Reduced speed zone; No passing zone sign; Markings on the road; School bus stopped with school bus signal lights flashing; School bus stopped with school bus signal lights not flashing; Railway crossing with signals, or signals and gates; Railway crossing with signs only; Control device not specified; No control present; Choice is other than the preceding values; unknown; not reported by jurisdiction.\\*
+\end{longtable}
+\endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-vehicle}\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  V\_ID & Vehicle sequence number & Number of vehicles: 1-98; Pedestrian sequence number: 99; unknown.\\
+V\_TYPE & Vehicle type & 21 levels: Light Duty Vehicle (Passenger car, Passenger van, Light utility vehicles and light duty pick up trucks); Panel/cargo van (<= 4536 KG GVWR   Panel or window type of van designed primarily for carrying goods); Other trucks and vans (<= 4536 KG GVWR); Unit trucks (> 4536 KG GVWR); Road tractor; School bus; Smaller school bus (< 25 passengers); Urban and Intercity Bus; Motorcycle and moped; Off road vehicles; Bicycle; Purpose-built motorhome; Farm equipment; Construction equipment; Fire engine; Snowmobile; Street car; Data element is not applicable  (e.g.
+dummy vehicle record created for pedestrian); Choice is other than the preceding values; unknown; not reported by jurisdiction.\\
+\rowcolor{gray!6}  V\_YEAR & Vehicle model year & Model year; dummy for pedestrians; unknown; not reported by jurisdiction.\\*
+\end{longtable}
+\endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
+\caption{\label{tab:ncdb-descriptives-person}\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endfirsthead
+\caption[]{\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables \textit{(continued)}}\\
+\toprule
+Variable & Description & Notes\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{3}{l}{\textit{Note: }}\\
+\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
+\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
+\endlastfoot
+\rowcolor{gray!6}  P\_ID & Person sequence number & Sequence number: 1-99; Not applicable (dummy for parked vehicles); not reported by jurisdiction.\\
+P\_SEX & Person sex & 5 levels: Male; Female; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_AGE & Person age & Age: less than 1 year; 1-98 years old; 99 years or older; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
+P\_PSN & Person position & Person position: Driver; Passenger front row, center; Passenger front row, right outboard (including motorcycle passenger in sidecar); Passenger second row, left outboard, including motorcycle passenger; Passenger second row, center; Passenger second row, right outboard; Passenger third row, left outboard;...; Position unknown, but the person was definitely an occupant; Sitting on someone’s lap; Outside passenger compartment; Pedestrian; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_ISEV & Medical treatment required & 6 levels: No Injury; Injury; Fatality; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\addlinespace
+P\_SAFE & Safety device used & 11 levels: No safety device used; Safety device used; Helmet worn; Reflective clothing worn; Both helmet and reflective clothing used; Other safety device used; No safety device equipped   (e.g.
+buses); Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
+\rowcolor{gray!6}  P\_USER & Road user class & 6 levels: Motor Vehicle Driver; Motor Vehicle Passenger; Pedestrian; Bicyclist; Motorcyclist; Not stated/Other/Unknown.\\*
+\end{longtable}
+\endgroup{}
 
@@
 
-\leavevmode\hypertarget{ref-Devlin2019road}{}%
-Devlin, A., Beck, B., Simpson, P.M., Ekegren, C.L., Giummarra, M.J.,
-Edwards, E.R., Cameron, P.A., Liew, S., Oppy, A., Richardson, M., Page,
-R., Gabbe, B.J., 2019. The road to recovery for vulnerable road users
-hospitalised for orthopaedic injury following an on-road crash. Accident
-Analysis and Prevention 132, 10.
-doi:\href{https://doi.org/10.1016/j.aap.2019.105279}{10.1016/j.aap.2019.105279}
-
 \leavevmode\hypertarget{ref-Dissanayake2002factors}{}%
@@
 
-\leavevmode\hypertarget{ref-Merlin2007stress}{}%
-Merlin, E.P.R., Gonzalez-Forteza, C., Lira, L.R., Tapia, J.A.J., 2007.
-Post-traumatic stress disorder in patients with non intentional injuries
-caused by road traffic accidents. Salud Mental 30, 43--48.
-
 \leavevmode\hypertarget{ref-Montella2013crash}{}%
@@
 
-\leavevmode\hypertarget{ref-Pelissier2019medical}{}%
-Pelissier, C., Fort, E., Fontana, L., Hours, M., n.d. Medical and
-socio-occupational predictive factors of psychological distress 5 years
-after a road accident: A prospective study. Social Psychiatry and
-Psychiatric Epidemiology 13.
-doi:\href{https://doi.org/10.1007/s00127-019-01780-0}{10.1007/s00127-019-01780-0}
-
 \leavevmode\hypertarget{ref-Penmetsa2017examining}{}%
@@
 
-\leavevmode\hypertarget{ref-Symons2019reduced}{}%
-Symons, J., Howard, E., Sweeny, K., Kumnick, M., Sheehan, P., 2019.
-Reduced road traffic injuries for young people: A preliminary investment
-analysis. Journal of Adolescent Health 65, S34--S43.
-doi:\href{https://doi.org/10.1016/j.jadohealth.2019.01.009}{10.1016/j.jadohealth.2019.01.009}
-
-\leavevmode\hypertarget{ref-Tarrao2014modeling}{}%
-Tarrao, G.A., Coelho, M.C., Rouphail, N.M., 2014. Modeling the impact of
-subject and opponent vehicles on crash severity in two-vehicle
-collisions. Transportation Research Record 53--64.
-doi:\href{https://doi.org/10.3141/2432-07}{10.3141/2432-07}
+\leavevmode\hypertarget{ref-Tang2019crash}{}%
+Tang, J.J., Liang, J., Han, C.Y., Li, Z.B., Huang, H.L., 2019. Crash
+injury severity analysis using a two-layer stacking framework. Accident
+Analysis and Prevention 122, 226--238.
+doi:\href{https://doi.org/10.1016/j.aap.2018.10.016}{10.1016/j.aap.2018.10.016}
 
@@
 doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
+
+\leavevmode\hypertarget{ref-Torrao2014modeling}{}%
+Torrao, G.A., Coelho, M.C., Rouphail, N.M., 2014. Modeling the impact of
+subject and opponent vehicles on crash severity in two-vehicle
+collisions. Transportation Research Record 53--64.
+doi:\href{https://doi.org/10.3141/2432-07}{10.3141/2432-07}
 
@@
 
-\leavevmode\hypertarget{ref-Wijnen2019analysis}{}%
-Wijnen, W., Weijermars, W., Schoeters, A., Berghe, W. van den, Bauer,
-R., Carnis, L., Elvik, R., Martensen, H., 2019. An analysis of official
-road crash cost estimates in european countries. Safety Science 113,
-318--327.
-doi:\href{https://doi.org/10.1016/j.ssci.2018.12.004}{10.1016/j.ssci.2018.12.004}
-
 \leavevmode\hypertarget{ref-WHO2019global}{}%

### 2020-05-10 Interactions-Crash-Severity-v1/Response-to-Reviewers-v2.Rmd
First round of edits to respond to reviewers after decision on original submission.

---
title: "Response to reviewers"
author: ""
date: "5/8/2020"
output: pdf_document
---

# Reviewer 1

The paper investigates three general modelling strategies to analyze crash data. Namely, use of opponent related factors, single and multi-level model approach, and different sampling approach.

>*Thank you for reading the paper and for your feedback to improve the paper.*

The combined application of those strategies is not completely new because it is part of a general methodological approach (and usually crash data contain information about opponent factors). The paper, in the present form, seems more an exercise, a good exercise, but not so much sufficient as to justify a paper. Why not to use other models besides ordered probit (row 525) (or ordered logit as in the abstract)?

>*Thank you for this perceptive comment. Indeed, the combined application of these strategies is not completely new, or new at all, since every single one of them has been used in the past by someone or other. Our argument is not that the strategies are new, but rather that they are used in a haphazard fashion (we document the different types of applications in the paper). Moreover, we argue that there is no systematic examination of how each of these strategies perform, and therefore little guidance in the literature in terms of which one to choose in applications. We believe that our paper provides new information that is not available elsewhere in the literature. With respect to the use of other models, this has been investigated by several others, who have compared multinomial to ordered models, ordered models to ordered models with random components, etc. We would like to draw your attention to this paragraph in the paper (p. 26, lines 497-510):*

>>*"We do not report results regarding other modelling strategies. On the one hand, more sophisticated modelling frameworks are generally capable of improving the performance of a model. On the other hand, there are well-known challenges in the estimation of more sophisticated models (see Lenguerrand et al., 2006, p. 47, for a discussion of convergence issues in models with mixed eﬀects; Mannering et al., 2016, p. 13, for some considerations regarding the complexity and cost of estimating more complex models; and Bogue et al., 2017, p. 27, on computational demands of models with random components). The additional cost and complexity of more sophisticated modelling approaches would, in our view, have greatly complicated our empirical assessment, particularly considering the large size of the sample involved in this research (a data set with over 164,000 records in the case of the full sample models). That said, we experimented with a model with random components using monthly subsets of data to ﬁnd that, indeed, estimation takes considerably longer, is more demanding in terms of ﬁxing potential estimation quirks, and in the end resulted in variance components that could not be reliably estimated as diﬀerent from zero (results can be consulted in the source R Notebook). For this reason, we choose to leave the application of more sophisticated models as a matter for future research."*

>*We _did_ try other models, but there is no squeezing out unobserved heterogeneity where none seems to exist.*

To support this argument, tables 9 and 10 (confusion matrices) show a very poor performance of the models, especially for fatality;  Models Ensemble improve it a little.

>*If you compare the average prediction errors (APEs) and weighted average prediction errors (WAPEs) of our models to those reported by Bogue, Paleti, and Balam (2017), you will notice that our models actually perform well according to these criteria. However, part of our argument is that a more complete comparison of the performance of models requires the use of other performance criteria, hence all the verification statistics.*

It is curious that its ‘percent correct by class’ is generally very high (close to or greater than 90%) when numbers are so bad; it is curious too that the ‘percent correct’ is always lower than the lowest value of each ‘percent correct by class’.

>*It is not that curious, really. Percent correct by class is heavily influenced by the dominant class (see note in Table 8). It is calculated as the sum of total hits and correct non-events divided by the number of cases. Accordingly (see for example Table 9, Model 1):*

Outcome | Pred. No Injury | Pred. Injury | Pred. Fatality |
--- | --- | --- | --- |
No Injury | 50652 | 22503 | 150 |
Injury | 28232 | 62121 | 797 |
Fatality | 2 | 51 | 3 |

>*The percent correct by class (No Injury) is calculated as follows:*

$$
PC_{No-Injury} = \frac{50652 + 62121 + 797 + 51 + 3}{164511} = 0.6907 = 69.07\%
$$

>*In the above, $50652$ is the number of total hits, and $62121 + 797 + 51 + 3$ is the number of correct non-events. Compare to the percent correct by class (Fatality):*

$$
PC_{Fatality} = \frac{3 + 50652 + 22503 + 28232 + 62121}{164511} = 0.9939 = 99.39\%
$$

>*In the above, $3$ is the number of total hits, and $50652 + 22503 + 28232 + 62121$ is the number of correct non-events. It is not surprising that the value is so high: in fact, the model correctly predicts non-fatalities most of the time! However, when we look at bias ($B$) we see that fatalities are severely underpredicted. This is despite APE and WAPE statistics that are similar or better than those reported in the literature. Our argument is that the use of verification statistics provides a more nuanced view of the performance of the models. Again, to the best of our knowledge, this more nuanced perspective is not currently available, and that it provides new insights into the performance of models.*

The availability of crash data  at a repository on the web to reproduce the research or to make other seems very interesting (not only for exercise) and should be encouraged.

>*Thank you. We also think that this should be encouraged. We did not find in our literature review a single paper on crash severity that was reproducible.*

Title is a little misleading since the role of ‘opponent’ variables is not really the main focus of the paper.

>*We have changed the title based on this and a related comment by Reviewer 2.*

## Minor comments

Reorganize section 5.3.1 by listing at the beginning the considered measures for verification statistics. Furthermore, all the section is difficult to read.

>*We followed this recommendation in the revision and reorganized the section to describe the verification statistics at the beginning. We also edited it, hopefully improving its readability.*

At row 271 in which unit is 4536?

>*It should be kg. We added the units.*

At row 302, “in two different” what?

>*In two different ways. This was corrected.*

At row 435, CSI requires  a note.

>*This paragraph was deleted as part of reorganizing the section.*

At row 533, there is an s too many in ‘s(e.g. …)’

>*Typo corrected.*

Acknowledgment: It is quite strange that authors did not add now their acknowledgments but they reserve to do it in the ‘final’ version.

>*This was an oversight caused by the practice of anonymizing papers. The revision includes the acknowledgments.*

>*Thanks again for your feedback and the time invested in helping us improve this paper. We hope that with these revisions the contributions of the paper will be more easily appreciated.*

\newpage

# Reviewer 2

An empirical assessment of strategies to model opponent effects in crash severity analysis – the paper examines the injury severity of crashes employing ordered model.

“Accident” should be avoided. Crash/collision should be used.

>*Thank you for this comment. We thought that since the name of the journal is Accident Analysis and Prevention the term "accident" was appropriate to refer to stochastic events with undesirable outcomes. We have changed most "accidents" to "collisions" and "crashes" to reflect the nuanced differences between them.*

“ways to deal with the way the different parties in a crash” – this sentence is confusing. Does it refer to methodological ways to accommodate for such correlation, if present? Please clarify.

>*The sentence was wordy and unclear, and it was edited for clarity. Now it reads:*

>>*Numerous efforts have been undertaken to understand the factors that affect road collisions in general, and the severity of crashes in particular. In this literature several strategies have been proposed to model interactions between parties in a crash, including the use of variables regarding the other party (or parties) in the collision, data subsetting, and estimating models with hierarchical components.*

“opponent effects” – this is misleading. What effects are you referring to and does opponent refer to 2 vehicles involved in a crash? After reading the paper it was possible to infer it, but, these terms should be clearly defined to avoid confusions and to improve readability.

>*"Opponent effects" are the effects on outcome severity of the attributes of the opposite party in the accident. For clarity, we now define these terms earlier in the paper.*

“participants in the crash”- how can someone participate in a crash?

>*We changed "participants" to "parties" and "individuals" to distinguish between groups of people who were party to a crash, and the individuals who composed those parties.*

Please clearly state what are the objectives of the study and how it is contributing to existing road safety research. It is not at all clear from introduction/background section. It was not even clear whether authors are focusing on motor vehicle crashes only.

>*The introduction was rewritten to remove some background information that was not strictly necessary, and to more clearly state the objectives and contributions of the paper.*

The four strategies identified in the methodology section; how different approaches contribute towards the empirical context? Please provide a detailed discussion first.

>*We identify three basic strategies, which can be combined. This is explained in sections 3.2 to 3.4.*

Line 125: the second part collect – what does collect represent here?

>*To bring together, as in, several items that are collected in the second summation in the formula. We changed this to "gather".*

Page 2, line 29 – “shamsunnahar and Eluru” should be “Yasmin and Eluru”.

>*Thank you for this clarification. This has been corrected.*

Line 224: Were pedestrian/bicycle involved crashes also considered in the severity analysis along with motor vehicle crashes? The mechanisms of injury severity for these two groups are different. How these could be modeled in one model?

>*Yes, they were, unless they were modelled separately. For example, Models 1 and 2 were re-estimated after subsetting the sample, so that only pedestrians were considered, or bicyclists, or light vehicles, etc. How could several be modelled in one model? By using the strategies described in the methods section.*

4.1.2 and 4.1.3: these procedures are standard in preparing crash severity data. Not sure why these are discussed in detail. For the data cleaning, the fundamental question is, other than deleting the cases with unknown information, how the statistical property of the dependent variable was maintained?

>*Data preprocessing is not as sexy as a new ranked ordered logit model with mixtures of copulas, etc. In fact, data preprocessing is seldom report in detail: everybody assumes that it is standard. As we were working on this research, what we found is that data preprocessing is as complex as data analysis and modelling, and we could not find good, complete discussions of this topic. Since data preprocessing is an essential component of two of the strategies discussed in the paper (data subsetting and opponent-effects), we would rather maintain this as part of the paper, unless the editor indicates otherwise.*

Line 262: virtually all crashes were fatal when the opponent is a driver – this statement is fundamentally wrong.

>*That is not what we said. This is the actual statement:*

>>*"In terms of outcomes, we observe that virtually all fatalities occur when the opponent is a driver, and only very rarely when the opponent is a motorcyclist."*

>*As seen in Table 2, where we split the outcomes by the roles of the opposite parties, the only fatalities recorded outside of Opponent: Driver where for Opponent: Motorcycle (7 fatalities). In other words, virtually all fatalities occur when the opponent is a driver. This is very different from "people virtually always die when the opponent is a driver", which seems to be your odd reading of our statement.*

The data sample is high. How is it ensured that the model is not over fitted with high number of records? Number of co-efficient in the models are an indication of over fitting.

>*No. The data sample is not "high", it is large. In the present case, we have a sample of size 164,511, and the models have at most 111 coefficients: this leaves tens of thousands of degrees of freedom. In fact, thinking that the number of coefficients is an indication of overfitting is overly-simplistic, since overfitting is a combination of sample size and the number of parameters. Evidence of overfitting only can be gleaned from evaluating the performance of the model when "new" information is used. That is the point of backcasting: to evaluate how the models behave in an out-of-sample situation.*

The study design is quite confusing. Isn’t it obvious that Model 3 and Model 4 are generalized version of the Model 1 and 2. More importantly, Model 3 and 4 can be easily combined to estimate a rather genralized version of severity model. What is the point of this modelling exercise? What is new in this exercise and why one should consider it?

>*It might be obvious to you that Models 3 and 4 can be easily combined, but we regret to inform you that there would be perfect multicollinearity if you tried to do so. The point is that even an expert such as yourself may fail to see this, and we consider that presenting a framework that explains the various strategies and compares them might be worthwhile for readers who eventually read the paper.*

The paper is merely a modelling exercise without any significant contribution to the state-of-the-art of severity model. The wordings of the paper are significantly misleading and confusing. If we are considering number of occupants in different vehicle units involved in a crash, the outcome is likely to be correlated. How such correlations were accommodated? What is the purpose of this modelling exercise rather than just estimating a set of models?

*We regret that you were not persuaded by the paper. We have taken into account those suggestions in your report that make sense, and hope that you will be persuaded to re-read the paper with attention to maybe see its value.*

\newpage

# Reviewer 3

The reviewer found the paper to be exceptionally well researched. It is quite long and so full of details, some quite nuanced, that reading the paper and digesting its many remarks and information is time consuming. However, the authors are praised on giving such detail, and especially on opening the code and data for others. It makes this paper highly useful as a base for further research on this important topic. The reviewer points out that Mannering has co-authored earlier work than Wang and Kockelman, which used direct opponent effects and classified models based on opponent types, in 2004 in AAP, although that particular paper did not focus on the opponent effects.

>*Thank you for your positive assessment of our paper. We realize that it is information-rich, and appreciate the time and effort you put into this review.*

One small item to improve would be to name the four models with descriptive titles rather than only model 1, model 2, model 3, and model 4, which requires readers to remember which is which and makes the tables unreadable for readers browsing the paper.

>*Thank you for this suggestion. We have labelled the models in the tables, which we think makes it easier to recall the main characteristics of each model, i.e., single-level, hierarchical, etc.*

>*Again, thank you for your thoughtful comments and suggestions.*

\newpage

# Reviewer 4

This study aims to evaluate approaches to model opponent effects in crashes involving two parties. A series of models are established and evaluated. The paper is well written and organized. Besides, authors make the data and code publicly available, which is great. Here are my comments:

>*Many thanks for your positive assessment of the paper.*

1. The data preprocessing section could be simplified to make the paper more concise.

>*Thank you for this comment. We collapsed three subsections of data preprocessing into one, and edited it to make it somewhat more concise. We are convinced that data preprocessing is a fundamental part of the process to implement the strategy of opponent-related variables. The process, as the paper probably shows, is not necessarily straightforward. And although not as sexy as presenting a new model with yet more unobserved heterogeneity terms, is as important. Our preference is to keep this aspect of the process well documented and transparent, but if the editor insists, we could reduce it even more.*

2. Ln 159, for any given coefficient q, you can have equation (7). I don’t quite understand it. Please clarify.

>*Thanks for catching this. The sentence should say "for any given coefficient $m$", since the nesting is of individual $l$ in traffic unit $m$. This was corrected.*

3. Table 3 is mentioned in section 4 but shown in section 5. It’s better to adjust the location of Table 3.

>*Done.*

4. Ln 376-377, as mentioned by authors, models in Table 6 are clearly overfitted to the 2017 dataset. They don’t generalize well to new data. To this end, the analysis based on these models might be biased.

>*It is true, and there is the perennial trade-off between bias and variance. That said, overfitting is not necessarily a bad thing, and an analyst might prefer somewhat biased predictions in exchange for lower variance. Here we do not take a strong stand on this issue because it really depends on the specific use of the models.*

5. As shown in Table 8, authors evaluated the outcomes of models by lots of different indicators. Normally, it’s difficult to compare model performances based on so many indicators. How did you manage that? BTW, what’s the difference between Percent Correct by Class and Probability of Detection?

>*We believe that using a fuller battery of indicators is a bit more challenging (it makes it somewhat more difficult to find the absolute _best_ model), but on the other hand it does give numerous new insights into how specific models perform well or not. As an example, Model 1 produces individual-level predictions that are less biased, but also has a higher Probability of False Detection. Does the analyst care more about reducing bias, or about reducing false detection? That might guide the choice of a model. The difference between the two verification statistics is that Percent Correct by Class is calculated with the hits and correct non-events, whereas the Probability of Detection is calculated using only the hits.*

>*Thank you for your suggestions. We hope that you will find the revised version of the paper much improved.*

### 2020-05-10 Interactions-Crash-Severity-v1/Cover-Letter-v2.Rmd
First round of edits to respond to reviewers after decision on original submission.

---
title: "Cover Letter"
author: ""
date: "5/8/2020"
output: pdf_document
---

Dear Editor:

We would like to express our gratitude to four anonymous reviewers for their generous reviews of our paper. Their comments have helped us to improve the paper. In this letter we detail our responses to the feedback received, and actions taken in this revision.

We trust that you and the expert reviewers will find that our responses address in a thorough and conscientious manner the points raised with respect to the previous version of the paper.

Looking forward to hearing back from you in due course, we remain sincerely yours.

The Authors

### 2020-05-10 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v2.Rmd
First round of edits to respond to reviewers after decision on original submission.

---
title: An empirical assessment of strategies to model party interactions in crash severity analysis
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: Corresponding Author
  - name: Hany Hassan
    email: hassan1@lsu.edu
    affiliation: Louisiana State University
  - name: Mark Ferguson
    email: fergumr@mcmaster.ca
    affiliation: McMaster University
  - name: Saiedeh Razavi
    email: razavi@mcmaster.ca
    affiliation: McMaster University    
address:
  - code: McMaster University
    address: McMaster Institute for Transportation and Logistics, McMaster University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1
  - code: Louisiana State University
    address: Department of Civil and Environmental Engineering, Louisiana State University, Baton Rouge, Louisiana, USA 70803
abstract: |
  Road crashes impose an important burden on health and the economy. Numerous efforts have been undertaken to understand the factors that affect road collisions in general, and the severity of crashes in particular. In this literature several strategies have been proposed to model interactions between parties in a crash, including the use of variables regarding the other party (or parties) in the collision, data subsetting, and estimating models with hierarchical components. Since no systematic assessment has been conducted of the performance of these strategies, they appear to be used in an ad-hoc fashion in the literature. The objective of this paper is to empirically evaluate ways to model party interactions in the context of crashes involving two parties. To this end, a series of models are estimated using data from Canada's National Collision Database. Three levels of crash severity (no injury/injury/fatality) are analyzed using ordered probit models and covariates for the parties in the crash and the conditions of the crash. The models are assessed using predicted shares and classes of outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis. The study also suggests that hierarchical (i.e., multi-level) specifications and subsetting do not necessarily perform better than a relatively simple single-level model with opponent-related factors. The results of this study provide insights regarding the performance of different modelling strategies, and should be informative to researchers in the field of crash severity.

journal: "Some Journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article

header-includes:
   - \usepackage[margin=1in]{geometry}
   - \usepackage{lineno}
   - \linenumbers
---

### 2020-05-10 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v2.tex
First round of edits to respond to reviewers after decision on original submission.

--- 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
@@
             pdfauthor={},
-            pdftitle={An empirical assessment of strategies to model opponent effects in crash severity analysis},
+            pdftitle={An empirical assessment of strategies to model party interactions in crash severity analysis},
             colorlinks=false,
@@
 \linenumbers
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
 
-  \title{An empirical assessment of strategies to model opponent effects in crash
-severity analysis}
-    \author[Some University]{Author 1\corref{Corresponding Author}}
-   \ead{author1@address.edu} 
-    \author[Another University]{Author 2}
-   \ead{author2@address.edu} 
-      \address[Some University]{Address}
-    \address[Another University]{Address}
+  \title{An empirical assessment of strategies to model party interactions in
+crash severity analysis}
+    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
+   \ead{paezha@mcmaster.ca} 
+    \author[Louisiana State University]{Hany Hassan}
+   \ead{hassan1@lsu.edu} 
+    \author[McMaster University]{Mark Ferguson}
+   \ead{fergumr@mcmaster.ca} 
+    \author[McMaster University]{Saiedeh Razavi}
+   \ead{razavi@mcmaster.ca} 
+      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
+University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
+    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
+University, Baton Rouge, Louisiana, USA 70803}
 
   \begin{abstract}
-  Road accidents impose an important burden on health and the economy.
+  Road crashes impose an important burden on health and the economy.
   Numerous efforts have been undertaken to understand the factors that
@@
   particular. In this literature several strategies have been proposed to
-  model opponent effects, that is, ways to deal with the way the different
-  parties in a crash interact to influence the severity of the individual
-  outcomes. These strategies include the use of opponent variables, data
-  subsetting, and estimating models with hierarchical components. Since,
-  no systematic assessment has been conducted of the performance of these
+  model interactions between parties in a crash, including the use of
+  variables regarding the other party (or parties) in the collision, data
+  subsetting, and estimating models with hierarchical components. Since no
+  systematic assessment has been conducted of the performance of these
   strategies, they appear to be used in an ad-hoc fashion in the
   literature. The objective of this paper is to empirically evaluate ways
-  to model opponent effects in the context of crashes involving two
+  to model party interactions in the context of crashes involving two
   parties. To this end, a series of models are estimated using data from
   Canada's National Collision Database. Three levels of crash severity (no
-  injury/injury/fatality) are analyzed using ordered logit models and
-  covariates for the participants in the crash and the conditions of the
-  crash. The models are assessed using predicted shares and classes of
-  outcomes, and the results highlight the importance of considering
-  opponent effects in crash severity analysis. The study also suggests
-  that hierarchical (i.e., multi-level) specifications and subsetting do
-  not necessarily perform better than a relatively simple single-level
-  model with opponent-related factors. The results of this study provide
-  insights regarding the performance of different modelling strategies,
-  and should be informative to researchers in the field of crash severity.
+  injury/injury/fatality) are analyzed using ordered probit models and
+  covariates for the parties in the crash and the conditions of the crash.
+  The models are assessed using predicted shares and classes of outcomes,
+  and the results highlight the importance of considering opponent effects
+  in crash severity analysis. The study also suggests that hierarchical
+  (i.e., multi-level) specifications and subsetting do not necessarily
+  perform better than a relatively simple single-level model with
+  opponent-related factors. The results of this study provide insights
+  regarding the performance of different modelling strategies, and should
+  be informative to researchers in the field of crash severity.
   \end{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Evidence from across the world suggests that the burden of road
-accidents is not borne evenly. There are important disparities at the
-international level, where the odds of death due to road crashes are
-three times higher in low-income countries compared to high-income
-countries (World Health Organization, 2019). In the case of high-income
-countries, where substantial gains in road safety have been observed for
-years, said gains have also been unevenly distributed; thus, while fatal
-crashes involving older adults in the United States and Great Britain
-declined between 1997 and 2010 (despite the graying of the population),
-the trend remained stable or increased slightly in Australia in roughly
-the same period (Thompson et al., 2018). The impact of road accidents
-also varies considerably. For example, in a study in the United States,
-Obeng (2011) reported that the factors affecting crash severity varied
-substantially in magnitude by gender. More recently, Regev et al.~(2018)
-used adjusted crash risk to find that the risk of fatal injuries for
-older drivers remained high even for minor collisions - which highlights
-the perils of accidents at older ages. Other studies have concentrated
-on the consequences of road accidents for the young (e.g., Peek-Asa et
-al., 2010), the old (e.g., Rakotonirainy et al., 2012), as well as
-pedestrians and cyclists (e.g., Hanson et al., 2013; McArthur et al.,
-2014).
-
-Given the relevance and cost of this matter, as well as the unevenness
-of impacts among different population segments, numerous efforts have
-been conducted to better understand the factors that affect road safety
-- including the probable consequences of crashes. Along these lines, a
-stream of research in the analysis of road accidents is concerned with
-the severity of crashes. In particular, multivariate analysis of crash
-severity is a useful way to clarify the way various factors can affect
-the outcome of an incident, and to discriminate between various levels
-of injury. This is an active area of research, and one where
-methodological developments have aimed at improving the reliability,
-accuracy, and precision of models (e.g., Savolainen et al., 2011; Bogue
-et al., 2017; Shamsunnahar and Eluru, 2013). Of interest in the
-literature on crash severity is the way different parties in a crash
-interact to influence the severity of individual outcomes. The
-importance of these interactions has been recognized in the existing
-literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al., 2017;
-Torrao et al., 2014), and a number of different modelling strategies
-have been proposed. These include the use of \emph{data subsetting},
-\emph{opponent variables}, and estimating models with \emph{hierarchical
-components}. A gap in the literature, however, is a systematic
-assessment of the performance of these various strategies.
-
-The objective of this paper is to empirically evaluate different
-strategies to model opponent effects in crash severity, in the context
-of incidents involving two parties. Therefore, we present a systematic
-assessment of several relevant modelling strategies. For the assessment
-we use data from Canada's National Collision Database, a database that
-collects all police-reported collisions in the country. Using the most
-recent version of the data set (2017). Three levels of crash severity
-(no injury/injury/fatality) are analyzed using ordered logit models, and
-covariates for the participants in the crash and the conditions of the
-crash. For model assessment, we conduct an in-sample prediction exercise
-using the estimation sample (i.e., \emph{nowcasting}), and also an
-out-of-sample prediction exercise using the data set corresponding to
-2016 (i.e., \emph{backcasting}). The models are assessed using predicted
-shares and predicted classes of outcomes at the individual level, using
-an extensive array of verification statistics.
-
-The rest of this paper is structured as follows. In Section
-\ref{sec:background} we discuss some background matters, and follow this
-with a concise review of the modelling strategies used to analyze crash
-severity \ref{sec:methods}. Section \ref{sec:application} describes the
-data requirements, data preprocessing, and the modelling strategies,
-along with the results of model estimation. The results of assessing the
-models and the discussion of these results is found in Section
-\ref{sec:assessment}. We then present some additional thoughts about the
-applicability of this approach in Section
-\ref{sec:further-considerations} before offering some concluding remarks
-in Section \ref{sec:concluding-remarks}.
-
-\hypertarget{sec:background}{%
-\section{Background}\label{sec:background}}
-
-Modelling the severity of injuries to participants in road crashes has
-been a preoccupation of transportation researchers, planners, auto
-insurance companies, governments, and the general public for decades.
-One of the earliest studies to investigate the severity of injuries
-conditional on an accident having occurred was by White and Clayton
-(1972). Kim et al.~(1995) later stated that the ``linkages between
-severity of injury and driver characteristics and behaviors have not
-been thoroughly investigated'' (p.~470). Nowadays, there is a burgeoning
-literature on this subject, including methodological developments, case
-studies, and more niche research with a focus on particular situations
-(e.g., crashes at intersections, Mussone et al., 2017; crashes in rural
-roads, Gong and Fan, 2017), and crashes involving special population
-groups (e.g., crashes involving motorcyclists or active travelers; see
-Shaheed et al., 2013; Salon and McIntyre, 2018).
-
-Crash severity is often modeled using models for discrete outcomes. An
-analyst interested in crash severity has at their disposal an ample
-repertoire of models to choose from, including classification techniques
-from machine learning (e.g., Iranitalab and Khattak, 2017; Chang and
-Wang, 2006; Effati et al., 2015; Khan et al., 2015), Poisson models for
-counts (e.g., Ma et al., 2008), unordered logit/probit models (e.g., Tay
-et al., 2011), as well as ordered logit/probit models (e.g., Rifaat and
-Chin, 2007), with numerous variants, such as random parameters/mixed
-logit (e.g., Aziz et al., 2013; Haleem and Gan, 2013), partial
-proportional odds models (e.g., Mooradian et al., 2013; Sasidharan and
-Menendez, 2014), and the use of copulas (e.g., Wang et al., 2015).
-Recent reviews of methods include Savolainen et al.~(2011), Shamsunnahar
-and Eluru (2013), and Mannering et al.~(2016).
-
-Irrespective of the modelling framework employed, models of crash
-severity often include variables in several categories, as shown with
-examples in Table \ref{tab:variable-categories} (also see Montella et
-al., 2013). Many crash databases and analyses also account for the
-multievent nature of many crashes. Participants may have had different
-roles in a crash depending on their context, with some acting as
-operators of a vehicle (i.e., drivers, bicyclists), while others were
-passengers. They also may differ depending on what type of traffic unit
-they were, for example pedestrians, or operators/passengers of a
-vehicle. The multiplicity of roles makes for complicated modelling
-decisions when trying to understand the severity of injuries; for
-example, what is the unit of analysis, the person, the traffic unit, or
-the collision? Not surprisingly, it is possible to find examples of
-studies that adopt different perspectives. A common simplifying strategy
-in model specification is to consider only \emph{drivers} and/or only
-\emph{single-vehicle} crashes (e.g., Kim et al., 2013; Gong and Fan,
-2017; Lee and Li, 2014; Osman et al., 2018). This strategy reduces the
-dimensions of the event, and it becomes possible, for example, to equate
-the traffic unit to the person for modelling purposes.
-
-The situation becomes more complex when dealing with events that involve
-two traffic units (e.g., Torrao et al., 2014; Wang et al., 2015) and
-multi-traffic unit crashes (e.g., Wu et al., 2014; Bogue et al., 2017).
-Different strategies have been developed to study these, more complex
-events. A number of studies advocate the estimation of separate models
-for different participants and/or situations. In this way, Wang and
-Kockelman (2005) estimated models for single-vehicle and two-vehicle
-crashes, while Savolainen and Mannering (2007) estimated models for
-single-vehicle and multi-vehicle crashes. More recently, Duddu et
-al.~(2018) and Penmetsa et al.~(2017) presented research that estimated
-separate models for at-fault and not-at-fault drivers. The strategy of
-estimating separate models relies on \emph{subsetting} the data set,
-although it is possible to link the relevant models more tightly by
-means of a common covariance structure, as is the case of bivariate
-models (e.g., Chiou et al., 2013; Chen et al., 2019) or models with
-copulas (e.g., Rana et al., 2010; Shamsunnahar et al., 2014; Wang et
-al., 2015).
-
-A related strategy to specify crash severity models is to organize the
-data in such a way that it is possible to introduce \emph{opponent
-effects}. There are numerous examples of studies that consider at least
-some characteristics of the opponents in two- or multi-vehicle crashes.
-For example, Wang and Kockelman (2005) considered the type of the
-opposing vehicle in their model for two-vehicle collisions. Similarly,
-Torrao et al.~(2014) included in their analysis the age, wheelbase,
-weight, and engine size of the opposing vehicle, while Bogue et
-al.~(2017) used the body type of the opposing vehicle. Penmetsa et
-al.~(2017) and Duddu et al.~(2018) are two of the most comprehensive
-examples of using opponent's information, as they used individual
-attributes of opponents (their physical condition, sex, and age), as
-well as characteristics of the opposing traffic unit (the vehicle type
-of the opponent). The twin strategies of subsetting the sample and using
-the attributes of the opponent are not mutually exclusive, but neither
-are they used consistently together, as a scan of the literature
-reveals. Another strategy is to introduce \emph{hierarchical components}
-in the model, a technique widely used in the hierarchical or multi-level
-modelling literature.
-
-In this paper we consider three general modelling strategies, as
-follows:
-
-\begin{itemize}
-\tightlist
-\item
-  Strategy 1. Introducing opponent-related factors
-\item
-  Strategy 2. Single-level model and multi-level (hierarchical) model
-  specifications
-\item
-  Strategy 3. Full sample and sample subsetting
-\end{itemize}
-
-These are discussed in more detail in the following section.
-
-\begin{table}
-
-\caption{\label{tab:table-variable-categories}\label{tab:variable-categories}Categories of variables used in the analysis of crash severity with examples}
-\centering
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{22em}}
-\toprule
-Category & Examples\\
-\midrule
-Human-factors & Attributes of participants in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
-Traffic unit-factors & Attributes of the traffic unit, e.g., type of traffic unit (pedestrian, car, motorcycle, etc.), maneuver, etc.\\
-Environmental-factors & Attributes of the crash, e.g., location, weather conditions, light conditions, number of parties, etc.\\
-Road-factors & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
-Opponent-related factors & Attributes of the opponent, e.g., age of opponent, gender of opponent, opponent vehicle type, etc.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{sec:methods}{%
-\section{Methods}\label{sec:methods}}
-
-\hypertarget{choice-of-model}{%
-\subsection{Choice of model}\label{choice-of-model}}
-
-Shamsunnahar and Eluru (2013) conducted an extensive comparison of
-models for discrete outcomes in crash severity analysis, and found only
-small differences in the performance of unordered models and ordered
-models; however, ordered models are usually more parsimonious since only
-one latent functions needs to be estimated for all outcomes, as opposed
-to one for each outcome in unordered models. Bogue et al.~(2017) also
-compared unordered and ordered models in the form of the mixed
-multinomial logit and a modified rank ordered logit, respectively, and
-found that the ordered model performed best. To keep the empirical
-assessment manageable, in this paper we will consider only the ordinal
-logit model, and will comment on potential extensions in Section
-\ref{sec:further-considerations}.
-
-The ordinal model is a latent-variable approach, whereby the severity of
-the crash (observed) is linked to an underlying latent variable that is
-a function of the variables of interest, as follows:
-
-\begin{equation}
-\label{eq:latent-function}
-y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
-\end{equation}
-
-The left-hand side of the expression above (\(y_{itk}^*\)) is a latent
-(unobservable) variable that is associated with the severity of crash
-\(k\) (\(k=1,\cdots,K\)) for participant \(i\) in traffic unit \(t\).
-The right-hand side of the expression is split in four parts. The first
-part collects \(l=1,\cdots,L\) human-factors \(p\) for participant \(i\)
-in traffic unit \(t\) and crash \(k\); these could relate to the person
-(e.g., age, gender, and road user class). The second part collects
-\(m=1,\cdots,M\) attributes \(u\) related to traffic unit \(t\) in crash
-\(k\); these could be items such as maneuver or vehicle type. The third
-part collects \(q=1,\cdots,Q\) attributes \(c\) related to the crash
-\(k\), including environmental-factors and road-factors, such as weather
-conditions, road alignment, and type of surface. Lastly, the fourth
-element is a random term specific to participant \(i\) in traffic unit
-\(t\) and crash \(k\). The function consists of a total of \(Z=L+M+Q\)
-covariates and associated parameters.
-
-For conciseness, in what follows we will abbreviate the function as
-follows:
-
-\begin{equation}
-\label{eq:latent-function-compact}
-y_{itk}^*=\sum_{z=1}^Z\theta_zx_{itkz} + \epsilon_{itk}
-\end{equation}
-
-The latent variable is not observed directly, but it is possible to
-posit a probabilistic relationship with the outcome \(y_{itk}\) (the
-severity of crash \(k\) for participant \(i\) in traffic unit \(t\)).
-Depending on the characteristics of the data and the assumptions made
-about the random component of the latent function different models can
-be obtained. For example, if crash severity is coded as a variable with
-three outcomes (e.g., property damage only/injury/fatal), we can relate
-the latent variable to the outcome as follows:
-
-\begin{equation}
-\label{eq:latent-function-ordered-outcomes}
-y_{itk} = 
-\begin{cases}
-\text{fatality} & \text{if } y_{itk}^*> \mu_2\\
-\text{injury} & \text{if } \mu_1< y_{itk}^*< \mu_2\\
-\text{PDO} & \text{if } y_{itk}^*< \mu_1
-\end{cases}
-\end{equation}
-
-\noindent where \(\mu_1\) and \(\mu_2\) are estimable thresholds. Due to
-the stochastic nature of the latent function, the outcome of the crash
-is not fully determined. However, we can make the following probability
-statements:
-
-\begin{equation}
-\label{eq:probability-ordered-outcomes}
-\begin{array}{rcl}\
-P(y_{itk} = \text{PDO}) &=& 1 - P(y_{itk} = \text{injury}) - P(y_{itk} = \text{fatality})\\ 
-P(y_{itk} = \text{injury}) &=& P(\mu_1 - \sum_{z=1}\theta_zp_{itkz} < \epsilon_{itk} < \mu_2 - \sum_{z=1}\theta_zp_{itkz})\\
-P(y_{itk} = \text{fatality}) &=& P(\epsilon_{itk} < \mu_1 - \sum_{z=1}\theta_zp_{itkz})
-\end{array}
-\end{equation}
-
-If the random terms are assumed to follow the logistic distribution, the
-ordered logit model is obtained; if the normal distribution, then the
-ordered probit model. Estimation methods for these models are very
-well-established (e.g., Maddala, 1986; Train, 2009). There are numerous
-variations of the basic modelling framework above, including
-hierarchical models, bivariate models, multinomial models, and Bayesian
-models, among others (see Savolainen et al., 2011 for a review of
-methods).
-
-\hypertarget{strategy-1-opponent-related-factors}{%
-\subsection{Strategy 1: opponent-related
-factors}\label{strategy-1-opponent-related-factors}}
-
-When opponent-related variables are included, the function is augmented
-as follows:
-
-\begin{equation}
-\label{eq:latent-function-with-opponent-variables}
-y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
-\end{equation}
-
-The additional part collects \(r=1,\cdots,R\) attributes \(o\) related
-to individual \(j\) in traffic unit \(v\) and crash \(k\) that opposed
-individual \(i\) in traffic unit \(t\) and crash \(k\). These could be
-individual characteristics of the opponent (such as age and gender)
-and/or characteristics of the opposing vehicle (such as vehicle type or
-weight). To qualify as an opponent, individual \(j\) must have been a
-participant in crash \(k\) but operating traffic unit \(v\ne t\).
-Sometimes the person \emph{is} the traffic unit, as is the case of a
-pedestrian. And we exclude passengers of vehicles as opponents, since
-they do not operate the traffic unit. In case the opponent attributes
-include only characteristics of the traffic unit, the condition for the
-traffic unit to be an opponent is that it participated in crash \(k\)
-and was different from \(t\). After introducing this new set of terms,
-the latent function now consists of a total of \(Z=L+M+Q+R\) covariates
-and associated parameters.
-
-\hypertarget{strategy-2-hierarchical-model-specification}{%
-\subsection{Strategy 2: hierarchical model
-specification}\label{strategy-2-hierarchical-model-specification}}
-
-We can choose to conceptualize the event leading to the outcome as a
-hierarchical process. There are a few different ways of doing this. For
-example, the hierarchy could be based on individuals in traffic units.
-In this case, we can rewrite the latent function as follows:
-
-\begin{equation}
-\label{eq:latent-function-with-hierarchical-traffic-unit}
-y_{itk}^*=\sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
-\end{equation}
-
-The coefficients of the traffic unit nest the individual attributes as
-follows. For any given coefficient \(q\):
-
-\begin{equation}
-\label{eq:hierarchical-traffic-unit-coefficients}
-\beta_{m}=\sum_{l=1}^L\beta_{ml}p_{itkl} 
-\end{equation}
-
-Therefore, the corresponding term in the latent function becomes
-(assuming that \(p_{itk1} = 1\), i.e., it is a constant term):
-
-\begin{equation}
-\label{eq:hierarchical-traffic-unit-coefficients}
-\begin{array}{rcl}\
-\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL}\big)u_{tkm}\\ 
-&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm}
-\end{array}
-\end{equation}
-
-As an alternative, the nesting unit could be the interaction
-person-opponent, in which case the opponent-level attributes are nested
-in the following fashion:
-
-\begin{equation}
-\label{eq:latent-function-with-opponent-variables}
-y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
-\end{equation}
-
-\noindent with any person-level coefficient \(l\) that we wish to expand
-defined as follows:
-
-\begin{equation}
-\label{eq:hierarchical-traffic-unit-coefficients}
-\alpha_{l}=\sum_{r=1}^R\alpha_{lr}o_{jvkr}
-\end{equation}
-
-\noindent with the same conditions as before, that \(j\ne i\) is the
-operator of traffic unit \(v\ne t\). The corresponding term in the
-latent function is now (assuming that \(o_{jvk1}=1\), i.e., it is a
-constant term):
-
-\begin{equation}
-\label{eq:hierarchical-traffic-unit-coefficients}
-\begin{array}{rcl}\
-\alpha_{l}p_{itkl} &=& \big(\alpha_{l1} + \alpha_{l2}o_{jvk2} + \cdots + \alpha_{lR}o_{jvkR} \big)p_{itkl}\\
-&=& \alpha_{l1}p_{itkl} + \alpha_{l2}o_{jvk2}p_{itkl} + \cdots + \alpha_{lR}o_{jvkR}p_{itkl}
-\end{array} 
-\end{equation}
-
-Discerning readers will identify this model specification strategy as
-Casetti's expansion method (Casetti, 1972; Roorda et al., 2010). This is
-a deterministic strategy for modelling contextual effects which, when
-augmented with random components, becomes the well-known multi-level
-modelling method (Hedeker and Gibbons, 1994, more on this in Section
-\ref{sec:further-considerations}). It is worthwhile to note that
-higher-order hierarchical effects are possible; for instance, individual
-attributes nested within traffic units, which in turn are nested within
-collisions. We do not explore higher-level hierarchies further in the
-current paper.
-
-\hypertarget{strategy-3-sample-subsetting}{%
-\subsection{Strategy 3: sample
-subsetting}\label{strategy-3-sample-subsetting}}
-
-The third model specification strategy that we will consider is
-subsetting the sample. This is applicable in conjunction with any of the
-other strategies discussed above. In essence, we define the latent
-function with restrictions as follows. Consider a continuous variable,
-e.g., age of person, and imagine that we wish to concentrate the
-analysis on older adults (e.g., Dissanayake and Lu, 2002). The latent
-function is defined as desired (see above), however, the following
-restriction is applied to the sample:
-
-\begin{equation}
-\label{eq:sampling-age}
-\text{Age of individual } i \text{ in traffic unit } t \text{ in crash } k = 
-\begin{cases}
-\ge 65 & \text{use record } itk\\
-< 65 & \text{do not use record } itk
-\end{cases}
-\end{equation}
-
-Suppose instead that we are interested in crashes by or against a
-specific type of traffic unit (e.g., pedestrians, Amoh-Gyimah et al.,
-2017):
-
-\begin{equation}
-\label{eq:sampling-pedestrian}
-\text{Road user class of individual } i \text{ in traffic unit } t \text{ in crash } k = 
-\begin{cases}
-\text{Pedestrian} & \text{use record } itk\\
-\text{Not pedestrian} & \text{do not use record } itk
-\end{cases}
-\end{equation}
-
-\noindent or:
-
-\begin{equation}
-\label{eq:sampling-pedestrian-opponent}
-\text{Road user class of individual } j \text{ in traffic unit } v \text{ in crash } k = 
-\begin{cases}
-\text{Pedestrian} & \text{use record } jvk\\
-\text{Not pedestrian} & \text{do not use record } jvk
-\end{cases}
-\end{equation}
-
-More generally, for any variable \(x\) of interest:
-
-\begin{equation}
-\label{eq:sampling-general}
-x_{itk} = 
-\begin{cases}
-\text{Condition: TRUE} & \text{use record } itk\\
-\text{Condition: FALSE} & \text{do not use record } itk
-\end{cases}
-\end{equation}
-
-Several conditions can be imposed to subset the sample in any way that
-the analyst deems appropriate or suitable.
-
-\hypertarget{sec:application}{%
-\section{Setting for empirical assessment}\label{sec:application}}
-
-In this section we present the setting for the empirical assessment of
-the modelling strategies discussed in Section \ref{sec:methods}, namely
-matters related to data and model estimation.
-
-\begin{quote}
-\textbf{Note:} this paper presents reproducible research. The source
-file is an R Markdown document. All code and data necessary to reproduce
-the analysis are available from the following anonymous Drive folder:
-\end{quote}
-
-\begin{quote}
-https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV
-\end{quote}
-
-\begin{quote}
-The source files, code, and data will be publicly available in a GitHub
-repository upon acceptance of the paper for publication
-\end{quote}
-
-\hypertarget{sec:data}{%
-\subsection{Data for empirical assessment}\label{sec:data}}
-
-To assess the performance of the various modelling strategies we use
-data from Canada's National Collision Database (NCDB). This database
-contains all police-reported motor vehicle collisions on public roads in
-Canada. Data are originally collected by provinces and territories, and
-shared with the federal government, that proceeds to combine, track, and
-analyze them for reporting deaths, injuries, and collisions in Canada at
-the national level. The NCDB is provided by Transport Canada, the agency
-of the federal government of Canada in charge of transportation policies
-and programs, under the Open Government License - Canada version 2.0
-{[}https://open.canada.ca/en/open-government-licence-canada{]}.
-
-The NCDB is available from 1999. For the purpose of this paper, we use
-the data corresponding to 2017, which is the most recent year available
-as of this writing. Furthermore, for assessment we also use the data
-corresponding to 2016. Similar to databases in other jurisdictions (see
-Montella et al., 2013), the NCDB contains information pertaining to the
-collision, the traffic unit(s), and the person(s) involved in a crash.
-The definitions of variables in this database are shown in the Appendix
-at the end of this document, in Tables
-\ref{tab:ncdb-descriptives-collision},
-\ref{tab:ncdb-descriptives-vehicle}, and
-\ref{tab:ncdb-descriptives-person}. Notice that, compared to Table
-\ref{tab:variable-categories}, environmental-factors variables and
-road-factors variables are collected under a single variable class,
-namely collision-related, since they are unique for each crash.
-
-Data are organized by person; in other words, there is one record per
-participant in a collision, be they drivers, passengers, pedestrians,
-etc. The only variable directly available with respect to opponents in a
-collision is the number of vehicles involved (see models in Bogue et
-al., 2017). Therefore, the data needs to be processed to obtain
-attributes of opponents for each participant in a collision. The
-protocol to do this is described next.
-
-\hypertarget{data-preprocessing-initial-filter}{%
-\subsubsection{Data preprocessing: initial
-filter}\label{data-preprocessing-initial-filter}}
-
-We apply an initial filter, whereby we scan the database to remove all
-records that are not a person (including parked cars and other objects)
-or that are missing information (as is the case when a participant in a
-crash is a runaway vehicle). Next, records missing at least one of the
-next variables are removed: P\_USER (the road user class), P\_SEX (sex),
-P\_AGE (age), and P\_ISEV (individual-level severity of crash). This
-initial filter ensures that all records are complete from the
-perspective of key information for analysis.
-
-\hypertarget{data-preprocessing-filter-two-vehicle-collisions}{%
-\subsubsection{Data preprocessing: filter two-vehicle
-collisions}\label{data-preprocessing-filter-two-vehicle-collisions}}
-
-After the initial filter, the database is summarized to find the number
-of individual-level records that correspond to each collision (C\_CASE).
-At this point, there are 32,298 collisions, involving only one (known)
-participant, there are 46,483 collisions involving two participants,
-19,433 collisions with three participants, 8,250 collisions involving
-four participants, 3,783 collisions with five participants, 1,789
-collisions with six participants, and 1,491 collisions involving seven
-or more participants. These participants were possibly occupants in
-different vehicles or were otherwise their own traffic units.
-Accordingly, the sample includes 174,741 drivers, 61,403 passengers,
-10,798 pedestrians, 5,286 bicyclists, and 6,564 motorcyclists.
-
-The next step in our data preprocessing protocol is to remove all
-collisions that involve only one participant. This still leaves numerous
-cases where multiple participants could have been in a single vehicle,
-for instance in a collision against a stationary object. Therefore, we
-proceed to use the vehicle sequence number to find the number of
-vehicles involved in each collision. This reveals that there are 20,732
-collisions that involve only one vehicle but possibly multiple
-participants (i.e., driver and one or more passengers). In addition,
-there are 165,520 collisions involving two vehicles (and possibly
-multiple participants). Finally, there are 40,242 collisions with three
-or more vehicles.
-
-Once we have identified the number of vehicles in each collision, we
-proceed to select all cases that involve only two vehicles. The most
-common cases in two-vehicle collisions are those that include drivers
-(40,297 collisions; this is reflective of the prevalence of
-single-occupant vehicles). This is followed by cases with passengers
-(14,120 collisions), pedestrians (5,204 collisions), bicyclists (2,238
-collisions), and motorcyclists (1,016 collisions). The distribution of
-individuals per traffic unit is as follows: 80,382 individuals are coded
-as being in V\_ID = 1, 76,523 individuals are coded as being in V\_ID =
-2, and 7,932 individuals are coded as pedestrians. In addition, 683
-individuals are coded as being in vehicles 3 through 9, despite our
-earlier filter to retain only collisions with two vehicles. We therefore
-proceed to select only individuals assigned to vehicles 1 or 2, as well
-as pedestrians. As a result of this filter a number of cases with only
-one known participant need to be removed.
-
-\hypertarget{data-preprocessing-extract-opponent-information-and-join-to-individual-records}{%
-\subsubsection{Data preprocessing: extract opponent information and join
-to individual
-records}\label{data-preprocessing-extract-opponent-information-and-join-to-individual-records}}
-
-Up to this point, the goal of data preprocessing has been to obtain a
-complete, workable sample of individual records of participants in
-two-vehicle collisions. There are two possible cases for the collisions,
-depending on the traffic units involved: 1) vehicle vs vehicle
-collisions (``vehicle'' is all motorized vehicles, including
-motorcycles/mopeds, as well as bicycles); and 2) vehicle vs pedestrian
-collisions. To identify opponents in a collision, it is convenient to
-classify collisions by pedestrian involvement. In this way, we find that
-the database includes 16,636 collisions that are vehicle vs pedestrian
-(possibly multiple pedestrians), and 147,594 collisions that involve two
-vehicles. After splitting the database according to pedestrian
-involvement, we can now extract relevant information about participants
-in the collision. This involves renaming the person-level variables so
-that we can distinguish each individual by their role as an individual
-or an opponent in a given record. Notice that when working with
-individuals in vehicles, only the driver is considered a legitimate
-opponent in a collision (i.e., passengers are never considered
-opponents).
-
-Once the personal attributes of individuals that can be considered
-opponents in a given collision have been extracted, their information is
-joined to the individual records by means of the collision unique
-identifier. As a result of this process, a new set of variables are now
-available for analysis: the age, sex, and road user class of the
-opponent, as well as the type of the opposing vehicle. A summary of
-opponent interactions and outcomes can be found in Table
-\ref{tab:summary-opponent-information}. The information there shows that
-the most common type of opponent for drivers is other drivers, followed
-by pedestrians. The only opponents of pedestrians, on the other hand,
-are drivers. Bicyclists and motorcyclists, on the other hand, are mostly
-opposed by drivers, but occasionally by other road users as well. In
-terms of outcomes, we observe that virtually all fatalities occur when
-the opponent is a driver, and only very rarely when the opponent is a
-motorcyclist. Injuries are also more common when the opponent is a
-driver, whereas no injuries are relatively more frequent when the
-opponent is a pedestrian or a bicyclist.
-
-\begin{table}
-
-\caption{\label{tab:summary-opponent-information}\label{tab:summary-opponent-information}Summary of opponent interactions and outcomes by road user class}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lrrrrrrrlll}
-\toprule
-\multicolumn{1}{c}{} & \multicolumn{4}{c}{Road User Class of Opponent} & \multicolumn{3}{c}{Outcome} & \multicolumn{3}{c}{Proportion by Road User Class} \\
-\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11}
-Road User Class & Driver & Pedestrian & Bicyclist & Motorcyclist & No Injury & Injury & Fatality & No Injury & Injury & Fatality\\
-\midrule
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{All opponents}}\\
-\hspace{1em}Driver & 97582 & 7880 & 3799 & 2498 & 59180 & 52143 & 436 & 0.52953 & 0.46657 & 0.003901\\
-\hspace{1em}Passenger & 35359 & 1282 & 667 & 818 & 19308 & 18667 & 151 & 0.50643 & 0.48961 & 0.003961\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
-\hspace{1em}Bicyclist & 3799 & 1 & 0 & 40 & 49 & 3760 & 31 & 0.01276 & 0.97917 & 0.008073\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 2498 & 30 & 40 & 338 & 204 & 2598 & 104 & 0.07020 & 0.89401 & 0.035788\\
-\addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Opponent: Driver}}\\
-\hspace{1em}Driver & 97582 & 0 & 0 & 0 & 45493 & 51657 & 432 & 0.46620 & 0.52937 & 0.004427\\
-\rowcolor{gray!6}  \hspace{1em}Passenger & 35359 & 0 & 0 & 0 & 16672 & 18536 & 151 & 0.47151 & 0.52422 & 0.004270\\
-Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist & 3799 & 0 & 0 & 0 & 43 & 3725 & 31 & 0.01132 & 0.98052 & 0.008160\\
-\hspace{1em}Motorcyclist & 2498 & 0 & 0 & 0 & 98 & 2299 & 101 & 0.03923 & 0.92034 & 0.040432\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Opponent: Pedestrian}}\\
-\hspace{1em}Driver & 0 & 7880 & 0 & 0 & 7693 & 187 & 0 & 0.97627 & 0.02373 & 0.000000\\
-\hspace{1em}Passenger & 0 & 1282 & 0 & 0 & 1246 & 36 & 0 & 0.97192 & 0.02808 & 0.000000\\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
-\hspace{1em}Bicyclist & 0 & 1 & 0 & 0 & 0 & 1 & 0 & 0.00000 & 1.00000 & 0.000000\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 30 & 0 & 0 & 11 & 19 & 0 & 0.36667 & 0.63333 & 0.000000\\
-\addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Opponent: Bicyclist}}\\
-\hspace{1em}Driver & 0 & 0 & 3799 & 0 & 3706 & 93 & 0 & 0.97552 & 0.02448 & 0.000000\\
-\rowcolor{gray!6}  \hspace{1em}Passenger & 0 & 0 & 667 & 0 & 649 & 18 & 0 & 0.97301 & 0.02699 & 0.000000\\
-\hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{1} -\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & -\\
-\hspace{1em}Motorcyclist & 0 & 0 & 40 & 0 & 16 & 24 & 0 & 0.40000 & 0.60000 & 0.000000\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Opponent: Motorcyclist}}\\
-\hspace{1em}Driver & 0 & 0 & 0 & 2498 & 2288 & 206 & 4 & 0.91593 & 0.08247 & 0.001601\\
-\hspace{1em}Passenger & 0 & 0 & 0 & 818 & 741 & 77 & 0 & 0.90587 & 0.09413 & 0.000000\\
-\rowcolor{gray!6}  Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
-\hspace{1em}Bicyclist & 0 & 0 & 0 & 40 & 6 & 34 & 0 & 0.15000 & 0.85000 & 0.000000\\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 0 & 0 & 338 & 79 & 256 & 3 & 0.23373 & 0.75740 & 0.008876\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\hypertarget{sec:model-estimation}{%
-\subsection{Model estimation}\label{sec:model-estimation}}
-
-Before model estimation, the variables are prepared as follows. First,
-age is scaled from years to decades. Secondly, new variables are defined
-to describe the vehicle type. Three classes of vehicle types are
-considered: 1) light duty vehicles (which in Canada include passenger
-cars, passenger vans, light utility vehicles, and light duty pick up
-trucks); 2) light trucks (all other vehicles \(\le\) 4536 kg in gross
-vehicle weight rating); and heavy vehicles (all other vehicles \(\ge\)
-4536 in gross vehicle weight rating). Furthermore, this typology of
-vehicle is combined with the road user class of the individual to
-distinguish between drivers and passengers of light duty vehicles, light
-trucks, and heavy vehicles, in addition to pedestrians, bicyclists, and
-motorcyclists. This is done for both the individual and the opponent.
-Variable interactions are calculated to produce hierarchical variables.
-For example, for a hierarchical definition of traffic unit-level
-variables, age (and the square of age to account for possible
-non-monotonic effects) are interacted with gender, road user class, and
-vehicle type. For hierarchical opponent variables, age (and the square
-of age) are interacted with the age of opponent (and the corresponding
-square). The variables thus obtained are shown in Table
-\ref{tab:model-specification-summary}. As seen in the table, Models 1
-and 2 are single-level models, and the difference between them is that
-Model 2 includes opponent variables. Models 3 and 4, in contrast, are
-hierarchical models. Model 3 considers the hierarchy on the basis of the
-traffic unit, while Model 4 considers the hierarchy on the basis of the
-opponent.
-
-Models 1 through 4 are estimated using the full sample. As discussed
-above, a related modelling strategy is to subset the sample (e.g., Islam
-et al., 2014; Lee and Li, 2014; Torrao et al., 2014; Wu et al., 2014).
-In this case we subset by a combination of traffic unit type of the
-individual (i.e., light duty vehicle, light truck, heavy vehicle,
-pedestrian, bicyclist, and motorcyclist) and vehicle type of the
-opponent (i.e., light duty vehicle, light truck, heavy vehicle). This
-leads to an ensemble of eighteen models to be estimated using subsets of
-data (see Table \ref{tab:model-specification-summary}). By subsetting
-the sample, at least \emph{some} opponent effects are incorporated
-implicitly. Models 1 and 2 are re-estimated using this strategy,
-dropping variables as necessary whenever they become irrelevant (for
-instance, after filtering for pedestrians, no other traffic unit types
-are present in the subset of data). In addition to variables that are no
-longer relevant in some data subsets, it is important to note that when
-using some data subsets a few variables had to be occasionally dropped
-to avoid convergence issues. This tended to happen particularly with
-smaller subsets where some particular combination of attributes was rare
-as a result of subsampling (e.g., in 2017 there were few or no
-collisions that involved a motorcyclist and a heavy vehicle in a bridge,
-or overpass, or viaduct). The process of estimation carefully paid
-attention to convergence issues to ensure the validity of the models
-reported here.
-
-The results of model estimation are discussed in the following section.
-
-\begin{table}
-
-\caption{\label{tab:model-specification-summary}\label{tab:model-specification-summary}Summary of variables and model specification}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{llllll}
-\toprule
-Variable & Notes & Model 1 & Model 2 & Model 3 & Model 4\\
-\midrule
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Individual-level variables}}\\
-\hspace{1em}Age & In decades & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Age Squared &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Sex & Reference: Female & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Use of Safety Devices & 7 levels; Reference: No Safety Device & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Traffic unit-level variables}}\\
-\hspace{1em}Passenger & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\hspace{1em}Pedestrian & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\hspace{1em}Motorcyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Light Truck & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\hspace{1em}Heavy Vehicle & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Opponent variables}}\\
-\hspace{1em}Age of Opponent & In decades &  & $\checkmark$ & $\checkmark$ & \\
-\hspace{1em}Age of Opponent Squared &  &  & $\checkmark$ & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Sex of Opponent & Reference: Female &  & $\checkmark$ & $\checkmark$ & \\
-\hspace{1em}Opponent: Light Duty Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Opponent: Light Truck & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Opponent: Heavy Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Hierarchical traffic unit variables}}\\
-\hspace{1em}Light Truck Driver:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Light Truck Driver:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Driver:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Heavy Vehicle Driver:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Light Truck Passenger:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Light Truck Passenger:Age Squared: &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Passenger:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Heavy Vehicle Passenger:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Pedestrian:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Pedestrian:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Bicyclist:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Bicyclist:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \hspace{1em}Motorcyclist:Age &  &  &  & $\checkmark$ & \\
-\hspace{1em}Motorcyclist:Age Squared &  &  &  & $\checkmark$ & \\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Hierarchical opponent variables}}\\
-\hspace{1em}Age:Age of Opponent &  &  &  &  & $\checkmark$\\
-\hspace{1em}Age:Age of Female Opponent &  &  &  &  & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Age:Age of Male Opponent Squared &  &  &  &  & $\checkmark$\\
-\hspace{1em}Age:Age of Female Opponent Squared &  &  &  &  & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Age Squared:Age of Male Opponent &  &  &  &  & $\checkmark$\\
-\hspace{1em}Age Squared:Age of Female Opponent &  &  &  &  & $\checkmark$\\
-\rowcolor{gray!6}  \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Collision-level variables}}\\
-\hspace{1em}Crash Configuration & 19 levels; Reference: Hit a moving object & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Road Configuration & 12 levels; Reference: Non-intersection & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Weather & 9 levels; Reference: Clear and sunny & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Surface & 11 levels; Reference: Dry & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Road Alignment & 8 levels; Reference: Straight and level & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\hspace{1em}Traffic Controls & 19 levels; Reference: Operational traffic signals & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\rowcolor{gray!6}  \hspace{1em}Month & 12 levels; Reference: January & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\hypertarget{sec:assessment}{%
-\section{Model assessment}\label{sec:assessment}}
-
-In this section we report an in-depth examination of the performance of
-the models. We begin by inspecting the statistical goodness of fit of
-the models by means of Akaike's Information Criterion (\(AIC\)). Next,
-we use the models to conduct in-sample predictions (i.e.,
-\emph{nowcasting}), using the same sample that was used to estimate the
-models, and out-of-sample predictions, using the data set corresponding
-to the year 2016 (i.e., \emph{backcasting}). Predictions are commonly
-evaluated in two different in the literature. Some researchers analyze
-the outcome shares based on the predicted probabilities (e.g., Bogue et
-al., 2017; Shamsunnahar and Eluru, 2013). This is a form of aggregate
-forecasting. Other researchers, in contrast, evaluate the classes of
-outcomes based on the individual-level predictions (e.g., Tang et al.,
-2019; Torrao et al., 2014). This is a form of disaggregate forecasting.
-
-\hypertarget{sec:goodness-of-fit}{%
-\subsection{Goodness of fit of models}\label{sec:goodness-of-fit}}
-
-We begin our empirical assessment by examining the results of estimating
-the models described above. Tables \ref{tab:model-full-sample-summary}
-and \ref{tab:model-subsample-summary} collect some key summary
-statistics of the estimated models. Of interest is the goodness of fit
-of the models, which in the case is measured with Akaike's Information
-Criterion (\(AIC\)). This criterion is calculated as follows:
-
-\begin{equation}
-\label{eq:aic}
-AIC = 2Z - 2\ln{\hat{L}}
-\end{equation}
-
-\noindent where \(Z\) is the number of coefficients estimated by the
-model, and \(\hat{L}\) the maximized likelihood of the model. Since
-\(AIC\) penalizes the model fit by means of the number of coefficients,
-this criterion gives preference to more parsimonious models. The
-objective is to minimize the \(AIC\), and therefore smaller values of
-this criterion represent better model fits. Model comparison can be
-conducted using the relative likelihood. Suppose that we have two
-models, say Model \(a\) and Model \(b\), with \(AIC_{a} \le AIC_{b}\).
-The relative likelihood is calculated as:
-
-\begin{equation}
-\label{eq:relative-likelihood}
-e^{(AIC_a - AIC_b)/2}
-\end{equation}
-
-The relative likelihood is interpreted as the probability that Model
-\(b\) minimizes the information loss as well as Model \(a\).
-
-Turning our attention to the models estimated using the full sample
-(Table \ref{tab:model-full-sample-summary}), it is possible to see that,
-compared to the base (single-level) model without opponent variables
-(Model 1), there are large and significant improvements in goodness of
-fit to be gained by introducing opponent effects. However, the gains are
-not as large when hierarchical specifications are used, even when the
-number of additional coefficients that need to be estimated is not
-substantially larger (recall that the penalty per coefficient in \(AIC\)
-is 2). The best model according to this measure of goodness of fit is
-Model 2 (single-level with opponent effects), followed by Model 4
-(hierarchical opponent variables), Model 3 (hierarchical traffic unit
-variables with opponent effects), and finally Model 1 (single-level
-without opponent effects).
-
-It is important to note that the likelihood function of a model, and
-therefore the value of its \(AIC\), both depend on the size of the
-sample, which is why \(AIC\) is not comparable across models estimated
-with different sample sizes. For this reason, the full sample models
-cannot be compared directly to the models estimated with subsets of
-data. The models in the ensembles, however, can be compared to each
-other (Table \ref{tab:model-subsample-summary}). As seen in the table,
-introducing opponent variables leads to a better fit in the case of
-most, but not all models. The simplest model (single-level without
-opponent effects) is clearly the best fitting candidate in the case of
-bicycle vs light truck collisions, bicycle vs heavy vehicle collisions,
-motorcyclist vs light duty vehicle collisions, and motorcyclist vs heavy
-vehicle collisions. Model 1 is a statistical toss for best performance
-with two competing models in the case of pedestrian vs heavy vehicle
-collisions. The relative likelihood of Model 1 compared to Models 2 and
-3 in this case is 0.56, which means that these two models are 0.56 times
-as probable as Model 1 to minimize the information loss.
-
-Model 2 is the best fit in the case of light truck vs heavy vehicle
-collisions. This model is also tied for best fit with Model 2 in the
-case of pedestrian vs light duty vehicle and pedestrian vs light truck
-collisions, and is a statistical toss with Model 4 in the case of heavy
-vehicle vs heavy vehicle collisions (relative likelihood is 0.592).
-Model 3 is the best fit in the case of light duty vehicle vs light duty
-vehicle collisions and heavy vehicle vs light duty vehicle. Model 4 is
-the best fit in the case of light duty vehicle vs light truck
-collisions, light duty vehicle vs heavy vehicle collisions, light truck
-vs light duty vehicle collisions, heavy vehicle vs light truck
-collisions, and motorcycle vs light truck collisions. This model is a
-statistical toss with Model 2 in the case of light truck vs light truck
-collisions, with a relative likelihood of 0.791.
-
-These results give some preliminary ideas about the relative performance
-of the different modelling strategies. In the next subsection we delve
-more deeply into this question by examining the predictive performance
-of the various modelling strategies. The results up to this point
-indicate that different model specification strategies might work best
-when combined with subsampling strategies. For space reasons, from this
-point onwards, we will consider the ensembles of models for predictions
-and will not compare individual models within the ensembles; this we
-suggest is a matter for future research.
-
-\begin{table}
-
-\caption{\label{tab:model-full-sample-summary}\label{tab:model-full-sample-summary}Summary of model estimation results: Full sample models}
-\centering
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lccc}
-\toprule
-Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC\\
-\midrule
-\rowcolor{gray!6}  Model 1 & 164,511 & 102 & 195,215\\
-Model 2 & 164,511 & 108 & 178,943\\
-\rowcolor{gray!6}  Model 3 & 164,511 & 118 & 181,333\\
-Model 4 & 164,511 & 111 & 179,018\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{landscape}\begin{table}
-
-\caption{\label{tab:model-subsample-summary}\label{tab:model-summary}Summary of model estimation results: Data Subset Models}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lccccccccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} & \multicolumn{2}{c}{Model 4} \\
-\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
-Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC\\
-\midrule
-\rowcolor{gray!6}  Light duty vehicle vs light duty vehicle & 114,841 & 94 & 145,390 & 97 & 143,903 & 100 & 143,896 & 100 & 144,004\\
-Light duty vehicle vs light truck & 3,237 & 79 & 3,943 & 82 & 3,927 & 85 & 3,937 & 85 & 3,922\\
-\rowcolor{gray!6}  Light duty vehicle vs heavy vehicle & 5,013 & 88 & 5,895 & 91 & 5,878 & 94 & 5,888 & 94 & 5,864\\
-Light truck vs light duty vehicle & 3,121 & 79 & 3,885 & 82 & 3,877 & 85 & 3,881 & 85 & 3,875\\
-\rowcolor{gray!6}  Light truck vs light truck & 809 & 67 & 1,170 & 70 & 1,156 & 73 & 1,162 & 73 & 1,155\\
-\addlinespace
-Light truck vs heavy vehicle & 198 & 64 & 288 & 67 & 281 & 70 & 287 & 70 & 286\\
-\rowcolor{gray!6}  Heavy vehicle vs light duty vehicle & 4,726 & 79 & 4,326 & 84 & 4,283 & 86 & 4,268 & 87 & 4,287\\
-Heavy vehicle vs light truck & 180 & 64 & 225 & 65 & 205 & 67 & 207 & 66 & 187\\
-\rowcolor{gray!6}  Heavy vehicle vs heavy vehicle & 779 & 74 & 1,147 & 77 & 1,136 & 80 & 1,141 & 80 & 1,137\\
-Pedestrian vs light duty vehicle & 7,176 & 88 & 2,826 & 91 & 2,821 & 91 & 2,821 & 93 & 2,827\\
-\addlinespace
-\rowcolor{gray!6}  Pedestrian vs light truck & 328 & 62 & 202 & 65 & 200 & 65 & 200 & 68 & 206\\
-Pedestrian vs heavy vehicle & 376 & 64 & 409 & 67 & 410 & 67 & 410 & 70 & 417\\
-\rowcolor{gray!6}  Bicyclist vs light duty vehicle & 3,521 & 80 & 654 & 83 & 659 & 83 & 659 & 85 & 657\\
-Bicyclist vs light truck & 116 & 42 & 84 & 57 & 114 & 57 & 114 & 54 & 108\\
-\rowcolor{gray!6}  Bicyclist vs heavy vehicle & - & - & - & - & - & - & - & - & -\\
-\addlinespace
-Motorcyclist vs light duty vehicle & 2,298 & 78 & 1,367 & 81 & 1,373 & 81 & 1,373 & 84 & 1,373\\
-\rowcolor{gray!6}  Motorcyclist vs light truck & 127 & 56 & 153 & 59 & 153 & 59 & 153 & 47 & 94\\
-Motorcyclist vs heavy vehicle & 62 & 43 & 88 & 45 & 90 & 46 & 92 & 51 & 102\\
-\bottomrule
-\multicolumn{10}{l}{\textit{Note: }}\\
-\multicolumn{10}{l}{There are zero cases of Bicyclist vs heavy vehicle in the sample}\\
-\end{tabular}}
-\end{table}
-\end{landscape}
-
-\hypertarget{sec:outcomes-shares}{%
-\subsection{Outcome shares based on predicted
-probabilities}\label{sec:outcomes-shares}}
-
-In this, and the following section, \emph{backcasting} refers to the
-prediction of probabilities and classes of outcomes using the 2016 data
-set. When conducting backcasting, the data set is preprocessed in
-identical manner as the 2017 data set. In addition, the variables used
-in backcasting match exactly those in the models. This means that some
-variables were dropped when they were present in the 2016 data set but
-not in the models. This tended to happen in the case of relatively rare
-outcomes (e.g., in 2016, there was at least one collision between a
-heavy vehicle and a light duty vehicle in a school crossing zone; no
-such event was observed in 2017).
-
-The shares of each outcome are calculated as the sum of the estimated
-probabilities for each observation:
-
-\begin{equation}
-\label{eq:predicted-shares}
-\begin{array}{rcl}
-\hat{S}_{\text{PDO}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{PDO})\\
-\hat{S}_{\text{injury}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{injury})\\
-\hat{S}_{\text{fatality}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{fatality})
-\end{array}
-\end{equation}
-
-\noindent where \(\hat{P}(y_{itk}=h_w)\) is the estimated probability of
-outcome \(h_w\) for individual \(i\) in traffic unit \(t\) and crash
-\(k\). The estimated share of outcome \(h\) is \(\hat{S}_{h_w}\).
-
-The estimated shares can be used to assess the ability of the model to
-forecast for the population the total number of cases of each outcome. A
-summary statistic useful to evaluate the performance is the Average
-Percentage Error, or \(APE\) (see Bogue et al., 2017, p. 31), which is
-calculated for each outcome as follows:
-
-\begin{equation}
-\label{eq:APE}
-APE_{h_w} = \Bigg|\frac{\hat{S}_{h_w} - S_{h_w}}{S_{h_w}}\Bigg|\times 100
-\end{equation}
-
-The Weighted Average Percentage Error (\(WAPE\)) aggregates the \(APE\)
-as follows:
-
-\begin{equation}
-\label{eq:WAPE}
-WAPE = \frac{\sum_w^WAPE_{h_w}\times S_{h_w}}{\sum_w^WS_{h_w}}
-\end{equation}
-
-The results of this exercise are reported in Table
-\ref{tab:ape-results}. Of the four full-sample models (Models 1-4), the
-\(APE\) of Model 2 is lowest in the nowcasting exercise for every
-outcome, with the exception of Fatality, where Model 4 produces a
-considerably lower \(APE\). When the results are aggregated by means of
-the \(WAPE\), Model 2 gives marginally better results than Model 4. It
-is interesting to see that the four ensemble models have lower \(APE\)
-values across the board in the nowcasting exercise, and much better
-\(WAPE\) than the full sample models. However, once we turn to the
-results of the backcasting exercise, these results do not hold, and it
-is possible to see that the Average Percentage Errors of the ensemble
-worsen considerably, particularly in the case of Fatality. The Weighted
-Average Prediction Error of the ensemble models in the backcasting
-exercise is also worse than for any of the full sample models. Excellent
-in-sample predictions but mediocre out-of-sample predictions are often
-evidence of overfitting, as in the case of the ensemble models here.
-
-In terms of backcasting, full sample Model 1 is marginally better than
-full sample Models 2 and 3, and better than full sample Model 4. The
-reason for this is the lower \(APE\) of Model 1 when predicting Injury,
-the most frequent outcome. However, the performance of Model 1 with
-respect to Fatality (the least frequent outcome) is the worst of all
-models. Whereas Model 4 has the best performance predicting Fatality,
-its performance with respect to other classes of outcomes is less
-impressive. Model 3 does better than Model 2 with respect to Injury, but
-performs relatively poorly when backcasting Fatality. Overall, Model 2
-appears to be the most balanced, with good in-sample performance and
-competitive out-of-sample performance that is also balanced with respect
-to the various classes of outcomes.
-
-\begin{table}
-
-\caption{\label{tab:table-ape-results}\label{tab:ape-results}Predicted shares and average prediction errors (APE) by model (percentages)}
-\centering
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lrrrrrrrrrr}
-\toprule
-\multicolumn{1}{c}{} & \multicolumn{3}{c}{No Injury} & \multicolumn{3}{c}{Injury} & \multicolumn{3}{c}{Fatality} & \multicolumn{1}{c}{} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
-Model & Observed & Predicted & APE & Observed & Predicted & APE & Observed & Predicted & APE & WAPE\\
-\rowcolor{gray!15}
-\midrule
-\addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{In-sample (nowcasting using 2017 data set, i.e., estimation data set)}}\\
-\hspace{1em}Model 1 & 78886 & 79029.00 & 0.18 & 84675 & 84533.74 & 0.17 & 950 & 948.26 & 0.18 & 0.17\\
-\rowcolor{gray!15}
-\hspace{1em}Model 2 & 78886 & 78928.98 & 0.05 & 84675 & 84641.94 & 0.04 & 950 & 940.08 & 1.04 & 0.05\\
-\hspace{1em}Model 3 & 78886 & 79027.29 & 0.18 & 84675 & 84512.50 & 0.19 & 950 & 971.21 & 2.23 & 0.20\\
-\rowcolor{gray!15}
-\hspace{1em}Model 4 & 78886 & 78939.18 & 0.07 & 84675 & 84622.54 & 0.06 & 950 & 949.28 & 0.08 & 0.06\\
-\hspace{1em}Model 1 Ensemble & 62413 & 62402.78 & 0.02 & 83564 & 83573.58 & 0.01 & 931 & 931.64 & 0.07 & 0.01\\
-\rowcolor{gray!15}
-\hspace{1em}Model 2 Ensemble & 62417 & 62407.00 & 0.02 & 83595 & 83604.14 & 0.01 & 931 & 931.86 & 0.09 & 0.01\\
-\rowcolor{gray!15}
-\hspace{1em}Model 3 Ensemble & 62411 & 62401.23 & 0.02 & 83596 & 83604.71 & 0.01 & 933 & 934.06 & 0.11 & 0.01\\
-\hspace{1em}Model 4 Ensemble & 62405 & 62395.28 & 0.02 & 83578 & 83586.75 & 0.01 & 932 & 932.97 & 0.10 & 0.01\\
-\rowcolor{gray!15}
-\addlinespace[0.3em]
-\multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 data set)}}\\
-\hspace{1em}Model 1 & 96860 & 96364.67 & 0.51 & 101605 & 102002.59 & 0.39 & 1109 & 1206.74 & 8.81 & 0.50\\
-\rowcolor{gray!15}
-\hspace{1em}Model 2 & 96860 & 96361.41 & 0.51 & 101605 & 102112.08 & 0.50 & 1109 & 1100.51 & 0.77 & 0.51\\
-\hspace{1em}Model 3 & 96860 & 96354.01 & 0.52 & 101605 & 102086.18 & 0.47 & 1109 & 1133.82 & 2.24 & 0.51\\
-\rowcolor{gray!15}
-\hspace{1em}Model 4 & 96860 & 96325.85 & 0.55 & 101605 & 102136.72 & 0.52 & 1109 & 1111.43 & 0.22 & 0.54\\
-\hspace{1em}Model 1 Ensemble & 77457 & 76822.49 & 0.82 & 100013 & 100580.60 & 0.57 & 1072 & 1138.91 & 6.24 & 0.71\\
-\rowcolor{gray!15}
-\hspace{1em}Model 2 Ensemble & 77459 & 76799.11 & 0.85 & 100049 & 100630.48 & 0.58 & 1071 & 1149.41 & 7.32 & 0.74\\
-\hspace{1em}Model 3 Ensemble & 77459 & 76786.76 & 0.87 & 100050 & 100644.29 & 0.59 & 1072 & 1149.95 & 7.27 & 0.75\\
-\rowcolor{gray!15}
-Model 4 Ensemble & 77461 & 76766.08 & 0.90 & 100029 & 100630.21 & 0.60 & 1070 & 1163.71 & 8.76 & 0.78\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{sec:outcomes-classes}{%
-\subsection{Outcome frequency based on predicted
-classes}\label{sec:outcomes-classes}}
-
-\(APE\) and \(WAPE\) are summary measures of the performance of models
-at the aggregated level. Aggregate-level predictions (i.e., shares of
-outcomes) are of interest from a population health perspective. In other
-cases, an analyst might be interested in the predicted outcomes at the
-individual level. In this section we examine the frequency of outcomes
-based on predicted classes, using the same two settings as above:
-nowcasting and backcasting.
-
-The individual-level outcomes are examined using an extensive array of
-verification statistics. Verification statistics are widely used in the
-evaluation of predictive approaches were the outcomes are categorical,
-and are often based on the analysis of \emph{confusion matrices} (e.g.,
-Provost and Kohavi, 1998; Beguería, 2006). Confusion matrices are
-cross-tabulations of \emph{observed} and \emph{predicted} classes. In a
-two-by-two confusion matrix there are four possible combinations of
-observed to predicted classes: hits, misses, false alarms, and correct
-non-events, as shown in Table \ref{tab:confusion-matrix}. When the
-outcome has more than two classes, the confusion matrix is converted to
-a two-by-two table to calculate verification statistics.
-
-\begin{table}[!h]
-
-\caption{\label{tab:example-confusion-matrix}\label{tab:confusion-matrix}Example of a two-by-two confusion matrix}
-\centering
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Observed} & \multicolumn{1}{c}{ } \\
-\cmidrule(l{3pt}r{3pt}){2-3}
-Predicted & Yes & No & Marginal Total\\
-\midrule
-\rowcolor{gray!6}  Yes & Hit & False Alarm & Predicted Yes\\
-No & Miss & Correct Non-event & Predicted No\\
-\rowcolor{gray!6}  Marginal Total & Observed Yes & Observed No & \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The statistics used in our assessment are summarized in Table
-\ref{tab:verification-statistics}, including brief descriptions of their
-interpretation (more on this below). The statistics evaluate different
-aspects of the performance of a model. Some are concerned with the
-ability of the model to be right. For example, Percent Correct (\(PC\))
-is the sum of hits and correct rejections divided by the number of
-cases. Others are concerned with the ability of the model to match the
-observed outcomes, including Bias (\(B\)), which measures whether a
-class has been over- or under-predicted. Whereas others measure the
-ability of the model to not be wrong, such as the probability of false
-detection (\(F\)) and the False Alarm Ration (\(FAR\)). We discuss the
-results of calculating the battery of verification statistics, first for
-the nowcasting case (are shown in (Table
-\ref{tab:nowcast-outcomes-assessment}) and subsequently for the
-backcasting case (Table \ref{tab:backcast-outcomes-assessment}).
-
-\begin{landscape}\begin{table}
-
-\caption{\label{tab:table-verification-statistics}\label{tab:verification-statistics}Verification statistics}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{30em}>{\raggedright\arraybackslash}p{30em}}
-\toprule
-Statistic & Description & Notes\\
-\midrule
-\rowcolor{gray!6}  Percent Correct ($PC$) & Total hits and correct rejections divided by number of cases & Strongly influenced by most common category\\
-Percent Correct by Class ($PC_c$) & Same as Percent Correct but by category & Strongly influenced by most common category\\
-\rowcolor{gray!6}  Bias ($B$) & Total predicted by category, divided by total observed by category & $B>1$: class is overpredicted; $B<1$: class is underpredicted\\
-Critical Success Index ($CSI$) & Total hits divided by total hits + false alarms + misses & $CSI = 1$: perfect score; $CSI = 0$: no skill\\
-\rowcolor{gray!6}  Probability of False Detection ($F$) & Proportion of no events forecast as yes; sensitive to false alarms but ignores misses & $F = 0$: perfect score\\
-\addlinespace
-Probability of Detection ($POD$) & Total hits divided by total observed by class & $POD = 1$: perfect score\\
-\rowcolor{gray!6}  False Alarm Ratio ($FAR$) & Total false alarms divided by total forecast yes by class; measures fraction of predicted yes that did not occur & $FAR = 0$: perfect score\\
-Heidke Skill Score ($HSS$) & Fraction of correct predictions after removing predictions attributable to chance; measures fractional improvement over random; tends to reward conservative forecasts & $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better\\
-\rowcolor{gray!6}  Peirce Skill Score ($PSS$) & Combines $POD$ and $F$; measures ability to separate yes events from no events; tends to reward conservative forecasts & $PSS = 1$: perfect score; $PSS = 0$: no skill\\
-Gerrity Score ($GS$) & Measures accuracy of predicting the correct category, relative to random; tends to reward correct forecasts of less likely category & $GS = 1$: perfect score; $GS = 0$: no skill\\
-\bottomrule
-\end{tabular}}
-\end{table}
-\end{landscape}
-
-\hypertarget{nowcasting-verification-statistics}{%
-\subsubsection{Nowcasting: verification
-statistics}\label{nowcasting-verification-statistics}}
-
-At first glance, the results of the verification statistics (Table
-\ref{tab:nowcast-outcomes-assessment}) make it clear that no model under
-comparison is consistently a top performer from every aspect of
-prediction. Recalling Box's aphorism, all models are wrong but some are
-useful - in this case it just so happens that some models are more wrong
-than others in subtly different ways. That said, it is noticeable that
-the worst scores across the board tend to accrue to Model 1 in its full
-sample and ensemble versions. On the other hand, Model 2 (full sample)
-concentrates most of the best scores and second best scores of all the
-models, but also some of the worst scores for Fatality. Model 4, in
-contrast, has most of the second best scores and a few top scores, but
-not a single worst score.
-
-Of all the models, Model 2 (full sample) performs best in terms of
-Percent Correct, followed by Model 4 (full sample). The worst performer
-from this perspective is Model 1 (full sample), with a \(PC\) score
-several percentage points below the top models. The second score is
-Percent Correct by Class (\(PC_c\)). This score is calculated
-individually for each outcome class. Model 2 (full sample), has the best
-performance for outcomes No Injury and Injury, and the second best score
-for Fatality. Model 4 (full sample) has the best score for Fatality, and
-is second best for No Injury and Injury. Model 1 (full sample) has worst
-scores for No Injury and Injury whereas its ensemble version has the
-worst score for Fatality. It is important to note that \(PC\) and
-\(PC_c\) are heavily influenced by the most common category, something
-that can be particularly appreciated in the scores for Fatality. The
-scores for this class of outcome are generally high, despite the fact
-that the number of hits are relatively low; the high values of \(PC_c\)
-in this case are due to the high occurrence of correct rejections
-elsewhere in the table.
-
-Bias (\(B\)) measures for each outcome class the proportion of total
-predictions by class (e.g., hits as well as false alarms) relative to
-the total number of cases observed for that class. For this reason, it
-is possible for predictions to have low bias (values closer to 1) but
-still do poorly in terms of hits. The models with the best performance
-in terms of \(B\) are Model 1 (full sample) for No Injury and Injury,
-and Model 3 (ensemble) for Fatality. Model 4 (full sample) is the second
-best performer for No Injury and Injury, and Model 4 (ensemble) is
-second best performer for Fatality. Model 1 (ensemble) has the worst
-bias for No Injury and Injury, whereas Model 2 (full sample) has the
-worst bias for Fatality.
-
-Critical Success Index (\(CSI\)) evaluates forecasting skill while
-assuming that the number of correct non-events is inconsequential skill.
-Accordingly, the statistic is calculated as the proportion of hits
-relative to the sum of hits plus false alarms plus misses. No model
-performs uniformly best from this perspective. Model 2 (full sample) has
-the best \(CSI\) for No Injury, Model 2 (ensemble) has the best score
-for Injury, and Model 4 (ensemble) the best score for Fatality. On the
-other hand, Model 1 (ensemble) has the worst score for No Injury, Model
-1 (full sample) the worst score for Injury, and Model 2 the worst score
-for Fatality. These scores indicates that the models are not
-particularly skilled at predicting the corresponding classes correctly,
-given the frequency with which they give false alarms or miss the class.
-
-The next statistic is Probability of False Detection (\(F\)). This is
-the proportion of false alarms relative to the total number of cases
-that the event is not observed. This statistic measure the frequency
-with which the model incorrectly predicts an event, but not when it
-incorrectly misses it. The lowest probability of false detection in the
-case of No Injury is 19.17\% for Model 2 (ensemble), with every other
-model having values lower than 21\%, with the exception of Model 1 (full
-sample) that has a score of 26.46\%. With respect to Injury, the lowest
-probabilities range 35.8\% and 35.29\% and 35.35\% for Models 4 (full
-sample) and 2 (full sample). In contrast, the highest probability of
-false detection for Injury is 45.12\% for Model 1 (ensemble). The scores
-for \(F\) for Fatality are all extremely low as a consequence of the
-very low frequency of this class of outcome in the sample.
-
-The Probability of Detection (\(POD\)) measures the frequency with which
-the model correctly predicts a class, relative to the number of cases of
-that class. As with some other verification statistics, no model is
-consistently a best performer. Model 4 (full sample) has the highest
-probability of detection for No Injury (65.38\%), followed by Model 2
-(65.32\%), whereas the worst probability of detection is by Model 1
-(ensemble) with a score of 55.54\%. In terms of Injury, all models have
-\(POD\) higher than 79\%, and the highest score is 80.67\% for Model 2
-(ensemble). The exception is Model 1 (full sample), which has a
-considerably lower \(POD\) of Injury with a score of 73.36\%. Lastly, in
-terms of Fatalities, all models have very low probabilities of
-detection, ranging from a high of 4.94\% in the case of Model 4
-(ensemble) to a worst score of 0.11\% in the case of Model 2 (full
-sample).
-
-The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
-that were false alarms, which means that lower scores are better. Model
-2 (full sample) has the best \(FAR\) statistic for No Injury, as only
-25.05\% of predictions for this class are false alarms. The next best
-score is by Model 4 (full sample), with only 25.23\% of No Injury
-predictions being false alarms. The worst performance in this class is
-by Model 1 (ensemble), which produces almost a third of false alarms in
-its predictions of No Injury. In the case of Injury, the False Alarm
-Ratio ranges from a low of 29.15\% by Model (ensemble), with every other
-model having scores lower than 30\% except Model 1 (full sample), that
-gives almost 32\% of false alarms. In terms of Fatality, the lowest
-\(FAR\) is also for Model 4 (ensemble) with only 17.86\% of false
-alarms, whereas the worst performance is by Model 2 (full sample), which
-produces over 95\% of false alarms.
-
-The next three verification statistics are skill scores that
-simultaneously consider different aspects of prediction, and are
-therefore overall indicators of prediction skill. Heidke's Skill Score
-(\(HSS\)) is the fraction of correct predictions above those that could
-be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
-Probability of Detection (\(POD\)) of a model and its Probability of
-False Detection (\(F\)) to measure the skill of a model to discriminate
-the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
-the model's skill predicting the correct classes that tends to reward
-correct forecasts of the least frequent class. These scores help to
-remove some of the ambiguity regarding the overall performance of a
-model. In this way, we know that Model 2 (full sample) does not do
-particularly well with the class Fatality - however, of all models, it
-tends to have the best overall performance. Its \(HSS\), for example,
-suggests that it achieves 44.74\% of correct predictions after removing
-correct predictions attributable to chance. In contrast, the lowest
-score is for Model 1 (ensemble), which only achieves 36.26\% correct
-predictions after removing those attributable to chance. Model 2 (full
-sample) also has the highest \(PSS\) and the second highest \(GS\).
-Model 4 (full sample) has the highest \(GS\) and the second highest
-\(HSS\) and \(PSS\).
-
-\begin{table}
-
-\caption{\label{tab:nowcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 data set, i.e., estimation data set)}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{lrrrllllllllll}
-\toprule
-\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
-Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 50652 & 22503 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62121 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
-
-Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.55}} & \textcolor{black}{99.39} & \textcolor{black}{0.0589} & \textcolor{black}{0.003} & \textcolor{red}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9464} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3725}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3696}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1902}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51531 & 17136 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67515 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
-
-Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{72.36}}} & \textcolor{black}{\underline{99.41}} & \textcolor{red}{0.0263} & \textcolor{red}{0.001} & \textcolor{black}{1e-04} & \textcolor{red}{0.0011} & \textcolor{red}{0.96} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4474}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4429}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2265}}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51102 & 17297 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
-
-\hspace{1em}Injury & 27784 & 67337 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7952} & \textcolor{black}{0.2985} &  &  & \\
-
-Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4356}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 51575 & 17318 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
-
-\hspace{1em}Injury & 27311 & 67334 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.1279}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
-
-Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{72.28}}} & \textcolor{black}{\textbf{99.41}} & \textcolor{black}{0.0274} & \textcolor{black}{0.0031} & \textcolor{black}{1e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.8846} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4458}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4414}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2268}}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 34664 & 16434 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1952} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\
-
-\hspace{1em}Injury & 27749 & 67120 & 829 &  & \textcolor{black}{69.35} & \textcolor{red}{1.1452} & \textcolor{black}{0.5985} & \textcolor{red}{0.4512} & \textcolor{black}{0.8032} & \textcolor{black}{0.2986} &  &  & \\
-
-Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.31}} & \textcolor{red}{99.39} & \textcolor{black}{0.0526} & \textcolor{black}{0.0414} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0419} & \textcolor{black}{\underline{0.2041}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3626}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3521}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.201}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35443 & 16146 & 60 &  & \textcolor{black}{70.61} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
-
-\hspace{1em}Injury & 26974 & 67436 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
-
-Fatality & 0 & 13 & 42 &  & \textcolor{black}{99.39} & \textcolor{black}{0.0591} & \textcolor{black}{0.0445} & \textcolor{black}{1e-04} & \textcolor{black}{0.0451} & \textcolor{black}{0.2364} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3784}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3678}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2106}}\\
-
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 35498 & 16203 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
-
-\hspace{1em}Injury & 26913 & 67380 & 828 &  & \textcolor{black}{70.09} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
-
-Fatality & 0 & 13 & 45 & \multirow{-6}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 35553 & 16297 & 59 &  & \textcolor{black}{70.59} & \textcolor{black}{0.8318} & \textcolor{black}{0.4514} & \textcolor{black}{0.1935} & \textcolor{black}{0.5697} & \textcolor{black}{0.3151} &  &  & \\
-
-\hspace{1em}Injury & 26852 & 67271 & 827 &  & \textcolor{black}{70.06} & \textcolor{black}{1.1361} & \textcolor{black}{0.6046} & \textcolor{black}{0.437} & \textcolor{black}{0.8049} & \textcolor{black}{\textbf{0.2915}} &  &  & \\
-
-Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.02}} & \textcolor{black}{99.39} & \textcolor{black}{\underline{0.0601}} & \textcolor{black}{\textbf{0.0488}} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{\textbf{0.0494}} & \textcolor{black}{\textbf{0.1786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3783}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3679}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2127}}\\
-\bottomrule
-\multicolumn{14}{l}{\textit{Note: }}\\
-\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
-\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
-\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
-\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
-\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
-\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
-\end{tabular}}
-\end{table}
-
-\hypertarget{backcasting-verification-statistics}{%
-\subsubsection{Backcasting: verification
-statistics}\label{backcasting-verification-statistics}}
-
-Table \ref{tab:backcast-outcomes-assessment} presents the results of the
-verification exercise for the case of our out-of-sample predictions
-(i.e., backasting). Qualitatively, the results are similar to those of
-the nowcasting experiments, but with a somewhat weaker performance of
-the ensemble models. This, again, supports the idea that these models
-might be overfitting the process, as discussed in reference to the
-aggregate forecasts (see Section \ref{sec:outcomes-shares}). Models 2
-(full sample) and 4 (full sample) are again identified as the best
-overall performers, and particularly Model 2 (full sample) performs
-somewhat more adroitly with respect to Fatality in backcasting than it
-did in nowcasting.
-
-\begin{table}
-
-\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 data set)}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{7}{9}\selectfont
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrrrllllllllll}
-\toprule
-\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
-Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 61683 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
-
-\hspace{1em}Injury & 35172 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3684} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
-
-Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.03}} & \textcolor{black}{99.4} & \textcolor{black}{0.0902} & \textcolor{black}{0.0083} & \textcolor{black}{5e-04} & \textcolor{black}{0.009} & \textcolor{red}{0.9} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3628}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3604}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1882}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 62736 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5318}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
-
-\hspace{1em}Injury & 34124 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
-
-Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4373}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4336}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 62249 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.862} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
-
-\hspace{1em}Injury & 34609 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
-
-Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{71.5}} & \textcolor{black}{99.43} & \textcolor{black}{0.0424} & \textcolor{red}{0.0052} & \textcolor{black}{2e-04} & \textcolor{red}{0.0054} & \textcolor{black}{0.8723} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.431}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4274}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2205}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 62788 & 21248 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8687}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2079} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80330 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
-
-Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4354}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
-\hspace{1em}No Injury & 42896 & 20230 & 95 &  & \textcolor{black}{69.26} & \textcolor{red}{0.8162} & \textcolor{red}{0.4387} & \textcolor{black}{0.2011} & \textcolor{red}{0.5538} & \textcolor{red}{0.3215} &  &  & \\
-
-\hspace{1em}Injury & 34546 & 79692 & 962 &  & \textcolor{black}{68.73} & \textcolor{red}{1.1519} & \textcolor{black}{0.588} & \textcolor{red}{0.4522} & \textcolor{black}{0.7968} & \textcolor{black}{0.3082} &  &  & \\
-
-Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{68.67}} & \textcolor{black}{99.35} & \textcolor{black}{0.1129} & \textcolor{black}{0.0127} & \textcolor{black}{6e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.876} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3539}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3447}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1831}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 43486 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.4461} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
-
-\hspace{1em}Injury & 33953 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
-
-Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3551}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 43526 & 20034 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8218} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
-
-\hspace{1em}Injury & 33915 & 79914 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7987}} & \textcolor{black}{0.3038} &  &  & \\
-
-Fatality & 18 & 102 & 16 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.13}} & \textcolor{black}{99.34} & \textcolor{black}{0.1269} & \textcolor{black}{\underline{0.0134}} & \textcolor{black}{7e-04} & \textcolor{black}{\underline{0.0149}} & \textcolor{black}{0.8824} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3639}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3546}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1885}}\\
-\cmidrule{1-14}
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 43560 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5623} & \textcolor{black}{0.3174} &  &  & \\
-
-\hspace{1em}Injury & 33876 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
-
-Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.07}} & \textcolor{red}{99.34} & \textcolor{black}{\textbf{0.1393}} & \textcolor{black}{\textbf{0.0141}} & \textcolor{red}{7e-04} & \textcolor{black}{\textbf{0.0159}} & \textcolor{black}{0.8859} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3629}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3538}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1886}}\\
-\bottomrule
-\multicolumn{14}{l}{\textit{Note: }}\\
-\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
-\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
-\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
-\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
-\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
-\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
-\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
-\end{tabular}}}
-\end{table}
-
-\hypertarget{sec:further-considerations}{%
-\section{Further considerations}\label{sec:further-considerations}}
-
-As discussed in Section \ref{sec:methods}, there is a rich selection of
-modelling approaches that are applicable to crash severity analysis.
-Based on the literature, we limited our empirical assessment of
-modelling strategies to only one model, namely the ordinal probit. On
-the other hand, since the modelling strategies discussed here all relate
-to the specification of the latent function and data subsetting, it is a
-relatively simple matter to extend them to other modelling approaches.
-For example, take Expression
-\ref{eq:latent-function-with-opponent-variables} and add a random
-component \(\mu_{k}\) as follows:
-
-\begin{equation}
-\label{eq:latent-function-with-opponent-variables-and-random-component}
-y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \mu_{k} + \epsilon_{itk}
-\end{equation}
-
-The addition of the random component in this fashion would help to
-capture, when appropriate, unobserved heterogeneity at the level of the
-crash (this is similar to the random intercepts approach in multi-level
-modelling; also see Mannering et al., 2016). As a second example, take
-Expressions \ref{eq:latent-function-with-hierarchical-traffic-unit} to
-\ref{eq:hierarchical-traffic-unit-coefficients} and add a random
-component to a hierarchical coefficient, to obtain:
-
-\begin{equation}
-\label{eq:hierarchical-traffic-unit-coefficients-with-random-component}
-\begin{array}{rcl}\
-\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL} + \mu_{mk}\big)u_{tkm}\\ 
-&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm} + \mu_{mk}u_{tkm}
-\end{array}
-\end{equation}
-
-This is similar to the random slopes strategy in multi-level modelling.
-
-We do not report results regarding other modelling strategies. On the
-one hand, more sophisticated modelling frameworks are generally capable
-of improving the performance of a model. On the other hand, there are
-well-known challenges in the estimation of more sophisticated models
-(see Lenguerrand et al., 2006, p. 47, for a discussion of convergence
-issues in models with mixed effects; Mannering et al., 2016, p. 13, for
-some considerations regarding the complexity and cost of estimating more
-complex models; and Bogue et al., 2017, p. 27, on computational demands
-of models with random components). The additional cost and complexity of
-more sophisticated modelling approaches would, in our view, have greatly
-complicated our empirical assessment, particularly considering the large
-size of the sample involved in this research (a data set with over
-164,000 records in the case of the full sample models). That said, we
-experimented with a model with random components using monthly subsets
-of data to find that, indeed, estimation takes considerably longer, is
-more demanding in terms of fixing potential estimation quirks, and in
-the end resulted in variance components that could not be reliably
-estimated as different from zero (results can be consulted in the source
-R Notebook). For this reason, we choose to leave the application of more
-sophisticated models as a matter for future research.
-
-\hypertarget{sec:concluding-remarks}{%
-\section{Concluding remarks}\label{sec:concluding-remarks}}
-
-The study of crash severity is an important component of accident
-research, as seen from a large and vibrant literature and numerous
-applications. Part of this literature covers different modelling
-strategies that can be used to model complex hierarchical, multievent
-outcomes such as the severity of injuries following a collision. In this
-paper, our objective has been to assess the performance of different
-strategies to model opponent effects in two-vehicle crashes. In broad
-terms, three strategies were considered: 1) incorporating opponent-level
-variables in the model; 2) single- versus multi-level model
-specifications; and 3) sample subsetting and estimation of separate
-models for different types of individual-opponent interactions. The
-empirical evaluation was based on data from Canada's National Crash
-Database and the application of ordered probit models. A suite of models
-that implemented the various strategies considered was estimated using
-data from 2017. We then assessed the performance of the models using one
-information criterion (AIC). Furthermore, the predictive performance of
-the models was assessed in terms of both nowcasting (in-sample
-predictions) and backcasting (out-of-sample predictions), the latter
-using data from 2016.
-
-The results of the empirical assessment strongly suggest that
-incorporating opponent effects can greatly improve the goodness-of-fit
-and predictive performance of a model. Two modelling strategies appear
-to outperform the rest: a relatively simple single-level modelling
-approach that incorporates opponent effects, and a hierarchical
-modelling approach with nested opponent effects. There was some evidence
-that subsetting the sample can improve the results in some cases s(e.g.,
-when modelling the severity of crashes involving active travelers or
-motorcyclists), but possibly at the risk of overfitting the process and
-leading to worse out-of-sample performance. In this paper we did not
-compare individual models in our ensemble approach, but we suggest that
-this is an avenue for future research.
-
-The results of this research should be informative to analysts
-interested in crashes involving two parties, since it provides some
-useful guidelines regarding the specification of opponent effects. Not
-only do opponent effects improve the goodness of fit and performance of
-models, they also add rich insights into their effects. The focus on
-this paper was on performance, and for space reasons it is not possible
-to include an examination of the best model without failing to do it
-justice. We plan to report the results of the best-fitting model in a
-future paper.
-
-The analysis also opens up a few avenues for future research. First, for
-reasons discussed in Section \ref{sec:further-considerations}, we did
-not consider more sophisticated modelling approaches, such as models
-with random components, partial proportional odds, ranked ordered
-models, or multinomial models, to mention just a few possibilities.
-Secondly, we only considered the performance of the models when making
-predictions for the full sample. That is, the submodels in the ensembles
-were not compared in detail, just their aggregate results when
-predicting the full sample. However, the goodness-of-fit was not
-uniformly better for any one modelling strategy when the data were
-subset, and it is possible that individual models perform better for a
-certain subset than competitors that are part of a better ensemble,
-overall. For this reason, we suggest that additional work with ensemble
-approaches is warranted. Finally, it is clear that the models do not
-generally do well when predicting the least frequent class of outcome,
-namely Fatality. It would be worthwhile to further investigate
-approaches for so-called imbalanced learning, a task where Torrao et
-al.~(2014) have already made some headway.
-
-Finally, as an aside, this paper is, to the best of our knowledge, the
-first example of reproducible research in crash severity analysis. By
-providing the data \emph{and} code for the analysis, it is our hope that
-this will allow other researchers to easily verify the results, and to
-extend them. A common practice in the machine learning community is to
-use canonical data set to demonstrate the performance of new techniques.
-Sharing code and data has remained relatively rare in transportation
-research, and we would like to suggest that the data sets used in this
-research could constitute one such canonical data set for future
-methodological developments.
-
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-To be added in the final version of the paper.
-
-\hypertarget{appendix}{%
-\section*{Appendix}\label{appendix}}
-\addcontentsline{toc}{section}{Appendix}
-
-Variable definitions in Canada's National Collision Database.
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-collision}\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  C\_CASE & Unique collision identifier & Unique identifier for collisions\\
-C\_YEAR & Year & Last two digits of year.\\
-\rowcolor{gray!6}  C\_MNTH & Month & 14 levels: January - December; unknown; not reported by jurisdiction.\\
-C\_WDAY & Day of week & 9 levels: Monday - Sunday; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_HOUR & Collision hour & 25 levels: hourly intervals; unknown; not reported by jurisdiction.\\
-\addlinespace
-C\_SEV & Collision severity & 4 levels: collision producing at least one fatality; collision producing non-fatal injury; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_VEHS & Number of vehicles involved in collision & Number of vehicles: 1-98 vehicles involved; 99 or more vehicles involved; unknown; not reported by jurisdiction.\\
-C\_CONF & Collision configuration & 21 levels: SINGLE VEHICLE: Hit a moving object (e.g.
-a person or an animal); Hit a stationary object (e.g.
-a tree); Ran off left shoulder; Ran off right shoulder; Rollover on roadway; Any other single vehicle collision configuration; TWO-VEHICLES SAME DIRECTION OF TRAVEL: Rear-end collision; Side swipe; One vehicle passing to the left of the other, or left turn conflict; One vehicle passing to the right of the other, or right turn conflict; Any other two vehicle - same direction of travel configuration; TWO-VEHICLES DIFFERENT DIRECTION OF TRAVEL: Head-on collision; Approaching side-swipe; Left turn across opposing traffic; Right turn, including turning conflicts; Right angle collision; Any other two-vehicle - different direction of travel configuration; TWO-VEHICLES, HIT A PARKED VEHICLE: Hit a parked motor vehicle; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_RCFG & Roadway configuration & 15 levels: Non-intersection; At an intersection of at least two public roadways; Intersection with parking lot entrance/exit, private driveway or laneway; Railroad level crossing; Bridge, overpass, viaduct; Tunnel or underpass; Passing or climbing lane; Ramp; Traffic circle; Express lane of a freeway system; Collector lane of a freeway system; Transfer lane of a freeway system; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-C\_WTHR & Weather condition & 10 levels: Clear and sunny; Overcast, cloudy but no precipitation; Raining; Snowing, not including drifting snow; Freezing rain, sleet, hail; Visibility limitation; Strong wind; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\addlinespace
-\rowcolor{gray!6}  C\_RSUR & Road surface & 12 levels: Dry, normal; Wet; Snow (fresh, loose snow); Slush, wet snow; Icy, packed snow; Debris on road (e.g., sand/gravel/dirt); Muddy; Oil; Flooded; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-C\_RALN & Road alignment & 9 levels: Straight and level; Straight with gradient; Curved and level; Curved with gradient; Top of hill or gradient; Bottom of hill or gradient; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
-\rowcolor{gray!6}  C\_TRAF & Traffic control & 21 levels: Traffic signals fully operational; Traffic signals in flashing mode; Stop sign; Yield sign; Warning sign; Pedestrian crosswalk; Police officer; School guard, flagman; School crossing; Reduced speed zone; No passing zone sign; Markings on the road; School bus stopped with school bus signal lights flashing; School bus stopped with school bus signal lights not flashing; Railway crossing with signals, or signals and gates; Railway crossing with signs only; Control device not specified; No control present; Choice is other than the preceding values; unknown; not reported by jurisdiction.\\*
-\end{longtable}
-\endgroup{}
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-vehicle}\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  V\_ID & Vehicle sequence number & Number of vehicles: 1-98; Pedestrian sequence number: 99; unknown.\\
-V\_TYPE & Vehicle type & 21 levels: Light Duty Vehicle (Passenger car, Passenger van, Light utility vehicles and light duty pick up trucks); Panel/cargo van (<= 4536 KG GVWR   Panel or window type of van designed primarily for carrying goods); Other trucks and vans (<= 4536 KG GVWR); Unit trucks (> 4536 KG GVWR); Road tractor; School bus; Smaller school bus (< 25 passengers); Urban and Intercity Bus; Motorcycle and moped; Off road vehicles; Bicycle; Purpose-built motorhome; Farm equipment; Construction equipment; Fire engine; Snowmobile; Street car; Data element is not applicable  (e.g.
-dummy vehicle record created for pedestrian); Choice is other than the preceding values; unknown; not reported by jurisdiction.\\
-\rowcolor{gray!6}  V\_YEAR & Vehicle model year & Model year; dummy for pedestrians; unknown; not reported by jurisdiction.\\*
-\end{longtable}
-\endgroup{}
-
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
-\caption{\label{tab:ncdb-descriptives-person}\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endfirsthead
-\caption[]{\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables \textit{(continued)}}\\
-\toprule
-Variable & Description & Notes\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\multicolumn{3}{l}{\textit{Note: }}\\
-\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
-\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
-\endlastfoot
-\rowcolor{gray!6}  P\_ID & Person sequence number & Sequence number: 1-99; Not applicable (dummy for parked vehicles); not reported by jurisdiction.\\
-P\_SEX & Person sex & 5 levels: Male; Female; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_AGE & Person age & Age: less than 1 year; 1-98 years old; 99 years or older; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
-P\_PSN & Person position & Person position: Driver; Passenger front row, center; Passenger front row, right outboard (including motorcycle passenger in sidecar); Passenger second row, left outboard, including motorcycle passenger; Passenger second row, center; Passenger second row, right outboard; Passenger third row, left outboard;...; Position unknown, but the person was definitely an occupant; Sitting on someone’s lap; Outside passenger compartment; Pedestrian; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_ISEV & Medical treatment required & 6 levels: No Injury; Injury; Fatality; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\addlinespace
-P\_SAFE & Safety device used & 11 levels: No safety device used; Safety device used; Helmet worn; Reflective clothing worn; Both helmet and reflective clothing used; Other safety device used; No safety device equipped   (e.g.
-buses); Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
-\rowcolor{gray!6}  P\_USER & Road user class & 6 levels: Motor Vehicle Driver; Motor Vehicle Passenger; Pedestrian; Bicyclist; Motorcyclist; Not stated/Other/Unknown.\\*
-\end{longtable}
-\endgroup{}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Amoh2017effect}{}%
-Amoh-Gyimah, R., Aidoo, E.N., Akaateba, M.A., Appiah, S.K., 2017. The
-effect of natural and built environmental characteristics on
-pedestrian-vehicle crash severity in ghana. International Journal of
-Injury Control and Safety Promotion 24, 459--468.
-doi:\href{https://doi.org/10.1080/17457300.2016.1232274}{10.1080/17457300.2016.1232274}
-
-\leavevmode\hypertarget{ref-Aziz2013exploring}{}%
-Aziz, H.M.A., Ukkusuri, S.V., Hasan, S., 2013. Exploring the
-determinants of pedestrian-vehicle crash severity in new york city.
-Accident Analysis and Prevention 50, 1298--1309.
-doi:\href{https://doi.org/10.1016/j.aap.2012.09.034}{10.1016/j.aap.2012.09.034}
-
-\leavevmode\hypertarget{ref-Begueria2006validation}{}%
-Beguería, S., 2006. Validation and evaluation of predictive models in
-hazard assessment and risk management. Natural Hazards 37, 315--329.
-doi:\href{https://doi.org/10.1007/s11069-005-5182-6}{10.1007/s11069-005-5182-6}
-
-\leavevmode\hypertarget{ref-Bogue2017modified}{}%
-Bogue, S., Paleti, R., Balan, L., 2017. A modified rank ordered logit
-model to analyze injury severity of occupants in multivehicle crashes.
-Analytic Methods in Accident Research 14, 22--40.
-doi:\href{https://doi.org/10.1016/j.amar.2017.03.001}{10.1016/j.amar.2017.03.001}
-
-\leavevmode\hypertarget{ref-Casetti1972generating}{}%
-Casetti, E., 1972. Generating models by the expansion method:
-Applications to geographic research. Geographical Analysis 4, 81--91.
-
-\leavevmode\hypertarget{ref-Chang2006analysis}{}%
-Chang, L.Y., Wang, H.W., 2006. Analysis of traffic injury severity: An
-application of non-parametric classification tree techniques. Accident
-Analysis and Prevention 38, 1019--1027.
-doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}
-
-\leavevmode\hypertarget{ref-Chen2019investigation}{}%
-Chen, F., Song, M.T., Ma, X.X., 2019. Investigation on the injury
-severity of drivers in rear-end collisions between cars using a random
-parameters bivariate ordered probit model. International Journal of
-Environmental Research and Public Health 16.
-doi:\href{https://doi.org/10.3390/ijerph16142632}{10.3390/ijerph16142632}
-
-\leavevmode\hypertarget{ref-Chiou2013modeling}{}%
-Chiou, Y.C., Hwang, C.C., Chang, C.C., Fu, C., 2013. Modeling
-two-vehicle crash severity by a bivariate generalized ordered probit
-approach. Accident Analysis and Prevention 51, 175--184.
-doi:\href{https://doi.org/10.1016/j.aap.2012.11.008}{10.1016/j.aap.2012.11.008}
-
-\leavevmode\hypertarget{ref-Dissanayake2002factors}{}%
-Dissanayake, S., Lu, J.J., 2002. Factors influential in making an injury
-severity difference to older drivers involved in fixed object-passenger
-car crashes. Accident Analysis and Prevention 34, 609--618.
-doi:\href{https://doi.org/10.1016/s0001-4575(01)00060-4}{10.1016/s0001-4575(01)00060-4}
-
-\leavevmode\hypertarget{ref-Duddu2018modeling}{}%
-Duddu, V.R., Penmetsa, P., Pulugurtha, S.S., 2018. Modeling and
-comparing injury severity of at-fault and not at-fault drivers in
-crashes. Accident Analysis and Prevention 120, 55--63.
-doi:\href{https://doi.org/10.1016/j.aap.2018.07.036}{10.1016/j.aap.2018.07.036}
-
-\leavevmode\hypertarget{ref-Effati2015geospatial}{}%
-Effati, M., Thill, J.C., Shabani, S., 2015. Geospatial and machine
-learning techniques for wicked social science problems: Analysis of
-crash severity on a regional highway corridor. Journal of Geographical
-Systems 17, 107--135.
-doi:\href{https://doi.org/10.1007/s10109-015-0210-x}{10.1007/s10109-015-0210-x}
-
-\leavevmode\hypertarget{ref-Gong2017modeling}{}%
-Gong, L.F., Fan, W.D., 2017. Modeling single-vehicle run-off-road crash
-severity in rural areas: Accounting for unobserved heterogeneity and age
-difference. Accident Analysis and Prevention 101, 124--134.
-doi:\href{https://doi.org/10.1016/j.aap.2017.02.014}{10.1016/j.aap.2017.02.014}
-
-\leavevmode\hypertarget{ref-Haleem2013effect}{}%
-Haleem, K., Gan, A., 2013. Effect of driver's age and side of impact on
-crash severity along urban freeways: A mixed logit approach. Journal of
-Safety Research 46, 67--76.
-doi:\href{https://doi.org/10.1016/j.jsr.2013.04.002}{10.1016/j.jsr.2013.04.002}
-
-\leavevmode\hypertarget{ref-Hanson2013severity}{}%
-Hanson, C.S., Noland, R.B., Brown, C., 2013. The severity of pedestrian
-crashes: An analysis using google street view imagery. Journal of
-Transport Geography 33, 42--53.
-doi:\href{https://doi.org/10.1016/j.jtrangeo.2013.09.002}{10.1016/j.jtrangeo.2013.09.002}
-
-\leavevmode\hypertarget{ref-Hedeker1994random}{}%
-Hedeker, D., Gibbons, R.D., 1994. A random-effects ordinal
-regression-model for multilevel analysis. Biometrics 50, 933--944.
-
-\leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
-Iranitalab, A., Khattak, A., 2017. Comparison of four statistical and
-machine learning methods for crash severity prediction. Accident
-Analysis and Prevention 108, 27--36.
-doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}
-
-\leavevmode\hypertarget{ref-Islam2014comprehensive}{}%
-Islam, S., Jones, S.L., Dye, D., 2014. Comprehensive analysis of single-
-and multi-vehicle large truck at-fault crashes on rural and urban
-roadways in alabama. Accident Analysis \& Prevention 67, 148--158.
-doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2014.02.014}{https://doi.org/10.1016/j.aap.2014.02.014}
-
-\leavevmode\hypertarget{ref-Khan2015exploring}{}%
-Khan, G., Bill, A.R., Noyce, D.A., 2015. Exploring the feasibility of
-classification trees versus ordinal discrete choice models for analyzing
-crash severity. Transportation Research Part C-Emerging Technologies 50,
-86--96.
-doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}
-
-\leavevmode\hypertarget{ref-Kim2013driver}{}%
-Kim, J.K., Ulfarsson, G.F., Kim, S., Shankar, V.N., 2013. Driver-injury
-severity in single-vehicle crashes in california: A mixed logit analysis
-of heterogeneity due to age and gender. Accident Analysis and Prevention
-50, 1073--1081.
-doi:\href{https://doi.org/10.1016/j.aap.2012.08.011}{10.1016/j.aap.2012.08.011}
-
-\leavevmode\hypertarget{ref-Kim1995personal}{}%
-Kim, K., Nitz, L., Richardson, J., Li, L., 1995. PERSONAL and behavioral
-predictors of automobile crash and injury severity. Accident Analysis
-and Prevention 27, 469--481.
-doi:\href{https://doi.org/10.1016/0001-4575(95)00001-g}{10.1016/0001-4575(95)00001-g}
-
-\leavevmode\hypertarget{ref-Lee2014analysis}{}%
-Lee, C., Li, X.C., 2014. Analysis of injury severity of drivers involved
-in single- and two-vehicle crashes on highways in ontario. Accident
-Analysis and Prevention 71, 286--295.
-doi:\href{https://doi.org/10.1016/j.aap.2014.06.008}{10.1016/j.aap.2014.06.008}
-
-\leavevmode\hypertarget{ref-Lenguerrand2006modelling}{}%
-Lenguerrand, E., Martin, J.L., Laumon, B., 2006. Modelling the
-hierarchical structure of road crash data - application to severity
-analysis. Accident Analysis and Prevention 38, 43--53.
-doi:\href{https://doi.org/10.1016/j.aap.2005.06.021}{10.1016/j.aap.2005.06.021}
-
-\leavevmode\hypertarget{ref-Li2017interplay}{}%
-Li, L., Hasnine, M.S., Habib, K.M.N., Persaud, B., Shalaby, A., 2017.
-Investigating the interplay between the attributes of at-fault and
-not-at-fault drivers and the associated impacts on crash injury
-occurrence and severity level. Journal of Transportation Safety \&
-Security 9, 439--456.
-doi:\href{https://doi.org/10.1080/19439962.2016.1237602}{10.1080/19439962.2016.1237602}
-
-\leavevmode\hypertarget{ref-Ma2008multivariate}{}%
-Ma, J.M., Kockelman, K.M., Damien, P., 2008. A multivariate
-poisson-lognormal regression model for prediction of crash counts by
-severity, using bayesian methods. Accident Analysis and Prevention 40,
-964--975.
-doi:\href{https://doi.org/10.1016/j.aap.2007.11.002}{10.1016/j.aap.2007.11.002}
-
-\leavevmode\hypertarget{ref-Maddala1986limited}{}%
-Maddala, G.S., 1986. Limited-dependent and qualitative variables in
-econometrics. Cambridge university press.
-
-\leavevmode\hypertarget{ref-Mannering2016unobserved}{}%
-Mannering, F., Shankar, V., Bhat, C.R., 2016. Unobserved heterogeneity
-and the statistical analysis of highway accident data. Analytic Methods
-in Accident Research 11, 1--16.
-doi:\href{https://doi.org/10.1016/j.amar.2016.04.001}{10.1016/j.amar.2016.04.001}
-
-\leavevmode\hypertarget{ref-McArthur2014spatial}{}%
-McArthur, A., Savolainen, P.T., Gates, T.J., 2014. Spatial analysis of
-child pedestrian and bicycle crashes development of safety performance
-function for areas adjacent to schools. Transportation Research Record
-57--63. doi:\href{https://doi.org/10.3141/2465-08}{10.3141/2465-08}
-
-\leavevmode\hypertarget{ref-Montella2013crash}{}%
-Montella, A., Andreassen, D., Tarko, A.P., Turner, S., Mauriello, F.,
-Imbriani, L.L., Romero, M.A., 2013. Crash databases in australasia, the
-european union, and the united states review and prospects for
-improvement. Transportation Research Record 128--136.
-doi:\href{https://doi.org/10.3141/2386-15}{10.3141/2386-15}
-
-\leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
-Mooradian, J., Ivan, J.N., Ravishanker, N., Hu, S., 2013. Analysis of
-driver and passenger crash injury severity using partial proportional
-odds models. Accident Analysis and Prevention 58, 53--58.
-doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}
-
-\leavevmode\hypertarget{ref-Mussone2017analysis}{}%
-Mussone, L., Bassani, M., Masci, P., 2017. Analysis of factors affecting
-the severity of crashes in urban road intersections. Accident Analysis
-and Prevention 103, 112--122.
-doi:\href{https://doi.org/10.1016/j.aap.2017.04.007}{10.1016/j.aap.2017.04.007}
-
-\leavevmode\hypertarget{ref-Obeng2011gender}{}%
-Obeng, K., 2011. Gender differences in injury severity risks in crashes
-at signalized intersections. Accident Analysis and Prevention 43,
-1521--1531.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.004}{10.1016/j.aap.2011.03.004}
-
-\leavevmode\hypertarget{ref-Osman2018injury}{}%
-Osman, M., Mishra, S., Paleti, R., 2018. Injury severity analysis of
-commercially-licensed drivers in single-vehicle crashes: Accounting for
-unobserved heterogeneity and age group differences. Accident Analysis
-and Prevention 118, 289--300.
-doi:\href{https://doi.org/10.1016/j.aap.2018.05.004}{10.1016/j.aap.2018.05.004}
-
-\leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
-Peek-Asa, C., Britton, C., Young, T., Pawlovich, M., Falb, S., 2010.
-Teenage driver crash incidence and factors influencing crash injury by
-rurality. Journal of Safety Research 41, 487--492.
-doi:\href{https://doi.org/10.1016/j.jsr.2010.10.002}{10.1016/j.jsr.2010.10.002}
-
-\leavevmode\hypertarget{ref-Penmetsa2017examining}{}%
-Penmetsa, P., Pulugurtha, S.S., Duddu, V.R., 2017. Examining injury
-severity of not-at-fault drivers in two-vehicle crashes. Transportation
-Research Record 164--173.
-doi:\href{https://doi.org/10.3141/2659-18}{10.3141/2659-18}
-
-\leavevmode\hypertarget{ref-Provost1998glossary}{}%
-Provost, F., Kohavi, R., 1998. Glossary of terms. Journal of Machine
-Learning 30, 271--274.
-
-\leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
-Rakotonirainy, A., Steinhardt, D., Delhomme, P., Darvell, M., Schramm,
-A., 2012. Older drivers' crashes in queensland, australia. Accident
-Analysis and Prevention 48, 423--429.
-doi:\href{https://doi.org/10.1016/j.aap.2012.02.016}{10.1016/j.aap.2012.02.016}
-
-\leavevmode\hypertarget{ref-Rana2010copula}{}%
-Rana, T.A., Sikder, S., Pinjari, A.R., 2010. Copula-based method for
-addressing endogeneity in models of severity of traffic crash injuries
-application to two-vehicle crashes. Transportation Research Record
-75--87. doi:\href{https://doi.org/10.3141/2147-10}{10.3141/2147-10}
-
-\leavevmode\hypertarget{ref-Regev2018crash}{}%
-Regev, S., Rolison, J.J., Moutari, S., 2018. Crash risk by driver age,
-gender, and time of day using a new exposure methodology. Journal of
-Safety Research 66, 131--140.
-doi:\href{https://doi.org/10.1016/j.jsr.2018.07.002}{10.1016/j.jsr.2018.07.002}
-
-\leavevmode\hypertarget{ref-Rifaat2007accident}{}%
-Rifaat, S.M., Chin, H.C., 2007. Accident severity analysis using ordered
-probit model. Journal of Advanced Transportation 41, 91--114.
-doi:\href{https://doi.org/10.1002/atr.5670410107}{10.1002/atr.5670410107}
-
-\leavevmode\hypertarget{ref-Roorda2010trip}{}%
-Roorda, M.J., Paez, A., Morency, C., Mercado, R., Farber, S., 2010. Trip
-generation of vulnerable populations in three canadian cities: A spatial
-ordered probit approach. Transportation 37, 525--548.
-doi:\href{https://doi.org/10.1007/s11116-010-9263-3}{10.1007/s11116-010-9263-3}
-
-\leavevmode\hypertarget{ref-Salon2018determinants}{}%
-Salon, D., McIntyre, A., 2018. Determinants of pedestrian and bicyclist
-crash severity by party at fault in san francisco, ca. Accident Analysis
-and Prevention 110, 149--160.
-doi:\href{https://doi.org/10.1016/j.aap.2017.11.007}{10.1016/j.aap.2017.11.007}
-
-\leavevmode\hypertarget{ref-Sasidharan2014partial}{}%
-Sasidharan, L., Menendez, M., 2014. Partial proportional odds model-an
-alternate choice for analyzing pedestrian crash injury severities.
-Accident Analysis and Prevention 72, 330--340.
-doi:\href{https://doi.org/10.1016/j.aap.2014.07.025}{10.1016/j.aap.2014.07.025}
-
-\leavevmode\hypertarget{ref-Savolainen2007probabilistic}{}%
-Savolainen, P., Mannering, F., 2007. Probabilistic models of
-motorcyclists' injury severities in single- and multi-vehicle crashes.
-Accident Analysis and Prevention 39, 955--963.
-doi:\href{https://doi.org/10.1016/j.aap.2006.12.016}{10.1016/j.aap.2006.12.016}
-
-\leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
-Savolainen, P.T., Mannering, F., Lord, D., Quddus, M.A., 2011. The
-statistical analysis of highway crash-injury severities: A review and
-assessment of methodological alternatives. Accident Analysis and
-Prevention 43, 1666--1676.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.025}{10.1016/j.aap.2011.03.025}
-
-\leavevmode\hypertarget{ref-Shaheed2013mixed}{}%
-Shaheed, M.S.B., Gkritza, K., Zhang, W., Hans, Z., 2013. A mixed logit
-analysis of two-vehicle crash seventies involving a motorcycle. Accident
-Analysis and Prevention 61, 119--128.
-doi:\href{https://doi.org/10.1016/j.aap.2013.05.028}{10.1016/j.aap.2013.05.028}
-
-\leavevmode\hypertarget{ref-Shamsunnahar2013evaluating}{}%
-Shamsunnahar, Y., Eluru, N., 2013. Evaluating alternate discrete outcome
-frameworks for modeling crash injury severity. Accident Analysis \&
-Prevention 59, 506--521.
-doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
-
-\leavevmode\hypertarget{ref-Shamsunnahar2014examining}{}%
-Shamsunnahar, Y., Eluru, N., Pinjari, A.R., Tay, R., 2014. Examining
-driver injury severity in two vehicle crashes - a copula based approach.
-Accident Analysis and Prevention 66, 120--135.
-doi:\href{https://doi.org/10.1016/j.aap.2014.01.018}{10.1016/j.aap.2014.01.018}
-
-\leavevmode\hypertarget{ref-Tang2019crash}{}%
-Tang, J.J., Liang, J., Han, C.Y., Li, Z.B., Huang, H.L., 2019. Crash
-injury severity analysis using a two-layer stacking framework. Accident
-Analysis and Prevention 122, 226--238.
-doi:\href{https://doi.org/10.1016/j.aap.2018.10.016}{10.1016/j.aap.2018.10.016}
-
-\leavevmode\hypertarget{ref-Tay2011multinomial}{}%
-Tay, R., Choi, J., Kattan, L., Khan, A., 2011. A multinomial logit model
-of pedestrian-vehicle crash severity. International Journal of
-Sustainable Transportation 5, 233--249.
-doi:\href{https://doi.org/10.1080/15568318.2010.497547}{10.1080/15568318.2010.497547}
-
-\leavevmode\hypertarget{ref-Thompson2018trends}{}%
-Thompson, J.P., Baldock, M.R.J., Dutschke, J.K., 2018. Trends in the
-crash involvement of older drivers in australia. Accident Analysis and
-Prevention 117, 262--269.
-doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
-
-\leavevmode\hypertarget{ref-Torrao2014modeling}{}%
-Torrao, G.A., Coelho, M.C., Rouphail, N.M., 2014. Modeling the impact of
-subject and opponent vehicles on crash severity in two-vehicle
-collisions. Transportation Research Record 53--64.
-doi:\href{https://doi.org/10.3141/2432-07}{10.3141/2432-07}
-
-\leavevmode\hypertarget{ref-Train2009discrete}{}%
-Train, K., 2009. Discrete choice methods with simulation, 2nd Edition.
-ed. Cambridge University Press, Cambridge.
-
-\leavevmode\hypertarget{ref-Wang2015copula}{}%
-Wang, K., Yasmin, S., Konduri, K.C., Eluru, N., Ivan, J.N., 2015.
-Copula-based joint model of injury severity and vehicle damage in
-two-vehicle crashes. Transportation Research Record 158--166.
-doi:\href{https://doi.org/10.3141/2514-17}{10.3141/2514-17}
-
-\leavevmode\hypertarget{ref-Wang2005use}{}%
-Wang, X.K., Kockelman, K.M., 2005. Use of heteroscedastic ordered logit
-model to study severity of occupant injury - distinguishing effects of
-vehicle weight and type, in: Statistical Methods; Highway Safety Data,
-Analysis, and Evaluation; Occupant Protection; Systematic Reviews and
-Meta-Analysis, Transportation Research Record. pp. 195--204.
-
-\leavevmode\hypertarget{ref-White1972effects}{}%
-White, S., Clayton, S., 1972. Some effects of alcohol, age of driver,
-and estimated speed on the likelihood of driver injury. Accident
-Analysis \& Prevention 4.
-
-\leavevmode\hypertarget{ref-WHO2019global}{}%
-World Health Organization, 2019. Global status report on road safety
-2018 (2018). Geneva.
-
-\leavevmode\hypertarget{ref-Wu2014mixed}{}%
-Wu, Q., Chen, F., Zhang, G.H., Liu, X.Y.C., Wang, H., Bogus, S.M., 2014.
-Mixed logit model-based driver injury severity investigations in single-
-and multi-vehicle crashes on rural two-lane highways. Accident Analysis
-and Prevention 72, 105--115.
-doi:\href{https://doi.org/10.1016/j.aap.2014.06.014}{10.1016/j.aap.2014.06.014}
-
 \end{document}

### 2020-05-10 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v2.tex
First round of edits to respond to reviewers after decision on original submission.

--- 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
@@
 
-  \journal{Some Journal} % Sets Journal name
+  \journal{Accident Analysis and Prevention} % Sets Journal name
 
@@
 \linenumbers
+\usepackage{setspace}
+\doublespacing
 \usepackage{booktabs}
@@
 
-Evidence from across the world suggests that the burden of road
-accidents is not borne evenly. There are important disparities at the
-international level, where the odds of death due to road crashes are
-three times higher in low-income countries compared to high-income
-countries (World Health Organization, 2019). In the case of high-income
-countries, where substantial gains in road safety have been observed for
-years, said gains have also been unevenly distributed; thus, while fatal
-crashes involving older adults in the United States and Great Britain
-declined between 1997 and 2010 (despite the graying of the population),
-the trend remained stable or increased slightly in Australia in roughly
-the same period (Thompson et al., 2018). The impact of road accidents
-also varies considerably. For example, in a study in the United States,
-Obeng (2011) reported that the factors affecting crash severity varied
-substantially in magnitude by gender. More recently, Regev et al.~(2018)
-used adjusted crash risk to find that the risk of fatal injuries for
-older drivers remained high even for minor collisions - which highlights
-the perils of accidents at older ages. Other studies have concentrated
-on the consequences of road accidents for the young (e.g., Peek-Asa et
-al., 2010), the old (e.g., Rakotonirainy et al., 2012), as well as
-pedestrians and cyclists (e.g., Hanson et al., 2013; McArthur et al.,
-2014).
-
-Given the relevance and cost of this matter, as well as the unevenness
-of impacts among different population segments, numerous efforts have
-been conducted to better understand the factors that affect road safety
-- including the probable consequences of crashes. Along these lines, a
-stream of research in the analysis of road accidents is concerned with
-the severity of crashes. In particular, multivariate analysis of crash
-severity is a useful way to clarify the way various factors can affect
-the outcome of an incident, and to discriminate between various levels
-of injury. This is an active area of research, and one where
+Modelling the severity of injuries to victims of road crashes has been a
+preoccupation of transportation researchers, planners, auto insurance
+companies, governments, and the general public for decades. One of the
+earliest studies to investigate the severity of injuries conditional on
+a collision having occurred was by White and Clayton (1972). Kim et
+al.~(1995) later stated that the ``linkages between severity of injury
+and driver characteristics and behaviors have not been thoroughly
+investigated'' (p.~470). Nowadays, there is a burgeoning literature on
+this subject, which often relies on multivariate analysis of crash
+severity to clarify the way various factors can affect the outcome of an
+incident, and to discriminate between various levels of injury.
+
+Crash severity is an active area of research, and one where
 methodological developments have aimed at improving the reliability,
 accuracy, and precision of models (e.g., Savolainen et al., 2011; Bogue
-et al., 2017; Shamsunnahar and Eluru, 2013). Of interest in the
-literature on crash severity is the way different parties in a crash
-interact to influence the severity of individual outcomes. The
-importance of these interactions has been recognized in the existing
-literature (e.g., Chiou et al., 2013; Lee and Li, 2014; Li et al., 2017;
-Torrao et al., 2014), and a number of different modelling strategies
-have been proposed. These include the use of \emph{data subsetting},
-\emph{opponent variables}, and estimating models with \emph{hierarchical
-components}. A gap in the literature, however, is a systematic
-assessment of the performance of these various strategies.
-
-The objective of this paper is to empirically evaluate different
-strategies to model opponent effects in crash severity, in the context
-of incidents involving two parties. Therefore, we present a systematic
-assessment of several relevant modelling strategies. For the assessment
-we use data from Canada's National Collision Database, a database that
-collects all police-reported collisions in the country. Using the most
-recent version of the data set (2017). Three levels of crash severity
-(no injury/injury/fatality) are analyzed using ordered logit models, and
-covariates for the participants in the crash and the conditions of the
-crash. For model assessment, we conduct an in-sample prediction exercise
-using the estimation sample (i.e., \emph{nowcasting}), and also an
-out-of-sample prediction exercise using the data set corresponding to
-2016 (i.e., \emph{backcasting}). The models are assessed using predicted
-shares and predicted classes of outcomes at the individual level, using
-an extensive array of verification statistics.
+et al., 2017; Yasmin and Eluru, 2013). Of interest in this literature is
+how different parties in a crash interact to influence the severity of
+individual outcomes. The importance of these interactions has been
+recognized by numerous authors (e.g., Chiou et al., 2013; Lee and Li,
+2014; Li et al., 2017; Torrao et al., 2014). Lee and Li (2014) for
+instance, note that ``{[}for{]} two-vehicle crashes, most studies only
+considered the effects of one vehicle on driver's injury severity or the
+highest injury severity of two drivers. However, it is expected that
+driver's injury severity is not only affected by characteristics of
+his/her own vehicle, but also characteristics of a partner vehicle.''
+More generally, the severity of the outcome depends, at least in part,
+on the characteristics of the parties, and a crash between two heavy
+vehicles is likely to have very different outcomes compared to crash
+where a heavy vehicle hits a pedestrian.
+
+For the purpose of this paper, we define a party as one or more
+individuals travelling in a traffic unit that becomes involved in a
+crash. Sometimes the traffic unit is a vehicle, and the party is a
+single individual (i.e.~a single occupant vehicle); but in other cases,
+a party could consist of several individuals (i.e., a driver and one or
+more passengers). Other times, the individual \emph{is} the traffic
+unit, for instance a pedestrian or a bicyclist. An opponent is a
+different party that is involved in the same collision. In the
+literature, interactions between parties in a collision are treated by
+means of different strategies, including the use of \emph{data
+subsetting}, \emph{opponent variables}, and estimating models with
+\emph{hierarchical components}. These strategies are not new, however, a
+systematic comparison between them is missing from the literature. For
+this reason, the objective of this paper is to empirically evaluate
+different strategies to model party interactions in crash severity in
+the context of incidents involving two parties. More concretely, this
+research aims to:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Systematize the analysis of party interactions in crash severity
+  models. Although every strategy considered here has been used in past
+  research, in this paper they are organized in a way that clarifies
+  their operation.
+\item
+  Present a data management workflow to prepare a dataset to implement
+  analysis of opponent effects.
+\item
+  Provide evidence of the performance of different modelling strategies.
+  In particular, the importance of considering opponent-level effects
+  and the suitability of single-level models.
+\item
+  Present an example of reproducible research in crash severity
+  analysis: all data and code are publicly available from the beginning
+  of the peer-review
+  process\footnote{The source file for this paper is an `R` Markdown document; all code and data necessary to reproduce the analysis are available from the following GitHub repository: \url{https://github.com/paezha/Modelling-Participant-Interactions-in-Crash-Severity}}.
+\end{enumerate}
+
+For the assessment we use data from Canada's National Collision
+Database, a database that collects all police-reported collisions in the
+country. Using the most recent version of the data set (2017). Three
+levels of crash severity (no injury/injury/fatality) are analyzed using
+ordered logit models, and covariates for the parties in the crash and
+the conditions of the crash. For model assessment, we conduct an
+in-sample prediction exercise using the estimation sample (i.e.,
+\emph{nowcasting}), and also an out-of-sample prediction exercise using
+the data set corresponding to 2016 (i.e., \emph{backcasting}). The
+models are assessed using predicted shares and predicted classes of
+outcomes at the individual level, using an extensive array of
+verification statistics.
 
@@
 
-Modelling the severity of injuries to participants in road crashes has
-been a preoccupation of transportation researchers, planners, auto
-insurance companies, governments, and the general public for decades.
-One of the earliest studies to investigate the severity of injuries
-conditional on an accident having occurred was by White and Clayton
-(1972). Kim et al.~(1995) later stated that the ``linkages between
-severity of injury and driver characteristics and behaviors have not
-been thoroughly investigated'' (p.~470). Nowadays, there is a burgeoning
-literature on this subject, including methodological developments, case
-studies, and more niche research with a focus on particular situations
-(e.g., crashes at intersections, Mussone et al., 2017; crashes in rural
-roads, Gong and Fan, 2017), and crashes involving special population
-groups (e.g., crashes involving motorcyclists or active travelers; see
-Shaheed et al., 2013; Salon and McIntyre, 2018).
-
-Crash severity is often modeled using models for discrete outcomes. An
-analyst interested in crash severity has at their disposal an ample
+Crash severity is often modeled using models for discrete outcomes.
+Analysts interested in crash severity have at their disposal an ample
 repertoire of models to choose from, including classification techniques
@@
 Menendez, 2014), and the use of copulas (e.g., Wang et al., 2015).
-Recent reviews of methods include Savolainen et al.~(2011), Shamsunnahar
-and Eluru (2013), and Mannering et al.~(2016).
+Recent reviews of methods include Savolainen et al.~(2011), Yasmin and
+Eluru (2013), and Mannering et al.~(2016).
 
@@
 al., 2013). Many crash databases and analyses also account for the
-multievent nature of many crashes. Participants may have had different
-roles in a crash depending on their context, with some acting as
+multi-event nature of many crashes. Individuals in the crash may have
+had different roles depending on their situation, with some acting as
 operators of a vehicle (i.e., drivers, bicyclists), while others were
@@
 events. A number of studies advocate the estimation of separate models
-for different participants and/or situations. In this way, Wang and
-Kockelman (2005) estimated models for single-vehicle and two-vehicle
+for different parties, individuals, and/or situations. In this way, Wang
+and Kockelman (2005) estimated models for single-vehicle and two-vehicle
 crashes, while Savolainen and Mannering (2007) estimated models for
@@
 effects}. There are numerous examples of studies that consider at least
-some characteristics of the opponents in two- or multi-vehicle crashes.
-For example, Wang and Kockelman (2005) considered the type of the
-opposing vehicle in their model for two-vehicle collisions. Similarly,
-Torrao et al.~(2014) included in their analysis the age, wheelbase,
-weight, and engine size of the opposing vehicle, while Bogue et
-al.~(2017) used the body type of the opposing vehicle. Penmetsa et
-al.~(2017) and Duddu et al.~(2018) are two of the most comprehensive
-examples of using opponent's information, as they used individual
-attributes of opponents (their physical condition, sex, and age), as
-well as characteristics of the opposing traffic unit (the vehicle type
-of the opponent). The twin strategies of subsetting the sample and using
-the attributes of the opponent are not mutually exclusive, but neither
-are they used consistently together, as a scan of the literature
-reveals. Another strategy is to introduce \emph{hierarchical components}
-in the model, a technique widely used in the hierarchical or multi-level
-modelling literature.
+some characteristics of the opposite party (or parties) in two- or
+multi-vehicle crashes. For example, Wang and Kockelman (2005) considered
+the type of the opposite vehicle in their model for two-vehicle
+collisions. Similarly, Torrao et al.~(2014) included in their analysis
+the age, wheelbase, weight, and engine size of the opposite vehicle,
+while Bogue et al.~(2017) used the body type of the opposite vehicle.
+Penmetsa et al.~(2017) and Duddu et al.~(2018) are two of the most
+comprehensive examples of using opponent's information, as they included
+attributes of individuals in the opposite party (their physical
+condition, sex, and age), as well as characteristics of the other
+party's traffic unit (the vehicle type of the opponent). The twin
+strategies of subsetting the sample and using the attributes of the
+opponent are not mutually exclusive, but neither are they used
+consistently together, as a scan of the literature reveals.
+
+Another strategy is to introduce \emph{hierarchical components} in the
+model, a technique widely used in the hierarchical or multi-level
+modelling literature. This involves considering observations as being
+nested at different levels: individuals nested in traffic units, which
+in turn are nested in accidents, as an example.
 
@@
 \begin{itemize}
-\tightlist
 \item
@@
 \midrule
-Human-factors & Attributes of participants in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
+Human-factors & Attributes of individuals in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
 Traffic unit-factors & Attributes of the traffic unit, e.g., type of traffic unit (pedestrian, car, motorcycle, etc.), maneuver, etc.\\
@@
 
-Shamsunnahar and Eluru (2013) conducted an extensive comparison of
-models for discrete outcomes in crash severity analysis, and found only
-small differences in the performance of unordered models and ordered
-models; however, ordered models are usually more parsimonious since only
-one latent functions needs to be estimated for all outcomes, as opposed
-to one for each outcome in unordered models. Bogue et al.~(2017) also
-compared unordered and ordered models in the form of the mixed
-multinomial logit and a modified rank ordered logit, respectively, and
-found that the ordered model performed best. To keep the empirical
-assessment manageable, in this paper we will consider only the ordinal
-logit model, and will comment on potential extensions in Section
-\ref{sec:further-considerations}.
+Before describing the modelling strategies, it is important to explain
+our choice of model. There have already been comparisons between
+different models. Yasmin and Eluru (2013), for instance, conducted an
+extensive comparison of models for discrete outcomes in crash severity
+analysis, and found only small differences in the performance of
+unordered models and ordered models; however, ordered models are usually
+more parsimonious since only one latent functions needs to be estimated
+for all outcomes, as opposed to one for each outcome in unordered
+models. Bogue et al.~(2017) also compared unordered and ordered models
+in the form of the mixed multinomial logit and a modified rank ordered
+logit, respectively, and found that the ordered model performed best. To
+keep the empirical assessment manageable, in this paper we will consider
+only the ordinal logit model, and will comment on potential extensions
+in Section \ref{sec:further-considerations}.
 
@@
 (unobservable) variable that is associated with the severity of crash
-\(k\) (\(k=1,\cdots,K\)) for participant \(i\) in traffic unit \(t\).
-The right-hand side of the expression is split in four parts. The first
-part collects \(l=1,\cdots,L\) human-factors \(p\) for participant \(i\)
-in traffic unit \(t\) and crash \(k\); these could relate to the person
-(e.g., age, gender, and road user class). The second part collects
+\(k\) (\(k=1,\cdots,K\)) for individual \(i\) in traffic unit \(t\). The
+right-hand side of the expression is split in four parts. The first part
+gathers \(l=1,\cdots,L\) human-factors \(p\) for individual \(i\) in
+traffic unit \(t\) and crash \(k\); these could relate to the person
+(e.g., age, gender, and road user class). The second part gathers
 \(m=1,\cdots,M\) attributes \(u\) related to traffic unit \(t\) in crash
 \(k\); these could be items such as maneuver or vehicle type. The third
-part collects \(q=1,\cdots,Q\) attributes \(c\) related to the crash
+part gathers \(q=1,\cdots,Q\) attributes \(c\) related to the crash
 \(k\), including environmental-factors and road-factors, such as weather
 conditions, road alignment, and type of surface. Lastly, the fourth
-element is a random term specific to participant \(i\) in traffic unit
+element is a random term specific to individual \(i\) in traffic unit
 \(t\) and crash \(k\). The function consists of a total of \(Z=L+M+Q\)
@@
 posit a probabilistic relationship with the outcome \(y_{itk}\) (the
-severity of crash \(k\) for participant \(i\) in traffic unit \(t\)).
+severity of crash \(k\) for individual \(i\) in traffic unit \(t\)).
 Depending on the characteristics of the data and the assumptions made
@@
 
-The additional part collects \(r=1,\cdots,R\) attributes \(o\) related
-to individual \(j\) in traffic unit \(v\) and crash \(k\) that opposed
-individual \(i\) in traffic unit \(t\) and crash \(k\). These could be
-individual characteristics of the opponent (such as age and gender)
-and/or characteristics of the opposing vehicle (such as vehicle type or
-weight). To qualify as an opponent, individual \(j\) must have been a
-participant in crash \(k\) but operating traffic unit \(v\ne t\).
-Sometimes the person \emph{is} the traffic unit, as is the case of a
-pedestrian. And we exclude passengers of vehicles as opponents, since
-they do not operate the traffic unit. In case the opponent attributes
-include only characteristics of the traffic unit, the condition for the
-traffic unit to be an opponent is that it participated in crash \(k\)
-and was different from \(t\). After introducing this new set of terms,
-the latent function now consists of a total of \(Z=L+M+Q+R\) covariates
-and associated parameters.
+This function includes one additional summation compared to Equation
+\ref{eq:latent-function-ordered-outcomes}. This summation gathers
+\(r=1,\cdots,R\) attributes \(o\) related to individual \(j\) in traffic
+unit \(v\) that was an opposite party to individual \(i\) in traffic
+unit \(t\) in crash \(k\). These attributes could be individual
+characteristics of the operator of the opposite traffic unit (such as
+age and gender) and/or characteristics of the opposite traffic unit
+(such as vehicle type or weight). To qualify as an opposite party,
+individual \(j\) must have been an individual in crash \(k\) but
+operating traffic unit \(v\ne t\). Sometimes the person \emph{is} the
+traffic unit, as is the case of a pedestrian. And we exclude passengers
+of vehicles as opponents, since they do not operate the traffic unit. In
+case the opponent attributes include only characteristics of the traffic
+unit, the condition for the traffic unit to be an opponent is that it
+was part of crash \(k\) and was different from \(t\). After introducing
+this new set of terms, the latent function now consists of a total of
+\(Z=L+M+Q+R\) covariates and associated parameters.
 
@@
 The coefficients of the traffic unit nest the individual attributes as
-follows. For any given coefficient \(q\):
+follows. For any given coefficient \(m\):
 
@@
 matters related to data and model estimation.
-
-\begin{quote}
-\textbf{Note:} this paper presents reproducible research. The source
-file is an R Markdown document. All code and data necessary to reproduce
-the analysis are available from the following anonymous Drive folder:
-\end{quote}
-
-\begin{quote}
-https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV
-\end{quote}
-
-\begin{quote}
-The source files, code, and data will be publicly available in a GitHub
-repository upon acceptance of the paper for publication
-\end{quote}
 
@@
 \ref{tab:variable-categories}, environmental-factors variables and
-road-factors variables are collected under a single variable class,
+road-factors variables are gathered under a single variable class,
 namely collision-related, since they are unique for each crash.
@@
 Data are organized by person; in other words, there is one record per
-participant in a collision, be they drivers, passengers, pedestrians,
+individual in a collision, be they drivers, passengers, pedestrians,
 etc. The only variable directly available with respect to opponents in a
@@
 al., 2017). Therefore, the data needs to be processed to obtain
-attributes of opponents for each participant in a collision. The
+attributes of the opposing party for each individual in a collision. The
 protocol to do this is described next.
 
-\hypertarget{data-preprocessing-initial-filter}{%
-\subsubsection{Data preprocessing: initial
-filter}\label{data-preprocessing-initial-filter}}
-
-We apply an initial filter, whereby we scan the database to remove all
-records that are not a person (including parked cars and other objects)
-or that are missing information (as is the case when a participant in a
-crash is a runaway vehicle). Next, records missing at least one of the
-next variables are removed: P\_USER (the road user class), P\_SEX (sex),
-P\_AGE (age), and P\_ISEV (individual-level severity of crash). This
-initial filter ensures that all records are complete from the
-perspective of key information for analysis.
-
-\hypertarget{data-preprocessing-filter-two-vehicle-collisions}{%
-\subsubsection{Data preprocessing: filter two-vehicle
-collisions}\label{data-preprocessing-filter-two-vehicle-collisions}}
+\hypertarget{data-preprocessing}{%
+\subsection{Data preprocessing}\label{data-preprocessing}}
+
+To prepare the data for analysis, in particular for Strategy 1
+(opponent-related factors), we apply an initial filter, whereby we scan
+the database to remove all records that are not a person (including
+parked cars and other objects) or that are missing information.
 
@@
 At this point, there are 32,298 collisions, involving only one (known)
-participant, there are 46,483 collisions involving two participants,
-19,433 collisions with three participants, 8,250 collisions involving
-four participants, 3,783 collisions with five participants, 1,789
-collisions with six participants, and 1,491 collisions involving seven
-or more participants. These participants were possibly occupants in
-different vehicles or were otherwise their own traffic units.
-Accordingly, the sample includes 174,741 drivers, 61,403 passengers,
-10,798 pedestrians, 5,286 bicyclists, and 6,564 motorcyclists.
-
-The next step in our data preprocessing protocol is to remove all
-collisions that involve only one participant. This still leaves numerous
-cases where multiple participants could have been in a single vehicle,
-for instance in a collision against a stationary object. Therefore, we
-proceed to use the vehicle sequence number to find the number of
-vehicles involved in each collision. This reveals that there are 20,732
-collisions that involve only one vehicle but possibly multiple
-participants (i.e., driver and one or more passengers). In addition,
-there are 165,520 collisions involving two vehicles (and possibly
-multiple participants). Finally, there are 40,242 collisions with three
-or more vehicles.
+individual, there are 46,483 collisions involving two parties, 19,433
+collisions with three parties, 8,250 collisions involving four parties,
+3,783 collisions with five parties, 1,789 collisions with six parties,
+and 1,491 collisions involving seven or more parties These parties were
+possibly occupants in different vehicles or were otherwise their own
+traffic units (e.g., pedestrians). Accordingly, the sample includes
+174,741 drivers, 61,403 passengers, 10,798 pedestrians, 5,286
+bicyclists, and 6,564 motorcyclists.
+
+The next step is to remove all collisions that involve only one party.
+This still leaves numerous cases where multiple parties could have been
+in a single vehicle, for instance in a collision against a stationary
+object. Therefore, we proceed to use the vehicle sequence number to find
+the number of vehicles involved in each collision. This reveals that
+there are 20,732 collisions that involve only one vehicle but possibly
+multiple individuals (i.e., driver and one or more passengers). In
+addition, there are 165,520 collisions involving two vehicles (and
+possibly multiple individuals). Finally, there are 40,242 collisions
+with three or more vehicles.
 
 Once we have identified the number of vehicles in each collision, we
-proceed to select all cases that involve only two vehicles. The most
-common cases in two-vehicle collisions are those that include drivers
-(40,297 collisions; this is reflective of the prevalence of
-single-occupant vehicles). This is followed by cases with passengers
-(14,120 collisions), pedestrians (5,204 collisions), bicyclists (2,238
+select all cases that involve only two vehicles. The most common cases
+in two-vehicle collisions are those that include drivers (40,297
+collisions; this is reflective of the prevalence of single-occupant
+vehicles). This is followed by cases with passengers (14,120
+collisions), pedestrians (5,204 collisions), bicyclists (2,238
 collisions), and motorcyclists (1,016 collisions). The distribution of
@@
 individuals are coded as being in vehicles 3 through 9, despite our
-earlier filter to retain only collisions with two vehicles. We therefore
-proceed to select only individuals assigned to vehicles 1 or 2, as well
-as pedestrians. As a result of this filter a number of cases with only
-one known participant need to be removed.
-
-\hypertarget{data-preprocessing-extract-opponent-information-and-join-to-individual-records}{%
-\subsubsection{Data preprocessing: extract opponent information and join
-to individual
-records}\label{data-preprocessing-extract-opponent-information-and-join-to-individual-records}}
-
-Up to this point, the goal of data preprocessing has been to obtain a
-complete, workable sample of individual records of participants in
-two-vehicle collisions. There are two possible cases for the collisions,
-depending on the traffic units involved: 1) vehicle vs vehicle
-collisions (``vehicle'' is all motorized vehicles, including
-motorcycles/mopeds, as well as bicycles); and 2) vehicle vs pedestrian
-collisions. To identify opponents in a collision, it is convenient to
-classify collisions by pedestrian involvement. In this way, we find that
-the database includes 16,636 collisions that are vehicle vs pedestrian
-(possibly multiple pedestrians), and 147,594 collisions that involve two
-vehicles. After splitting the database according to pedestrian
-involvement, we can now extract relevant information about participants
-in the collision. This involves renaming the person-level variables so
-that we can distinguish each individual by their role as an individual
-or an opponent in a given record. Notice that when working with
-individuals in vehicles, only the driver is considered a legitimate
-opponent in a collision (i.e., passengers are never considered
-opponents).
-
-Once the personal attributes of individuals that can be considered
-opponents in a given collision have been extracted, their information is
-joined to the individual records by means of the collision unique
-identifier. As a result of this process, a new set of variables are now
-available for analysis: the age, sex, and road user class of the
-opponent, as well as the type of the opposing vehicle. A summary of
-opponent interactions and outcomes can be found in Table
-\ref{tab:summary-opponent-information}. The information there shows that
-the most common type of opponent for drivers is other drivers, followed
-by pedestrians. The only opponents of pedestrians, on the other hand,
-are drivers. Bicyclists and motorcyclists, on the other hand, are mostly
-opposed by drivers, but occasionally by other road users as well. In
-terms of outcomes, we observe that virtually all fatalities occur when
-the opponent is a driver, and only very rarely when the opponent is a
-motorcyclist. Injuries are also more common when the opponent is a
-driver, whereas no injuries are relatively more frequent when the
-opponent is a pedestrian or a bicyclist.
-
-\begin{table}
+earlier filter to retain only collisions with two vehicles. At this
+point we select only individuals assigned to vehicles 1 or 2, as well as
+pedestrians. As a result of this filter a number of cases with only one
+known individual need to be removed.
+
+\begin{table}[!h]
 
@@
 
+At this point we have a complete, workable sample of individual records
+of parties in two-vehicle collisions. There are two possible cases for
+the collisions, depending on the traffic units involved: 1) vehicle vs
+vehicle collisions (``vehicle'' is all motorized vehicles, including
+motorcycles/mopeds, as well as bicycles); and 2) vehicle vs pedestrian
+collisions. To identify the opposite parties in each collision it is
+convenient to classify collisions by pedestrian involvement. In this
+way, we find that the database includes 16,636 collisions that are
+vehicle vs pedestrian (possibly multiple pedestrians), and 147,594
+collisions that involve two vehicles. After splitting the database
+according to pedestrian involvement, we can now extract relevant
+information about the different parties in the collision. This involves
+renaming the person-level variables so that we can distinguish each
+individual by their party in a given record. Notice that when working
+with individuals in vehicles, only drivers are considered opposites in a
+collision.
+
+Once the personal attributes of opposite operators in a given collision
+are extracted, their information is joined to the individual records by
+means of the collision unique identifier. As a result of this process, a
+new set of variables are now available for analysis: the age, sex, and
+road user class of the opposite driver, as well as the type of the
+opposite vehicle. A summary of opponent interactions and outcomes can be
+found in Table \ref{tab:summary-opponent-information}. The information
+there shows that the most common type of opponent for drivers are other
+drivers, followed by pedestrians. The only opponents of pedestrians, on
+the other hand, are drivers. Bicyclists and motorcyclists are mostly
+opposed by drivers, but occasionally by other road users as well. In
+terms of outcomes, we observe that virtually all fatalities occur when
+the opponent is a driver, and only very rarely when the opponent is a
+motorcyclist. Injuries are also more common when the opponent is a
+driver, whereas ``no injury'' is a relatively more frequent outcome when
+the opponent is a pedestrian or a bicyclist.
+
 \hypertarget{sec:model-estimation}{%
@@
 vehicle weight rating); and heavy vehicles (all other vehicles \(\ge\)
-4536 in gross vehicle weight rating). Furthermore, this typology of
+4536 kg in gross vehicle weight rating). Furthermore, this typology of
 vehicle is combined with the road user class of the individual to
@@
 
-The results of model estimation are discussed in the following section.
-
-\begin{table}
+\begin{table}[!h]
 
@@
 \fontsize{7}{9}\selectfont
-\begin{tabular}[t]{llllll}
+\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}}
 \toprule
-Variable & Notes & Model 1 & Model 2 & Model 3 & Model 4\\
+Variable & Notes & Model 1
+Single-level
+/No opponent & Model 2
+Single-level
+/Opponent attributes & Model 3
+Hierarchical: Traffic unit & Model 4
+Hierarchical: Opponent attributes\\
 \midrule
@@
 to the year 2016 (i.e., \emph{backcasting}). Predictions are commonly
-evaluated in two different in the literature. Some researchers analyze
-the outcome shares based on the predicted probabilities (e.g., Bogue et
-al., 2017; Shamsunnahar and Eluru, 2013). This is a form of aggregate
+evaluated in two different ways in the literature. Some researchers
+analyze the outcome shares based on the predicted probabilities (e.g.,
+Bogue et al., 2017; Yasmin and Eluru, 2013). This is a form of aggregate
 forecasting. Other researchers, in contrast, evaluate the classes of
@@
 the models described above. Tables \ref{tab:model-full-sample-summary}
-and \ref{tab:model-subsample-summary} collect some key summary
+and \ref{tab:model-subsample-summary} present some key summary
 statistics of the estimated models. Of interest is the goodness of fit
@@
 \midrule
-\rowcolor{gray!6}  Model 1 & 164,511 & 102 & 195,215\\
-Model 2 & 164,511 & 108 & 178,943\\
-\rowcolor{gray!6}  Model 3 & 164,511 & 118 & 181,333\\
-Model 4 & 164,511 & 111 & 179,018\\
+\rowcolor{gray!6}  Model 1. Single-level/No opponent & 164,511 & 102 & 195,215\\
+Model 2. Single-level/Opponent attributes & 164,511 & 108 & 178,943\\
+\rowcolor{gray!6}  Model 3. Hierarchical: Traffic unit & 164,511 & 118 & 181,333\\
+Model 4. Hierarchical: Opponent) & 164,511 & 111 & 179,018\\
 \bottomrule
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} & \multicolumn{2}{c}{Model 4} \\
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{\makecell[c]{Model 1\\Single-level/No opponent}} & \multicolumn{2}{c}{\makecell[c]{Model 2\\Single-level/Opponent attributes}} & \multicolumn{2}{c}{\makecell[c]{Model 3\\Hierarchical: Traffic unit}} & \multicolumn{2}{c}{\makecell[c]{Model 4\\Hierarchical: Opponent}} \\
 \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
@@
 \bottomrule
+\multicolumn{11}{l}{\textit{Note: }}\\
+\multicolumn{11}{l}{Model 1. Single-level/No opponent}\\
+\multicolumn{11}{l}{Model 2. Single-level/Opponent attributes}\\
+\multicolumn{11}{l}{Model 3. Hierarchical: Traffic unit}\\
+\multicolumn{11}{l}{Model 4. Hierarchical: Opponent}\\
 \end{tabular}
@@
 
-The individual-level outcomes are examined using an extensive array of
+The individual-level outcomes are examined using an array of
 verification statistics. Verification statistics are widely used in the
@@
 \ref{tab:verification-statistics}, including brief descriptions of their
-interpretation (more on this below). The statistics evaluate different
-aspects of the performance of a model. Some are concerned with the
-ability of the model to be right. For example, Percent Correct (\(PC\))
-is the sum of hits and correct rejections divided by the number of
-cases. Others are concerned with the ability of the model to match the
-observed outcomes, including Bias (\(B\)), which measures whether a
-class has been over- or under-predicted. Whereas others measure the
-ability of the model to not be wrong, such as the probability of false
-detection (\(F\)) and the False Alarm Ration (\(FAR\)). We discuss the
-results of calculating the battery of verification statistics, first for
-the nowcasting case (are shown in (Table
+interpretation. The statistics evaluate different aspects of the
+performance of a model. Some are concerned with the ability of the model
+to be right, others are concerned with the ability of the model to match
+the observed outcomes, and yet others measure the ability of the model
+to not be wrong. These verification statistics are discussed briefly
+next.
+
+Percent correct and percent correct by class \(PC\) and \(PC_c\) are
+relatively simple statistics, and are calculated as the proportion of
+correct predictions (i.e., hits and correct non-events) relative to the
+number of events, for the whole table \(PC\) or for one class only
+\(PC_c\).
+
+Bias (\(B\)) measures for each outcome class the proportion of total
+predictions by class (e.g., hits as well as false alarms) relative to
+the total number of cases observed for that class. For this reason, it
+is possible for predictions to have low bias (values closer to 1) but
+still do poorly in terms of hits.
+
+Critical Success Index (\(CSI\)) evaluates forecasting skill while
+assuming that the number of correct non-events is inconsequential for
+demonstrating skill. Accordingly, the statistic is calculated as the
+proportion of hits relative to the sum of hits plus false alarms plus
+misses.
+
+Probability of False Detection (\(F\)) is the proportion of false alarms
+relative to the total number of times that the event is not observed.
+This statistic measures the frequency with which the model incorrectly
+predicts an event, but not when it incorrectly misses it. The
+Probability of Detection (\(POD\)), in contrast, measures the frequency
+with which the model correctly predicts a class, relative to the number
+of cases of that class.
+
+The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
+that were false alarms evaluate a different way in which a model can
+make equivocal predictions. In this case, lower scores are better.
+
+The last three verification statistics that we consider are skill scores
+that simultaneously consider different aspects of prediction, and are
+therefore overall indicators of prediction skill. Heidke's Skill Score
+(\(HSS\)) is the fraction of correct predictions above those that could
+be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
+Probability of Detection (\(POD\)) of a model and its Probability of
+False Detection (\(F\)) to measure the skill of a model to discriminate
+the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
+the model's skill predicting the correct classes that tends to reward
+correct forecasts of the least frequent class.
+
+We discuss the results of calculating this battery of verification
+statistics, first for the nowcasting case (Table
 \ref{tab:nowcast-outcomes-assessment}) and subsequently for the
@@
 \midrule
-\rowcolor{gray!6}  Percent Correct ($PC$) & Total hits and correct rejections divided by number of cases & Strongly influenced by most common category\\
+\rowcolor{gray!6}  Percent Correct ($PC$) & Total hits and correct non-events divided by number of cases & Strongly influenced by most common category\\
 Percent Correct by Class ($PC_c$) & Same as Percent Correct but by category & Strongly influenced by most common category\\
@@
 that the number of hits are relatively low; the high values of \(PC_c\)
-in this case are due to the high occurrence of correct rejections
+in this case are due to the high occurrence of correct non-events
 elsewhere in the table.
 
-Bias (\(B\)) measures for each outcome class the proportion of total
-predictions by class (e.g., hits as well as false alarms) relative to
-the total number of cases observed for that class. For this reason, it
-is possible for predictions to have low bias (values closer to 1) but
-still do poorly in terms of hits. The models with the best performance
-in terms of \(B\) are Model 1 (full sample) for No Injury and Injury,
-and Model 3 (ensemble) for Fatality. Model 4 (full sample) is the second
-best performer for No Injury and Injury, and Model 4 (ensemble) is
-second best performer for Fatality. Model 1 (ensemble) has the worst
-bias for No Injury and Injury, whereas Model 2 (full sample) has the
-worst bias for Fatality.
-
-Critical Success Index (\(CSI\)) evaluates forecasting skill while
-assuming that the number of correct non-events is inconsequential skill.
-Accordingly, the statistic is calculated as the proportion of hits
-relative to the sum of hits plus false alarms plus misses. No model
-performs uniformly best from this perspective. Model 2 (full sample) has
-the best \(CSI\) for No Injury, Model 2 (ensemble) has the best score
-for Injury, and Model 4 (ensemble) the best score for Fatality. On the
-other hand, Model 1 (ensemble) has the worst score for No Injury, Model
-1 (full sample) the worst score for Injury, and Model 2 the worst score
-for Fatality. These scores indicates that the models are not
-particularly skilled at predicting the corresponding classes correctly,
-given the frequency with which they give false alarms or miss the class.
-
-The next statistic is Probability of False Detection (\(F\)). This is
-the proportion of false alarms relative to the total number of cases
-that the event is not observed. This statistic measure the frequency
-with which the model incorrectly predicts an event, but not when it
-incorrectly misses it. The lowest probability of false detection in the
-case of No Injury is 19.17\% for Model 2 (ensemble), with every other
-model having values lower than 21\%, with the exception of Model 1 (full
-sample) that has a score of 26.46\%. With respect to Injury, the lowest
-probabilities range 35.8\% and 35.29\% and 35.35\% for Models 4 (full
-sample) and 2 (full sample). In contrast, the highest probability of
-false detection for Injury is 45.12\% for Model 1 (ensemble). The scores
-for \(F\) for Fatality are all extremely low as a consequence of the
-very low frequency of this class of outcome in the sample.
-
-The Probability of Detection (\(POD\)) measures the frequency with which
-the model correctly predicts a class, relative to the number of cases of
-that class. As with some other verification statistics, no model is
-consistently a best performer. Model 4 (full sample) has the highest
-probability of detection for No Injury (65.38\%), followed by Model 2
-(65.32\%), whereas the worst probability of detection is by Model 1
-(ensemble) with a score of 55.54\%. In terms of Injury, all models have
-\(POD\) higher than 79\%, and the highest score is 80.67\% for Model 2
-(ensemble). The exception is Model 1 (full sample), which has a
-considerably lower \(POD\) of Injury with a score of 73.36\%. Lastly, in
-terms of Fatalities, all models have very low probabilities of
-detection, ranging from a high of 4.94\% in the case of Model 4
+The models with the best performance in terms of \(B\) are Model 1 (full
+sample) for No Injury and Injury, and Model 3 (ensemble) for Fatality.
+Model 4 (full sample) is the second best performer for No Injury and
+Injury, and Model 4 (ensemble) is second best performer for Fatality.
+Model 1 (ensemble) has the worst bias for No Injury and Injury, whereas
+Model 2 (full sample) has the worst bias for Fatality.
+
+No model performs uniformly best from the perspective of Critical
+Success Index (\(CSI\)). Model 2 (full sample) has the best \(CSI\) for
+No Injury, Model 2 (ensemble) has the best score for Injury, and Model 4
+(ensemble) the best score for Fatality. On the other hand, Model 1
+(ensemble) has the worst score for No Injury, Model 1 (full sample) the
+worst score for Injury, and Model 2 the worst score for Fatality. These
+scores indicates that the models are not particularly skilled at
+predicting the corresponding classes correctly, given the frequency with
+which they give false alarms or miss the class.
+
+The lowest probability of false detection \(F\) in the case of No Injury
+is 19.17\% for Model 2 (ensemble), with every other model having values
+lower than 21\%, with the exception of Model 1 (full sample) that has a
+score of 26.46\%. With respect to Injury, the lowest probabilities range
+35.8\% and 35.29\% and 35.35\% for Models 4 (full sample) and 2 (full
+sample). In contrast, the highest probability of false detection for
+Injury is 45.12\% for Model 1 (ensemble). The scores for \(F\) for
+Fatality are all extremely low as a consequence of the very low
+frequency of this class of outcome in the sample.
+
+As with some other verification statistics, no model is consistently a
+best performer in terms of \(POD\). Model 4 (full sample) has the
+highest probability of detection for No Injury (65.38\%), followed by
+Model 2 (65.32\%), whereas the worst probability of detection is by
+Model 1 (ensemble) with a score of 55.54\%. In terms of Injury, all
+models have \(POD\) higher than 79\%, and the highest score is 80.67\%
+for Model 2 (ensemble). The exception is Model 1 (full sample), which
+has a considerably lower \(POD\) of Injury with a score of 73.36\%.
+Lastly, in terms of Fatalities, all models have very low probabilities
+of detection, ranging from a high of 4.94\% in the case of Model 4
 (ensemble) to a worst score of 0.11\% in the case of Model 2 (full
@@
 
-The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
-that were false alarms, which means that lower scores are better. Model
-2 (full sample) has the best \(FAR\) statistic for No Injury, as only
-25.05\% of predictions for this class are false alarms. The next best
-score is by Model 4 (full sample), with only 25.23\% of No Injury
+Model 2 (full sample) has the best \(FAR\) statistic for No Injury, as
+only 25.05\% of predictions for this class are false alarms. The next
+best score is by Model 4 (full sample), with only 25.23\% of No Injury
 predictions being false alarms. The worst performance in this class is
@@
 
-The next three verification statistics are skill scores that
-simultaneously consider different aspects of prediction, and are
-therefore overall indicators of prediction skill. Heidke's Skill Score
-(\(HSS\)) is the fraction of correct predictions above those that could
-be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
-Probability of Detection (\(POD\)) of a model and its Probability of
-False Detection (\(F\)) to measure the skill of a model to discriminate
-the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
-the model's skill predicting the correct classes that tends to reward
-correct forecasts of the least frequent class. These scores help to
-remove some of the ambiguity regarding the overall performance of a
-model. In this way, we know that Model 2 (full sample) does not do
-particularly well with the class Fatality - however, of all models, it
-tends to have the best overall performance. Its \(HSS\), for example,
-suggests that it achieves 44.74\% of correct predictions after removing
-correct predictions attributable to chance. In contrast, the lowest
-score is for Model 1 (ensemble), which only achieves 36.26\% correct
-predictions after removing those attributable to chance. Model 2 (full
-sample) also has the highest \(PSS\) and the second highest \(GS\).
-Model 4 (full sample) has the highest \(GS\) and the second highest
-\(HSS\) and \(PSS\).
+The skill scores help to remove some of the ambiguity regarding the
+overall performance of a model. In this way, we know that Model 2 (full
+sample) does not do particularly well with the class Fatality - however,
+of all models, it tends to have the best overall performance. Its
+\(HSS\), for example, suggests that it achieves 44.74\% of correct
+predictions after removing correct predictions attributable to chance.
+In contrast, the lowest score is for Model 1 (ensemble), which only
+achieves 36.26\% correct predictions after removing those attributable
+to chance. Model 2 (full sample) also has the highest \(PSS\) and the
+second highest \(GS\). Model 4 (full sample) has the highest \(GS\) and
+the second highest \(HSS\) and \(PSS\).
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1}}\\
-\hspace{1em}No Injury & 50652 & 22503 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62121 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
+\hspace{1em}No Injury & 50652 & 22504 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62120 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2}}\\
-\hspace{1em}No Injury & 51531 & 17136 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67515 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
+\hspace{1em}No Injury & 51531 & 17137 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
+
+\hspace{1em}Injury & 27355 & 67514 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 51102 & 17297 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
-
-\hspace{1em}Injury & 27784 & 67337 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7952} & \textcolor{black}{0.2985} &  &  & \\
-
-Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4356}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
+\multicolumn{14}{l}{\textbf{Model 3. Hierarchical: Traffic unit}}\\
+\hspace{1em}No Injury & 51103 & 17296 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
+
+\hspace{1em}Injury & 27783 & 67338 & 868 &  & \textcolor{black}{72.05} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7953} & \textcolor{black}{0.2985} &  &  & \\
+
+Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4357}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
 \cmidrule{1-14}
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 51575 & 17318 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
-
-\hspace{1em}Injury & 27311 & 67334 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.1279}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
+\hspace{1em}No Injury & 51574 & 17316 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
+
+\hspace{1em}Injury & 27312 & 67336 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.128}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
+\multicolumn{14}{l}{\textbf{Model 1 Ensemble. Single-level/No opponent}}\\
 \hspace{1em}No Injury & 34664 & 16434 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1952} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 35443 & 16146 & 60 &  & \textcolor{black}{70.61} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
-
-\hspace{1em}Injury & 26974 & 67436 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
-
-Fatality & 0 & 13 & 42 &  & \textcolor{black}{99.39} & \textcolor{black}{0.0591} & \textcolor{black}{0.0445} & \textcolor{black}{1e-04} & \textcolor{black}{0.0451} & \textcolor{black}{0.2364} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3784}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3678}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2106}}\\
-
-\addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 35498 & 16203 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
-
-\hspace{1em}Injury & 26913 & 67380 & 828 &  & \textcolor{black}{70.09} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
-
-Fatality & 0 & 13 & 45 & \multirow{-6}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
+\multicolumn{14}{l}{\textbf{Model 2 Ensemble. Single-level/Opponent attributes}}\\
+\hspace{1em}No Injury & 35443 & 16145 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\
+
+\hspace{1em}Injury & 26974 & 67437 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\
+
+Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.05}} & \textcolor{black}{99.39} & \textcolor{black}{0.0591} & \textcolor{black}{0.0445} & \textcolor{black}{1e-04} & \textcolor{black}{0.0451} & \textcolor{black}{0.2364} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3784}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3678}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2106}}\\
 \cmidrule{1-14}
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
+\multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
+\hspace{1em}No Injury & 35498 & 16204 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
+
+\hspace{1em}Injury & 26913 & 67379 & 828 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
+
+Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
+\cmidrule{1-14}
+\addlinespace[0.3em]
+\multicolumn{14}{l}{\textbf{Model 4 Ensemble. Hierarchical: Opponent}}\\
 \hspace{1em}No Injury & 35553 & 16297 & 59 &  & \textcolor{black}{70.59} & \textcolor{black}{0.8318} & \textcolor{black}{0.4514} & \textcolor{black}{0.1935} & \textcolor{black}{0.5697} & \textcolor{black}{0.3151} &  &  & \\
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1}}\\
+\multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
 \hspace{1em}No Injury & 61683 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2}}\\
+\multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
 \hspace{1em}No Injury & 62736 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5318}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3}}\\
-\hspace{1em}No Injury & 62249 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.862} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
-
-\hspace{1em}Injury & 34609 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 3. Hierarchical: Traffic unit}}\\
+\hspace{1em}No Injury & 62248 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.8619} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\
+
+\hspace{1em}Injury & 34610 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4}}\\
-\hspace{1em}No Injury & 62788 & 21248 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8687}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2079} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80330 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
+\hspace{1em}No Injury & 62788 & 21247 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80331 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 1 Ensemble}}\\
+\multicolumn{14}{l}{\textbf{Model 1 Ensemble. Single-level/No opponent}}\\
 \hspace{1em}No Injury & 42896 & 20230 & 95 &  & \textcolor{black}{69.26} & \textcolor{red}{0.8162} & \textcolor{red}{0.4387} & \textcolor{black}{0.2011} & \textcolor{red}{0.5538} & \textcolor{red}{0.3215} &  &  & \\
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 2 Ensemble}}\\
-\hspace{1em}No Injury & 43486 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.4461} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
-
-\hspace{1em}Injury & 33953 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
-
-Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3551}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
+\multicolumn{14}{l}{\textbf{Model 2 Ensemble. Single-level/Opponent attributes}}\\
+\hspace{1em}No Injury & 43485 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.446} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
+
+\hspace{1em}Injury & 33954 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
+
+Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.355}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
 \cmidrule{1-14}
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 3 Ensemble}}\\
-\hspace{1em}No Injury & 43526 & 20034 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8218} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
-
-\hspace{1em}Injury & 33915 & 79914 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7987}} & \textcolor{black}{0.3038} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
+\hspace{1em}No Injury & 43526 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8217} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
+
+\hspace{1em}Injury & 33915 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\
 
@@
 \addlinespace[0.3em]
-\multicolumn{14}{l}{\textbf{Model 4 Ensemble}}\\
-\hspace{1em}No Injury & 43560 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5623} & \textcolor{black}{0.3174} &  &  & \\
-
-\hspace{1em}Injury & 33876 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
+\multicolumn{14}{l}{\textbf{Model 4 Ensemble. Hierarchical: Opponent}}\\
+\hspace{1em}No Injury & 43561 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5624} & \textcolor{black}{0.3174} &  &  & \\
+
+\hspace{1em}Injury & 33875 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
 
@@
 Based on the literature, we limited our empirical assessment of
-modelling strategies to only one model, namely the ordinal probit. On
-the other hand, since the modelling strategies discussed here all relate
-to the specification of the latent function and data subsetting, it is a
+modelling strategies to only one model, namely the ordinal logit. On the
+other hand, since the modelling strategies discussed here all relate to
+the specification of the latent function and data subsetting, it is a
 relatively simple matter to extend them to other modelling approaches.
@@
 modelling approach with nested opponent effects. There was some evidence
-that subsetting the sample can improve the results in some cases s(e.g.,
+that subsetting the sample can improve the results in some cases (e.g.,
 when modelling the severity of crashes involving active travelers or
@@
 
-To be added in the final version of the paper.
+This research was supported by a Research Excellence grant from McMaster
+University. The authors wish to express their gratitude to the Office of
+the Provost, the Faculty of Science, and the Faculty of Engineering for
+their generous support. The comments of four anonymous reviewers are
+also duly acknowledged: their feedback helped to improve the quality and
+clarity of presentation of this paper.
 
@@
 
-\leavevmode\hypertarget{ref-Hanson2013severity}{}%
-Hanson, C.S., Noland, R.B., Brown, C., 2013. The severity of pedestrian
-crashes: An analysis using google street view imagery. Journal of
-Transport Geography 33, 42--53.
-doi:\href{https://doi.org/10.1016/j.jtrangeo.2013.09.002}{10.1016/j.jtrangeo.2013.09.002}
-
 \leavevmode\hypertarget{ref-Hedeker1994random}{}%
@@
 
-\leavevmode\hypertarget{ref-McArthur2014spatial}{}%
-McArthur, A., Savolainen, P.T., Gates, T.J., 2014. Spatial analysis of
-child pedestrian and bicycle crashes development of safety performance
-function for areas adjacent to schools. Transportation Research Record
-57--63. doi:\href{https://doi.org/10.3141/2465-08}{10.3141/2465-08}
-
 \leavevmode\hypertarget{ref-Montella2013crash}{}%
@@
 
-\leavevmode\hypertarget{ref-Mussone2017analysis}{}%
-Mussone, L., Bassani, M., Masci, P., 2017. Analysis of factors affecting
-the severity of crashes in urban road intersections. Accident Analysis
-and Prevention 103, 112--122.
-doi:\href{https://doi.org/10.1016/j.aap.2017.04.007}{10.1016/j.aap.2017.04.007}
-
-\leavevmode\hypertarget{ref-Obeng2011gender}{}%
-Obeng, K., 2011. Gender differences in injury severity risks in crashes
-at signalized intersections. Accident Analysis and Prevention 43,
-1521--1531.
-doi:\href{https://doi.org/10.1016/j.aap.2011.03.004}{10.1016/j.aap.2011.03.004}
-
 \leavevmode\hypertarget{ref-Osman2018injury}{}%
@@
 
-\leavevmode\hypertarget{ref-Peek-Asa2010teenage}{}%
-Peek-Asa, C., Britton, C., Young, T., Pawlovich, M., Falb, S., 2010.
-Teenage driver crash incidence and factors influencing crash injury by
-rurality. Journal of Safety Research 41, 487--492.
-doi:\href{https://doi.org/10.1016/j.jsr.2010.10.002}{10.1016/j.jsr.2010.10.002}
-
 \leavevmode\hypertarget{ref-Penmetsa2017examining}{}%
@@
 
-\leavevmode\hypertarget{ref-Rakotonirainy2012older}{}%
-Rakotonirainy, A., Steinhardt, D., Delhomme, P., Darvell, M., Schramm,
-A., 2012. Older drivers' crashes in queensland, australia. Accident
-Analysis and Prevention 48, 423--429.
-doi:\href{https://doi.org/10.1016/j.aap.2012.02.016}{10.1016/j.aap.2012.02.016}
-
 \leavevmode\hypertarget{ref-Rana2010copula}{}%
@@
 
-\leavevmode\hypertarget{ref-Regev2018crash}{}%
-Regev, S., Rolison, J.J., Moutari, S., 2018. Crash risk by driver age,
-gender, and time of day using a new exposure methodology. Journal of
-Safety Research 66, 131--140.
-doi:\href{https://doi.org/10.1016/j.jsr.2018.07.002}{10.1016/j.jsr.2018.07.002}
-
 \leavevmode\hypertarget{ref-Rifaat2007accident}{}%
@@
 doi:\href{https://doi.org/10.1007/s11116-010-9263-3}{10.1007/s11116-010-9263-3}
-
-\leavevmode\hypertarget{ref-Salon2018determinants}{}%
-Salon, D., McIntyre, A., 2018. Determinants of pedestrian and bicyclist
-crash severity by party at fault in san francisco, ca. Accident Analysis
-and Prevention 110, 149--160.
-doi:\href{https://doi.org/10.1016/j.aap.2017.11.007}{10.1016/j.aap.2017.11.007}
 
@@
 
-\leavevmode\hypertarget{ref-Shaheed2013mixed}{}%
-Shaheed, M.S.B., Gkritza, K., Zhang, W., Hans, Z., 2013. A mixed logit
-analysis of two-vehicle crash seventies involving a motorcycle. Accident
-Analysis and Prevention 61, 119--128.
-doi:\href{https://doi.org/10.1016/j.aap.2013.05.028}{10.1016/j.aap.2013.05.028}
-
-\leavevmode\hypertarget{ref-Shamsunnahar2013evaluating}{}%
-Shamsunnahar, Y., Eluru, N., 2013. Evaluating alternate discrete outcome
-frameworks for modeling crash injury severity. Accident Analysis \&
-Prevention 59, 506--521.
-doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
-
 \leavevmode\hypertarget{ref-Shamsunnahar2014examining}{}%
@@
 doi:\href{https://doi.org/10.1080/15568318.2010.497547}{10.1080/15568318.2010.497547}
-
-\leavevmode\hypertarget{ref-Thompson2018trends}{}%
-Thompson, J.P., Baldock, M.R.J., Dutschke, J.K., 2018. Trends in the
-crash involvement of older drivers in australia. Accident Analysis and
-Prevention 117, 262--269.
-doi:\href{https://doi.org/10.1016/j.aap.2018.04.027}{10.1016/j.aap.2018.04.027}
 
@@
 
-\leavevmode\hypertarget{ref-WHO2019global}{}%
-World Health Organization, 2019. Global status report on road safety
-2018 (2018). Geneva.
-
 \leavevmode\hypertarget{ref-Wu2014mixed}{}%
@@
 
+\leavevmode\hypertarget{ref-Yasmin2013evaluating}{}%
+Yasmin, S., Eluru, N., 2013. Evaluating alternate discrete outcome
+frameworks for modeling crash injury severity. Accident Analysis \&
+Prevention 59, 506--521.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}
+
 \end{document}

### 2020-05-16 Interactions-Crash-Severity-v1/Response-to-Reviewers-v2.Rmd
Revisions after feedback from Hany, Mark, and Saiedeh. Resubmission to Accident Analysis and Prevention.

@@
 
+We wish to thank four anonymous reviewers for their feedback on our paper.
+In this letter we respond to their comments and describe the changes made to the paper in response.
+Reviewer comments are in black and our responses in blue.
+
+\newpage
+
 # Reviewer 1
@@
 
->*Thank you for reading the paper and for your feedback to improve the paper.*
+>\textcolor{blue}{Thank you for reading the paper and for your feedback to improve the paper.}
 
@@
 
->*Thank you for this perceptive comment.
-Indeed, the combined application of these strategies is not completely new, or new at all, since every single one of them has been used in the past by someone or other.
+>\textcolor{blue}{Thank you for this perceptive comment.
+Indeed, the combined application of these strategies is not new, since every single one of them has been used in the past by someone or other.
 Our argument is not that the strategies are new, but rather that they are used in a haphazard fashion (we document the different types of applications in the paper).
@@
 We would like to draw your attention to this paragraph in the paper (p.
-26, lines 497-510):*
+26, lines 497-510):}
 
->>*"We do not report results regarding other modelling strategies.
+>>\textcolor{blue}{"We do not report results regarding other modelling strategies.
 On the one hand, more sophisticated modelling frameworks are generally capable of improving the performance of a model.
@@
 That said, we experimented with a model with random components using monthly subsets of data to ﬁnd that, indeed, estimation takes considerably longer, is more demanding in terms of ﬁxing potential estimation quirks, and in the end resulted in variance components that could not be reliably estimated as diﬀerent from zero (results can be consulted in the source R Notebook).
-For this reason, we choose to leave the application of more sophisticated models as a matter for future research."*
+For this reason, we choose to leave the application of more sophisticated models as a matter for future research."}
 
->*We _did_ try other models, but there is no squeezing out unobserved heterogeneity where none seems to exist.*
+>\textcolor{blue}{We tried other models, but there is no squeezing out unobserved heterogeneity where none seems to exist.}
 
@@
 
->*If you compare the average prediction errors (APEs) and weighted average prediction errors (WAPEs) of our models to those reported by Bogue, Paleti, and Balam (2017), you will notice that our models actually perform well according to these criteria.
-However, part of our argument is that a more complete comparison of the performance of models requires the use of other performance criteria, hence all the verification statistics.*
+>\textcolor{blue}{If you compare the average prediction errors (APEs) and weighted average prediction errors (WAPEs) of our models to those reported by Bogue, Paleti, and Balam (2017), you will notice that our models actually perform well according to these criteria.
+However, part of our argument is that a more complete comparison of the performance of models requires the use of other performance criteria, hence all the verification statistics.}
 
@@
 
->*It is not that curious, really.
-Percent correct by class is heavily influenced by the dominant class (see note in Table 8).
+>\textcolor{blue}{Percent correct by class is heavily influenced by the dominant class (see note in Table 8).
 It is calculated as the sum of total hits and correct non-events divided by the number of cases.
-Accordingly (see for example Table 9, Model 1):*
+Accordingly (see for example Table 9, Model 1):}
 
@@
 
->*The percent correct by class (No Injury) is calculated as follows:*
+>\textcolor{blue}{The percent correct by class (No Injury) is calculated as follows:}
 
@@
 
->*In the above, $50652$ is the number of total hits, and $62121 + 797 + 51 + 3$ is the number of correct non-events. Compare to the percent correct by class (Fatality):*
+>\textcolor{blue}{In the above, $50652$ is the number of total hits, and $62121 + 797 + 51 + 3$ is the number of correct non-events. Compare to the percent correct by class (Fatality):}
 
@@
 
->*In the above, $3$ is the number of total hits, and $50652 + 22503 + 28232 + 62121$ is the number of correct non-events.
+>\textcolor{blue}{In the above, $3$ is the number of total hits, and $50652 + 22503 + 28232 + 62121$ is the number of correct non-events.
 It is not surprising that the value is so high: in fact, the model correctly predicts non-fatalities most of the time!
@@
 Our argument is that the use of verification statistics provides a more nuanced view of the performance of the models.
-Again, to the best of our knowledge, this more nuanced perspective is not currently available, and that it provides new insights into the performance of models.*
+Again, to the best of our knowledge, this more nuanced perspective is not currently available, and it provides new insights into the performance of models.}
 
@@
 
->*Thank you. We also think that this should be encouraged. We did not find in our literature review a single paper on crash severity that was reproducible.*
+>\textcolor{blue}{Thank you. We also think that this should be encouraged. We did not find in our literature review a single paper on crash severity that was reproducible.}
 
@@
 
->*We have changed the title based on this and a related comment by Reviewer 2.*
+>\textcolor{blue}{We have changed the title based on this and a related comment by Reviewer 2.}
 
@@
 
->*We followed this recommendation in the revision and reorganized the section to describe the verification statistics at the beginning. We also edited it, hopefully improving its readability.*
+>\textcolor{blue}{We followed this recommendation in the revision and reorganized the section to describe the verification statistics at the beginning.
+We also edited it, hopefully improving its readability.}
 
@@
 
->*It should be kg. We added the units.*
+>\textcolor{blue}{It should be kg. We added the units.}
 
@@
 
->*In two different ways. This was corrected.*
+>\textcolor{blue}{In two different ways. This was corrected.}
 
@@
 
->*This paragraph was deleted as part of reorganizing the section.*
+>\textcolor{blue}{This paragraph was deleted as part of reorganizing the section.}
 
@@
 
->*Typo corrected.*
+>\textcolor{blue}{Typo corrected.}
 
@@
 
->*This was an oversight caused by the practice of anonymizing papers. The revision includes the acknowledgments.*
+>\textcolor{blue}{This was an oversight caused by the practice of anonymizing papers. The revision includes the acknowledgments.}
 
->*Thanks again for your feedback and the time invested in helping us improve this paper. We hope that with these revisions the contributions of the paper will be more easily appreciated.*
+>\textcolor{blue}{Thanks again for your feedback and the time invested in helping us improve this paper.
+We hope that with these revisions the contributions of the paper will be more easily appreciated.}
 
@@
 
+>\textcolor{blue}{Thank you for your thoughtful comments on our paper.}
+
 “Accident” should be avoided. Crash/collision should be used.
 
->*Thank you for this comment.
-We thought that since the name of the journal is Accident Analysis and Prevention the term "accident" was appropriate to refer to stochastic events with undesirable outcomes.
-We have changed most "accidents" to "collisions" and "crashes" to reflect the nuanced differences between them.*
+>\textcolor{blue}{Thank you for this comment. We have changed "accidents" to "collisions" and "crashes".}
 
@@
 
->*The sentence was wordy and unclear, and it was edited for clarity. Now it reads:*
+>\textcolor{blue}{The sentence was wordy and unclear, and it was edited for clarity. Now it reads:}
 
->>*Numerous efforts have been undertaken to understand the factors that affect road collisions in general, and the severity of crashes in particular.
-In this literature several strategies have been proposed to model interactions between parties in a crash, including the use of variables regarding the other party (or parties) in the collision, data subsetting, and estimating models with hierarchical components.*
+>>\textcolor{blue}{Numerous efforts have been undertaken to understand the factors that affect road collisions in general, and the severity of crashes in particular.
+In this literature several strategies have been proposed to model interactions between parties in a crash, including the use of variables regarding the other party (or parties) in the collision, data subsetting, and estimating models with hierarchical components.}
+
+>\textcolor{blue}{Furthermore, the rewritten introduction now includes this:}
+
+>>\textcolor{blue}{For the purpose of this paper, we define a party as one or more individuals travelling in a traffic unit that becomes involved in a crash.
+Sometimes the traffic unit is a vehicle, and the party is a single individual (i.e.
+a single occupant vehicle); but in other cases, a party could consist of several individuals (i.e., a driver and one or more passengers).
+Other times, the individual is the traffic unit, for instance a pedestrian or a bicyclist.
+An opponent is a different party that is involved in the same collision.}
 
@@
 
->*"Opponent effects" are the effects on outcome severity of the attributes of the opposite party in the accident. For clarity, we now define these terms earlier in the paper.*
+>\textcolor{blue}{"Opponent effects" are the effects on outcome severity of the attributes of the opposite party in the accident. As noted above, we now define these terms earlier in the paper.}
 
@@
 
->*We changed "participants" to "parties" and "individuals" to distinguish between groups of people who were party to a crash, and the individuals who composed those parties.*
+>\textcolor{blue}{We changed "participants" to "parties" and "individuals" to distinguish between groups of people who were party to a crash, and the individuals who composed those parties.}
 
@@
 
->*The introduction was rewritten to remove some background information that was not strictly necessary, and to more clearly state the objectives and contributions of the paper.*
+>\textcolor{blue}{Thank you for this comment.
+To address it, the introduction was substantially rewritten to remove some background information and to more clearly state the objectives and contributions of the paper.}
 
@@
 
->*We identify three basic strategies, which can be combined. This is explained in sections 3.2 to 3.4.*
+>\textcolor{blue}{We identify three basic strategies, which can be combined.
+This is explained in sections 3.2 to 3.4.
+The strategies contribute to the empirical context by systematizing the different approaches to specify a model and prepare the data for the analysis of crashes involving two parties.}
 
@@
 
->*To bring together, as in, several items that are collected in the second summation in the formula. We changed this to "gather".*
+>\textcolor{blue}{To bring together several items in the second summation in the formula. We changed the description to avoid confusion.}
 
@@
 
->*Thank you for this clarification. This has been corrected.*
+>\textcolor{blue}{Thank you for this clarification. This has been corrected.}
 
@@
 
->*Yes, they were, unless they were modelled separately.
+>\textcolor{blue}{Yes, they were, unless they were modelled separately.
 For example, Models 1 and 2 were re-estimated after subsetting the sample, so that only pedestrians were considered, or bicyclists, or light vehicles, etc.
-How could several be modelled in one model?
-By using the strategies described in the methods section.*
+We might ask why the mechanisms of injury for the two groups are different, and our answer lies, in part, in the opposite party: it is not the same thing when to pedestrians collide than when a heavy vehicle collides with a pedestrian.
+By considering the atributes of the opponent explicitly, we provide more contextual information and account for the differences in mechanisms of injury severity.
+This is done implicitly when subsetting the data, since then we consider exclusively one type of collision, say, a heavy vehicle hitting a pedestrian.}
 
@@
 
->*Data preprocessing is not as sexy as a new ranked ordered logit model with mixtures of copulas, etc.
-In fact, data preprocessing is seldom report in detail: everybody assumes that it is standard.
-As we were working on this research, what we found is that data preprocessing is as complex as data analysis and modelling, and we could not find good, complete discussions of this topic.
-Since data preprocessing is an essential component of two of the strategies discussed in the paper (data subsetting and opponent-effects), we would rather maintain this as part of the paper, unless the editor indicates otherwise.*
+>\textcolor{blue}{Data preprocessing is seldom reported in detail since most everybody assumes that it is standard.
+As we were working on this research, we found this aspect of the project is as complex as data analysis and modelling, and as important.
+Since we could not find good, complete discussions of this topic, and data preprocessing is an essential component of the strategies discussed in the paper, we would rather maintain this as part of the paper.
+We feel that this is particularly important given our push for reproducible research.}
 
@@
 
->*That is not what we said. This is the actual statement:*
+>\textcolor{blue}{The statement actually says:}
 
->>*"In terms of outcomes, we observe that virtually all fatalities occur when the opponent is a driver, and only very rarely when the opponent is a motorcyclist."*
+>>\textcolor{blue}{"In terms of outcomes, we observe that virtually all fatalities occur when the opponent is a driver, and only very rarely when the opponent is a motorcyclist."}
 
->*As seen in Table 2, where we split the outcomes by the roles of the opposite parties, the only fatalities recorded outside of Opponent: Driver where for Opponent: Motorcycle (7 fatalities).
+>\textcolor{blue}{As seen in Table 2, where we split the outcomes by the roles of the opposite parties, the only fatalities recorded outside of Opponent: Driver where for Opponent: Motorcycle (7 fatalities).
 In other words, virtually all fatalities occur when the opponent is a driver.
-This is very different from "people virtually always die when the opponent is a driver", which seems to be your odd reading of our statement.*
+This is very different from "people virtually always die when the opponent is a driver".}
 
@@
 
->*No.
-The data sample is not "high", it is large.
+>\textcolor{blue}{Overfitting is a combination of sample size and the number of parameters.
 In the present case, we have a sample of size 164,511, and the models have at most 111 coefficients: this leaves tens of thousands of degrees of freedom.
-In fact, thinking that the number of coefficients is an indication of overfitting is overly-simplistic, since overfitting is a combination of sample size and the number of parameters.
-Evidence of overfitting only can be gleaned from evaluating the performance of the model when "new" information is used.
-That is the point of backcasting: to evaluate how the models behave in an out-of-sample situation.*
+It is challenging to look at the number of coefficients and conclude from that alone that there is overfitting.
+Instead, evidence of overfitting is gleaned from evaluating the performance of the model when "new" information is used.
+That is the point of backcasting: to evaluate how the models behave in an out-of-sample situation.
+The results of our assessment strongly suggest that there is some overfitting when we take an ensemble approach, since the performance of this approach deteriorates more when backcasting.
+But this is relative: even when "overfitting" the models still produce good results according to APE and WAPE.
+The set of verification statistics shows that the ensemble models do not perform as well, but also their performance is not abysmal.}
 
@@
 
->*It might be obvious to you that Models 3 and 4 can be easily combined, but we regret to inform you that there would be perfect multicollinearity if you tried to do so.
-The point is that even an expert such as yourself may fail to see this, and we consider that presenting a framework that explains the various strategies and compares them might be worthwhile for readers who eventually read the paper.*
+>\textcolor{blue}{This is a tempting idea. Unfortunately, Models 3 and 4 cannot be combined due to perfect multicollinearity.}
 
@@
 
-*We regret that you were not persuaded by the paper.
-We have taken into account those suggestions in your report that make sense, and hope that you will be persuaded to re-read the paper with attention to maybe see its value.*
+>\textcolor{blue}{We hope that after taking into account your thoughtful suggestions you will find that the paper has improved.}
 
@@
 
->*Thank you for your positive assessment of our paper. We realize that it is information-rich, and appreciate the time and effort you put into this review.*
+>\textcolor{blue}{Thank you for your positive assessment of our paper. We realize that it is information-rich, and appreciate the time and effort you put into this review.}
 
@@
 
->*Thank you for this suggestion. We have labelled the models in the tables, which we think makes it easier to recall the main characteristics of each model, i.e., single-level, hierarchical, etc.*
+>\textcolor{blue}{Thank you for this suggestion.
+We have labelled the models in the tables, which we think makes it easier to recall the main characteristics of each model, i.e., single-level, hierarchical, etc.}
 
->*Again, thank you for your thoughtful comments and suggestions.*
+>\textcolor{blue}{Again, thank you for your thoughtful comments and suggestions.}
 
@@
 
->*Many thanks for your positive assessment of the paper.*
+>\textcolor{blue}{Many thanks for your positive assessment of the paper.}
 
@@
 
->*Thank you for this comment.
+>\textcolor{blue}{Thank you for this comment.
 We collapsed three subsections of data preprocessing into one, and edited it to make it somewhat more concise.
-We are convinced that data preprocessing is a fundamental part of the process to implement the strategy of opponent-related variables.
-The process, as the paper probably shows, is not necessarily straightforward.
-And although not as sexy as presenting a new model with yet more unobserved heterogeneity terms, is as important.
-Our preference is to keep this aspect of the process well documented and transparent, but if the editor insists, we could reduce it even more.*
+We are convinced that data preprocessing is a fundamental part of the process to implement the strategies discussed in the paper.
+The process, as the paper hopefully illustrates, is not necessarily straightforward.
+Our preference is to keep this aspect of the process well documented and transparent, which is also in line with our push for reproducible research.}
 
@@
 
->*Thanks for catching this. The sentence should say "for any given coefficient $m$", since the nesting is of individual $l$ in traffic unit $m$. This was corrected.*
+>\textcolor{blue}{Thanks for catching this. The sentence should say "for any given coefficient $m$", since the nesting is of individual $l$ in traffic unit $m$. This was corrected.}
 
@@
 
->*Done.*
+>\textcolor{blue}{Done.}
 
@@
 
->*It is true, and there is the perennial trade-off between bias and variance.
-That said, overfitting is not necessarily a bad thing, and an analyst might prefer somewhat biased predictions in exchange for lower variance.
-Here we do not take a strong stand on this issue because it really depends on the specific use of the models.*
+>\textcolor{blue}{The models actually do not do a poor job when backcasting.
+See Table 6: the Weighted Average Prediction Errors (although higher than for the nowcasting case) are still less than one percent for every model/ensemble of models.}
 
@@
 
->*We believe that using a fuller battery of indicators is a bit more challenging (it makes it somewhat more difficult to find the absolute _best_ model), but on the other hand it does give numerous new insights into how specific models perform well or not.
+>\textcolor{blue}{We believe that using a fuller battery of indicators is a bit more challenging (it makes it somewhat more difficult to find the absolute _best_ model), but on the other hand it does give numerous new insights into how specific models perform well or not.
 As an example, Model 1 produces individual-level predictions that are less biased, but also has a higher Probability of False Detection.
@@
 That might guide the choice of a model.
-The difference between the two verification statistics is that Percent Correct by Class is calculated with the hits and correct non-events, whereas the Probability of Detection is calculated using only the hits.*
+The difference between the two verification statistics is that Percent Correct by Class is calculated with the hits and correct non-events, whereas the Probability of Detection is calculated using only the hits.}
 
->*Thank you for your suggestions. We hope that you will find the revised version of the paper much improved.*
+>\textcolor{blue}{Thank you for your suggestions. We hope that you will find the revised version of the paper much improved.}

### 2020-05-16 Interactions-Crash-Severity-v1/Cover-Letter-v2.Rmd
Revisions after feedback from Hany, Mark, and Saiedeh. Resubmission to Accident Analysis and Prevention.

@@
 
-We trust that you and the expert reviewers will find that our responses address in a thorough and conscientious manner the points raised with respect to the previous version of the paper.
+We hope that you and the expert reviewers will find that our responses address in a thorough and conscientious manner the points raised with respect to the previous version of the paper.
 
-Looking forward to hearing back from you in due course, we remain sincerely yours.
+Looking forward to hearing back from you in due course with your opinion about this revision, we remain sincerely yours.
 

### 2020-05-16 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v2.tex
Revisions after feedback from Hany, Mark, and Saiedeh. Resubmission to Accident Analysis and Prevention.

@@
             pdfauthor={},
-            pdftitle={An empirical assessment of strategies to model party interactions in crash severity analysis},
+            pdftitle={A systematic assessment of the use of opponent variables, data subsetting and hierarchical specification in two-party crash severity analysis},
             colorlinks=false,
@@
 
-  \title{An empirical assessment of strategies to model party interactions in
-crash severity analysis}
+  \title{A systematic assessment of the use of opponent variables, data
+subsetting and hierarchical specification in two-party crash severity
+analysis}
     \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
@@
 exercise is also worse than for any of the full sample models. Excellent
-in-sample predictions but mediocre out-of-sample predictions are often
-evidence of overfitting, as in the case of the ensemble models here.
+in-sample predictions but not as good out-of-sample predictions are
+often evidence of overfitting, as in the case of the ensemble models
+here.
 

### 2020-06-06 Interactions-Crash-Severity/Response-to-Reviewers-v3.Rmd
Changed name of folder with paper before working on revisions for v3

--- 2020-05-16 Interactions-Crash-Severity-v1/Response-to-Reviewers-v2.Rmd
@@
 author: ""
-date: "5/8/2020"
+date: "6/6/2020"
 output: pdf_document
@@
 
-We wish to thank four anonymous reviewers for their feedback on our paper.
+We wish to thank three anonymous reviewers for their feedback on the second version of our paper.
 In this letter we respond to their comments and describe the changes made to the paper in response.
@@
 
-The paper investigates three general modelling strategies to analyze crash data. Namely, use of opponent related factors, single and multi-level model approach, and different sampling approach.
+No comment to authors.
 
->\textcolor{blue}{Thank you for reading the paper and for your feedback to improve the paper.}
+>\textcolor{blue}{Thank you for reading the paper again.}
 
-The combined application of those strategies is not completely new because it is part of a general methodological approach (and usually crash data contain information about opponent factors).
-The paper, in the present form, seems more an exercise, a good exercise, but not so much sufficient as to justify a paper.
-Why not to use other models besides ordered probit (row 525) (or ordered logit as in the abstract)?
+\newpage
 
->\textcolor{blue}{Thank you for this perceptive comment.
-Indeed, the combined application of these strategies is not new, since every single one of them has been used in the past by someone or other.
-Our argument is not that the strategies are new, but rather that they are used in a haphazard fashion (we document the different types of applications in the paper).
-Moreover, we argue that there is no systematic examination of how each of these strategies perform, and therefore little guidance in the literature in terms of which one to choose in applications.
-We believe that our paper provides new information that is not available elsewhere in the literature.
-With respect to the use of other models, this has been investigated by several others, who have compared multinomial to ordered models, ordered models to ordered models with random components, etc.
-We would like to draw your attention to this paragraph in the paper (p.
-26, lines 497-510):}
+# Reviewer 3
 
->>\textcolor{blue}{"We do not report results regarding other modelling strategies.
-On the one hand, more sophisticated modelling frameworks are generally capable of improving the performance of a model.
-On the other hand, there are well-known challenges in the estimation of more sophisticated models (see Lenguerrand et al., 2006, p.
-47, for a discussion of convergence issues in models with mixed eﬀects; Mannering et al., 2016, p.
-13, for some considerations regarding the complexity and cost of estimating more complex models; and Bogue et al., 2017, p.
-27, on computational demands of models with random components).
-The additional cost and complexity of more sophisticated modelling approaches would, in our view, have greatly complicated our empirical assessment, particularly considering the large size of the sample involved in this research (a data set with over 164,000 records in the case of the full sample models).
-That said, we experimented with a model with random components using monthly subsets of data to ﬁnd that, indeed, estimation takes considerably longer, is more demanding in terms of ﬁxing potential estimation quirks, and in the end resulted in variance components that could not be reliably estimated as diﬀerent from zero (results can be consulted in the source R Notebook).
-For this reason, we choose to leave the application of more sophisticated models as a matter for future research."}
+This paper is very well done and the revisions have improved it even further.
 
->\textcolor{blue}{We tried other models, but there is no squeezing out unobserved heterogeneity where none seems to exist.}
+>\textcolor{blue}{Thank you for your suggestions to improve the paper.}
 
-To support this argument, tables 9 and 10 (confusion matrices) show a very poor performance of the models, especially for fatality;  Models Ensemble improve it a little.
+\newpage
 
->\textcolor{blue}{If you compare the average prediction errors (APEs) and weighted average prediction errors (WAPEs) of our models to those reported by Bogue, Paleti, and Balam (2017), you will notice that our models actually perform well according to these criteria.
-However, part of our argument is that a more complete comparison of the performance of models requires the use of other performance criteria, hence all the verification statistics.}
+# Reviewer 4
 
-It is curious that its ‘percent correct by class’ is generally very high (close to or greater than 90%) when numbers are so bad; it is curious too that the ‘percent correct’ is always lower than the lowest value of each ‘percent correct by class’.
+1.
+Authors mentioned that the WAPE are still less than one percent for every model/ensemble of models in Table 6.
+This is true.
+But it does not change the fact that the models are overfitting.
+If you evaluate the model performances based on other criteria, the in-sample and out-of-sample performances might be quite different.
+As such, these models might be biased.
+
+2.
+I’m very confused by the calculation process of “percent correct”.
+Not sure if it is a proper performance measure.
+As the authors mentioned, that crash data is highly imbalanced.
+In this case, there are other performance metrics which are more adapted into imbalanced data problem, such as ROC, G-mean, or F-measure.
+You can refer to the following paper
+
+Learning from class-imbalanced data: Review of methods and applications
+
+https://www.sciencedirect.com/science/article/pii/S0957417416307175
+
+>>It is curious that its ‘percent correct by class’ is generally very high (close to or greater than 90%) when numbers are so bad; it is curious too that the ‘percent correct’ is always lower than the lowest value of each ‘percent correct by class’.
 
@@
 Again, to the best of our knowledge, this more nuanced perspective is not currently available, and it provides new insights into the performance of models.}
-
-The availability of crash data  at a repository on the web to reproduce the research or to make other seems very interesting (not only for exercise) and should be encouraged.
-
->\textcolor{blue}{Thank you. We also think that this should be encouraged. We did not find in our literature review a single paper on crash severity that was reproducible.}
-
-Title is a little misleading since the role of ‘opponent’ variables is not really the main focus of the paper.
-
->\textcolor{blue}{We have changed the title based on this and a related comment by Reviewer 2.}
-
-## Minor comments
-
-Reorganize section 5.3.1 by listing at the beginning the considered measures for verification statistics. Furthermore, all the section is difficult to read.
-
->\textcolor{blue}{We followed this recommendation in the revision and reorganized the section to describe the verification statistics at the beginning.
-We also edited it, hopefully improving its readability.}
-
-At row 271 in which unit is 4536?
-
->\textcolor{blue}{It should be kg. We added the units.}
-
-At row 302, “in two different” what?
-
->\textcolor{blue}{In two different ways. This was corrected.}
-
-At row 435, CSI requires  a note.
-
->\textcolor{blue}{This paragraph was deleted as part of reorganizing the section.}
-
-At row 533, there is an s too many in ‘s(e.g. …)’
-
->\textcolor{blue}{Typo corrected.}
-
-Acknowledgment: It is quite strange that authors did not add now their acknowledgments but they reserve to do it in the ‘final’ version.
-
->\textcolor{blue}{This was an oversight caused by the practice of anonymizing papers. The revision includes the acknowledgments.}
-
->\textcolor{blue}{Thanks again for your feedback and the time invested in helping us improve this paper.
-We hope that with these revisions the contributions of the paper will be more easily appreciated.}
-
-\newpage
-
-# Reviewer 2
-
-An empirical assessment of strategies to model opponent effects in crash severity analysis – the paper examines the injury severity of crashes employing ordered model.
-
->\textcolor{blue}{Thank you for your thoughtful comments on our paper.}
-
-“Accident” should be avoided. Crash/collision should be used.
-
->\textcolor{blue}{Thank you for this comment. We have changed "accidents" to "collisions" and "crashes".}
-
-“ways to deal with the way the different parties in a crash” – this sentence is confusing. Does it refer to methodological ways to accommodate for such correlation, if present? Please clarify.
-
->\textcolor{blue}{The sentence was wordy and unclear, and it was edited for clarity. Now it reads:}
-
->>\textcolor{blue}{Numerous efforts have been undertaken to understand the factors that affect road collisions in general, and the severity of crashes in particular.
-In this literature several strategies have been proposed to model interactions between parties in a crash, including the use of variables regarding the other party (or parties) in the collision, data subsetting, and estimating models with hierarchical components.}
-
->\textcolor{blue}{Furthermore, the rewritten introduction now includes this:}
-
->>\textcolor{blue}{For the purpose of this paper, we define a party as one or more individuals travelling in a traffic unit that becomes involved in a crash.
-Sometimes the traffic unit is a vehicle, and the party is a single individual (i.e.
-a single occupant vehicle); but in other cases, a party could consist of several individuals (i.e., a driver and one or more passengers).
-Other times, the individual is the traffic unit, for instance a pedestrian or a bicyclist.
-An opponent is a different party that is involved in the same collision.}
-
-“opponent effects” – this is misleading.
-What effects are you referring to and does opponent refer to 2 vehicles involved in a crash?
-After reading the paper it was possible to infer it, but, these terms should be clearly defined to avoid confusions and to improve readability.
-
->\textcolor{blue}{"Opponent effects" are the effects on outcome severity of the attributes of the opposite party in the accident. As noted above, we now define these terms earlier in the paper.}
-
-“participants in the crash”- how can someone participate in a crash?
-
->\textcolor{blue}{We changed "participants" to "parties" and "individuals" to distinguish between groups of people who were party to a crash, and the individuals who composed those parties.}
-
-Please clearly state what are the objectives of the study and how it is contributing to existing road safety research.
-It is not at all clear from introduction/background section.
-It was not even clear whether authors are focusing on motor vehicle crashes only.
-
->\textcolor{blue}{Thank you for this comment.
-To address it, the introduction was substantially rewritten to remove some background information and to more clearly state the objectives and contributions of the paper.}
-
-The four strategies identified in the methodology section; how different approaches contribute towards the empirical context? Please provide a detailed discussion first.
-
->\textcolor{blue}{We identify three basic strategies, which can be combined.
-This is explained in sections 3.2 to 3.4.
-The strategies contribute to the empirical context by systematizing the different approaches to specify a model and prepare the data for the analysis of crashes involving two parties.}
-
-Line 125: the second part collect – what does collect represent here?
-
->\textcolor{blue}{To bring together several items in the second summation in the formula. We changed the description to avoid confusion.}
-
-Page 2, line 29 – “shamsunnahar and Eluru” should be “Yasmin and Eluru”.
-
->\textcolor{blue}{Thank you for this clarification. This has been corrected.}
-
-Line 224: Were pedestrian/bicycle involved crashes also considered in the severity analysis along with motor vehicle crashes?
-The mechanisms of injury severity for these two groups are different.
-How these could be modeled in one model?
-
->\textcolor{blue}{Yes, they were, unless they were modelled separately.
-For example, Models 1 and 2 were re-estimated after subsetting the sample, so that only pedestrians were considered, or bicyclists, or light vehicles, etc.
-We might ask why the mechanisms of injury for the two groups are different, and our answer lies, in part, in the opposite party: it is not the same thing when to pedestrians collide than when a heavy vehicle collides with a pedestrian.
-By considering the atributes of the opponent explicitly, we provide more contextual information and account for the differences in mechanisms of injury severity.
-This is done implicitly when subsetting the data, since then we consider exclusively one type of collision, say, a heavy vehicle hitting a pedestrian.}
-
-4.1.2 and 4.1.3: these procedures are standard in preparing crash severity data.
-Not sure why these are discussed in detail.
-For the data cleaning, the fundamental question is, other than deleting the cases with unknown information, how the statistical property of the dependent variable was maintained?
-
->\textcolor{blue}{Data preprocessing is seldom reported in detail since most everybody assumes that it is standard.
-As we were working on this research, we found this aspect of the project is as complex as data analysis and modelling, and as important.
-Since we could not find good, complete discussions of this topic, and data preprocessing is an essential component of the strategies discussed in the paper, we would rather maintain this as part of the paper.
-We feel that this is particularly important given our push for reproducible research.}
-
-Line 262: virtually all crashes were fatal when the opponent is a driver – this statement is fundamentally wrong.
-
->\textcolor{blue}{The statement actually says:}
-
->>\textcolor{blue}{"In terms of outcomes, we observe that virtually all fatalities occur when the opponent is a driver, and only very rarely when the opponent is a motorcyclist."}
-
->\textcolor{blue}{As seen in Table 2, where we split the outcomes by the roles of the opposite parties, the only fatalities recorded outside of Opponent: Driver where for Opponent: Motorcycle (7 fatalities).
-In other words, virtually all fatalities occur when the opponent is a driver.
-This is very different from "people virtually always die when the opponent is a driver".}
-
-The data sample is high. How is it ensured that the model is not over fitted with high number of records? Number of co-efficient in the models are an indication of over fitting.
-
->\textcolor{blue}{Overfitting is a combination of sample size and the number of parameters.
-In the present case, we have a sample of size 164,511, and the models have at most 111 coefficients: this leaves tens of thousands of degrees of freedom.
-It is challenging to look at the number of coefficients and conclude from that alone that there is overfitting.
-Instead, evidence of overfitting is gleaned from evaluating the performance of the model when "new" information is used.
-That is the point of backcasting: to evaluate how the models behave in an out-of-sample situation.
-The results of our assessment strongly suggest that there is some overfitting when we take an ensemble approach, since the performance of this approach deteriorates more when backcasting.
-But this is relative: even when "overfitting" the models still produce good results according to APE and WAPE.
-The set of verification statistics shows that the ensemble models do not perform as well, but also their performance is not abysmal.}
-
-The study design is quite confusing.
-Isn’t it obvious that Model 3 and Model 4 are generalized version of the Model 1 and 2.
-More importantly, Model 3 and 4 can be easily combined to estimate a rather genralized version of severity model.
-What is the point of this modelling exercise?
-What is new in this exercise and why one should consider it?
-
->\textcolor{blue}{This is a tempting idea. Unfortunately, Models 3 and 4 cannot be combined due to perfect multicollinearity.}
-
-The paper is merely a modelling exercise without any significant contribution to the state-of-the-art of severity model.
-The wordings of the paper are significantly misleading and confusing.
-If we are considering number of occupants in different vehicle units involved in a crash, the outcome is likely to be correlated.
-How such correlations were accommodated?
-What is the purpose of this modelling exercise rather than just estimating a set of models?
-
->\textcolor{blue}{We hope that after taking into account your thoughtful suggestions you will find that the paper has improved.}
-
-\newpage
-
-# Reviewer 3
-
-The reviewer found the paper to be exceptionally well researched.
-It is quite long and so full of details, some quite nuanced, that reading the paper and digesting its many remarks and information is time consuming.
-However, the authors are praised on giving such detail, and especially on opening the code and data for others.
-It makes this paper highly useful as a base for further research on this important topic.
-The reviewer points out that Mannering has co-authored earlier work than Wang and Kockelman, which used direct opponent effects and classified models based on opponent types, in 2004 in AAP, although that particular paper did not focus on the opponent effects.
-
->\textcolor{blue}{Thank you for your positive assessment of our paper. We realize that it is information-rich, and appreciate the time and effort you put into this review.}
-
-One small item to improve would be to name the four models with descriptive titles rather than only model 1, model 2, model 3, and model 4, which requires readers to remember which is which and makes the tables unreadable for readers browsing the paper.
-
->\textcolor{blue}{Thank you for this suggestion.
-We have labelled the models in the tables, which we think makes it easier to recall the main characteristics of each model, i.e., single-level, hierarchical, etc.}
-
->\textcolor{blue}{Again, thank you for your thoughtful comments and suggestions.}
-
-\newpage
-
-# Reviewer 4
-
-This study aims to evaluate approaches to model opponent effects in crashes involving two parties.
-A series of models are established and evaluated.
-The paper is well written and organized.
-Besides, authors make the data and code publicly available, which is great.
-Here are my comments:
-
->\textcolor{blue}{Many thanks for your positive assessment of the paper.}
-
-1. The data preprocessing section could be simplified to make the paper more concise.
-
->\textcolor{blue}{Thank you for this comment.
-We collapsed three subsections of data preprocessing into one, and edited it to make it somewhat more concise.
-We are convinced that data preprocessing is a fundamental part of the process to implement the strategies discussed in the paper.
-The process, as the paper hopefully illustrates, is not necessarily straightforward.
-Our preference is to keep this aspect of the process well documented and transparent, which is also in line with our push for reproducible research.}
-
-2. Ln 159, for any given coefficient q, you can have equation (7). I don’t quite understand it. Please clarify.
-
->\textcolor{blue}{Thanks for catching this. The sentence should say "for any given coefficient $m$", since the nesting is of individual $l$ in traffic unit $m$. This was corrected.}
-
-3. Table 3 is mentioned in section 4 but shown in section 5. It’s better to adjust the location of Table 3.
-
->\textcolor{blue}{Done.}
-
-4.
-Ln 376-377, as mentioned by authors, models in Table 6 are clearly overfitted to the 2017 dataset.
-They don’t generalize well to new data.
-To this end, the analysis based on these models might be biased.
-
->\textcolor{blue}{The models actually do not do a poor job when backcasting.
-See Table 6: the Weighted Average Prediction Errors (although higher than for the nowcasting case) are still less than one percent for every model/ensemble of models.}
-
-5.
-As shown in Table 8, authors evaluated the outcomes of models by lots of different indicators.
-Normally, it’s difficult to compare model performances based on so many indicators.
-How did you manage that?
-BTW, what’s the difference between Percent Correct by Class and Probability of Detection?
-
->\textcolor{blue}{We believe that using a fuller battery of indicators is a bit more challenging (it makes it somewhat more difficult to find the absolute _best_ model), but on the other hand it does give numerous new insights into how specific models perform well or not.
-As an example, Model 1 produces individual-level predictions that are less biased, but also has a higher Probability of False Detection.
-Does the analyst care more about reducing bias, or about reducing false detection?
-That might guide the choice of a model.
-The difference between the two verification statistics is that Percent Correct by Class is calculated with the hits and correct non-events, whereas the Probability of Detection is calculated using only the hits.}
-
->\textcolor{blue}{Thank you for your suggestions. We hope that you will find the revised version of the paper much improved.}

### 2020-06-06 Interactions-Crash-Severity/Cover-Letter-v1.Rmd
Changed name of folder with paper before working on revisions for v3

--- 2020-02-17 Interactions-Crash-Severity-v1/Cover-Letter-v1.Rmd
@@
 Please find attached our jointly authored submission entitled "An empirical assessment of strategies to model opponent effects in crash severity analysis".
-We wish to submit this paper for publication consideration in Analytical Methods in Accident Research.
+We wish to submit this paper for publication consideration in Accident Analysis & Prevention.
 
@@
 
-It is our hope that you will find the topic of the research to be a good fit to Analytical Methods in Accident Research, and that your expert reviewers will be able to provide valuable feedback to improve our research.
+It is our hope that you will find the topic of the research to be a good fit to Accident Analysis & Prevention, and that your expert reviewers will be able to provide valuable feedback to improve our research.
 

### 2020-06-06 Interactions-Crash-Severity/Interactions-Crash-Severity-Title-Page-v1.tex
Changed name of folder with paper before working on revisions for v3

--- 2020-01-16 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.tex
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{lineno}
+\linenumbers
 

### 2020-06-06 Interactions-Crash-Severity/Interactions-Crash-Severity-v1.tex
Changed name of folder with paper before working on revisions for v3

--- 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v1.tex
@@
 severity analysis}
-    \author[Some University]{Author 1\corref{Corresponding Author}}
-   \ead{author1@address.edu} 
-    \author[Another University]{Author 2}
-   \ead{author2@address.edu} 
-      \address[Some University]{Address}
-    \address[Another University]{Address}
+    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
+   \ead{paezha@mcmaster.ca} 
+    \author[Louisiana State University]{Hany Hassan}
+   \ead{hassan1@lsu.edu} 
+    \author[McMaster University]{Mark Ferguson}
+   \ead{fergumr@mcmaster.ca} 
+    \author[McMaster University]{Saiedeh Razavi}
+   \ead{razavi@mcmaster.ca} 
+      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
+University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
+    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
+University, Baton Rouge, Louisiana, USA 70803}
 
@@
 file is an R Markdown document. All code and data necessary to reproduce
-the analysis are available from the following anonymous Drive folder:
+the analysis are available from the following GitHub repository:
 \end{quote}
@@
 \begin{quote}
-https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV
-\end{quote}
-
-\begin{quote}
-The source files, code, and data will be publicly available in a GitHub
-repository upon acceptance of the paper for publication
+https://github.com/paezha/Modelling-Participant-Interactions-in-Crash-Severity
 \end{quote}

### 2020-06-07 Interactions-Crash-Severity/Response-to-Reviewers-v3.Rmd
v3 Revisions after second round of reviews.

@@
 ---
+
+```{r include=FALSE}
+library(tidyverse)
+library(kableExtra)
+```
 
@@
 
->\textcolor{blue}{Thank you for your suggestions to improve the paper.}
+>\textcolor{blue}{Thank you for your earlier suggestions to improve the paper.}
 
@@
 
-1.
 Authors mentioned that the WAPE are still less than one percent for every model/ensemble of models in Table 6.
@@
 
-2.
-I’m very confused by the calculation process of “percent correct”.
-Not sure if it is a proper performance measure.
+>\textcolor{blue}{We recognize that the models might be biased, although the amount of bias is unknowable in this case since we do not know the true data generation process.
+See the following in the concluding remarks:}
+
+>>\textcolor{blue}{"There was some evidence that subsetting the sample can improve the results in some cases (e.g., when modelling the severity of crashes involving active travelers or motorcyclists), but possibly at the risk of overfitting the process.
+It is well known that overfitting can increase the accuracy of in-sample predictions at the expense of bias in out-of-sample predictions.
+Alas, since the true data generating process is unknowable in this empirical research, it is not possible to assess the extent of estimator bias."}
+
+I’m very confused by the calculation process of “percent correct”. Not sure if it is a proper performance measure.
+
+>\textcolor{blue}{Thank you for this comment. We suspect that the confusion is caused by the differences in terminology. The terminology we are using is as follows:}
+
+```{r example-confusion-matrix, echo=FALSE}
+data.frame(Predicted = c("Yes",
+                        "No",
+                        "Marginal Total"),
+           Yes = c("Hit", "Miss", "Observed Yes"),
+           No = c("False Alarm", "Correct Non-event", "Observed No"),
+           Margin = c("Predicted Yes", "Predicted No", " ")) %>%
+  kable(booktabs = TRUE,
+        align = c("l", "c", "c", "c"),
+        col.names = c("Predicted", "Yes", "No", "Marginal Total"),
+        caption = "\\label{tab:confusion-matrix}Example of a two-by-two confusion matrix") %>%
+  kable_styling(font_size = 7, latex_options =  c("striped", "hold_position")) %>%
+  add_header_above(c(" " = 1, "Observed" = 2, " " = 1))
+```
+
+>\textcolor{blue}{An alternative terminology is used by He and Garcia in their 2009 paper "Learning from Imbalanced Data" (see Fig. 9 in their paper):}
+
+```{r example-confusion-matrix-2, echo=FALSE}
+data.frame(Predicted = c("Yes",
+                        "No",
+                        "Marginal Total"),
+           Yes = c("True Positive (TP)", "False Positive (FP)", "P_c"),
+           No = c("False Negative (FN)", "True Negative (TN)", "N_c")) %>%
+  kable(booktabs = TRUE,
+        align = c("l", "c", "c"),
+        col.names = c("Hypothesis", "Yes", "No"),
+        caption = "\\label{tab:confusion-matrix}Example of a two-by-two confusion matrix") %>%
+  kable_styling(font_size = 7, latex_options =  c("striped", "hold_position")) %>%
+  add_header_above(c(" " = 1, "True" = 2))
+```
+
+>\textcolor{blue}{Percent correct in our paper is simply the accuracy of the forecast, in the terminology of He and Garcia (see formula 13 in their paper):}
+
+$$
+PC = \frac{\text{Hit} + \text{Correct Non-event}}{\text{Observed Yes} + \text{Observed No}} = Accuracy = \frac{\text{TP} + \text{TN}}{P_c + N_c}
+$$
+
 As the authors mentioned, that crash data is highly imbalanced.
 In this case, there are other performance metrics which are more adapted into imbalanced data problem, such as ROC, G-mean, or F-measure.
-You can refer to the following paper
 
-Learning from class-imbalanced data: Review of methods and applications
+>\textcolor{blue}{Other performance metrics for imbalanced data in He and Garcia include the Precision (formula 14 in He and Garcia):}
+
+$$
+Precision = \frac{TP}{TP + TF}
+$$
+
+>\textcolor{blue}{The Precision is equivalent to the Post-Agreement, which we don't report, but that is the complement of the False Alarm Ratio, which we do report:}
+
+$$
+PA = \frac{\text{Hit}}{\text{Hit} + \text{False Alarm}} = 1 - FAR
+$$
+
+>\textcolor{blue}{Another example: Recall in He and Garcia is our Probability of Detection:}
+
+$$
+Recall = \frac{TP}{TP + FN} = POD = \frac{\text{Hit}}{\text{Hit} + \text{Miss}}
+$$
+
+>\textcolor{blue}{As you can see, the metrics that we use are suitable for imbalanced data.
+Other measures, such as ROC, G-mean, or F-measure, represent different ways of combining the same metrics (Hits, False Alarm, Miss, Correct Non-event), similar to the Skill Scores that we report.
+Hopefully this will clear the matter.}
+
+You can refer to the following paper. Learning from class-imbalanced data: Review of methods and applications
 
@@
 
->>It is curious that its ‘percent correct by class’ is generally very high (close to or greater than 90%) when numbers are so bad; it is curious too that the ‘percent correct’ is always lower than the lowest value of each ‘percent correct by class’.
-
->\textcolor{blue}{Percent correct by class is heavily influenced by the dominant class (see note in Table 8).
-It is calculated as the sum of total hits and correct non-events divided by the number of cases.
-Accordingly (see for example Table 9, Model 1):}
-
-Outcome | Pred. No Injury | Pred. Injury | Pred. Fatality |
---- | --- | --- | --- |
-No Injury | 50652 | 22503 | 150 |
-Injury | 28232 | 62121 | 797 |
-Fatality | 2 | 51 | 3 |
-
->\textcolor{blue}{The percent correct by class (No Injury) is calculated as follows:}
-
-$$
-PC_{No-Injury} = \frac{50652 + 62121 + 797 + 51 + 3}{164511} = 0.6907 = 69.07\%
-$$
-
->\textcolor{blue}{In the above, $50652$ is the number of total hits, and $62121 + 797 + 51 + 3$ is the number of correct non-events. Compare to the percent correct by class (Fatality):}
-
-$$
-PC_{Fatality} = \frac{3 + 50652 + 22503 + 28232 + 62121}{164511} = 0.9939 = 99.39\%
-$$
-
->\textcolor{blue}{In the above, $3$ is the number of total hits, and $50652 + 22503 + 28232 + 62121$ is the number of correct non-events.
-It is not surprising that the value is so high: in fact, the model correctly predicts non-fatalities most of the time!
-However, when we look at bias ($B$) we see that fatalities are severely underpredicted.
-This is despite APE and WAPE statistics that are similar or better than those reported in the literature.
-Our argument is that the use of verification statistics provides a more nuanced view of the performance of the models.
-Again, to the best of our knowledge, this more nuanced perspective is not currently available, and it provides new insights into the performance of models.}
+>\textcolor{blue}{Thank you for bringing this paper to our attention. We now cite it (as well as He and Garcia) in the concluding remarks.}

### 2020-06-07 Interactions-Crash-Severity/Cover-Letter-v3.Rmd
v3 Revisions after second round of reviews.

--- 2020-05-16 Interactions-Crash-Severity-v1/Cover-Letter-v2.Rmd
@@
 
-We would like to express our gratitude to four anonymous reviewers for their generous reviews of our paper.
+We would like to express our gratitude to three anonymous reviewers for their generous reviews of our paper.
 Their comments have helped us to improve the paper.

### 2020-06-07 Interactions-Crash-Severity/Interactions-Crash-Severity-v3.tex
v3 Revisions after second round of reviews.

--- 2020-05-16 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v2.tex
@@
 \multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
-\hspace{1em}No Injury & 50652 & 22504 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
-
-\hspace{1em}Injury & 28232 & 62120 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
+\hspace{1em}No Injury & 50652 & 22503 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\
+
+\hspace{1em}Injury & 28232 & 62121 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
-\hspace{1em}No Injury & 51531 & 17137 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
-
-\hspace{1em}Injury & 27355 & 67514 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
+\hspace{1em}No Injury & 51530 & 17136 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\
+
+\hspace{1em}Injury & 27356 & 67515 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 3. Hierarchical: Traffic unit}}\\
-\hspace{1em}No Injury & 51103 & 17296 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
-
-\hspace{1em}Injury & 27783 & 67338 & 868 &  & \textcolor{black}{72.05} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7953} & \textcolor{black}{0.2985} &  &  & \\
-
-Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4357}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
+\hspace{1em}No Injury & 51102 & 17297 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\
+
+\hspace{1em}Injury & 27784 & 67337 & 868 &  & \textcolor{black}{72.04} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7952} & \textcolor{black}{0.2985} &  &  & \\
+
+Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4356}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
-\hspace{1em}No Injury & 51574 & 17316 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
-
-\hspace{1em}Injury & 27312 & 67336 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.128}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
+\hspace{1em}No Injury & 51574 & 17317 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\
+
+\hspace{1em}Injury & 27312 & 67335 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.128}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 1 Ensemble. Single-level/No opponent}}\\
-\hspace{1em}No Injury & 34664 & 16434 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1952} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\
-
-\hspace{1em}Injury & 27749 & 67120 & 829 &  & \textcolor{black}{69.35} & \textcolor{red}{1.1452} & \textcolor{black}{0.5985} & \textcolor{red}{0.4512} & \textcolor{black}{0.8032} & \textcolor{black}{0.2986} &  &  & \\
+\hspace{1em}No Injury & 34663 & 16435 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1953} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\
+
+\hspace{1em}Injury & 27750 & 67119 & 829 &  & \textcolor{black}{69.35} & \textcolor{red}{1.1452} & \textcolor{black}{0.5985} & \textcolor{red}{0.4512} & \textcolor{black}{0.8032} & \textcolor{black}{0.2986} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
-\hspace{1em}No Injury & 35498 & 16204 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
-
-\hspace{1em}Injury & 26913 & 67379 & 828 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
-
-Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
+\hspace{1em}No Injury & 35499 & 16203 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\
+
+\hspace{1em}Injury & 26912 & 67380 & 828 &  & \textcolor{black}{70.09} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\
+
+Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3787}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
-\hspace{1em}No Injury & 61683 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
-
-\hspace{1em}Injury & 35172 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3684} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
+\hspace{1em}No Injury & 61684 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\
+
+\hspace{1em}Injury & 35171 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3683} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
-\hspace{1em}No Injury & 62736 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5318}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
-
-\hspace{1em}Injury & 34124 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
-
-Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4373}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4336}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
+\hspace{1em}No Injury & 62735 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5317}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\
+
+\hspace{1em}Injury & 34125 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\
+
+Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4372}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4335}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
-\hspace{1em}No Injury & 62788 & 21247 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
-
-\hspace{1em}Injury & 34071 & 80331 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
-
-Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4354}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
+\hspace{1em}No Injury & 62788 & 21246 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\
+
+\hspace{1em}Injury & 34071 & 80332 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5878} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\
+
+Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4355}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 2 Ensemble. Single-level/Opponent attributes}}\\
-\hspace{1em}No Injury & 43485 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.446} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
-
-\hspace{1em}Injury & 33954 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
-
-Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.355}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
+\hspace{1em}No Injury & 43486 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.4461} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\
+
+\hspace{1em}Injury & 33953 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\
+
+Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3551}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
 \cmidrule{1-14}
@@
 \multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
-\hspace{1em}No Injury & 43526 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8217} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
-
-\hspace{1em}Injury & 33915 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\
+\hspace{1em}No Injury & 43527 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8218} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\
+
+\hspace{1em}Injury & 33914 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\
 
@@
 \multicolumn{14}{l}{\textbf{Model 4 Ensemble. Hierarchical: Opponent}}\\
-\hspace{1em}No Injury & 43561 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5624} & \textcolor{black}{0.3174} &  &  & \\
-
-\hspace{1em}Injury & 33875 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
+\hspace{1em}No Injury & 43560 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5623} & \textcolor{black}{0.3174} &  &  & \\
+
+\hspace{1em}Injury & 33876 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\
 
@@
 when modelling the severity of crashes involving active travelers or
-motorcyclists), but possibly at the risk of overfitting the process and
-leading to worse out-of-sample performance. In this paper we did not
-compare individual models in our ensemble approach, but we suggest that
-this is an avenue for future research.
+motorcyclists), but possibly at the risk of overfitting the process. It
+is well known that overfitting can increase the accuracy of in-sample
+predictions at the expense of bias in out-of-sample predictions. Alas,
+since the true data generating process is unknowable in this empirical
+research, it is not possible to assess the extent of estimator bias. It
+is also worthwhile noting that in this paper we did not compare
+individual models in our ensemble approach, but we suggest that this is
+an avenue for future research.
 
@@
 namely Fatality. It would be worthwhile to further investigate
-approaches for so-called imbalanced learning, a task where Torrao et
-al.~(2014) have already made some headway.
+approaches for so-called imbalanced learning, a task that has received
+attention in the machine learning community (e.g., Haixiang et al.,
+2017; He and Garcia, 2009), and where Torrao et al.~(2014) have already
+made some headway in crash severity analysis.
 
@@
 
+\leavevmode\hypertarget{ref-Haixiang2017learning}{}%
+Haixiang, G., Yijing, L., Shang, J., Mingyun, G., Yuanyue, H., Bing, G.,
+2017. Learning from class-imbalanced data: Review of methods and
+applications. Expert Systems with Applications 73, 220--239.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.eswa.2016.12.035}{https://doi.org/10.1016/j.eswa.2016.12.035}
+
 \leavevmode\hypertarget{ref-Haleem2013effect}{}%
@@
 doi:\href{https://doi.org/10.1016/j.jsr.2013.04.002}{10.1016/j.jsr.2013.04.002}
+
+\leavevmode\hypertarget{ref-He2009learning}{}%
+He, H., Garcia, E.A., 2009. Learning from imbalanced data. IEEE
+Transactions on knowledge and data engineering 21, 1263--1284.
 
